### 2025-01-10 scripts/08.figures_markdown/Fig2_GRCCAres.Rmd
All analysis code

---
title: "Figure 2: GRCCA results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig2_GRCCAres/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig2_GRCCAres/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load DE res (to compare to GRCCA)
load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
```

### A | Model results

```{r Fig2a.plot}
fig2a <- df_results %>% 
    mutate(significant = ifelse(pval == 0.001, "yes", "no")) %>% 
    ggplot(aes(x = factor(varx*100), y = -log10(pval))) +
    geom_segment(aes(y = 0, yend = -log10(pval))) +
    geom_point(aes(size = correl, shape = significant, fill = -log10(pval))) +
    geom_hline(yintercept = 0, color = "darkgrey") +
    scale_shape_manual(values = c("yes" = 23, "no" = 21), guide = "none") +
    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0, 3), guide = "none") +
    guides(size = guide_legend(title = "X-Y LV correlation", override.aes = list(shape = 21))) +
    coord_cartesian(clip = "off") +
    labs(x = "X variance explained (%)", y = "-log10(model p-value)",
         title = "A | Optimal variance explained \nby model"
    ) +
    theme(
        legend.position = c(0.075, 0.6),
        legend.key.size = unit(0.1, "cm")
    )
```

```{r Fig2a.save, fig.width = 2.5}
fig2a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.optimal_var_explained", .x), width = fig_width, height = fig_height)
)
```

### B | Covariate (Y) structure correlations

```{r Fig2b.plot}
fig2b <- df_y_res %>% 
    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(covariate, pearsons_r))) +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_text(aes(label = significant), size = 6, vjust = 0.75, hjust = -1.5, color = "black") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    xlim(c(-1, 1)) +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[y] * ")"), 
         title = "B | Covariate structure correlations") +
    theme(legend.position = c(0.68, 0.35), 
          legend.box = "horizontal",
          #legend.justification = "center",
          #legend.key.size = unit(0.2, 'cm')
          legend.key.height = unit(0.29, "cm"),
          legend.key.width = unit(0.30, "cm")
    )
```

```{r Fig2b.save, fig.width = 4.0}
fig2b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.covariate_structCor", .x), width = 4.0, height = fig_height)
)
```

### C | Top n genes (X) by structure correlation

```{r Fig2c.prep}
df_fig2c <- df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) >= 2) %>%
    top_n(n = 20, wt = abs(pearsons_r)) %>%
    mutate(direction = ifelse(pearsons_r < 0, "2neg", "1pos"))
```

```{r Fig2c.plot}
fig2c <- df_fig2c %>% 
    ggplot(aes(x = pearsons_r, y = reorder(gene_symbol, pearsons_r))) +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    facet_wrap(vars(direction), ncol = 1, scales = "free_y") +
    force_panelsizes(rows = c(16, 4)) +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         limits = c(-0.57, 0.57), guide = "none") +
    scale_size_continuous(range = c(2, 5)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    xlim(c(-1, 1)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[x] * ")"), 
         title = "C | Top 20 genes by \nstructure correlation"
    ) +
    theme(strip.text = element_blank(), 
          panel.spacing = unit(2, "lines"),
          legend.position = c(0.05, 0.7), 
          legend.box = "horizontal",
          legend.key.size = unit(0.35, "cm")
    )
```

```{r Fig2c.save, fig.height = 4}
fig2c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.gene_structCor", .x), width = fig_width, height = 4)
)
```

### D | Expression directions of top positively- and negatively-weighted genes

```{r Fig2d.prep}
top_pos_gene <- df_x_res %>% dplyr::filter(pearsons_r > 0) %>% arrange(-abs(pearsons_r)) %>% head(1) %>% pull(ensembl_gene_id) # BCL7A
top_neg_gene <- df_x_res %>% dplyr::filter(pearsons_r < 0) %>% arrange(-abs(pearsons_r)) %>% head(1) %>% pull(ensembl_gene_id) # CFAP46
df_fig2d <- df_vsd_regress %>% 
    dplyr::select(sample, all_of(c(top_neg_gene, top_pos_gene))) %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expr") %>% 
    left_join(df_ensembl_to_symbol) %>% 
    mutate(dx = case_when(
        str_detect(sample, "control") ~ "Control",
        str_detect(sample, "bipolar") ~ "BD",
        str_detect(sample, "mdd") ~ "MDD",
        str_detect(sample, "schizo") ~ "SCZ"
    ) %>% 
        factor(levels = names(dx_colors)))
```

```{r Fig2d.stats}
load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
df_de_res %>% dplyr::filter(ensembl_gene_id %in% c(top_pos_gene, top_neg_gene))
```

```{r Fig2d.plot}
fig2d <- df_fig2d %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.5) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "darkgrey") +
    facet_wrap(vars(gene_symbol), scales = "free_y", nrow = 1) +
    #force_panelsizes(cols = c(1, 1)) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    labs(x = NULL, y = "Corrected expression value",
         title = "D | Gene expression values") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

```{r Fig2d.save, fig.width = 4.0}
fig2d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "D.top_gene_expression_values", .x), width = 4, height = fig_height)
)
```

### E | GRCCA structure correlations (r) as they relate to expression values

```{r Fig2e.prep}
df_fig2e <- df_vsd_regress %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expression_value") %>% 
    mutate(dx = case_when(
        str_detect(sample, "control") ~ "Control",
        str_detect(sample, "bipolar") ~ "BD",
        str_detect(sample, "mdd") ~ "MDD",
        str_detect(sample, "schizo") ~ "SCZ"
    ) %>% factor(levels = names(dx_colors))) %>% 
    group_by(ensembl_gene_id, dx) %>% 
    summarise(mean_expr = mean(expression_value)) %>% 
    left_join(df_x_res)
```

```{r Fig2e.plot}
fig2e <- df_fig2e %>% 
    ggplot(aes(x = mean_expr, y = pearsons_r)) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(color = "black", size = 0.25) +
    geom_smooth(aes(color = dx), method = "lm", linewidth = 0.75) +
    stat_cor(aes(label = after_stat(r.label), color = dx), label.sep = "\n", size = 3,
             label.y.npc = "top", hjust = -0.1, vjust = -1.75, face = "bold") +
    facet_wrap(vars(dx), nrow = 1) +
    scale_color_manual(values = dx_colors) +
    scale_x_continuous(breaks = c(-0.4, 0, 0.4)) +
    coord_cartesian(clip = "off") +
    labs(x = "Mean expression (corrected)", y = expression("Structure correlation (" * italic(r)[x] * ")"),
         title = "E | Relationship between gene \nstructure correlation & expression"
    ) +
    theme(legend.position = "none")
```

```{r Fig2e.save, fig.width = 4.0}
fig2e
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "E.gene_strucCor_vs_Expr", .x), width = 4.0, height = fig_height)
)
```

### 2025-01-10 scripts/08.figures_markdown/archive/FigS11_rareVariants.Rmd
All analysis code

---
title: "Figure S11: Role of rare variant enrichment"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Load data & functions
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
supplement_figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/supplement/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
#analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig4_characterizeGRCCA/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load DGE data
load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res (generated in A.DGE_analysis.R)

## Load GRCCA data
load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res

## Pull rare variants for plots
rare_variants <- benchmarking_lists[["SCZ (rare)"]]
```

```{r setup_labels}
## Create dataframe of labels for rare variants on plots
df_rare_variant_labs <- df_x_res %>% 
    filter(ensembl_gene_id %in% rare_variants) %>% 
    left_join(df_de_res) %>% 
    mutate(
        p_adj_lab = ifelse(p_adj < 0.01, format(p_adj, digits = 3), round(p_adj, 2)),
        GRCCA_lab = paste0("R=", round(pearsons_r, 2), "; FDR=", p_adj_lab),
        DGE_lab = paste0("L2FC=", round(log2fold_change*10, 2), "; p=", round(pvalue, 2))
    ) %>% 
    dplyr::select(ensembl_gene_id, gene_symbol, GRCCA_lab, DGE_lab)
```

### A | Expression direction of rare variants

```{r FigS11a.prep}
## Set factor levels for rare variant genes using structure correlation
rare_variant_order <- df_x_res %>% 
    filter(ensembl_gene_id %in% rare_variants) %>% 
    arrange(-pearsons_r) %>% 
    pull(gene_symbol)

## Pull normalized and corrected expression values for each rare variant-associated gene
df_figS11a <- df_vsd_regress %>% 
    dplyr::select(sample, any_of(rare_variants)) %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expr") %>% 
    left_join(df_ensembl_to_symbol, by = join_by(ensembl_gene_id)) %>% 
    left_join(df_rare_variant_labs, by = join_by(ensembl_gene_id, gene_symbol)) %>% 
    mutate(
        dx = case_when(
            str_detect(sample, "control") ~ "Control",
            str_detect(sample, "bipolar") ~ "BD",
            str_detect(sample, "mdd") ~ "MDD",
            str_detect(sample, "schizo") ~ "SCZ"
        ) %>% 
            factor(levels = names(dx_colors)),
        gene_symbol = factor(gene_symbol, levels = rare_variant_order)
    )
```

```{r FigS11a.plot}
figS11a <- df_figS11a %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.3) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "black", lty = 2) +
    geom_text(aes(label = GRCCA_lab), x = Inf, y = -Inf, vjust = -0.5, hjust = 1,
              check_overlap = TRUE, size = 2.75) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    facet_wrap(vars(gene_symbol), scales = "free_y", nrow = 3) +
    labs(x = NULL, y = "Corrected expression value",
         title = "A | Rare variant-associated gene expression values") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 0.9),
        strip.text = element_text(size = 9)
    )
```

```{r FigS11a.save, fig.height = 4, fig.width = 5}
figS11a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.rareVariant_gene_expression_values", .x), width = 5, height = 4)
)
```

### B | Rare variants in GRCCA structure correlation and DGE L2FC distributions

```{r FigS11b.prep}
## Create df of all struct cor & L2FC for each gene
df_figS11b <- df_x_res %>% 
    left_join(df_de_res, by = join_by(ensembl_gene_id, gene_symbol)) %>%

    # pivot DGE and GRCCA stats for facet wrapping
    pivot_longer(cols = c(pearsons_r, log2fold_change), names_to = "analysis", values_to = "value") %>% 
    mutate(analysis = ifelse(analysis == "pearsons_r", "GRCCA structure correlation", "DGE L2FC")) %>%

    # identify significant genes in each analysis
    mutate(significant = case_when(
        analysis == "GRCCA" & p_adj < 0.05 ~ 1,
        analysis == "DGE" & pvalue < 0.05 ~ 1,
        TRUE ~ 0
    ) %>% 
        factor
    ) %>% 
    arrange(significant, abs(value))

## Create separate df that only contains points to plot
df_figS11b_points <- df_figS11b %>% 
    dplyr::filter(ensembl_gene_id %in% rare_variants)

## Plotting specificities
scale_max <- df_x_res %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)
```

```{r FigS11b.function}
## Write function to plot distributions
f_plot_gene_points_distribution <- function(df, df_points) {
    df %>%
        ggplot(aes(x = value)) +
        geom_density() +
        geom_vline(xintercept = 0, color = "darkgrey") +
        geom_point(data = df_points, 
                   mapping = aes(x = value, y = -1, fill = value, color = significant), 
                   pos = jitter_pos, shape = 21, size = 2
        ) +
        geom_text_repel(data = df_points,
                        mapping = aes(x = value, y = -1, label = gene_symbol),
                        pos = jitter_pos, 
                        size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
        ) +
        #facet_wrap(vars(analysis), scales = "free") +

        scale_fill_gradientn(colors = gene_weight_color_scale, guide = "none") +
        scale_color_manual(values = c("gray", "black"), guide = "none") +
        labs(x = NULL, y = NULL, 
             title = NULL) +
        theme(axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),
              axis.line.y = element_blank()
        )
}
```

```{r FigS11b.plot}
f_plot_gene_points_distribution(df = df_figS11b %>% filter(analysis == "DGE L2FC"), 
                                df_points = df_figS11b_points %>% filter(analysis == "DGE L2FC")
) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.27, 0.27), guide = "none")

```

```{r FigS11b.save, fig.height = 2, fig.width = 4}
figS11b
```

### 2025-01-10 scripts/08.figures_markdown/archive/FigS12_GRCCAvsRCCA.Rmd
All analysis code

---
title: "Figure S12: Compare GRCCA and RCCA results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/FigS11_GRCCAvsRCCA/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/FigS11_GRCCAvsRCCA/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Also load GRCCA results for comparison
GRCCAres_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig3_GRCCAres/")
GRCCAenrich_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig4_characterizeGRCCA/")
for (dir in c(GRCCAres_dir, GRCCAenrich_dir)) {
    objects <- list.files(dir)
    objects <- objects[!str_detect(objects, "null|archive")]
    for (obj in objects){
      print(paste0("loading ", obj, "...."))
      load(paste0(dir, obj))
    }
}
```

### A | Correlate GRCCA and RCCA covariate (y) structure correlations

```{r FigS11a.plot}
figS11a <- df_y_res_all %>% 
    pivot_wider(id_cols = covariate, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(fill = rcca + grcca), shape = 21, size = 2.5) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    geom_text_repel(aes(label = covariate), size = 2.5, max.overlaps = 20, box.padding = 1.25,
                     nudge_y = 0.3, force = 150) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.84, 1.84), guide = "none") +
    labs(title = "A | Covariate structure \ncorrelations",
         x = "RCCA", y = "GRCCA")
```

```{r FigS11a.save}
figS11a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.covariate_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### B | Correlate GRCCA and RCCA gene (x) structure correlations

```{r FigS11b.plot}
figS11b <- df_x_res_all %>% 
    pivot_wider(id_cols = ensembl_gene_id, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(color = rcca + grcca), size = 1) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    stat_cor(label.sep = "\n") +
    #geom_label_repel(aes(label = gene_symbol), size = 3, max.overlaps = 50, box.padding = 1.5) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-1.14, 1.14), guide = "none") +
    labs(title = "B | Gene structure \ncorrelations",
         x = "RCCA", y = "GRCCA")
```

```{r FigS11b.save}
figS11b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.gene_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### C | Venn diagram of significant gene overlap (made in PowerPoint)

### D | RCCA results risk gene enrichment

```{r FigS11d.prep}
df_figS11d <- df_rcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))
```

```{r FigS11d.plot}
figS11d <- df_figS11d %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES)) +
    geom_text(aes(label = sig), size = 6, color = "white", vjust = 0.75) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradient(low = "white", high = "#67001F") +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "D | RCCA risk gene \nenrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
```

```{r FigS11d.save}
figS11d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
)
```

### E | GRCCA vs RCCA hypergeometric overlap across structure correlation thresholds

```{r FigS11e.plot}
df_rcca_grcca_hypergeometric_benchmark %>%

    ggplot(aes(x = threshold, y = -log10(p_value))) +
    geom_hline(aes(yintercept = -log10(0.05)), color = "gray") +
    #geom_vline(aes(xintercept = length(sig_rcca_genes) / length(ensembl_ids)), lty = 2, color = "#00BFC4") +
    #geom_vline(aes(xintercept = length(sig_grcca_genes) / length(ensembl_ids)), lty = 2, color = "#F8766D") +
    geom_point(aes(color = analysis)) +
    geom_line(aes(group = analysis, color = analysis)) +
    #annotate(geom = "text", label = "GRCCA sig θ", x = 0.04, y = 3.1, size = 3) +
    #annotate(geom = "text", label = "RCCA sig θ", x = 0.13, y = 3.1, size = 3) +

    facet_wrap(vars(benchmark_list), scales = "free_y") +
    #facet_grid2(benchmark_type ~ benchmark_list, independent = "all", scales = "free") +
    labs(title = "Hypergeometric overlaps with risk gene lists across structure correlation thresholds",
         x = "Proportion of genes included (ordered by structure correlation)", y = "-log10(hypergeometric p-value)") +
    theme(legend.position = "top",
          legend.justification = "right")
```

### 2025-01-10 scripts/08.figures_markdown/archive/FigS2_drugMCA.Rmd
All analysis code

---
title: "Figure S2: Drug Multiple Correspondence Analysis (MCA)"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/supplement/")

source(paste0(base_dir, "sscripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### A | Matrix of drug use across 185 samples

```{r FigS2a.prep}
## Format drug names for publications
df_drugs_numeric <- df_covariates_numeric %>% 
  dplyr::select(sample, dx, all_of(drug_covariates)) %>%

  # change drug names for figures
  dplyr::rename_all(~str_replace_all(.x, "anti_", "anti")) %>% 
  dplyr::rename_all(~str_replace_all(.x, "_", " ")) %>% 
  dplyr::rename(
    "nicotine/cotinine" =  "nicotine cotinine",
    "sedatives, hypnotics, anxiolitics" =  "sedative hypnotic anxiolitics",
    "major stimulants (cocaine included)" =  "major stimulants cocaine included",
    "benzodiazepines" = "benzos",
    "other psychotropic drug" = "other psychotropic drug"
  )

## Cluster samples to arrange for plot
samples_order <- df_drugs_numeric %>% 
  group_by(dx) %>% 
  nest() %>% 
  mutate(order = map(
    .x = data,
    .f = ~ .x %>% 
      column_to_rownames("sample") %>% 
      dist %>% 
      hclust %>% 
      .$order
  )) %>% 
  unnest(cols = c(data, order)) %>% 
  dplyr::select(dx, order, sample) %>% 
  mutate(dx = case_when(
    str_detect(sample, "control") ~ 1,
    str_detect(sample, "bipolar") ~ 2,
    str_detect(sample, "mdd") ~ 3,
    str_detect(sample, "schizo") ~ 4
  )) %>% 
  arrange(dx, order) %>% 
  pull(sample)

## Cluster drugs by sample use
drugs_order <- tibble(drug = df_drugs_numeric %>% dplyr::select(-sample, -dx) %>% colnames,
       order = df_drugs_numeric %>% 
         dplyr::select(-dx) %>% 
         column_to_rownames("sample") %>% 
         t() %>% 
         dist %>% 
         hclust %>% 
         .$order
) %>% 
  arrange(order) %>% 
  pull(drug)

## Format tibble with arrange rows & columns
df_figS2a <- df_drugs_numeric %>% 
    pivot_longer(3:ncol(.), names_to = "drug", values_to = "value") %>% 
    mutate(value = case_when(
        value == 1 ~ "detected/reported",
        value == 0 ~ "not detected/reported",
        is.na(value) ~ "unknown"
    ), 
    sample = factor(sample, levels = samples_order),
    drug = factor(drug, levels = drugs_order),
    dx = case_when(
        dx == 0 ~ "BD",
        dx == 1 ~ "Control",
        dx == 2 ~ "MDD",
        dx == 3 ~ "SCZ"
    ) %>% factor(levels = names(dx_colors))
    )

## Set matrix colors
matrix_colors <- c(
    "detected/reported" = "white",
    "not detected/reported" = "black",
    "unknown" = "gray"
)
```

```{r FigS2a.plot}
figS2a <- df_figS2a %>%
  ggplot(aes(x = sample)) +
  geom_tile(aes(y = drug, fill = value)) +
  geom_tile(aes(y = 0, fill = dx)) +
  geom_vline(xintercept = c(55.5, 90.5, 141.5), color = "maroon", linewidth = 0.75) +
  scale_fill_manual(values = c(matrix_colors, dx_colors),
                    na.value = "gray") +
  guides(fill = guide_legend(title = NULL)) +
  labs(y = NULL, title = "A | Drug use across 185 samples") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = c(-0.7, 0.92),
        legend.key.size = unit(0.2, "cm"),
        legend.text = element_text(size = 8),
        axis.line = element_blank()
        )
```

```{r FigS2a.save, fig.height = 2.25, fig.width = 5}
figS2a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.drug_use_matrix", .x), width = 5, height = 2.25)
)
```

### B | Variance explained of each drug MCA dimension

```{r FigS2b.plot}
figS2b <- df_eig %>% 
    mutate(dim = factor(dim),
           label = paste0(round(cum_var, 2), "%")) %>%

    ggplot(aes(x = dim)) +
    geom_col(aes(y = var, fill = var), color = "black") +
    #geom_point(aes(y = cum_var)) +
    # geom_text(aes(y = cum_var, label = label),
    #           size = 3, vjust = -1) +
    geom_hline(yintercept = 5, color = "maroon") +
    scale_fill_gradient(low = "white", high = "black", guide = "none") +
    labs(title = "B | MCA dimension \nvariance explained",
         x = "Dimension", y = "% variance explained") +
    theme(axis.text.x = element_text(angle = 45)
    )
```

```{r figS2b.save}
figS2b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.MCA_var_explained", .x), width = fig_width, height = fig_height)
)
```

### C | MCA dimension - covariate correlations

```{r FigS2c.prep}
## Set order
m_covariates <- df_covariates_numeric %>% column_to_rownames("sample") %>% t()
covariate_order <- m_covariates[hclust(dist(m_covariates))$order,] %>% rownames

## Arrange covariates and format for plotting
df_figS2c <- df_mca_covar_cor %>% 
    mutate(covariate = factor(covariate, levels = covariate_order),
           dimension = str_remove(dimension, "dim") %>% as.numeric %>% as.factor,
           #significant = ifelse(p_adj < 0.05, "yes", "no")
           significance = case_when(
               p_adj < 0.05 & p_adj > 0.01 ~ "*",
               p_adj < 0.01 & p_adj > 0.001 ~ "**",
               p_adj < 0.001 ~ "***",
               TRUE ~ ""
           )
    ) %>%

    # remove technical covariates
    filter(!(covariate %in% technical_covariates))
```

```{r FigS2c.plot}
figS2c <- df_figS2c %>% 
    ggplot(aes(x = dimension, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), linewidth = 0.5) +
    geom_text(aes(label = significance), vjust = 0.75, size = 2.5) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-1.0, 1.0)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.position = "left", title.hjust = 0.5)) +
    labs(x = "Dimension", y = NULL, 
         title = "C | MCA dimension correlations with \nknown covariates") +
    theme(legend.key.height = unit(1.5, "cm"),
          legend.key.width = unit(0.1, "cm"),
          legend.title = element_text(angle = 90),
          legend.margin = margin(l = -15)
    )
```

```{r FigS2c.save, fig.height = 5, fig.width = 5.5}
figS2c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.MCA_dimension_covariate_correlations", .x), width = 5.5, height = 5)
)
```

### 2025-01-10 scripts/08.figures_markdown/archive/TRY-transcript_GRCCA_res.Rmd
All analysis code

---
title: "Potential supplement figures: Transcript-level GRCCA results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig5_GRCCAtranscriptsRes/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WTCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### Y weights

```{r, fig.height = 3.5, fig.width = 10}
p_covariates <- df_y_res %>% 
    pivot_longer(c(z_score, pearsons_r), names_to = "metric", values_to = "value") %>% 
    mutate(metric = ifelse(metric == "z_score", "weight z-score", "structure correlation")) %>% 
    mutate(significance = case_when(
        metric == "weight z-score" & abs(value) > 2 ~ "*",
        metric == "structure correlation" & p_adj < 0.05 & p_adj > 0.01 ~ "*",
        metric == "structure correlation" & p_adj < 0.01 & p_adj > 0.001 ~ "**",
        metric == "structure correlation" & p_adj < 0.001 ~ "***",
        TRUE ~ ""
    )
    ) %>%

    ggplot(aes(x = value, y = reorder_within(covariate, value, model))) +
    geom_point(aes(size = -log10(p_adj), fill = value), shape = 21) +
    geom_vline(xintercept = 0, lty = 2) +
    geom_text(aes(label = significance), size = 4, vjust = -0.15, color = "black") +
    facet_wrap(model ~ metric, scales = "free", nrow = 1) +

    scale_fill_gradientn(colors = gene_weight_color_scale) +
    scale_size_continuous(range = c(2, 6)) +
    scale_y_reordered() +
    guides(fill = guide_colorbar(title = NULL),
           size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = NULL, 
         title = "Covariate results") +
    theme(legend.position = "bottom", 
          legend.justification = "center",
          legend.box = "horizontal",
          #legend.justification = "center",
          #legend.key.size = unit(0.2, 'cm')
          legend.key.height = unit(0.20, "cm"),
          legend.key.width = unit(1.0, "cm")
    )
p_covariates
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "covariate_weight_structCor", .x), width = 10, height = 3.5)
)
```

### X weights

```{r}
df_x_res %>% 
    pivot_wider(id_cols = c(ensembl_transcript_id, transcript_symbol), names_from = model, values_from = pearsons_r) %>%

    ggplot(aes(x = includeDim2, y = regressDim2)) +
    geom_point() +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm") +
    stat_cor()
```

### Risk gene GSEA

```{r, fig.height = 3, fig.width = 5}
p_risk_genes <- df_grcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(label = round(NES, 3)) %>%

    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES)) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(class ~ model, scales = "free", nrow = 2) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-2, 2)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
p_risk_genes
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "risk_gene_matrices", .x), width = 5, height = 3)
)
```

### GO GSEA

```{r, fig.height = 8, fig.width = 9}
p_go_gsea <- df_grcca_fgsea_go %>% 
    dplyr::filter(padj < 0.05 & model == "regressDim2") %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(padj < 0.031, term, "")) %>% 
    #mutate(label = ifelse(term %in% fig4c_labels, str_wrap(term, 25), "")) %>%

    ggplot(aes(x = padj, y = NES)) +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 3, force = 5,
                    box.padding = 0.5, min.segment.length = 0.1, max.overlaps = 30) +
    geom_hline(yintercept = 0, color = "gray") +
    facet_wrap(vars(ontology), nrow = 2) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3, 3), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.025, 0.05)) +
    coord_cartesian(clip = "off") +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    labs(x = "Enrichment FDR", y = "NES",
         title = "GRCCA transcript enrichments (regressDim2)") +
    theme(
        legend.position = "bottom",
        legend.key.size = unit(0.3, "cm")
    )
p_go_gsea
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "GO_GSEA", .x), width = 9, height = 8)
)
```

### Example transcript plot (relationship with Dim2 & relationship with SCZ)

```{r}
load(paste0(base_dir, "outputs/objects/FigS2_drugMCA/drug_MCA_results.Rdata")) # df_var_loadings, df_ind_loadings, df_eig

## Identify transcript to plot
top_transcript <- df_x_res %>% 
    filter(model == "includeDim2") %>% 
    top_n(10, pearsons_r) %>% 
    pull(ensembl_transcript_id)

## Combine transcript expression with Dim 2 scores & diagnosis
df_dim2_tx <- df_ind_loadings %>% 
    dplyr::select(sample, dim2) %>% 
    left_join(
        df_vsd_regress_filt %>% 
            dplyr::select(sample, all_of(top_transcript))
    ) %>% 
    left_join(
        df_covariates %>% dplyr::select(sample, dx)
    ) %>% 
    pivot_longer(contains("ENST"), names_to = "ensembl_transcript_id", values_to = "expr") %>% 
    left_join(df_transcript_to_gene) %>% 
    distinct() %>% 
    mutate(transcript_symbol = ifelse(is.na(transcript_symbol), ensembl_transcript_id, transcript_symbol))

## Plot correlation between Dim2 score and transcript expression
df_dim2_tx %>% 
    ggplot(aes(x = dim2, y = expr)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_cor() +
    facet_wrap(vars(transcript_symbol), scales = "free")

## Plot expression levels across diagnostic groups
df_dim2_tx %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.5) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "black", lty = 2) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    facet_wrap(vars(transcript_symbol), scales = "free_y") +
    labs(x = NULL, y = "Corrected expression value",
         title = "Transcript expression values") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9))

```

```{r}
df_dim2_txAll <- df_ind_loadings %>% 
    dplyr::select(sample, dim2) %>% 
    left_join(df_vsd_regress_filt)
# Run all correlations
df_x_res_dim2cor <- df_x_res %>% 
    mutate(
        dim2_cor = map(
            .x = ensembl_transcript_id,
            .f = ~ cor.test(df_dim2_txAll$dim2, df_dim2_txAll[[.x]])$estimate
        )
    ) %>% 
    unnest(cols = c(dim2_cor))
# Plot
df_x_res_dim2cor %>% 
    ggplot(aes(x = dim2_cor, y = pearsons_r)) +
    geom_point() +
    stat_cor()
```

### Gene & transcript correlation

```{r, fig.height = 3, fig.width = 3.5}
load(paste0(analysis_objects_dir, "../Fig3_GRCCAres/GRCCA_results.Rdata")) # df_x_res
df_x_res_genes <- df_x_res

objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Map transcript to gene and correlate
p <- df_x_res %>% 
    filter(model == "regressDim2" & !is.na(ensembl_gene_id)) %>% 
    group_by(ensembl_gene_id) %>% 
    summarise(pearsons_r = max(pearsons_r)) %>% 
    dplyr::rename("transcript_res" = "pearsons_r") %>%

    left_join(df_x_res_genes %>% 
                  dplyr::select(ensembl_gene_id, gene_symbol, pearsons_r) %>% 
                  dplyr::rename("gene_res" = "pearsons_r")
    ) %>% 
    mutate(
        risk_gene = ifelse(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, prioritized)"]] | ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]], "yes", "no"),
        label = ifelse(risk_gene == "yes", gene_symbol, "")
    ) %>% 
    arrange(risk_gene) %>%

    ggplot(aes(x = gene_res, y = transcript_res)) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_hline(yintercept = 0, color = "gray") +
    geom_point(aes(fill = gene_res + transcript_res, color = risk_gene, size = risk_gene), shape = 21) +
    geom_text_repel(aes(label = label), size = 2.5, max.overlaps = 50) +
    stat_cor() +
    geom_smooth(method = "lm", color = "black", linewidth = 0.75) +
    xlim(c(-0.6, 0.6)) +
    ylim(c(-0.6, 0.6)) +
    scale_fill_gradientn(colors = gene_weight_color_scale, guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_size_manual(values = c("yes" = 1.5, "no" = 0.5)) +
    labs(x = "Gene GRCCA structure correlations", y = "Transcript GRCCA structure correlations",
         title = "Correspondence between gene & \ntranscript GRCCA results")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0("~/Downloads/gene_transcript_GRCCA_cor", .x), width = 3.5, height = 3)
)
```

```{r}
## Risk gene transcripts
df_x_res %>% 
    filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]])

## Sig GRCCA transcripts
sig_grcca_genes <- df_x_res_genes %>% 
    filter(abs(z_score) >= 2 & p_adj < 0.05) %>% 
    pull(ensembl_gene_id)
df_x_res %>% 
    filter(ensembl_gene_id %in% sig_grcca_genes & ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]])
```

### 2025-01-10 scripts/08.figures_markdown/supplement.Rmd
All analysis code

---
title: "Supplemental figures for RLSmith Transl Psych 2024"
author: "Rachel L Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

source(paste0(base_dir, "scripts/load_WGCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load Nirmala's DEG results
df_akula_res <- read_xlsx(path = paste0(base_dir, "data/Akula2021_geneLevel_21kgenes_baseMeanGe5_cqn-Rin-Race-GC_DESeq2-pvalsCorrected_betaPriorT-lfcShrink_092518.xlsx"),
          sheet = "SczVsCtrls") %>% 
    clean_names %>% 
    dplyr::select(ensembl_gene_id, hgnc_symbol, log2fold_change, fold_change, lfc_se, stat, pvalue, padj)
```

# TABLES #

```{r TableS1}
write_xlsx(df_covariates, path = paste0(tables_dir, "TableS1.covariate_information.xlsx"))
```

```{r TableS2}
write_xlsx(list("A.GRCCA results" = df_results,
                "B.Covariate (Y) results" = df_y_res,
                "C.Gene (X) results" = df_x_res), 
           path = paste0(tables_dir, "TableS2.GRCCA_results.xlsx")
)
```

```{r TableS3}
write_xlsx(
    list(
        "A.Risk genes" = df_grcca_fgsea_benchmark %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "B.Cell-type" = df_grcca_fgsea_cell %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "C.GO" = df_grcca_fgsea_go %>% 
            dplyr::filter(padj < 0.05) %>%
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes))
    ), 
    path = paste0(tables_dir, "TableS3.GRCCA_GSEA_enrichments.xlsx")
)
```

```{r TableS4}
write_xlsx(df_de_res, path = paste0(tables_dir, "TableS4.DEGs.xlsx"))
```

```{r TableS5}
### Export enrichment results as table ###
write_xlsx(
    list(
        "A.Cell-type" = df_de_fgsea_cell %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "B.GO" = df_de_fgsea_go %>% 
            dplyr::filter(padj < 0.05) %>%
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "C.Risk genes" = df_de_fgsea_benchmark %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes))
    ), 
    path = paste0(tables_dir, "TableS5.DEG_GSEA_enrichments.xlsx")
)
```

```{r TableS6}
write_xlsx(list("A.GRCCA results" = df_results_transcripts,
                "B.Covariate (Y) results" = df_y_res_transcripts,
                "C.Gene (X) results" = df_x_res_transcripts), 
           path = paste0(tables_dir, "TableS6.GRCCA_transcript_results.xlsx")
)
```

```{r TableS7}
write_xlsx(list(
    "A. Module gene assignments" = df_modules_filt %>% left_join(df_ensembl_to_symbol) %>% dplyr::select(module, color, ensembl_gene_id, gene_symbol) %>% as.data.frame,
    "B. Module GO results" = df_mods_go %>% dplyr::filter(pvalue < 0.05 & module != 0) %>% as.data.frame
),
path = paste0(tables_dir, "TableS7.WGCNA_module_results.xlsx")
)
```

```{r TableS8}
write_xlsx(df_lm_dx %>% dplyr::select(-data) %>% dplyr::filter(p_value < 0.05),
           path = paste0(tables_dir, "TableS8.kME_analysis_significant_results.xlsx")
)
```

```{r TableS9}
write_xlsx(list("A.RCCA results" = df_results_rcca,
                "B.Covariate (Y) results" = df_y_res_rcca,
                "C.Gene (X) results" = df_x_res_rcca), 
           path = paste0(tables_dir, "TableS9.RCCA_results.xlsx")
)
```

```{r TableS10}
write_xlsx(
    list(
        "A.Cell-type" = df_rcca_fgsea_cell %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "B.GO" = df_rcca_fgsea_go %>% 
            dplyr::filter(padj < 0.05) %>%
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes)),
        "C.Risk genes" = df_rcca_fgsea_benchmark %>% 
            mutate(genes = map_chr(leadingEdge, ~ f_collapse_genes(.x, index = 30000))) %>% 
            dplyr::select(-leadingEdge) %>% unnest(cols = c(genes))
    ), 
    path = paste0(tables_dir, "TableS10.RCCA_GSEA_enrichments.xlsx")
)
```

# FIG S1: Transcript filtering by coefficient of variation #

### A | Relationship between mean and standard deviation
```{r FigS1a.plot, fig.width = 2.6}
df_mean_sd %>% 
  ggplot(aes(x = mean %>% log10, y = stdev %>% log10)) +
  geom_point(shape = 21, size = 0.75) +
  #geom_density2d(linewidth = 0.25) +
  geom_smooth(method = "lm", color = "maroon", linewidth = 0.5) +
  stat_cor(label.sep = "\n", vjust = 1, size = 3.5) +
  labs(x = "log10(mean)", y = "log10(standard deviation)",
       title = "A | Mean vs standard deviation \nin transcript raw counts")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S1A.transcript_raw_counts_mean_stdev", .x), width = 2.6, height = fig_height)
)
```

### B | Relationship between mean and coefficient of variation
```{r FigS1b.plot, fig.width = 2.6}
df_mean_sd %>% 
    ggplot(aes(x = log10(mean), y = CV %>% log10)) +
    geom_point(shape = 21, size = 0.75) +
    geom_density2d() +
    geom_smooth(method = "lm", color = "maroon", linewidth = 0.5) +
    stat_cor(label.sep = "\n", label.x.npc = "right",
             vjust = 1, hjust = 1, size = 3.5
    ) +
    labs(x = "log10(mean)", y = "log10(coefficient of variation)",
         title = "B | Mean vs coefficient of \nvariation in transcript counts")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S1B.transcript_raw_counts_mean_CV", .x), width = 2.6, height = fig_height)
)
```

### C | Cutoff for coefficient of variation
```{r FigS1c.plot}
df_mean_sd %>% 
  ggplot(aes(x = CV)) +
  geom_hline(yintercept = 0) +
  geom_point(aes(y = 3.5), position = position_jitter(0.001), alpha = 0.5, size = 0.75, shape = 21) +
  geom_density(color = "blue", fill = "transparent") +
  geom_boxplot(aes(y = -0.1), color = "blue", width = 0.2, outlier.shape = NA) +
  geom_vline(xintercept = cutoff, lty = 2, color = "maroon") +
  labs(x = "Coefficient of variation (CV)", y = NULL,
       title = "C | Coefficient of variation \nacross transcripts") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
  ) 
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S1C.CV_distribution", .x), width = fig_width, height = fig_height)
)
```

# FIG S2: Raw count preprocessing: qSV and known covariate regression #

### A | Variance explained by qSVs
```{r FigS2a.plot}
df_qsv_var_expl %>% #dplyr::count(feature)
    dplyr::filter(feature == "Proportion of Variance" & qSV %in% 1:16) %>%

    ggplot(aes(x = qSV, y = value*100)) +
    geom_col(aes(fill = value), color = "black") +
    geom_hline(yintercept = 0.02*100, color = "maroon") +
    scale_fill_gradient(low = "white", high = "black", 
                        limits = c(0, 0.06), guide = "none") +
    #scale_y_continuous(breaks = seq(0, 0.06, 0.01)) +
    coord_cartesian(clip = "off") +
    labs(y = "% variance explained",
         title = "A | qSV variance explained") +
    theme(axis.text.x = element_text(angle = 45))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S2A.qSV_variance_explained", .x), width = fig_width, height = fig_height)
)
```

### B | Correlate qSVs with known covariates

```{r FigS2b.prep}
## Define covariate order based on clustering within each type
f_define_order <- function(covariate_matrix, covariate_type = c("technical", "biological", "toxicological")) {
    if (covariate_type == "technical") {
        subset <- technical_covariates
    } else if (covariate_type == "biological") {
        subset <- biological_covariates
    } else if (covariate_type == "toxicological") {
        subset <- drug_covariates
    } else (stop("Covariate type must be technical, biological, or toxicological"))
    matrix_tmp <- covariate_matrix[subset,]
   order <- matrix_tmp[hclust(dist(matrix_tmp))$order,] %>% rownames
   return(order)
}
covariate_order <- c(
    f_define_order(m_covariates, "biological"),
    f_define_order(m_covariates, "toxicological"),
    f_define_order(m_covariates, "technical")
)
covariate_order

## Switch major_stimulants_cocaine_included and cannabinoids for the plot
swtch <- function(vec, value1, value2) {
  # Ensure the values exist in the vector
  if (!(value1 %in% vec) || !(value2 %in% vec)) {
    stop("Both values must exist in the vector.")
  }

  # Find the positions of the values
  pos1 <- which(vec == value1)
  pos2 <- which(vec == value2)

  # Swap the elements
  vec[pos1] <- value2
  vec[pos2] <- value1

  return(vec)
}
covariate_order <- swtch(covariate_order, "major_stimulants_cocaine_included", "cannabinoids")

## Create dataframe defining covariate type & order
df_figS2b <- df_covariate_qsv_cor %>% 
    mutate(significance = case_when(
        p_adj < 0.05 & p_adj > 0.01 ~ "*",
        p_adj < 0.01 ~ "**",
        TRUE ~ ""
    ),
           #covariate = factor(covariate, levels = covariate_order),
           type = case_when(
               covariate %in% technical_covariates ~ "Technical",
               covariate %in% biological_covariates ~ "Biological",
               covariate %in% drug_covariates ~ "Toxicological",
               TRUE ~ "Diagnostic group"
           ) %>% 
               factor(levels = c("Diagnostic group", "Technical", "Toxicological", "Biological"))
    ) %>% 
    filter(type != "Diagnostic group") %>% 
    arrange(type, qsv, pearsons_r) %>% 
    mutate(covariate = factor(covariate, levels = covariate_order))
```

```{r FigS2b.plot, fig.height = 6.5, fig.width = 4}
df_figS2b %>% 
    ggplot(aes(x = qsv, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), width = 0.98, height = 0.98) +
    geom_text(aes(label = significance), vjust = 0.75) +
    facet_wrap(vars(type), scales = "free", ncol = 1) +
    force_panelsizes(rows = c(length(biological_covariates), length(drug_covariates), length(technical_covariates))) +

    scale_fill_gradientn(colors = rev(brewer.rdbu(100))) +
    coord_cartesian(clip = "off") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.hjust = 0.5, title.position = "left")) +
    labs(x = "qSV", y = NULL, 
         title = "B | qSVs and known \ncovariate correlations"
    ) +
    theme(legend.position = c(-1.1, 1.0),
          legend.key.height = unit(0.6, "cm"),
          legend.key.width = unit(0.15, "cm"),
          legend.title = element_text(angle = 90),
          axis.text.x = element_text(angle = 45),
          strip.background = element_rect(fill = "white", color = "black")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S2B.qSV_covariate_correlations", .x), width = 4, height = 6.5)
)
```

### C | PC-covariate correlations before and after covariate regressions

```{r FigS2c.prep}
df_figS2c <- df_pca1_covar_cor %>% 
    mutate(regression = "Before", .before = 1) %>% 
    bind_rows(
        df_pca3_covar_cor %>% 
            mutate(regression = "After", .before = 1) 
    ) %>% 
    mutate(regression = factor(regression, levels = c("Before", "After"))) %>% 
    filter(covariate != "dx") %>% 
    mutate(covariate = factor(covariate, levels = c(covariate_order, paste0("qSV", 1:16))),
           significance = case_when(
               p_adj < 0.05 & p_adj > 0.01 ~ "*",
               p_adj < 0.01 ~ "**",
               TRUE ~ ""
           ),
           type = case_when(
               covariate %in% technical_covariates ~ "Technical",
               covariate %in% biological_covariates ~ "Biological",
               covariate %in% drug_covariates ~ "Toxicological",
               TRUE ~ "qSV"
           ) %>% 
               factor(levels = c("qSV", "Biological", "Toxicological", "Technical"))
    )
```

```{r FigS2c.plot, fig.width = 6.5, fig.height = 4}
df_figS2c %>% 
    ggplot(aes(x = pc, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), width = 0.98, height = 0.98, linewidth = 0.5) +
    geom_text(aes(label = significance), vjust = 0.75, angle = 90) +

    facet_grid2(regression ~ type, scales = "free") +
    force_panelsizes(rows = c(1, 4), cols = c(16, length(biological_covariates), length(drug_covariates), length(technical_covariates))) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-0.43, 0.43)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    coord_flip(clip = "off") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.hjust = 0.5, title.position = "top")) +
    labs(x = "PC (> 2% explained var)", y = NULL,
         title = "C | PC-covariate correlations before and after regression") +
    theme(legend.position = c(0, -0.4),
          legend.direction = "horizontal",
          legend.key.width = unit(0.6, "cm"),
          legend.key.height = unit(0.15, "cm"),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8),
          strip.background = element_rect(fill = "white", color = "black"),
          strip.clip = "off",
          panel.spacing = unit(0.1, "lines")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S2C.PCA_covariate_cor", .x), width = 6.5, height = 4)
)
```

# FIG S3: Drug Multiple Correspondence Analysis results #

### A | Matrix of drug use across 185 samples

```{r FigS3a.prep}
## Format drug names for publications
df_drugs_numeric <- df_covariates_numeric %>% 
  dplyr::select(sample, dx, all_of(drug_covariates)) #%>%

  # # change drug names for figures
  # dplyr::rename_all(~str_replace_all(.x, "anti_", "anti")) %>% 
  # dplyr::rename_all(~str_replace_all(.x, "_", " ")) %>% 
  # dplyr::rename(
  #   "nicotine/cotinine" =  "nicotine cotinine",
  #   "sedatives, hypnotics, anxiolitics" =  "sedative hypnotic anxiolitics",
  #   "major stimulants (cocaine included)" =  "major stimulants cocaine included",
  #   "benzodiazepines" = "benzos",
  #   "other psychotropic drug" = "other psychotropic drug"
  # )

## Cluster samples to arrange for plot
samples_order <- df_drugs_numeric %>% 
  group_by(dx) %>% 
  nest() %>% 
  mutate(order = map(
    .x = data,
    .f = ~ .x %>% 
      column_to_rownames("sample") %>% 
      dist %>% 
      hclust %>% 
      .$order
  )) %>% 
  unnest(cols = c(data, order)) %>% 
  dplyr::select(dx, order, sample) %>% 
  mutate(dx = case_when(
    str_detect(sample, "control") ~ 1,
    str_detect(sample, "bipolar") ~ 2,
    str_detect(sample, "mdd") ~ 3,
    str_detect(sample, "schizo") ~ 4
  )) %>% 
  arrange(dx, order) %>% 
  pull(sample)

## Cluster drugs by sample use
drugs_order <- f_define_order(m_covariates, "toxicological")

## Format tibble with arrange rows & columns
df_figS3a <- df_drugs_numeric %>% 
    pivot_longer(3:ncol(.), names_to = "drug", values_to = "value") %>% 
    mutate(value = case_when(
        value == 1 ~ "detected",
        value == 0 ~ "not detected",
        is.na(value) ~ "unknown"
    ), 
    sample = factor(sample, levels = samples_order),
    drug = factor(drug, levels = drugs_order),
    dx = case_when(
        dx == 0 ~ "BD",
        dx == 1 ~ "Control",
        dx == 2 ~ "MDD",
        dx == 3 ~ "SCZ"
    ) %>% factor(levels = names(dx_colors))
    )

## Set matrix colors
matrix_colors <- c(
    "detected" = "white",
    "not detected" = "black",
    "unknown" = "gray"
)
```

```{r FigS3a.plot, fig.height = 2.25, fig.width = 4.5}
df_figS3a %>%
  ggplot(aes(x = sample)) +
  geom_tile(aes(y = drug, fill = value)) +
  geom_tile(aes(y = 0, fill = dx)) +
  geom_vline(xintercept = c(55.5, 90.5, 141.5), color = "maroon", linewidth = 0.5) +
  scale_fill_manual(values = c(matrix_colors, dx_colors),
                    na.value = "gray") +
  guides(fill = guide_legend(title = NULL)) +
  labs(y = NULL, title = "A | Sample toxicology data") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = c(-0.85, 0.92),
        legend.key.size = unit(0.2, "cm"),
        legend.text = element_text(size = 8),
        axis.line = element_blank()
        )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3A.sample_toxicology_matrix", .x), width = 4.5, height = 2.25)
)
```

### B | Variance explained of each drug MCA dimension

```{r FigS3b.plot, fig.width = 2.1}
df_eig %>% 
    mutate(dim = factor(dim),
           label = paste0(round(cum_var, 2), "%")) %>%

    ggplot(aes(x = dim)) +
    geom_col(aes(y = var, fill = var), color = "black") +
    #geom_point(aes(y = cum_var)) +
    # geom_text(aes(y = cum_var, label = label),
    #           size = 3, vjust = -1) +
    geom_hline(yintercept = 5, color = "maroon") +
    scale_fill_gradient(low = "white", high = "black", guide = "none") +
    labs(title = "B | MCA dimension \nvariance explained",
         x = "Dimension", y = "% variance explained") +
    theme(axis.text.x = element_text(angle = 45, size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3B.MCA_variance_explained", .x), width = 2.1, height = 2)
)
```

### C | MCA variable loadings

```{r FigS3C.prep}
df_figS3c <- df_var_loadings %>% 
  pivot_longer(3:ncol(.), names_to = "dimension", values_to = "score") %>% 
  filter(use == "yes") %>% 
  mutate(drug = factor(drug, levels = drugs_order),
         dimension = str_replace(dimension, "dim", "") %>% factor(levels = 1:17),
         label = ifelse(abs(score) > 0.015, "*", NA)
  )
```

```{r FigS3C.plot, fig.width = 3.75, fig.height = 3}
df_figS3c %>% 
    ggplot(aes(x = drug, y = dimension)) +
    geom_tile(aes(fill = score)) +
    geom_text(aes(label = label), vjust = 0.75) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.065, 0.065)) +
    guides(fill = guide_colorbar(title = "Loading", title.position = "left", title.hjust = 0.5)) +
    coord_flip() +
    labs(y = "Dimension", x = NULL, 
         title = "C | Variable loadings on \n MCA dimensions") +
    theme(legend.position = c(-1.15, 0.92),
          legend.key.width = unit(0.15, "cm"),
          legend.key.height = unit(0.60, "cm"),
          legend.title = element_text(angle = 90),
          axis.text.y = element_text(size = 8.5),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3C.MCA_variable_loadings", .x), width = 3.75, height = 3)
)
```

### D | MCA dimension - covariate correlations

```{r FigS3d.prep}
## Arrange covariates and format for plotting
df_figS3d <- df_mca_covar_cor %>% 
    # remove technical covariates & dx
    filter(!(covariate %in% technical_covariates) & covariate != "dx") %>% 
    mutate(
        covariate = factor(covariate, levels = covariate_order),
        dimension = str_remove(dimension, "dim") %>% as.numeric %>% as.factor,
        #significant = ifelse(p_adj < 0.05, "yes", "no")
        significance = case_when(
            p_adj < 0.05 & p_adj > 0.01 ~ "*",
            p_adj < 0.01 ~ "**",
            TRUE ~ ""
        ),
        type = case_when(
               #covariate %in% technical_covariates ~ "Technical",
               covariate %in% biological_covariates ~ "Biological",
               covariate %in% drug_covariates ~ "Toxicological",
               TRUE ~ "Diagnostic group"
           ) %>% 
               factor(levels = c("Diagnostic group", "Biological", "Toxicological"))
    )
```

```{r FigS3d.plot, fig.width = 3.75, fig.height = 5}
df_figS3d %>% 
    ggplot(aes(x = dimension, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), linewidth = 0.5) +
    geom_text(aes(label = significance), vjust = 0.75, size = 2.5) +
    facet_wrap(vars(type), scales = "free_y", ncol = 1) +
    force_panelsizes(rows = c(length(biological_covariates), length(drug_covariates))) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-1.0, 1.0)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.position = "left", title.hjust = 0.5)) +
    labs(x = "Dimension", y = NULL, 
         title = "D | MCA dimension \ncorrelations with \nknown covariates") +
    theme(
        legend.position = c(-0.80, 1),
        legend.key.height = unit(0.6, "cm"),
        legend.key.width = unit(0.15, "cm"),
        legend.title = element_text(angle = 90),
        axis.text.y = element_text(size = 8.5),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3D.MCA_dimension_covariate_correlations", .x), width = 3.75, height = 5)
)
```

# FIG S4: WGCNA module sizes across parameter space #

### A | Soft-threshold power selection

```{r FigS4a.plot, fig.width = 5, fig.height = 2}
## Model fit of each soft thresholding power
p1 <- df_sft %>% 
    ggplot(aes(x = power, y = -sign(slope)*sft_r_sq)) +
    geom_point(aes(color = power == 3), size = 1) +
    #geom_text(aes(label = power), size = 3) +
    geom_hline(yintercept = 0.9, lty = 2, color = "maroon") +
    scale_color_manual(values = c("TRUE" = "maroon", "FALSE" = "black"), guide = "none") +
    scale_x_continuous(breaks = seq(2, 16, by = 2)) +
    labs(
        x = "Soft threshold (power)",
        y = "Scale free topology model fit \n(signed R^2)"
    ) +
    theme(
        panel.grid.major.x = element_line(color = "darkgrey", size = 0.2)#,
        #panel.grid.minor.x = element_line(color = "lightgray")
    )

## Median connectivity as a function of the soft-thresholding power (want to remain reasonably high (in the 100s) or above)
p2 <- df_sft %>% 
    ggplot(aes(x = power, y = median_k)) +
    geom_point(aes(color = power == 3), size = 1) +
    #geom_text(aes(label = power), size = 3) +
    geom_hline(yintercept = 100, lty = 2, color = "maroon") +
    scale_color_manual(values = c("TRUE" = "maroon", "FALSE" = "black"), guide = "none") +
    scale_x_continuous(breaks = seq(2, 16, by = 2)) +
    labs(
        x = "Soft threshold (power)",
        y = "Median connectivity"
    ) +
    theme(
        panel.grid.major.x = element_line(color = "darkgrey", size = 0.2)#,
        #panel.grid.minor.x = element_line(color = "lightgray")
    )

## Combine and save
p1 + p2 + plot_annotation(title = "A | Soft-thresholding power fit to data")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S4A.SFT_fit", .x), width = 5, height = fig_height)
)
```

### B | Module stability across minimum module size & cut height thresholds

```{r FigS4b.plot, fig.width = 5, fig.height = 6.5}
df_modules %>%
    group_by(sft, min_size, cut_height) %>%
    dplyr::count(module, color) %>% 
    #filter(cut_height != 1) %>%

    ggplot(aes(x = module)) +
    geom_segment(aes(y = 0, yend = log10(n), color = color)) +
    geom_point(aes(y = log10(n), fill = color), shape = 21, size = 0.75) +
    geom_vline(aes(xintercept = 21), lty = 2, color = "maroon") +
    scale_fill_identity() +
    scale_color_identity() +
    scale_x_discrete(breaks = seq(0, 50, 10)) +
    facet_grid(cut_height ~ min_size) +
    labs(y = "log10(N genes)", 
         title = "B | Module assignments across WGCNA parameters"
    ) +
    theme(
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S4B.module_sizes", .x), width = 5, height = 6.5)
)
```

# FIG S5: WGCNA module enrichments #

### A | Module sizes

```{r FigS5a.plot, fig.width = 3.5, fig.height = 2.5}
df_modules_filt %>% 
    dplyr::count(module, color) %>% 
    ggplot(aes(x = module, y = n)) +
    geom_col(aes(fill = module), color = "black", linewidth = 0.25) +
    geom_text(aes(label = n), size = 3, angle = 45, hjust = 0, vjust = 0) +
    scale_fill_manual(values = module_colors) +
    coord_cartesian(clip = "off") +
    labs(title = "A | Module sizes",
         x = NULL, y = "Number of genes") +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 0.9, size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S5A.Module_sizes", .x), width = 3.5, height = 2.5)
)

```

### B | Semantic themes across module GO results

```{r FigS5b.prep}
## Extract topic descriptions (frequent words)
# FREX weights words by frequency and exclusivity to the topic
df_word_beta <- tidy(topic_model, matrix = "beta") %>%
    group_by(topic) %>%
    arrange(topic, -beta) %>% 
    top_n(n = 10, wt = beta) %>%
    mutate(topic = paste0("topic ", topic)) %>% 
    dplyr::filter(!str_detect(term, "cell")) %>% 
    group_by(topic) %>% 
    mutate(top_word = ifelse(beta == max(beta), "yes", "no"))

## Extract topic probabilities for each module (document)
df_group_gamma <- tidy(
    topic_model, 
    matrix = "gamma",
    document_names = rownames(go_sparse)
) %>% 
    mutate(module = factor(document)) %>% 
    dplyr::select(module, topic, gamma) %>% 
    mutate(topic = factor(topic, levels = seq(n_topics, 1, by = -1)),
           module = paste0("geneM", module) %>% factor(levels = paste0("geneM", levels(module)))
    ) %>% 
    dplyr::filter(gamma > 0.01)
```

```{r FigS3b.plot, fig.width = 3, fig.height = 6}
## Plot word strength of association with each topic
p_figS5b_beta <- df_word_beta %>% 
    ggplot(aes(x = reorder_within(term, beta, topic), y = beta, fill = beta)) +
    geom_col(color = "#222222") +
    geom_text(aes(label = term, y = 0.001, size = beta, color = top_word), angle = 90, hjust = 0) +
    facet_wrap(vars(topic), ncol = 1, scales = "free_x") +
    scale_x_reordered() +
    scale_y_continuous(limits = c(0, 0.06), breaks = c(0, 0.05)) +
    scale_size_continuous(range = c(2, 3), guide = "none") +
    scale_color_manual(values = c("yes" = "white", "no" = "black"), guide = "none") +
    scale_fill_gradient(low = "white", high = "blue", guide = "none", limits = c(0, 0.058)) +
    labs(x = NULL, y = "Word strength of association with each topic") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill = "white", color = "black"),
          strip.text = element_text(size = 9)
    )

## Plot module strength of association with each topic
p_figS5b_gamma <- df_group_gamma %>%
    ggplot(aes(x = gamma, y = topic, color = module)) +
    geom_jitter(aes(alpha = gamma), position = position_jitter(seed = 2, width = 0.15)) +
    geom_text_repel(aes(alpha = gamma, label = module), size = 2.75, box.padding = 0.2, min.segment.length = 0,
                    max.overlaps = 15, position = position_jitter(seed = 2, width = 0.15)) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), color = "black", linewidth = 0.25) +
    scale_color_manual(values = module_colors) +
    labs(x = "Module strength of association \nwith each topic", y = NULL) +
    theme(legend.position = "none",
          axis.text.y = element_blank()
    )

## Combine & save
p_figS5b_beta + p_figS5b_gamma + plot_layout(widths = c(1, 2)) + plot_annotation(title = "B | GO semantic themes")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S5B.topic_modeling", .x), width = 3, height = 6)
)
```

### C | Cell-type enrichment of GO modules

```{r FigS5c.prep}
df_figS5c <- df_mods_cell_type_hypergeometric %>%
    mutate(
        p_adj = ifelse(p_adj == 0, 10^-15, p_adj),
        color = ifelse(p_value < 0.05, paste0(module), NA) %>% factor,
        label = case_when(
            p_adj < 0.05 & p_adj > 0.01 ~ "*",
            p_adj < 0.01 & p_adj > 0.001 ~ "**",
            p_adj < 0.001 ~ "***",
            TRUE ~ ""
        )
    )
```

```{r FigS5c.plot, fig.width = 3.35, fig.height = 3.4}
df_figS5c %>% 
    ggplot(aes(x = cell_type, y = module)) +
    geom_tile(aes(fill = color, alpha = -log10(p_adj), color = color), 
              width = 0.95, height = 0.95, linewidth = 0.5) +
    geom_text(aes(label = label), vjust = 0.75, size = 3.5) +
    #annotate(geom = "text", label = "* = FDR < 0.05", x = 6.5, y = 1, size = 3) +
    scale_alpha_continuous(range = c(0.1, 1), na.value = 1) +
    scale_fill_manual(values = module_colors, na.value = "transparent", guide = "none") +
    scale_color_manual(values = module_colors, na.value = "transparent", guide = "none") +
    guides(alpha = guide_legend(title = "-log10(FDR)", title.position = "top", title.hjust = 0.5)) +
    labs(title =  "C | Cell-type enrichment", 
         x = NULL, y = NULL
    ) +
    theme(legend.position = "bottom",
          legend.margin = margin(t = -12),
          legend.key.size = unit(0.3, "cm"),
          legend.justification = "center",
          axis.text.y = element_text(size = 8.5),
          axis.text.x = element_text(angle = 45, hjust = 0.9)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S5C.cell_type_enrichment", .x), width = 3.35, height = 3.4)
)
```

### D | Developmental trajectory of each module per PsychENCODE data

```{r FigS5d.plot, fig.width = 6.5, fig.height = 3}
df_expr_mod_median %>% 
    ggplot(aes(x = window, y = median_expr, color = module, fill = module)) +
    geom_smooth(method = "loess") +
    geom_vline(xintercept = 5, linewidth = 0.2) +
    facet_wrap( ~ module, scales = "free_y", nrow = 3) +
    scale_fill_manual(values = module_colors) +
    scale_color_manual(values = module_colors) +
    labs(x = "Developmental window", y = "Median expression (PsychENCODE)", 
         title = "D | Developmental expression trajectory of each module") +
    theme(legend.position = "none",
          strip.text.x = element_text(size = 9),
          strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),
          panel.spacing = unit(0.05, "cm"),
          strip.clip = "off"
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S5D.developmental_trajectories", .x), width = 6.5, height = 3)
)
```

# FIG S6: RCCA and GRCCA setup (generated in PowerPoint) #

# FIG S7: Gene-level GRCCA covariate weights #

```{r FigS7.plot}
df_y_res %>% 
    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%

    ggplot(aes(x = reorder(covariate, weight), y = weight)) +
    geom_col(aes(fill = covariate, alpha = abs(weight)), color = "black") +
    geom_errorbar(aes(ymin = weight - sd, ymax = weight + sd), width = 0.2) +
    geom_text(aes(label = significant, y = 0), size = 6, vjust = 1.5) +
    geom_hline(aes(yintercept = 0), color = "black") +
    scale_fill_manual(values = c(dx_colors, rep("gray", 6))) +
    scale_alpha_continuous(range = c(0.1, 1)) +
    scale_x_discrete(labels = function(x) {str_replace(x, "_", " ")}) +
    ylim(c(-0.10, 0.10)) +
    labs(x = "", y = "weight", 
         #caption = "* = abs(z-score) >= 2 & p-adj < 0.05; \nError bar shows +/- 1 sd",
         title = "Covariate weights",
         #title = bquote(bold("A")~" | Covariate (y) coefficient weights")
    ) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 0.9))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S7.geneLevel_covariate_weights", .x), width = fig_width, height = fig_height)
)
```

# FIG S8: Vector of significant GRCCA genes (FDR_r < 0.05 & z-score > 2) #

```{r FigS8.plot, fig.height = 3}
df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) >= 2) %>%

    ggplot(aes(x = pearsons_r, y = reorder(ensembl_gene_id, pearsons_r))) +
    geom_col(aes(fill = pearsons_r)) +
    geom_vline(xintercept = 0, color = "gray") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1)) +
    guides(fill = guide_colorbar(title = expression(italic(r)[x]))) +
    #scale_color_manual(values = c("*" = "black", "-" = "transparent")) +
    xlim(c(-1, 1)) +
    labs(y = "Gene", x = expression("Structure correlation (" * italic(r)[x] * ")"), 
         title = "Significant GRCCA genes"
         #title = bquote(bold("D")~" | Significant gene structure correlations"),
    ) +
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(),
          legend.position = c(0.15, 0.70), 
          legend.box = "horizontal",
          legend.key.width = unit(0.15, "cm"),
          legend.key.height = unit(0.6, "cm")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S8.significant_GRCCA_genes_all", .x), width = fig_width, height = 3)
)
```

# FIG S9: Compare current DEGs to Akula et al 2021 DEGs #

```{r FigS9.prep}
## Combine current study results with Akula 2021
df_de_akula_res <- df_de_res %>% 
    mutate(analysis = "current", .before = 1) %>%

    bind_rows(df_akula_res %>% 
                  mutate(analysis = "akula") %>% 
                  dplyr::rename("gene_symbol" = "hgnc_symbol")
    )
```

### A | Venn diagram of significant DEG overlap

```{r FigS9a.prep}
## Venn diagram (overlap with Nirmala's results)
my_degs <- df_de_res %>% dplyr::filter(pvalue < 0.05) %>% pull(ensembl_gene_id)
length(my_degs) # N = 1690

akula_degs <- df_akula_res %>% 
    dplyr::filter(pvalue < 0.05) %>% 
    pull(ensembl_gene_id)
length(akula_degs) # N = 1373

sig_grcca_genes <- df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
    pull(ensembl_gene_id)
length(sig_grcca_genes) # N = 1211

# Akula DEGs vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, akula_degs, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 433; p < 0.0001

# GRCCA vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 211; p < 0.0001

# GRCCA vs Akula DEGs
hypergeometric_overlap <- f_hypergeometric(akula_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 285; p < 0.0001

# Combine the data into a named list for ggvenn
FigS9a_data <- list(
  `Current study DEGs` = my_degs,
  `Akula et al 2021 DEGs` = akula_degs,
  `Significant GRCCA genes` = sig_grcca_genes
)
```

```{r FigS9a.plot}
ggvenn(
    FigS9a_data, 
    fill_color = c("#E0E0E0", "#E0E0E0", "#E0E0E0"),
    stroke_size = 0.75,
    text_size = 3,
    set_name_size = 4,
    show_percentage = FALSE
) +
    coord_fixed(clip = "off") +
    theme_void() +
    theme(
        text = element_text(size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S9A.DEG_GRCCA_venn_diagram", .x), width = fig_width, height = fig_height)
)
```

### B | Correlation plot of L2FC

```{r FigS9b.prep}
## Create df for plot
df_figS9b <- df_de_akula_res %>%

    # pivot wider to plot
    pivot_wider(id_cols = c(ensembl_gene_id, gene_symbol), names_from = analysis, values_from = stat) %>% 
    dplyr::filter(!is.na(akula)) %>%

    # normalize values to compare across analyses
    #mutate(akula = scale(akula), current = scale(current)) %>%

    # add gene labels
    mutate(label = ifelse(abs(current) > 1.5 & abs(akula) > 1.5, gene_symbol, "")) %>%

    # arrange to plot
    arrange(-abs(current))
```

```{r FigS3b.plot}
df_figS9b %>% 
    ggplot(aes(x = akula, y = current)) +
    geom_point(aes(color = akula), size = 0.75) +
    geom_hline(aes(yintercept = 0), color = "grey") +
    geom_vline(aes(xintercept = 0), color = "grey") +
    geom_smooth(method = "lm", color = "black", linewidth = 0.75) +
    geom_abline(lty = 2) +
    #geom_text(x = 2.1, y = 2.6, label = "y = x", check_overlap = TRUE) +
    geom_text_repel(aes(label = label), min.segment.length = 0, max.overlaps = 15, size = 2.5) +
    stat_cor(label.sep = "\n", vjust = 1, size = 3.5) +
    #stat_regline_equation(label.y.npc = "top", vjust = 3) +
    scale_color_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-6.2, 6.2), guide = "none") +
    #scale_color_manual(values = c("transparent", "black")) +
    #guides(color = guide_legend(title = "FDR < 0.05", title.vjust = -1.5)) +
    labs(x = "Akula DEGs t-statistic", y = "Current DEGs t-statistic", 
         title = "B | Correlation between \nDGE analysis results") +
    theme(legend.position = c(0.05, 0.7),
          legend.spacing.x = unit(0.1, "mm"),
          legend.background = element_blank(),
          legend.box.background = element_rect(colour = "black")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S9B.DEG_correlation", .x), width = fig_width, height = fig_height)
)
```

### C | Significant DEG overlap with significant GRCCA genes

```{r FigS9c.prep}
sig_grcca_genes <- df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
    pull(ensembl_gene_id)
length(sig_grcca_genes) # N = 1211
length(my_degs) # N = 1690
length(akula_degs) # N = 1373

# GRCCA vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 211; p < 0.0001

# GRCCA vs Akula DEGs
hypergeometric_overlap <- f_hypergeometric(akula_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 285; p < 0.0001

```

# FIG S10: GRCCA results are related to single-cell expression gradients implicated in SCZ #

```{r FigS10.prep}
df_figS10 <- df_lf4_grcca %>% 
    left_join(df_lake_cell_type) %>% 
    dplyr::filter((type == "Astro" & cell_type == "astrocyte") |
                      (type == "Neuro-Ex" & cell_type == "glutamatergic") |
                      (type == "Neuro-In" & cell_type == "gabaergic")
    )
```

```{r FigS10.plot, fig.width = 6}
df_figS10 %>% 
    ggplot(aes(x = PEER_4, y = pearsons_r)) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_point(aes(color = pearsons_r), size = 0.5) +
    geom_smooth(color = "black", method = "lm", linewidth = 0.75) +
    stat_cor(label.sep = "\n", size = 3, 
             label.y.npc = "top", label.x.npc = "right", hjust = 1, vjust = 1) +
    facet_wrap(vars(cell_type), scales = "free") +
    scale_color_gradientn(colors = gene_weight_color_scale, guide = "none") +
    scale_x_continuous(breaks = c(-2000, 0, 2000)) +
    labs(x = "Ling et al LF4", y = expression("Structure correlation (" * italic(r)[x] * ")"),
         title = "GRCCA correlation with SNAP program by cell type")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S10.correlation_with_SNAP", .x), width = 6, height = fig_height)
)
```

# FIG S11: SCZ-WGCNA module associations determined through traditional module eigengene analysis #

### A | Modules significantly associated with SCZ per module eigengene analysis

```{r FigS11a.prep}
df_figS11a <- df_lm_dx %>% 
    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
    unnest(cols = c(data)) %>% 
    mutate(label = case_when(
        dx == "SCZ" & p_value < 0.05 & p_adj > 0.05 ~ "*",
        dx == "SCZ" & p_value < 0.05 & p_adj < 0.05 ~ "**",
        TRUE ~ ""
    )
    )
```

```{r FigS11a.plot, fig.height = 5}
df_figS11a %>% 
    ggplot(aes(x = kme_resids, y = dx, color = dx, fill = dx)) +
    geom_point(position = position_jitter(width = 0.1), size = 0.5) +
    geom_boxplot(alpha = 0.1, outlier.shape = NA) +
    # geom_text(aes(label = label, x = -Inf), color = "black", size = 6,
    #           vjust = 0.75, hjust = -0.1, check_overlap = TRUE) +
    facet_wrap(vars(module), ncol = 1, scales = "free_x") +
    scale_fill_manual(values = dx_colors) +
    scale_color_manual(values = dx_colors) +
    labs(x = "kME (corrected for covariates)", y = NULL,
         #caption = "* = p < 0.05; ** = FDR < 0.05",
         title = "A | SCZ kME-diagnosis \nassociations"
    ) +
    theme(legend.position = "none",
          axis.text.y = element_text(size = 10)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
)
```

### B | Significant SCZ module enrichments

```{r FigS11b.prep}
## Identify significant modules
sig_modules <- df_lm_dx %>% 
    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
    pull(module)

# Filter for modules of interest & take top 5 pathways for each
df_figS11b <- df_mods_go %>% 
    dplyr::filter(module %in% sig_modules) %>% 
    group_by(module) %>% 
    top_n(5, wt = -log10(pvalue)) %>% 
    mutate(Description = str_wrap(Description, width = 40))
```

```{r FigS11b.bubbleplot, fig.height = 5, fig.width = 3.5}
df_figS11b %>% 
    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
    facet_wrap(vars(module), scales = "free_y", ncol = 1) +
    scale_fill_manual(values = module_colors, guide = "none") +
    scale_y_reordered() +
    scale_x_reverse(breaks = c(0.015, 0.00)) +
    scale_size_continuous(range = c(1, 4)) +
    coord_cartesian(clip = "off") +
    labs(x = "Enrichment FDR", y = NULL,
         title = "B | SCZ module enrichments") +
    theme(legend.position = c(-1.9, 0.85))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11B.sig_SCZ_kME_modules", .x), width = 3.5, height = 5)
)
```

# FIG S12: WGCNA module enrichment for risk gene lists #

```{r FigS12.prep}
df_figS12 <- df_mods_risk_gene_hypergeometric %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(sig = case_when(p_adj < 0.001 ~ "***", p_adj < 0.01 ~ "**", p_adj < 0.05 ~ "*", TRUE ~ ""))
```

```{r FigS12.plot, fig.width = 5, fig.height = 4}
df_figS12 %>% 
    ggplot(aes(x = benchmark_list, y = module)) +
    geom_tile(aes(fill = module, alpha = -log10(p_adj))) +
    geom_text(aes(label = sig), size = 6, color = "black", vjust = 0.75) +
    facet_wrap(vars(class), scales = "free_x", nrow = 1) +
    scale_fill_manual(values = module_colors, guide = "none") +
    guides(alpha = guide_legend(title = "-log10(FDR)")) +
    labs(x = NULL, y = NULL, 
         title = "Risk gene list enrichment in WGCNA modules")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S12.risk_gene_module_hypergeometric", .x), width = 5, height = 4)
)
```

# FIG S13: Functional enrichment of GRCCA overrepresented WGCNA modules #

### A | Hypergeometric overlap of WGCNA modules and significant GRCCA genes

```{r FigS13a.prep}
## Identify significant GRCCA genes (FDR < 0.05 & |z| > 2)
sig_grcca_genes <- df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
    pull(ensembl_gene_id)
# n = 1211

## Hypergeometric test for significant overlap with modules
df_grcca_module_hypergeometric <- map_dfr(
    .x = names(module_colors),
    .f = ~ f_module_hypergeometric(
        gene_list = sig_grcca_genes, 
        gene_list_name = "", 
        wgcna_module = .x, 
        gene_universe = ensembl_gene_ids
    ) %>% 
        dplyr::select(-gene_list)
) %>% 
    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>% 
    arrange(p_adj) %>% 
    mutate(module = factor(module, levels = names(module_colors))) %>%

    # specify modules to label
    mutate(module_label = ifelse(p_adj < 0.05, paste0("N = ", overlap_n), ""))
```

```{r FigS10a.plot, fig.height = 5, fig.width = 2.15}
f_plot_module_hypergeometric(df_grcca_module_hypergeometric %>% mutate(p_adj = ifelse(p_adj == 0, 10^-5, p_adj)) %>% dplyr::filter(module != "geneM0"),
                             x_axis = "p-adj",
                             plot_title = "A | GRCCA significant modules",
                             #subset_x_labs = TRUE,
                             include_module_labels = TRUE,
                             module_text_size = 3,
                             legend_position = "none",
                             p_value_threshold = 0.05,
                             include_threshold_label = TRUE
)
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S13A.GRCCA_module_hypergeometric", .x), width = 2.15, height = 5)
)
```

### B | Direction of significant genes in top modules

```{r FigS13b.prep}
## Identify significant modules
sig_modules <- df_grcca_module_hypergeometric %>% dplyr::filter(module != "geneM0" & p_adj < 0.05) %>% pull(module) %>% sort()

## Identify panel sizes (number of significant genes in each module)
panel_widths <- df_x_res %>% 
    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
    dplyr::count(module) %>% 
    pull(n)
```

```{r FigS13b.plot, fig.height = 5, fig.width = 2.15}
df_x_res %>% 
    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
    mutate(weight = ifelse(significant == 1, weight, 0)) %>%

    ggplot(aes(x = reorder_within(ensembl_gene_id, within = module, by = pearsons_r), y = pearsons_r)) +
    geom_col(aes(fill = module, alpha = abs(pearsons_r))) +
    facet_wrap(vars(module), ncol = 1, scales = "free_y") +
    force_panelsizes(rows = panel_widths) +
    scale_fill_manual(values = module_colors) +
    scale_alpha_continuous(range = c(0.3, 1)) +
    labs(title = "B | Significant module genes",
         x = "Gene", y = expression("Structure correlation (" * italic(r)[x] * ")")) +
    guides(fill = guide_legend(nrow = 2)) +
    coord_flip() +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          strip.text = element_blank(),
          legend.position = "none",
          panel.spacing = unit(0.3, "lines")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S13B.sig_module_direction", .x), width = 2.15, height = 5)
)

```

### C | GO BP enrichments of significant modules

```{r FigS13c.prep}
# Filter for GRCCA modules of interest & take top 5 pathways for each
df_figS13c <- df_mods_go %>% 
    dplyr::filter(module %in% sig_modules) %>% 
    group_by(module) %>% 
    top_n(5, wt = -log10(pvalue)) %>% 
    mutate(Description = str_wrap(Description, width = 23.5))
```

```{r FigS13c.bubbleplot, fig.height = 5, fig.width = 2.15}
df_figS13c %>% 
    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
    facet_wrap(vars(module), scales = "free_y", ncol = 1) +
    scale_fill_manual(values = module_colors, guide = "none") +
    scale_y_reordered() +
    scale_x_reverse(breaks = c(0.00005, 0.00)) +
    scale_size_continuous(range = c(1, 4)) +
    coord_cartesian(clip = "off") +
    labs(x = "Enrichment FDR", y = NULL,
         title = "C | Module enrichments") +
    theme(
        legend.position = c(-1.6, 1.05),
        legend.key.size = unit(0.2, "cm"),
        axis.text = element_text(size = 8.5)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S13C.sig_GRCCA_module_enrichments", .x), width = 2.15, height = 5)
)
```

# FIG S14: Compare GRCCA and RCCA directly #

### A | Correlate GRCCA and RCCA covariate (y) structure correlations

```{r FigS14a.plot}
df_y_res_all %>% 
    pivot_wider(id_cols = covariate, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(fill = rcca + grcca), shape = 21, size = 2.5) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    geom_text_repel(aes(label = covariate), size = 2.5, max.overlaps = 20, box.padding = 1.25,
                     nudge_y = 0.3, force = 150) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.84, 1.84), guide = "none") +
    labs(title = "A | Covariate structure correlations",
         x = "RCCA", y = "GRCCA")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S14A.covariate_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### B | Correlate GRCCA and RCCA gene (x) structure correlations

```{r FigS14b.plot}
df_x_res_all %>% 
    pivot_wider(id_cols = ensembl_gene_id, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(color = rcca + grcca), size = 1) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    stat_cor(label.sep = "\n") +
    #geom_label_repel(aes(label = gene_symbol), size = 3, max.overlaps = 50, box.padding = 1.5) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-1.14, 1.14), guide = "none") +
    labs(title = "B | Gene structure correlations",
         x = "RCCA", y = "GRCCA")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S14B.gene_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### C | Venn diagram of significant gene overlap (made in PowerPoint/Inkscape)

```{r FigS14c.prep}
## Venn diagram (overlap with Nirmala's results)
sig_rcca_genes <- df_x_res_rcca %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
    pull(ensembl_gene_id)
length(sig_grcca_genes) # N = 1211
length(sig_rcca_genes) # N = 2026
hypergeometric_overlap <- f_hypergeometric(sig_grcca_genes, sig_rcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 855; P < 0.0001
```

### D | RCCA results risk gene enrichment

```{r FigS14d.prep}
df_fig13d <- df_rcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))
```

```{r FigS14d.plot}
df_fig13d %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "D | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S14D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
)
```

# GRAVEYARD #

## Fig S4B alt: Module stability (alluvial)

```{r FigS4b.alluvial, fig.width = 6.5, fig.height = 4}
df_summary <- df_modules %>%
  group_by(min_size, cut_height, module, color) %>%
  summarise(n_genes = n(), .groups = "drop")

# Create the alluvial plot
library(ggalluvial)
df_summary %>% 
    mutate(
        cut_height = factor(cut_height),
        min_size = factor(min_size)
    ) %>%

    ggplot(aes(x = min_size, y = n_genes, stratum = module, alluvium = module, fill = color)) +
    geom_flow(stat = "alluvium", lode.guidance = "frontback", aes(fill = color)) +
    geom_stratum(width = 0.2, color = "black") +
    scale_fill_identity() +  # Use the hex codes from the 'color' column
    facet_grid(cols = vars(cut_height)) +  # Facet into rows by min_size
    theme_minimal() +
    labs(x = "Cut Height", 
         y = "Number of Genes",
         title = "Module stability across parameters") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 10)
    )
```

## Fig SXX: Full WGCNA module functional enrichments

```{r FigS5.prep}
df_figS5 <- df_mods_go %>% 
    clean_names %>% 
    dplyr::filter(!is.na(description) & module != 0) %>%

    # take top 5 paths by p-value per module to plot
    dplyr::group_by(module) %>% 
    arrange(module, -p_adjust) %>% 
    mutate(row = row_number(),
           gene_ratio = parse(text = gene_ratio) %>% eval
    ) %>% 
    top_n(n = 4, wt = row) %>% 
    #slice_max(order_by = -weight_fisher, n = 5) %>% 
    mutate(`survives FDR` = ifelse(p_adjust < 0.05, "yes", "no")) %>% 
    mutate(description = str_wrap(description, width = 10))
```

```{r FigS5.plot, fig.height = 8.75, fig.width = 6.5}
df_figS5 %>% 
    ggplot(aes(x = p_adjust, y = reorder_within(description, within = module, by = -p_adjust))) +
    #y = reorder_within(str_wrap(description, width = 35), -pvalue, module))) +
    geom_point(aes(size = gene_ratio, fill = ontology, color = `survives FDR`),
               shape = 21, stroke = 1) +
    geom_vline(aes(xintercept = 0.05), linewidth = 0.75, color = "gray") +
    facet_wrap(vars(module), scales = "free", ncol = 5) +
    scale_size_continuous(range = c(0.5, 3)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_y_reordered() +
    scale_x_reverse(breaks = c(0.25, 0.15, 0.05, 0)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = "Enrichment FDR",
         title = "WGCNA module GO enrichments"
    ) +
    guides(size = guide_legend(title = "Gene ratio")) +
    theme(legend.position = c(0.85, 0.0),
          legend.direction = "vertical",
          legend.title.position = "top",
          #legend.justification = "center",
          strip.background = element_rect(fill = "white"),
          strip.clip = "off",
          axis.text = element_text(size = 7.0)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S5.full_GO_enrichments", .x), width = 6.5, height = 8.75)
)
```

Figure SXX. Full GO enrichments of gene-level modules. Top 4 Gene Ontology pathways associated with each module, determined by p-value, across all three ontologies (BP = biological process, CC = cellular component, MF = molecular function). The x-axis shows the enrichment FDR, while the y-axis shows pathway names. Each point is colored by its ontology and sized by the gene ratio (number of genes in the module in the pathway divided by the total number of genes in the pathway); points that survived correction for multiple comparisons (FDR < 0.05; vertical gray line) are outlined in black.

## Fig S10B alternate (word clouds)

```{r FigS10b.wordcloud, fig.height = 5}
## Identify uninteresting words
data(stop_words)
vague_terms <- c("positive", "negative", "regulation", "response", "protein", "activity",
                 "pathway", "process", "involved", "signaling")

## Identify top words in each significant (FDR < 0.05) module pathway
module_tokens_freq <- df_sig_go_res %>% 
    dplyr::filter(p.adjust < 0.05) %>% 
    group_by(module) %>% 
    unnest_tokens(word, Description) %>% 
    anti_join(stop_words) %>% 
    dplyr::filter(!(word %in% vague_terms)) %>% 
    dplyr::count(word, sort = TRUE)

## Create wordcloud plot for each module
l_figS10b <- map(
    .x = as.character(sig_modules),
    .f = ~ module_tokens_freq %>% 
        dplyr::filter(module == .x) %>% 
        top_n(10, wt = n) %>% 
        ggplot(aes(label = word, size = n, color = module)) +
        geom_text_wordcloud() +
        facet_wrap(vars(module)) +
        scale_color_manual(values = module_colors) +
        scale_size_continuous(range = c(2.5, 5)) +
        theme_void() +
        theme(strip.text = element_blank(),
              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
        )
)

## Combine and save
wrap_plots(l_figS10b, ncol = 1) & plot_annotation(title = "B | SCZ module enrichments")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S10B.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
)
```

GO biological process pathway enrichments for each module. The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module. Words are colored by respective module.

## Fig S12C alternate (word clouds)
```{r FigS12c.prep}
## Load module GO enrichment data
#load(paste0(base_dir, "objects/", prefix, sep = "_", "SIGNED_SFT", soft_power, "_GO_RES.RDS")) # df_mods_go

## Identify words to not include
data(stop_words)
vague_terms <- c("positive", "negative", "regulation", "response", "activity",
                 "pathway", "process", "involved", "signaling")

# Filter for modules of interest
df_go_res <- df_mods_go %>% 
    dplyr::filter(module %in% sig_modules)

## Identify top words in each significant (FDR < 0.05) module pathway
module_tokens_freq <- df_go_res %>% 
    dplyr::filter(p.adjust < 0.05) %>% 
    group_by(module) %>% 
    unnest_tokens(word, Description) %>% 
    anti_join(stop_words) %>% 
    dplyr::filter(!(word %in% vague_terms)) %>% 
    dplyr::count(word, sort = TRUE)
```

```{r FigS12c.wordcloud}
l_module_wordclouds <- map(
    .x = as.character(sig_modules),
    .f = ~ module_tokens_freq %>% 
        dplyr::filter(module == .x) %>% 
        top_n(10, wt = n) %>% 
        ggplot(aes(label = word, size = n, color = module)) +
        geom_text_wordcloud() +
        facet_wrap(vars(module)) +
        scale_color_manual(values = module_colors) +
        scale_size_continuous(range = c(2.5, 5)) +
        theme_void() +
        theme(strip.text = element_blank(),
              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
        )
)
```

GO biological process pathway enrichments for each module. The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module. Words are colored by respective module.

## Fig SXX: Cell-type enrichment of DEGs

```{r FigSXX.plot}
df_de_fgsea_cell %>% 
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>%

    ggplot(aes(x = NES, y = reorder(cell_type, NES))) +
    geom_col(aes(fill = NES), color = "gray", shape = 21) +
    geom_text(aes(label = sig, x = NES), vjust = 0.75, size = 4, check_overlap = TRUE) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-2.61, 2.61)) +
    coord_cartesian(clip = "off") +
    labs(x = "NES", y = NULL,
         title = "D | Cell-type enrichment") +
    theme(legend.position = "none")
```

## Fig S13E: GRCCA vs RCCA hypergeometric overlap across structure correlation thresholds

```{r FigS13e.plot}
df_rcca_grcca_hypergeometric_benchmark %>%

    ggplot(aes(x = threshold, y = -log10(p_value))) +
    geom_hline(aes(yintercept = -log10(0.05)), color = "gray") +
    #geom_vline(aes(xintercept = length(sig_rcca_genes) / length(ensembl_ids)), lty = 2, color = "#00BFC4") +
    #geom_vline(aes(xintercept = length(sig_grcca_genes) / length(ensembl_ids)), lty = 2, color = "#F8766D") +
    geom_point(aes(color = analysis)) +
    geom_line(aes(group = analysis, color = analysis)) +
    #annotate(geom = "text", label = "GRCCA sig θ", x = 0.04, y = 3.1, size = 3) +
    #annotate(geom = "text", label = "RCCA sig θ", x = 0.13, y = 3.1, size = 3) +

    facet_wrap(vars(benchmark_list), scales = "free_y") +
    #facet_grid2(benchmark_type ~ benchmark_list, independent = "all", scales = "free") +
    labs(title = "Hypergeometric overlaps with risk gene lists across structure correlation thresholds",
         x = "Proportion of genes included (ordered by structure correlation)", y = "-log10(hypergeometric p-value)") +
    theme(legend.position = "top",
          legend.justification = "right")
```

## Figs SXX - SXX: Rare-variant related figures

### Rare variant positions in GRCCA structure correlation distribution
```{r FigSXX.GRCCA}
## Subset X res for genes in SCZ common broad list for bottom panel
df_SXX <- df_x_res_genes %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]]) %>% 
    mutate(significant = ifelse(p_adj < 0.05, 1, 0) %>% factor) %>% 
    arrange(significant, abs(pearsons_r)) %>% 
    mutate(label = gene_symbol)
    #mutate(label = ifelse(ensembl_gene_id %in% top_genes, gene_symbol, ""))

## Plot aesthetics
scale_max <- df_x_res_genes %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)

## Plot
figSXX <- df_x_res_genes %>%
    ggplot(aes(x = pearsons_r)) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density() +
    geom_point(data = df_SXX, 
               mapping = aes(x = pearsons_r, y = -1, fill = pearsons_r, color = significant), 
               pos = jitter_pos, shape = 21, size = 2
    ) +
    geom_text_repel(data = df_SXX,
                    mapping = aes(x = pearsons_r, y = -1, label = label),
                    pos = jitter_pos, 
                    size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
    ) +

    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-scale_max, scale_max), guide = "none") +
    scale_color_manual(values = c("gray", "black"), guide = "none") +
    labs(x = expression("Structure correlation (" * italic(r)[x] * ")"), y = NULL, 
         title = expression("SCZ rare variants in " * italic(r)[x] * " distribution")) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_rect(fill = "transparent", color = "black", size = 1),
          axis.line = element_blank()
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "SXX.rare_variants_in_distribution", .x), width = fig_width, height = fig_height)
)

```

### Rare variant positions in DGE distribution
```{r FigSXX.DGE}
## Subset X res for genes in SCZ common broad list for bottom panel
df_SXX_dge <- df_de_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]]) %>% 
    mutate(significant = ifelse(pvalue < 0.05, 1, 0) %>% factor) %>% 
    arrange(significant, abs(stat)) %>% 
    mutate(label = gene_symbol)
    #mutate(label = ifelse(ensembl_gene_id %in% top_genes, gene_symbol, ""))

## Plot aesthetics
scale_max <- df_de_res %>% pull(stat) %>% abs %>% max
jitter_pos <- position_jitter(width = 0, seed = 123)

## Plot
figSXX_dge <- df_de_res %>%
    ggplot(aes(x = stat)) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density() +
    geom_point(data = df_SXX_dge, 
               mapping = aes(x = stat, y = 0, fill = stat, color = significant), 
               #pos = jitter_pos, 
               shape = 21, size = 2
    ) +
    geom_text_repel(data = df_SXX_dge,
                    mapping = aes(x = stat, y = 0, label = label),
                    #pos = jitter_pos, 
                    size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
    ) +

    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-scale_max, scale_max), guide = "none") +
    scale_color_manual(values = c("gray", "black"), guide = "none") +
    labs(x = "DGE t-statistic", y = NULL, 
         title = "SCZ rare variants in DGE t-statistic distribution") +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_rect(fill = "transparent", color = "black", size = 1),
          axis.line = element_blank()
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "SXX.rare_variants_in_DGE_distribution", .x), width = fig_width, height = fig_height)
)

```

### Expression directions of rare variants

```{r FigSXX.prep}
df_figSXX <- df_vsd_regress %>% 
    dplyr::select(sample, any_of(benchmarking_lists[["SCZ (rare)"]])) %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expr") %>% 
    left_join(df_ensembl_to_symbol) %>% 
    mutate(dx = case_when(
        str_detect(sample, "control") ~ "Control",
        str_detect(sample, "bipolar") ~ "BD",
        str_detect(sample, "mdd") ~ "MDD",
        str_detect(sample, "schizo") ~ "SCZ"
    ) %>% 
        factor(levels = names(dx_colors)))
```

```{r FigSXX.plot, fig.height = 5, fig.width = 6}
figSXX <- df_figSXX %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.5) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "darkgrey") +
    facet_wrap(vars(gene_symbol), scales = "free_y", nrow = 3) +
    #force_panelsizes(cols = c(1, 1)) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    labs(x = NULL, y = "Corrected expression value",
         title = "Gene expression values of rare variants") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
figSXX
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "SXX.rare_variants_expression_directions", .x), width = 6, height = 5)
)
```

### Transcript derivatives of rare variants

```{r FigSXX}
df_x_res_transcripts %>% 
    filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]]) %>% 
    mutate(significant = ifelse(p_adj < 0.05, "yes", "no")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(transcript_symbol, pearsons_r))) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(fill = pearsons_r, size = -log10(p_adj), color = significant), shape = 21) +
    facet_wrap(vars(gene_symbol), scales = "free_y") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.17, 0.17), guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "gray"), guide = "none") +
    scale_size_continuous(range = c(1, 4)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    labs(x = expression("Transcript " * italic(r)[x]), y = NULL)
```

### 2025-01-10 scripts/zarchive/FigS5_6_moduleEigengeneRes.Rmd
All analysis code

---
title: "Figure S5: WGCNA module characterization"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/FigS5_6_moduleEigengeneRes/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/FigS5_6_moduleEigengeneRes/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load GO enrichment data
load(paste0(base_dir, "objects/", prefix, sep = "_", "SIGNED_SFT", soft_power, "_GO_RES.RDS")) # df_mods_go
```

### A | Modules significantly associated with SCZ per module eigengene analysis

```{r FigS5a.prep}
df_figS5a <- df_lm_dx %>% 
    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
    unnest(cols = c(data)) %>% 
    mutate(label = case_when(
        dx == "SCZ" & p_value < 0.05 & p_adj > 0.05 ~ "*",
        dx == "SCZ" & p_value < 0.05 & p_adj < 0.05 ~ "**",
        TRUE ~ ""
    )
    )
```

```{r FigS5a.plot}
figS5a <- df_figS5a %>% 
    ggplot(aes(x = kme_resids, y = dx, color = dx, fill = dx)) +
    geom_point(position = position_jitter(width = 0.1), size = 0.5) +
    geom_boxplot(alpha = 0.1, outlier.shape = NA) +
    # geom_text(aes(label = label, x = -Inf), color = "black", size = 6,
    #           vjust = 0.75, hjust = -0.1, check_overlap = TRUE) +
    facet_wrap(vars(module), ncol = 1, scales = "free_x") +
    scale_fill_manual(values = dx_colors) +
    scale_color_manual(values = dx_colors) +
    labs(x = "kME (corrected for covariates)", y = NULL,
         #caption = "* = p < 0.05; ** = FDR < 0.05",
         title = "A | SCZ kME-diagnosis \nassociations"
    ) +
    theme(legend.position = "none",
          axis.text.y = element_text(size = 10)
    )
```

```{r FigS5a.save, fig.height = 5, fig.width = 2.5}
figS5a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
)
```

### B | Significant SCZ module enrichments

```{r FigS5b.prep}
data(stop_words)
vague_terms <- c("positive", "negative", "regulation", "response", "protein", "activity",
                 "pathway", "process", "involved", "signaling")

## Identify significant modules
sig_modules <- df_lm_dx %>% 
    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
    pull(module)

# Filter for modules of interest
df_sig_go_res <- df_mods_go %>% 
    dplyr::filter(module %in% sig_modules)

## Identify top words in each significant (FDR < 0.05) module pathway
module_tokens_freq <- df_sig_go_res %>% 
    dplyr::filter(p.adjust < 0.05) %>% 
    group_by(module) %>% 
    unnest_tokens(word, Description) %>% 
    anti_join(stop_words) %>% 
    dplyr::filter(!(word %in% vague_terms)) %>% 
    dplyr::count(word, sort = TRUE)
```

```{r FigS5b.plot}
l_figS5b <- map(
    .x = as.character(sig_modules),
    .f = ~ module_tokens_freq %>% 
        dplyr::filter(module == .x) %>% 
        top_n(10, wt = n) %>% 
        ggplot(aes(label = word, size = n, color = module)) +
        geom_text_wordcloud() +
        facet_wrap(vars(module)) +
        scale_color_manual(values = module_colors) +
        scale_size_continuous(range = c(2.5, 5)) +
        theme_void() +
        theme(strip.text = element_blank(),
              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
        )
)
```

```{r FigS5b.save, fig.height = 5, fig.width = 2.5}
wrap_plots(l_figS5b, ncol = 1) & plot_annotation(title = "B | SCZ module enrichments")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
)
```

### S6 | WGCNA module enrichment for risk gene lists

```{r FigS6.prep}
df_figS6 <- df_mods_risk_gene_hypergeometric %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(sig = case_when(p_adj < 0.001 ~ "***", p_adj < 0.01 ~ "**", p_adj < 0.05 ~ "*", TRUE ~ ""))
```

```{r FigS6.plot}
figS6 <- df_figS6 %>% 
    ggplot(aes(x = benchmark_list, y = module)) +
    geom_tile(aes(fill = module, alpha = -log10(p_adj))) +
    geom_text(aes(label = sig), size = 6, color = "black", vjust = 0.75) +
    facet_wrap(vars(class), scales = "free_x", nrow = 1) +
    scale_fill_manual(values = module_colors, guide = "none") +
    guides(alpha = guide_legend(title = "-log10(FDR)")) +
    labs(x = NULL, y = NULL, 
         title = "S6 | Risk gene list enrichment in WGCNA modules")
```

```{r figS6.save, fig.width = 4.5, fig.height= = 4}
figS6
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "S6.risk_gene_module_hypergeometric", .x), width = 5, height = 4)
)
```

### 2025-01-10 scripts/zarchive/topic_modeling_demo/topic_modeling_demo.Rmd
All analysis code

---
title: "Topic modeling demo"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

### Load necessary libraries
```{r libraries}
library(tidyverse)
library(tidymodels)
library(tidylo)
library(stm)
library(purrr)
library(RColorBrewer)
library(ggwordcloud)
library(ggrepel)
library(patchwork)

#devtools::install_github("juliasilge/tidytext", force = TRUE) 
library(tidytext)

theme_set(theme_bw())
```

### Load data: publicly available text datasets for the Office dialogue across all seasons:
```{r data}
library(schrute)
df_office <- schrute::theoffice
speaking_characters <- df_office %>% 
  count(character, sort = TRUE) %>%

  # select only characters with at least 500 lines across all seasons
  filter(n > 500) %>% 
  pull(character)
```

### Unnest tokens: restructure as one-token-per-row (each word is a separate row in a tibble)
```{r unnest_tokens}
df_tokens <- df_office %>%

  # filter for our speaking characters
  filter(character %in% speaking_characters) %>%

  group_by(character) %>% 
  unnest_tokens(word, text) %>%

  # remove stop words (uninteresting words like of, from, and, the, etc.)
  anti_join(get_stopwords(), by = join_by(word))

head(df_tokens)
```

### Create a sparse matrix: each row is a corpus (book) and each column is a word
```{r sparse_matrix}
sparse_mat <- df_tokens %>%

  # count number of times a word is used by each character
  dplyr::count(character, word) %>%

  # filter for words that appear at least 3 times
  filter(n > 3) %>%

  # create sparse matrix
  cast_sparse(character, word, n)

dim(sparse_mat)
sparse_mat[1:10, 1:10]
```

### Fit topic model using Structural Topic Models (http://www.structuraltopicmodel.com/)
```{r topic_model}
set.seed(0306)
topic_model <- stm(sparse_mat, 
                   K = 4, # K is the number of 'topics' (clusters)
                   verbose = FALSE)

summary(topic_model)
```

**lift** = frequency divided by frequency in other topics
**FREX** weights words by frequency and exclusivity to the topic

## Ta-da!!

But.... how can we determine the optimal number of topics for our dataset?

```{r many_models}
df_many_models <- tibble(K = seq(3, 10, by = 1)) %>% 
  mutate(topic_model = map (

    .x = K,
    .f = ~ stm(sparse_mat, 
               K = .x, 
               verbose = FALSE)

  ))

head(df_many_models)
```

Now that we’ve fit all these topic models with different numbers of topics, we can explore how many topics are appropriate/good/“best”

```{r diagnostics, fig.height = 3}
df_diagnostics <- df_many_models %>%
  mutate(exclusivity = map(topic_model, exclusivity),
         semantic_coherence = map(topic_model, semanticCoherence, sparse_mat),
         residual = map(topic_model, checkResiduals, sparse_mat),
         bound =  map_dbl(topic_model, function(x) max(x$convergence$bound)),
         lfact = map_dbl(topic_model, function(x) lfactorial(x$settings$dim$K)),
         lbound = bound + lfact,
         iterations = map_dbl(topic_model, function(x) length(x$convergence$bound)))

head(df_diagnostics)

df_diagnostics %>%
  transmute(K,
            `Lower bound` = lbound,
            Residuals = map_dbl(residual, "dispersion"),
            `Semantic coherence` = map_dbl(semantic_coherence, mean)) %>%
  gather(Metric, Value, -K) %>%
  ggplot(aes(K, Value, color = Metric)) +
  geom_line(linewidth = 1.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(x = "K (number of topics)",
       y = NULL,
       title = "Model diagnostics by number of topics")
```

Semantic coherence is maximized when the most probable words in a given topic frequently co-occur together, and it’s a metric that correlates well with human judgment of topic quality. Having high semantic coherence is relatively easy, though, if you only have a few topics dominated by very common words, so you want to look at both semantic coherence and exclusivity of words to topics. It’s a tradeoff. Read more about semantic coherence in the original paper about it (https://dl.acm.org/citation.cfm?id=2145462.

```{r semantic_coherence, fig.height = 3}
df_diagnostics %>%
  select(K, exclusivity, semantic_coherence) %>%
  unnest(cols = c(exclusivity, semantic_coherence)) %>%
  filter(K %in% c(3, 5, 7)) %>% 
  mutate(K = as.factor(K)) %>%
  ggplot(aes(semantic_coherence, exclusivity, color = K)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(x = "Semantic coherence",
       y = "Exclusivity",
       title = "Comparing exclusivity and semantic coherence")
```

In choosing our final topic model, we should go with a K value that is statistically sound, but also contextually (in this case, biologically) meaningful and informative

### Finalize model
```{r final_model, fig.height = 6, fig.width = 8}
topic_model <- stm(sparse_mat, 
                   K = 6, # K is the number of 'topics' (clusters)
                   verbose = FALSE)

p_beta <- tidy(topic_model, matrix = "frex") %>%
  left_join(tidy(topic_model)) %>% 
  group_by(topic) %>%
  slice_head(n = 20) %>%
  mutate(topic = paste0("topic ", topic)) %>%

  ggplot(aes(x = beta, y = reorder(term, beta), fill = topic)) +
  geom_col() +
  labs(y = "Word") +
  facet_wrap(vars(topic), scales = "free", nrow = 1)

# Topic probabilities
group_gamma <- tidy(
  topic_model, 
  matrix = "gamma",
  document_names = rownames(sparse_mat)) %>% 
  mutate(chapter = factor(document)) %>% 
  dplyr::select(chapter, topic, gamma) %>% 
  mutate(topic = factor(topic)
  )

p_gamma <- group_gamma %>%
  filter(gamma > 0.01 ) %>% 
  ggplot(aes(x = topic, y = gamma, color = topic)) +
  geom_point(aes(alpha = gamma)) +
  geom_text_repel(aes(alpha = gamma, label = chapter), size = 3, 
                  max.overlaps = 50) +

  labs(y = expression(gamma), x = "") +
  ggtitle("Corpus (character) strength of association with each topic") +
  theme(legend.position = "none",
        axis.text.x = element_blank())

(p_gamma / p_beta)
```

### 2025-01-10 functions/f_load_CCA_res.R

@@ -1,0 +1,200 @@
+
+########################################################################################
+
+## Function that loads results from the RCCA toolkit and formats result into a table ##
+
+## INPUTS
+## - cca_directory = directory that contains CCA data & framework folders
+## - type = specify if genes or transcripts (character)
+## - analysis_type = CCA algorithm used, either rcca or grcca
+## - mu = mu (x group regularization parameter) value used in analysis (numeric or character ok; set to NULL if RCCA was run (though any value in there is fine, it won't be used))
+## - VARx = was variance explained in the X matrix optimized ("0.1_1") or set at 0.99?
+## - include_Cmat = logical indicating whether Cmat was included in the analysis (were confounds regressed?)
+
+########################################################################################
+
+f_load_cca_res <- function(cca_directory = cca_dir,
+                           level = c("gene", "transcript"),
+                           analysis_type = c("rcca", "grcca"),
+                           mu,
+                           lambda = "Nfeat",
+                           VARx = c("0.1_1", "0.99"),
+                           include_Cmat = c(TRUE, FALSE),
+                           rename_covariates = c(TRUE, FALSE)
+                           ) {
+    
+### Set paths & data split to load results ###
+    
+    print("Setting paths")
+    
+    ## Identify correct framework analysis directory
+    if (include_Cmat) {cmat <- "withCmat"} else {cmat <- "noCmat"}
+    if (analysis_type == "grcca") {group <- paste0("_groupL2x", as.character(mu))} else {group <- ""}
+    analysis_dir <- paste0(analysis_type, "_permutation_VARx", VARx, "_L2x", lambda, group, "_", cmat, "_allEffects")
+    
+    ## Combine paths to point to results (and check to make sure directory exists)
+    results_dir <- paste0(cca_dir, "framework/", analysis_dir)
+    if (!dir.exists(results_dir)) {stop("Analysis directory does not exist\n", results_dir)}
+    
+    ## Identify final model path (weights)
+    model_path <- paste0(results_dir, "/res/level1/model_1.mat")
+    
+    # Identify bath to bootstrapped weights (to calculate sd)
[…]

### 2025-01-10 functions/hypergeometric_functions.R

@@ -1,0 +1,127 @@
+
+# General hypergeometric function -----------------------------------------
+
+#=====================================================================================================================#
+
+### GOAL: Test hypergeometric overlap between two gene lists
+
+### INPUTS ###
+
+## gene_list1 = list of genes (can be any type of ID, e.g., Ensembl ID or HGNC symbol)
+## gene_list2 = list of genes (must be same type as gene_list1)
+## gene_universe = list of all the genes that were included in the analysis (same ID type as gene_list1 & gene_list2)
+
+### OUTPUTS ###
+
+## overlap_genes = list of genes that were in both gene_list1 and gene_list2
+## p_value = hypergeometric p-value (significant of overlap given total pool of genes)
+
+#=====================================================================================================================#
+
+f_hypergeometric <- function(gene_list1, gene_list2, gene_universe) {
+    
+    # Include only genes that were part of universe
+    gene_list1_in_universe <- intersect(gene_universe, gene_list1) %>% unique
+    gene_list2_in_universe <- intersect(gene_universe, gene_list2) %>% unique
+    
+    # Identify the overlap between the published gene list (including only those in our universe) and the gene list of interest
+    overlap_genes <- intersect(gene_list1_in_universe, gene_list2_in_universe)
+    
+    # Run hypergeometric test
+    phyper <- 1 - phyper(q = length(overlap_genes), 
+                         m = length(gene_list2_in_universe), 
+                         n = length(gene_universe) - length(gene_list2_in_universe), 
+                         k = length(gene_list1_in_universe)
+    ) 
+    
+    # Return a list of common genes and the hypergeometric p-value
+    return(list("overlap_genes" = overlap_genes, "p_value" = phyper))
+    
+}
+
+
+
+# WGCNA module hypergeometric ----------------------------------------
+
+#=====================================================================================================================#
+
+### GOAL: Specific case of hypergeometric function to test overlap of a gene list with WGCNA module
+
+### INPUTS ###
+
+## gene_list = list of genes (can be any type of ID, e.g., Ensembl ID or HGNC symbol)
+## gene_list_name = name of the gene list that you want to include for the output table
+## wgcna_module = name of the WGCNA-derived module of interest
+## gene_universe = list of all the genes that were included in the analysis (same ID type as gene_list1 & gene_list2)
[…]

### 2025-01-10 functions/plot_functions.R

@@ -1,0 +1,302 @@
+
+###################################################################
+
+# Functions to generate all plots from these analyses
+
+###################################################################
+
+
+# Plot theme --------------------------------------------------------------
+
+## Color themes
+dx_colors <- c("Control" = "#0072B2", "BD" = "#9966FF", "MDD" = "#009E73", "SCZ" = "#E69F00")
+ontology_colors <- c("BP" = "#F8766D", "CC" = "#00BA38", "MF" = "#619CFF")
+gene_weight_color_scale <- brewer.rdbu(100) %>% rev
+
+## Plot aesthetics
+theme_set(
+    theme_cowplot() +
+        theme(plot.title = element_text(size = 11),
+              strip.text = element_text(size = 10, color = "black"),
+              axis.title = element_text(size = 10),
+              axis.text = element_text(size = 9),
+              legend.title = element_text(size = 9),
+              legend.text = element_text(size = 8),
+              legend.margin = margin(t = 0, r = 0, b = 0, l = 0),  # Remove legend margin
+              plot.margin = margin(t = 0, r = 5, b = 0, l = 0)
+        )
+)
+
+
+# plot GO results by ontology ---------------------------------------------
+
+f_plot_go_by_ontology <- function(df_go_res,
+                                  n_top_pathways = 10,
+                                  pathway_text_width = 40,
+                                  pathway_text_size = 9,
+                                  plot_title = "GO pathway results",
+                                  n_facet_rows = 2,
+                                  legend_position = c(0.7, 0.25)
+) {
+    df_go_res %>% 
+        
+        # clean dataframe (if not already)
+        clean_names %>% 
+        
+        # take top 10 paths by p-value per ontology to plot
+        group_by(ontology) %>% 
+        arrange(-p_adjust) %>% 
+        mutate(row = row_number(),
+               gene_ratio = parse(text = gene_ratio) %>% eval
+        ) %>% 
+        top_n(n = n_top_pathways, wt = row) %>% 
+        
+        # indicate whether pathway survived FDR correction
+        mutate(`survives FDR` = ifelse(p_adjust < 0.05, "yes", "no")) %>% 
+        mutate(description = str_wrap(description, width = pathway_text_width)) %>%
+        
+        # plot
+        ggplot(aes(x = -log10(p_adjust), y = reorder_within(description, within = ontology, by = -p_adjust))) +
[…]

### 2025-01-10 functions/zarchive/cell_type_hypergeometric_overlap.R

@@ -1,0 +1,39 @@
+f_cell_type_overlap <- function(df,
+                                module_no, 
+                                cell_type) {
+  
+  # pull genes in module
+  module_genes <- df %>% 
+    dplyr::filter(module == module_no) %>% 
+    pull(ensembl_gene_id)
+  
+  # pull genes in cell_type
+  cell_type_genes <- df %>% 
+    dplyr::filter(type == cell_type) %>% 
+    pull(ensembl_gene_id)
+  
+  # find intersect of module with gene_universe_intersect genes
+  module_intersect <- intersect(gene_universe_intersect, module_genes)
+  
+  # find intersect of cell_type with gene_universe_intersect genes
+  cell_type_intersect <- intersect(gene_universe_intersect, cell_type_genes)
+  
+  # find intersect of module intersects
+  module_cell_type_intersect <- intersect(module_intersect, cell_type_intersect)
+  
+  # set variables for hypergeometric function
+  q <- (module_cell_type_intersect %>% length) - 1 # overlap between two modules - 1
+  m <- module_intersect %>% length # module 1 
+  n <- (gene_universe_intersect %>% length) - m # pop size - module 1
+  k <- cell_type_intersect %>% length # module 2
+  
+  # return tibble
+  tibble(module = module_no,
+         type = cell_type,
+         q = q,
+         m = m,
+         n = n,
+         k = k,
+         p_val = phyper(q, m, n, k, lower.tail = FALSE, log.p = FALSE))
+  
+}

### 2025-01-10 scripts/00.preprocessing/00.clean_covariates.R

@@ -1,0 +1,163 @@
+
+######################################################################################
+
+# Tidy covariate data for downstream analyses
+
+######################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(readxl)
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+
+# complete covariates from Ajeet
+df_complete_covariates <- read_xlsx(paste0(base_dir, "data/14Nov2023_all_covariates.xlsx"))
+
+# other covariates from Nirmala (includes additional technical covariates)
+df_nirmala_covariates <- read.table(paste0(base_dir, "data/185samples_multinomialDx_46covariates_10evecs.txt"), 
+                                    sep = "\t", header = TRUE) %>% 
+  dplyr::select(-contains("evec")) %>% 
+  as_tibble() %>% 
+  mutate_if(is.character, ~ ifelse(.x == "", NA_character_, .x)) %>% 
+  clean_names %>% 
+  dplyr::rename("rna_extraction_batch" = "rn_aextraction_batch", 
+                "gc_percent" = "g_cpercent",
+                "dx" = "multi_nomial_dx") %>% 
+  dplyr::select(-contains("_1")) %>% # remove duplicate columns
+  mutate(sample = tolower(sample))
+
+# define covariate types
+technical_covariates <- c("source", "pmi", "ph", "pmi_confidence", "max_rin", "max_rine", "mapped_percent", "five_prime_three_prime_bias",
+                          "rna_extraction_batch", "library_batch", "gc_percent")
+biological_covariates <- c("age_death", "sex_at_birth", "race", "marital_status", "education", "manner_death", "suicide", "brain_weight",
+                           "height", "weight", "bmi")
+drug_covariates <- c("smoker", "nicotine_cotinine", "alcohol", "sedative_hypnotic_anxiolitics", "opioids", "cannabinoids",
+                     "major_stimulants_cocaine_included", "minor_stimulants", "anticholinergics", "antidepressants",
+                     "anti_epileptics", "anti_histamines", "antipsychotics", "mood_stabilizers", "non_psychiatric",
[…]

### 2025-01-10 scripts/00.preprocessing/02.gene_raw_count_preprocessing.R

@@ -1,0 +1,433 @@
+
+######################################################################################
+
+### GENES ###
+# Read in raw count data, normalize using DESeq2, qSVA correction & regress covariates
+
+######################################################################################
+
+
+# Setup -------------------------------------------------------------------
+
+
+### SET DIRECTORIES ###
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+### LOAD LIBRARIES ###
+
+library(tidyverse)
+library(DESeq2)
+
+### LOAD DATA ###
+
+## Cleaned covariates
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in 00.clean_covariates.R)
+
+## Gene raw counts
+df_raw_counts <- read.table(paste0(base_dir, "data/185samples_allGenes_auto-PARs_noSexChrs_noFilters_KoryAnalysis_21Kgenes.txt")) %>% 
+  as_tibble(rownames = "ensembl_gene_id") %>% 
+  rename_at(2:ncol(.), ~substr(.x, start = 2, stop = nchar(.x))) %>% 
+  clean_names() %>% 
+  dplyr::rename_at(2:ncol(.), ~str_remove(.x, "x"))
+
+## Filter for samples that are in gene data
+df_raw_counts <- df_raw_counts %>% 
+  dplyr::select(ensembl_gene_id, contains(df_covariates_numeric$sample))
+
+## Load QSVs (generated in 01.qSVA.R)
+load(paste0(base_dir, "objects/06March2024_qSVs.Rdata"))
+
+# COMBINE qSVS WITH COVARIATE DATA  
+df_covar_qsv <- df_covariates_numeric %>% left_join(df_qsvs)
+
+# set order
+m_covariates <- df_covariates_numeric %>% column_to_rownames("sample") %>% t()
+covariate_order <- m_covariates[hclust(dist(m_covariates))$order,] %>% rownames
+
+
+# Remove low count genes --------------------------------------------------
+
[…]

### 2025-01-10 scripts/00.preprocessing/04.transcript_filtering.R

@@ -1,0 +1,90 @@
+
+################################################################################################################
+
+# TRANSCRIPTS: Filter transcripts based on coefficient of variation; i.e., how variable is the transcript across
+# samples given its expression levels?
+
+################################################################################################################
+
+
+# Setup -------------------------------------------------------------------
+
+## Set directories
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+## Load regressed transcript data (all transcripts)
+prefix <- "08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC"
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress (generated in 03.transcript_raw_count_preprocessing.R)
+
+## Load raw transcript counts
+df_transcript_raw_counts <- read.csv(paste0(base_dir, "data/transcript_count_matrix.csv")) %>% 
+    as_tibble() %>% 
+    dplyr::rename("ensembl_transcript_id" = "X") %>% 
+    rename_all(~str_remove(.x, "_MergedBam_stringtieOutput")) %>% 
+    clean_names() %>% 
+    rename_all(~str_remove(.x, "x"))
+
+# Calculate transcript coefficient of variation ----------------------------------------------
+
+
+## "Selecting information features and reducing dimensionality"
+# https://www.nature.com/articles/s41576-023-00586-w
+# https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-10-193 (pick arbitrary threshold?)
+
+# However, we don't just want to filter on variance because variance scales with mean and some transcripts have very low expression
+# So we can use coef of variance, which filters on variance relative to mean
+
+
+### Filter by sd/mean (CV) of raw counts across samples ###
+
+## Calculate CV of each transcript
+df_mean_sd <- df_transcript_raw_counts %>% 
+    filter(ensembl_transcript_id %in% colnames(df_vsd_regress)) %>% 
+    pivot_longer(2:ncol(.), names_to = "sample", values_to = "raw_counts") %>% 
+    group_by(ensembl_transcript_id) %>% 
+    summarise(
[…]

### 2025-01-10 scripts/02.DGE/00.DGE_analysis.R

@@ -1,0 +1,79 @@
+
+######################################################################
+
+# Updated DGE analyis including all covariates used in WGCNA and GRCCA
+
+######################################################################
+
+## Set directories & load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig2_DGEres/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_genes.R"))
+
+
+# Convert count data to matrix for DESeq2; start by running only SCZ-control analyses ----------------------------
+
+m_vsd_regress <- df_vsd_regress %>% 
+    dplyr::filter(str_detect(sample, "schizo|control")) %>% # subset for only schizophrenia-control comparisons
+    column_to_rownames("sample") %>% 
+    t() + 10 # add pseudo-count of 10 so there are no negative values in the matrix
+m_vsd_regress <- round(m_vsd_regress*10^5) # multiply by 10000 and round to convert to integers but maintain precision
+
+
+# Generate covariates matrix ----------------------------------------------
+
+## Combine drug MCs with known covariates
+df_covariates_mca <- df_covariates %>% 
+    left_join(
+        df_ind_loadings %>% 
+            dplyr::rename_at(vars(contains("dim")), ~ str_replace(.x, "dim", "MC")),
+        by = join_by(sample)
+    )
+
+## Convert to matrix
+n_mcs <- 8
+paste0("MC", 1:n_mcs)
+m_covariates <- df_covariates_mca %>% 
+    dplyr::filter(dx %in% c("SCZ", "Control")) %>% # SCZ vs controls only
+    mutate(dx = factor(dx, levels = c("Control", "SCZ"))) %>% 
+    dplyr::select(sample, dx, all_of(paste0("MC", 1:n_mcs)), brain_weight) %>% 
+    column_to_rownames("sample")
+
+
+# Run DESeq ---------------------------------------------------------------
+
+## Create DESeq2 object
+dds <- DESeqDataSetFromMatrix(countData = m_vsd_regress, 
+                              colData = m_covariates, 
+                              design = ~ dx + 
[…]

### 2025-01-10 scripts/02.DGE/01.characterize_DEGs.R

@@ -1,0 +1,62 @@
+
+######################################################################
+
+# Gene ontology and gene set enrichment analysis of current study DEGS
+
+######################################################################
+
+### Set directories & load data ###
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig2_DGEres/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_genes.R"))
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res (generated in A.DGE_analysis.R)
+
+
+### Get ranked gene list ###
+my_degs_ranked <- df_de_res %>% 
+    arrange(-stat) %>% 
+    dplyr::select(ensembl_gene_id, log2fold_change) %>% 
+    deframe
+
+
+### Cell-type GSEA ###
+df_de_fgsea_cell <- fgsea(pathways = cell_types, 
+                               stats = my_degs_ranked, 
+                               eps = 0
+)  %>% 
+    as_tibble() %>% 
+    dplyr::rename("cell_type" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### GO GSEA ###
+df_de_fgsea_go <- fgsea(pathways = go_pathways, 
+                        stats = my_degs_ranked, 
+                        eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("go_id" = "pathway") %>% 
+    arrange(-abs(NES)) %>% 
+    left_join(df_go_terms) %>% # add names of GO terms
+    dplyr::select(go_id, term, ontology, everything())
+
+
+### Benchmark lists ###
+df_de_fgsea_benchmark <- fgsea(pathways = benchmarking_lists, 
+                               stats = my_degs_ranked, 
+                               eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("benchmark_list" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### Save objects for figures ###
+save(df_de_fgsea_cell, df_de_fgsea_go, df_de_fgsea_benchmark,
+     file = paste0(analysis_objects_dir, "DE_enrichment_benchmark_res.Rdata")
+)
+

### 2025-01-10 scripts/02.DGE/DEPR.benchmark_DEGs.R

@@ -1,0 +1,109 @@
+
+###################################################################
+
+# Quantify DEG overlap with published SCZ gene lists                        
+
+###################################################################
+
+## Set directories & load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig2_DGEres/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig2_DGEres/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_genes.R"))
+
+
+# Identify SCZ-control DEGs -------------------------------
+
+## DEGs
+my_degs <- df_de_res %>% 
+    dplyr::filter(pvalue < 0.05) %>% 
+    pull(ensembl_gene_id)
+length(my_degs) # 1690
+
+### DEGs (ranked) ###
+my_degs_ranked <- df_de_res %>% 
+    arrange(-log2fold_change) %>% 
+    dplyr::select(ensembl_gene_id, log2fold_change) %>% 
+    deframe
+
+## Identify max L2FC (for plot color scales)
+scale_max <- df_de_res %>% pull(log2fold_change) %>% abs %>% max
+
+
+
+# Run hypergeometric test (testing for overall enrichment) on DEGs and each benchmark list -----------------
+
+df_hypergeometric_res <- tibble(
+    benchmark_list = names(benchmarking_lists)
+) %>% 
+    mutate(
+        benchmark_list_n = map(
+          .x = benchmark_list,
+          .f = ~ length(benchmarking_lists[[.x]])
+        ),
+        benchmark_list_n_in_universe = map(
+            .x = benchmark_list,
+            .f = ~ intersect(benchmarking_lists[[.x]], ensembl_gene_ids) %>% length
+        ),
+        hypergeometric_test = map(
+            .x = benchmark_list,
+            .f = ~ f_hypergeometric(gene_list1 = my_degs, 
+                                    gene_list2 = benchmarking_lists[[.x]],
+                                    gene_universe = ensembl_gene_ids
+            )
+        ),
+        overlap_n = map(
+            .x = hypergeometric_test,
+            .f = ~ length(.x$overlap_genes)
+        ),
[…]

### 2025-01-10 scripts/03.WGCNA/00.case_control_WGCNA.R

@@ -1,0 +1,159 @@
+
+########################################################################################
+
+# Select soft-thresholding power, calculate adjacency & TOM, cluster, and assign modules
+
+########################################################################################
+
+
+### Set directories & load data ###
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/FigS3_WGCNAenrichment/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_genes.R"))
+
+
+# Select soft thresholding power ------------------------------------------
+
+  
+# CONVERT REGRESSED VSD COUNTS TO TRANSPOSED MATRIX
+  m_vsd_regress <- df_vsd_regress %>% column_to_rownames("sample")
+  
+# CHOOSE A SET OF SOFT-THRESHOLDING POWERS TO TEST
+  powers <- 1:16
+
+# CALL THE NETWORK TOPOLOGY ANALYSIS FUNCTION
+  sft <- pickSoftThreshold(m_vsd_regress, powerVector = powers, verbose = 5)
+  
+  # convert results to tibble
+  df_sft <- sft$fitIndices %>% as_tibble %>% clean_names
+  
+  # save
+  save(df_sft, file = paste0(base_dir, "objects/", prefix, "_SFT.RDS")) # for downstream analyses
+  save(df_sft, file = paste0(analysis_objects_dir, prefix, "_SFT.RDS")) # for supplement
+
+
+  
+# Calculate adjacency matrix, TOM, clustering ------------------------------------
+
+  doParallel::registerDoParallel()
+  
+# CONVERT REGRESSED VSD COUNTS TO TRANSPOSED MATRIX
+  m_vsd_regress <- df_vsd_regress %>% column_to_rownames("sample")
+  
+# SELECT SFT
+  soft_power <- 3
+
+# CALCULATE CO-EXPRESSION SIMILARITY AND ADJACENCY
+  set.seed(20240306)
+  adjacency <- adjacency(m_vsd_regress, type = "signed hybrid", power = soft_power)
+
+# TOPOLOGICAL OVERLAP MATRIX (TOM)
+  set.seed(20240306)
+  TOM <- TOMsimilarity(adjacency, TOMType = "signed")
[…]

### 2025-01-10 scripts/03.WGCNA/01.characterize_module_enrichment.R

@@ -1,0 +1,147 @@
+
+########################################################################################
+
+# GENE LEVEL: run GO and cell-type enrichment on WGCNA modules
+
+########################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+# Run GO on each WGCNA module --------------------------------------------
+
+  
+## Run GO pathway overrepresentation analysis on each module
+modules <- df_modules_filt %>% pull(module) %>% unique
+doParallel::registerDoParallel()
+df_mods_go <- map_dfr(
+  .x = modules,
+  .f = ~ enrichGO(gene = df_modules_filt %>% dplyr::filter(module == .x) %>% pull(ensembl_gene_id),
+                  OrgDb = "org.Hs.eg.db",
+                  universe = ensembl_gene_ids,
+                  keyType = "ENSEMBL",
+                  ont = "ALL",
+                  pAdjustMethod = "BH",
+                  pvalueCutoff = 1,
+                  qvalueCutoff = 1,
+                  readable = TRUE
+  ) %>% 
+    as_tibble() %>% 
+    mutate(module = .x, .before = 1)
+)
+
+
+## Save results for figures & tables ##
+
+
+# Figures
+save(df_mods_go, file = paste0(analysis_objects_dir, "WGCNA_module_GO.RDS"))
+
+# Downstream analyses
+save(df_mods_go, file = paste0(base_dir, "objects/", prefix, sep = "_", "SIGNED_SFT", soft_power, "_GO_RES.RDS"))
+  
+
+
+
+# Run topic modeling on GO results for each module ------------------------------------------------------
+
+#load(paste0(analysis_objects_dir, "WGCNA_module_GO.RDS"))
+
+## Unnest tokens to get GO terms within each module (group)
+df_go_tokens <- df_mods_go %>% 
+    dplyr::filter(pvalue < 0.05) %>% 
+    dplyr::filter(!is.na(Description)) %>% 
+    group_by(module) %>% 
+    unnest_tokens(word, Description)
+
+## Create a sparse matrix using only meaningful words that appear at least twice within module
[…]

### 2025-01-10 scripts/05b.GRCCA/01.GRCCA_results.R

@@ -1,0 +1,66 @@
+
+################################################################################
+
+# GENE-LEVEL: Characterize GRCCA results (gene weights & enrichments)
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      paste0(dx_groups_to_analyze, collapse = ""), "_",
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "gene",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    VARx = VARx,
+    include_Cmat = FALSE,
+    rename_covariates = TRUE
+)
+
+## Assign results to tibbles; negate everything so SCZ is positive
+df_results <- l_grcca_res$model_results
+df_lvs <- l_grcca_res$latent_variables %>% mutate_at(vars(c(lvx, lvy)), ~ -.x)
+df_y_res <- l_grcca_res$y_res %>% mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x)
+df_x_res <- l_grcca_res$x_res %>% 
+    mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x) %>% 
+    left_join(df_ensembl_to_symbol) %>% 
+    left_join(df_modules_filt %>% dplyr::select(-c(mod_set, color)), by = join_by(ensembl_gene_id)) %>% 
[…]

### 2025-01-10 scripts/05b.GRCCA/02.characterize_GRCCA_res_GSEA.R

@@ -1,0 +1,64 @@
+
+################################################################################
+
+# GENE-LEVEL: Validate GRCCA results using external data
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig4_characterizeGRCCA/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+### Get ranked gene list ###
+my_grcca_ranked <- df_x_res %>% 
+    arrange(-pearsons_r) %>% 
+    dplyr::select(ensembl_gene_id, pearsons_r) %>% 
+    deframe
+
+
+### Cell-type GSEA ###
+df_grcca_fgsea_cell <- fgsea(pathways = cell_types, 
+                          stats = my_grcca_ranked, 
+                          eps = 0
+)  %>% 
+    as_tibble() %>% 
+    dplyr::rename("cell_type" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### GO GSEA ###
+df_grcca_fgsea_go <- fgsea(pathways = go_pathways, 
+                        stats = my_grcca_ranked, 
+                        eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("go_id" = "pathway") %>% 
+    arrange(-abs(NES)) %>% 
+    left_join(df_go_terms) %>% # add names of GO terms
+    dplyr::select(go_id, term, ontology, everything())
+
+
+### Benchmark lists ###
+df_grcca_fgsea_benchmark <- fgsea(pathways = benchmarking_lists, 
+                               stats = my_grcca_ranked, 
+                               eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("benchmark_list" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### Save objects for figures ###
+save(df_grcca_fgsea_cell, df_grcca_fgsea_go, df_grcca_fgsea_benchmark,
+     file = paste0(analysis_objects_dir, "GRCCA_enrichment_benchmark_res.Rdata")
+)
+

### 2025-01-10 scripts/05b.GRCCA/03.developmental_trajectories.R

@@ -1,0 +1,50 @@
+
+################################################################################
+
+# Characterize normative developmental expression of GRCCA results deciles
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig4_characterizeGRCCA/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/oad_WGCNA_res.R"))
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+
+# Developmental trajectories (data) ----------------------------------------------
+
+load(paste0(base_dir, "objects/psychencode_development_expr_data.Rdata")) # df_psychencode_metadata, df_psychencode_expr
+
+## Identify samples from region of interest
+df_sample_window <- df_psychencode_metadata %>% 
+    dplyr::filter(region_broad == "frontal_cortex")
+frontal_cortex_samples <- df_sample_window%>% 
+    pull(id)
+
+## Filter expression data for samples and genes of interest
+df_psychencode_expr_filt <- df_psychencode_expr %>% 
+    dplyr::filter(GENE %in% ensembl_gene_ids) %>% 
+    dplyr::select(GENE, all_of(frontal_cortex_samples)) %>% 
+    dplyr::rename("ensembl_gene_id" = "GENE")
+
+## Combine developmental expression data with GRCCA results (split into deciles)
+df_psychencode_grcca <- df_psychencode_expr_filt %>% 
+    pivot_longer(starts_with("HSB"), names_to = "id", values_to = "dev_expression") %>% 
+    left_join(df_sample_window, by = join_by(id)) %>% 
+    dplyr::select(window, id, regioncode, ensembl_gene_id, dev_expression) %>% 
+    left_join(df_x_res %>% 
+                  mutate(decile = ntile(-pearsons_r, 10) %>% factor)
+    )
+
+## Save for figures
+save(df_psychencode_grcca,
+     file = paste0(analysis_objects_dir, "GRCCA_developmental_expression.RDS")
+)
+

### 2025-01-10 scripts/05b.GRCCA/04.compare_to_literature.R

@@ -1,0 +1,65 @@
+
+################################################################################
+
+# Relate GRCCA results vector to SCZ literature
+
+################################################################################
+
+
+### SETUP ###
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+### SNAP PAPER DATA ###
+
+## Load SNAP paper LF loadings
+df_loading_factors <- read_xlsx("~/Downloads/41586_2024_7109_MOESM6_ESM.xlsx", sheet = "Gene loadings")
+
+## Subset for LF 4
+df_lf4 <- df_loading_factors %>% 
+    dplyr::select(id, PEER_4) %>% 
+    separate(id, into = c("gene_symbol", "cell_type"), sep = "_")
+
+## Combine with GRCCA results
+df_lf4_grcca <- df_lf4 %>% 
+    left_join(df_x_res, by = join_by(gene_symbol))
+
+
+### SAVE FOR FUTURE ANALYSES ###
+save(df_lf4_grcca, file = paste0(analysis_objects_dir, "SNAP_LF4.RDS"))
+
+
+
+### PLOT ###
+df_lf4 %>% 
+    left_join(df_de_res, by = join_by(gene_symbol)) %>% 
+    left_join(df_lake_cell_type) %>% 
+    dplyr::filter((type == "Astro" & cell_type == "astrocyte") |
+                      (type == "Neuro-Ex" & cell_type == "glutamatergic") |
+                      (type == "Neuro-In" & cell_type == "gabaergic")
+    ) %>% 
+    
+    ggplot(aes(x = PEER_4, y = stat)) +
+    geom_vline(aes(xintercept = 0), color = "gray") +
+    geom_hline(aes(yintercept = 0), color = "gray") +
+    geom_point(aes(color = stat), size = 0.5) +
+    geom_smooth(color = "black", method = "lm", linewidth = 0.75) +
+    stat_cor(label.sep = "\n", size = 3, 
[…]

### 2025-01-10 scripts/05b.GRCCA/DEPR-load_GRCCA_results.R

@@ -1,0 +1,181 @@
+
+########################################################################################
+
+# Load gene-level GRCCA results to analyze
+
+########################################################################################
+
+# Setup -------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+# Set paths & data split to load results -----------------------------------------------------
+
+## Set paths to read in RCCA res (generated in cluster)
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/") # _WITHGRAY, _withGray
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+analysis_dir <- "grcca_permutation_VARx0.1_1_L2xNfeat_groupL2x0.1_noCmat_allEffects"
+model_path <- paste0(cca_dir, "framework/", analysis_dir, "/res/level1/model_1.mat")
+boot_path <- paste0(cca_dir, "framework/", analysis_dir, "/boot/level1/allboot_1.mat")
+
+## Read results table output
+df_results <- read_table(paste0(cca_dir, "framework/", analysis_dir, "/res/level1/results_table.txt"))
+
+## Select the split (defined by variance explained) with the lowest p-value &
+## highest X-Y latent variable correlation
+best_split <- df_results %>% arrange(pval, -correl) %>% pull(set) %>% .[[1]]
+
+
+# Load X.mat & Y.mat ------------------------------------------------------
+
+## Y.mat
+Y.mat <- read.table(paste0(cca_dir, "data/Y.txt"))
+df_y <- Y.mat %>% 
+    rownames_to_column("sample") %>% 
+    as_tibble()
+covariates <- df_y %>% dplyr::select(-sample) %>% colnames
+
+# rename covariates for plotting
+covariates_renamed <- case_when(
+    covariates == "control" ~ "Control",
+    covariates == "bipolar" ~ "BD",
+    covariates == "mdd" ~ "MDD",
[…]

### 2025-01-10 scripts/05b.GRCCA/TRY-GRCCA_BDMDD_only.R

@@ -1,0 +1,53 @@
+
+################################################################################
+
+# Remove SCZ from Y mat to see if we get a BD or MDD signal?
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig3_GRCCAres/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig3_GRCCAres/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      paste0(dx_groups_to_analyze, collapse = ""), "_",
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "gene",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    VARx = VARx,
+    include_Cmat = FALSE,
+    rename_covariates = FALSE
+)
+
+## Assign results to tibbles; negate everything so SCZ is positive
+df_results <- l_grcca_res$model_results
+df_lvs <- l_grcca_res$latent_variables %>% mutate_at(vars(c(lvx, lvy)), ~ -.x)
+df_y_res <- l_grcca_res$y_res %>% mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x)
+df_x_res <- l_grcca_res$x_res %>% 
+    mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x) %>% 
+    left_join(df_ensembl_to_symbol) %>% 
[…]

### 2025-01-10 scripts/05b.GRCCA/TRY-compare_DEGs_GRCCA.R

@@ -1,0 +1,106 @@
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig4_characterizeGRCCA/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig4_characterizeGRCCA/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+## Load DEGs
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
+
+## Combine DE & GRCCA res
+df_grcca_de_res <- df_de_res %>% 
+    left_join(df_x_res)
+
+
+
+# Astrocytes? -------------------------------------------------------------
+
+cell_types[["Astro"]]
+df_astrocyte_genes <- df_grcca_de_res %>% 
+    dplyr::filter(ensembl_gene_id %in% cell_types[["Astro"]])
+
+df_astrocyte_genes %>% 
+    mutate(label = ifelse(abs(log2fold_change) > 0.075, gene_symbol, "") ) %>% 
+    ggplot(aes(x = log2fold_change, y = pearsons_r)) +
+    geom_point() +
+    geom_text_repel(aes(label = label)) +
+    stat_cor()
+
+high_FC_low_r <- df_astrocyte_genes %>% 
+    dplyr::filter(abs(log2fold_change) > 0.075) %>% 
+    pull(ensembl_gene_id)
+df_vsd_regress %>% 
+    dplyr::select(sample, all_of(high_FC_low_r)) %>% 
+    left_join(df_covariates) %>% 
+    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expression_value") %>% 
+    left_join(df_ensembl_to_symbol, by = join_by(ensembl_gene_id)) %>% 
+
+    ggplot(aes(x = mood_stabilizers, y = expression_value)) +
+    geom_point(position = position_jitter(width = 0.2), size = 0.75) +
+    geom_boxplot(fill = "transparent", outlier.shape = NA) +
[…]

### 2025-01-10 scripts/05b.GRCCA/TRY-other_stuff.R

@@ -1,0 +1,240 @@
+
+################################################################################
+
+# GENE-LEVEL: Other things to enhance GRCCA results...
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig4_characterizeGRCCA/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig4_characterizeGRCCA/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+
+# gene-eQTL mapping -------------------------------------------------------
+
+
+df_eqtls <- read.table("~/Downloads/DER-08a_hg38_eQTL.significant.txt") %>% 
+    row_to_names(1) %>%
+    as_tibble()
+
+
+df_eqtls_grcca <- df_eqtls %>% 
+    mutate(ensembl_gene_id = str_remove(gene_id, "\\..*")) %>% 
+    dplyr::select(ensembl_gene_id, SNP_id, nominal_pval, FDR, regression_slope, top_SNP) %>% 
+    left_join(df_x_res) %>% 
+    dplyr::filter(!is.na(pearsons_r))
+
+## Summarise genes by taking median regression slope across eQTLs
+df_eqtls_grcca_avg <- df_eqtls_grcca %>%
+    group_by(ensembl_gene_id) %>% 
+    mutate(regression_slope = as.numeric(regression_slope)) %>% 
+    summarise(
+        regression_slope = median(regression_slope),
+        pearsons_r = median(pearsons_r)
+    )
+
+## Summarise genes by taking only the top SNP
+df_eqtls_grcca_top <- df_eqtls_grcca %>%
+    dplyr::filter(top_SNP == 1) %>% 
+    mutate(regression_slope = as.numeric(regression_slope))
+
+## Plot
+df_eqtls_grcca_top %>% 
+    #mutate(eqtl_decile = ntile(-regression_slope, 10) %>% factor) %>% 
+    
+    #dplyr::filter(regression_slope > 0) %>% 
+    ggplot(aes(x = regression_slope, y = pearsons_r)) +
+    geom_point() +
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/00.case_control_WTCNA.R

@@ -1,0 +1,263 @@
+
+##### !!!!! NOTE: transcripts have already been filtered pre-WTCNA! !!!!
+
+#########################################################################################################################
+
+# TRANSCRIPTS: Select soft-thresholding power, calculate adjacency & TOM, cluster, and assign modules
+
+#########################################################################################################################
+
+####### ****************** RUN ON CLUSTER ********************************* #######
+
+### Setup ###
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_transcripts.R"))
+
+# Set directories to save objects
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+
+# (CLUSTER) Select soft thresholding power ------------------------------------------
+
+
+# # CHOOSE A SET OF SOFT-THRESHOLDING POWERS TO TEST
+# powers <- 1:16
+# 
+# # CALL THE NETWORK TOPOLOGY ANALYSIS FUNCTION
+# sft <- pickSoftThreshold(m_vsd_regress_t, powerVector = powers, verbose = 5)
+# 
+# # convert results to tibble
+# df_sft <- sft$fitIndices %>% as_tibble %>% clean_names
+# 
+# # save
+# save(df_sft, file = paste0(base_dir, "objects/", prefix2, "_SFT.RDS"))
+
+# PLOT RESULTS   
+
+#### CODE FOR CLUSTER
+
+# IN LOCAL SPACE: transfer matrix to cluster
+# cd ~/Documents/PhD/projects/sgacc_wgcna_grcca/objects
+# rsync 08Feb2024_OVERALLsubs_regress_qSV1.7_ageDeath_RNAbatch_GCperc_matrix.RDS smithral@biowulf.nih.gov:/data/smithral/sgacc_wgcna/transcripts/data
+
+# ON CLUSTER
+# ssh smithral@biowulf.nih.gov
+# cd /data/smithral/sgacc_wgcna/transcripts
+# spersist --mem=500g --gres=lscratch:800
+# module load rstudio R
+# Rscript scripts/00.calc_sft.R
+
+# IN LOCAL SPACE
+# cd ~/Documents/PhD/projects/sgacc_wgcna_grcca/objects
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/01.define_input_matrices.R

@@ -1,0 +1,148 @@
+
+################################################################################
+
+# Define X.mat, Y.mat, and C.mat to include in GRCCA analysis
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_WTCNA_res.R"))
+
+
+### SELECT samples to include ###
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+samples_to_include <- df_covariates %>% 
+    filter(dx %in% dx_groups_to_analyze) %>% 
+    pull(sample) %>% 
+    unique
+
+### X.mat ###
+
+# NEW: filter transcripts for risk genes derivatives and/or genes that were interesting in GRCCA
+load(paste0(base_dir, "outputs/objects/Fig2_GRCCAres/GRCCA_results.Rdata")) # df_x_res
+df_x_res_genes <- df_x_res
+rm(list = "df_x_res")
+
+grcca_genes_to_include <- df_x_res_genes %>% 
+    filter(p_adj < 0.05) %>% # significantly associated with latent variable (not necessarily non-zero)
+    #filter(abs(z_score) >= 1.96) %>% # genes with significant weights (not necessarily related to SCZ signal)
+    pull(ensembl_gene_id) %>%
+    unique
+risk_genes_to_include <- unlist(benchmarking_lists) %>% unique # all psychiatric risk genes
+genes_to_include <- unique(c(grcca_genes_to_include, risk_genes_to_include))
+length(genes_to_include) # n = 5130 (FDR < 0.05 & risk genes); n = 2714 (FDR < 0.05 & |z| >= 2); n = 4491 (|z| >= 1.96)
+
+transcripts_to_include <- df_transcript_to_gene %>% 
+    filter(ensembl_gene_id %in% genes_to_include) %>% 
+    pull(ensembl_transcript_id) %>% 
+    unique()
+length(transcripts_to_include) # n = 6705 (&); n = 1876 (broad & rare SCZ risk genes); 
+# n = 12986 for GRCCA FDR < 0.05 and psychiatric risk genes
+# n = 10344 (significant GRCCA genes FDR < 0.05); 
+# n = 3610 (all psychiatric risk genes)
+# 11040 for abs(z) >= 1.96
+# 48213 for only genes included at gene-level
+
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/02.GRCCA_transcript_results.R

@@ -1,0 +1,76 @@
+
+########################################################################################
+
+# Load transcript-level GRCCA results
+
+########################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WTCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+# 08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1_sft2_minSize35_cutHeight0.988_ControlBDMDDSCZ_8MCA_ControlBDMDDSCZ_regressBrainWeight_WITHGRAY
+# 08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1_sft2_minSize35_cutHeight0.988_ControlBDMDDSCZ_8MCA_regressDim2_WITHGRAY
+
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "OVERALLsubs/"
+project_dir <- paste0(prefix, # date & preprocessing info
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", # WTCNA parameters
+                      paste0(dx_groups_to_analyze, collapse = ""), "_", # diagnostic groups included
+                      n_mca_dim, "MCA_regressDim2_", # covariates included
+                      "broadRareSCZBDMDDASD_sigGRCCAfdr0_05/" # define subset of transcripts included (based on gene-level results)
+                      #"WITHGRAY" # includes all transcripts
+)
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "transcript",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    lambda = "Nfeat",
+    VARx = VARx,
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/03.characterize_GRCCA_transcript_results.R

@@ -1,0 +1,94 @@
+
+########################################################################################
+
+# Characterize transcript-level GRCCA results with risk genes & functional enrichments
+
+########################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WTCNA_res.R"))
+
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_transcript_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+### Get ranked gene list ###
+my_grcca_ranked <- df_x_res_transcripts %>% 
+    group_by(ensembl_gene_id) %>% 
+    summarise(pearsons_r = median(pearsons_r)) %>% 
+    arrange(-pearsons_r) %>% 
+    dplyr::select(ensembl_gene_id, pearsons_r) %>% 
+    deframe
+
+
+### Cell-type GSEA ###
+df_grcca_fgsea_cell <- fgsea(pathways = cell_types, 
+                             stats = my_grcca_ranked, 
+                             eps = 0
+)  %>% 
+    as_tibble() %>% 
+    dplyr::rename("cell_type" = "pathway") %>% 
+    arrange(-abs(NES))
+df_grcca_fgsea_cell %>% 
+    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>% 
+    
+    ggplot(aes(x = NES, y = reorder(cell_type, NES))) +
+    geom_col(aes(fill = NES), color = "gray", shape = 21) +
+    geom_text(aes(label = sig, x = NES), vjust = 0.75, size = 4, check_overlap = TRUE) +
+    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3.39, 3.39)) +
+    coord_cartesian(clip = "off") +
+    labs(x = "NES", y = NULL,
+         title = "B | Cell-type enrichment") +
+    theme(legend.position = "none")
+
+
+### GO GSEA ###
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/TRY-alternative_splicing.R

@@ -1,0 +1,754 @@
+
+################################################################################
+
+# Characterize role of alternative splicing in GRCCA results
+# Distinguish results from gene- vs transcript-level
+
+################################################################################
+
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/transcripts/overall_subset/GRCCA/04.3.load_GRCCA_results.R"))
+
+# Load gene & transcript-level GRCCA results --------------------------------------------------------------------
+
+# LOAD GENE GRCCA RESULTS (X RES ONLY)
+df_x_res_genes <- read_excel(paste0(base_dir, "outputs/tables/for_manuscript/TableS3_GENES_GRCCA_res_withGray.xlsx"), sheet = 3) %>% 
+    mutate(module = as.numeric(module %>% str_remove("geneM")),
+           module = factor(paste0("geneM", module), levels = paste0("geneM", 0:max(module)))
+           )
+
+# LOAD TRANSCRIPT GRCCA RESULTS
+df_y_res_transcripts <- read_excel(paste0(base_dir, "outputs/tables/for_manuscript/TableS7_TRANSCRIPTS_GRCCA_res_withGray.xlsx"), sheet = 2)
+df_x_res_transcripts <- read_xlsx(paste0(base_dir, "outputs/tables/for_manuscript/TableS7_TRANSCRIPTS_GRCCA_res_withGray.xlsx"), sheet = 3) %>% 
+    mutate(module = as.numeric(module %>% str_remove("transcriptM")),
+           module = factor(paste0("transcriptM", module), levels = paste0("transcriptM", 0:max(module)))
+    )
+
+# TRANSCRIPT RAW COUNTS
+df_transcript_raw_counts <- read.csv(paste0(base_dir, "data/transcript_count_matrix.csv")) %>% 
+  as_tibble() %>% 
+  dplyr::rename("ensembl_transcript_id" = "X") %>% 
+  rename_all(~str_remove(.x, "_MergedBam_stringtieOutput")) %>% 
+  clean_names() %>% 
+  rename_all(~str_remove(.x, "x")) 
+
+
+# Identify rare transcripts -----------------------------------------------
+
+# DEFINE TRANSCRIPT UNIVERSE AS ALL TRANSCRIPTS IN GRCCA ANALYSIS
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/TRY-rare_vs_common.R

@@ -1,0 +1,508 @@
+
+########################################################################################
+
+# OVERALL SUBSET transcript level 
+# characterize role of rare and common transcripts in GRCCA results
+
+########################################################################################
+
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig5_GRCCAtranscriptsRes/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig5_GRCCAtranscriptsRes/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+
+## Load GRCCA data
+load(paste0(base_dir, "objects/GRCCA_transcript_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+
+# Identify rare vs common transcripts -------------------------------------
+
+## Load raw count data
+df_transcript_raw_counts <- read.csv(paste0(base_dir, "data/transcript_count_matrix.csv")) %>% 
+    as_tibble() %>% 
+    dplyr::rename("ensembl_transcript_id" = "X") %>% 
+    rename_all(~str_remove(.x, "_MergedBam_stringtieOutput")) %>% 
+    clean_names() %>% 
+    rename_all(~str_remove(.x, "x")) 
+
+## Identify rare transcripts (At least 10 counts and less than 100 counts in at least 80% of samples = rare)
+rare_transcripts <- df_transcript_raw_counts %>% 
+    dplyr::filter(ensembl_transcript_id %in% ensembl_transcript_ids) %>% 
+    mutate_if(is.numeric, ~ifelse(.x >= 10 & .x < 100, 1, 0)) %>% 
+    mutate(thresh = rowSums(dplyr::select(., -ensembl_transcript_id))/(ncol(.) - 1), .before = 2) %>% 
+    dplyr::filter(thresh >= 0.80) %>% 
+    pull(ensembl_transcript_id) 
+length(rare_transcripts) # 23211
+
+
+# Are implicated transcripts rare or common? ------------------------------
+
+
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/TRY-subset_transcripts_GRCCA.R

@@ -1,0 +1,298 @@
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig5_GRCCAtranscriptsRes/")
+tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig5_GRCCAtranscriptsRes/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WTCNA_res.R"))
+
+
+
+# load GRCCA results ------------------------------------------------------
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "OVERALLsubs/"
+project_dir <- paste0(prefix, # date & preprocessing info
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", # WTCNA parameters
+                      paste0(dx_groups_to_analyze, collapse = ""), "_", # diagnostic groups included
+                      n_mca_dim, "MCA_regressDim2_", # covariates included
+                      #"broadRareSCZ/"
+                      "broadRareSCZBDMDDASD_sigGRCCAfdr0_05z2/" # define subset of transcripts included (based on gene-level results)
+)
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "transcript",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    lambda = "Nfeat", #"GeneLevel",
+    VARx = VARx,
+    include_Cmat = TRUE,
+    rename_covariates = TRUE # FALSE if newer analysis where disorders are abbreviated, TRUE for old analysis with old labels
+)
+l_grcca_res
+
+## Assign results to tibbles (negate if SCZ is negative to make SCZ positive)
+df_results_transcripts <- l_grcca_res$model_results
+df_lvs_transcripts <- l_grcca_res$latent_variables
[…]

### 2025-01-10 scripts/06.GRCCA_transcripts/TRY-with_without_regressDim2.R

@@ -1,0 +1,166 @@
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+# Load transcript-level GRCCA results (with & without Cmat, or Dim2 regression)
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig5_GRCCAtranscriptsRes/")
+
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WTCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+# 08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1_sft2_minSize35_cutHeight0.988_8MCA_ControlBDMDDSCZ_regressDim2_WITHGray
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "OVERALLsubs/"
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+
+# with Cmat --> DOES NOT include Dim 2 drugs in Y results! (regress!)
+project_dir <- paste0(
+    prefix,
+    "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_",
+    paste0(dx_groups_to_analyze, collapse = ""), "_",
+    n_mca_dim, "MCA_regressDim2_WITHGRAY/"
+)
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+l_grcca_res_withCmat <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "transcript",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    VARx = VARx,
[…]

### 2025-01-10 scripts/07.sensitivity_RCCA/00.RCCA_results.R

@@ -1,0 +1,63 @@
+
+########################################################################################
+
+# Load gene-level RCCA results
+
+########################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+## Identify CCA directory
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in RCCA analysis results
+analysis_type <- "rcca"
+VARx <- "0.1_1"
+l_rcca_res <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "gene",
+    analysis_type = analysis_type,
+    mu = 0.1,
+    VARx = VARx,
+    include_Cmat = FALSE
+)
+
+## Assign results to tibbles; negate everything so SCZ is positive
+df_results_rcca <- l_rcca_res$model_results
+df_lvs_rcca <- l_rcca_res$latent_variables %>% mutate_at(vars(c(lvx, lvy)), ~ -.x)
+df_y_res_rcca <- l_rcca_res$y_res %>% mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x)
+df_x_res_rcca <- l_rcca_res$x_res %>% 
+    mutate_at(vars(c(weight, z_score, pearsons_r)), ~ -.x) %>% 
+    left_join(df_ensembl_to_symbol) %>% 
+    left_join(df_modules_filt %>% dplyr::select(-c(mod_set, color)), by = join_by(ensembl_gene_id)) %>% 
+    dplyr::select(ensembl_gene_id, gene_symbol, module, everything()) %>% 
+    mutate(gene_symbol = ifelse(is.na(gene_symbol), ensembl_gene_id, gene_symbol))
+
+
[…]

### 2025-01-10 scripts/07.sensitivity_RCCA/01.characterize_RCCA_res_GSEA.R

@@ -1,0 +1,64 @@
+
+########################################################################################
+
+# Assess functional, cell-type, and risk gene enrichments of RCCA results
+
+########################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+## Load RCCA data
+load(paste0(base_dir, "objects/RCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res
+
+
+### Get ranked gene list ###
+my_rcca_ranked <- df_x_res_rcca %>% 
+    arrange(-pearsons_r) %>% 
+    dplyr::select(ensembl_gene_id, pearsons_r) %>% 
+    deframe
+
+
+### Cell-type GSEA ###
+df_rcca_fgsea_cell <- fgsea(pathways = cell_types, 
+                          stats = my_rcca_ranked, 
+                          eps = 0
+)  %>% 
+    as_tibble() %>% 
+    dplyr::rename("cell_type" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### GO GSEA ###
+df_rcca_fgsea_go <- fgsea(pathways = go_pathways, 
+                        stats = my_rcca_ranked, 
+                        eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("go_id" = "pathway") %>% 
+    arrange(-abs(NES)) %>% 
+    left_join(df_go_terms) %>% # add names of GO terms
+    dplyr::select(go_id, term, ontology, everything())
+
+
+### Benchmark lists ###
+df_rcca_fgsea_benchmark <- fgsea(pathways = benchmarking_lists, 
+                               stats = my_rcca_ranked, 
+                               eps = 0
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("benchmark_list" = "pathway") %>% 
+    arrange(-abs(NES))
+
+
+### Save objects for figures ###
+save(df_rcca_fgsea_cell, df_rcca_fgsea_go, df_rcca_fgsea_benchmark,
+     file = paste0(analysis_objects_dir, "RCCA_enrichment_benchmark_res.Rdata")
+)
+

### 2025-01-10 scripts/08.figures_markdown/archive/analysis_pipeline_figure.R

@@ -1,0 +1,382 @@
+
+################################################################################
+
+# Put together figures for Fig 1 analysis pipeline
+
+################################################################################
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+## DEFINE WGCNA PARAMETERS
+soft_power <- 3
+minimum_size <- 40
+tree_cut_height <- 0.98
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+load(paste0(base_dir, "objects/hsapiens_genome_v110.RDS")) # df_hsapiens_genome
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+    filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module)
+
+
+# WGCNA heatmap & dendrogram -------------------------------------------------------------------
+
+
+# CONVERT REGRESSED VSD COUNTS TO TRANSPOSED MATRIX
+m_vsd_regress <- df_vsd_regress %>% column_to_rownames("sample")
+
+# CALCULATE CO-EXPRESSION SIMILARITY AND ADJACENCY
+set.seed(20240306)
+adjacency <- adjacency(m_vsd_regress, type = "signed hybrid", power = soft_power)
+
+# TOPOLOGICAL OVERLAP MATRIX (TOM)
+set.seed(20240306)
+TOM <- TOMsimilarity(adjacency, TOMType = "signed")
+
+# subset for plotting
+baby_TOM <- TOM[1:2000, 1:2000]
+baby_diss_TOM <- 1 - baby_TOM
+
+# HIERARCHICAL CLUSTERING ON TOM  
+baby_gene_tree <- hclust(as.dist(baby_diss_TOM))
+
+# plot the resulting clustering tree
[…]

### 2025-01-10 scripts/load_WTCNA_res.R

@@ -1,0 +1,32 @@
+
+### NOTE: This script can only be run after scripts/06.GRCCA_transcripts/00.case_control_WTCNA.R is run! ###
+
+################################################################################
+
+# Load transcript-level module assignments
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_transcripts.R"))
+
+## Define parameters for module set of interest
+soft_power <- 2
+minimum_size <- 35
+tree_cut_height <- 0.988
+
+## Load module data
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+## Filter module data for correct module set
+df_modules_filt <- df_modules %>% 
+    dplyr::filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    tidyr::unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module) %>% 
+    
+    # rename modules to specify gene-level
+    mutate(module = paste0("transcriptM", module) %>% factor(levels = paste0("transcriptM", levels(module)))
+    )
+
+## Define module colors for plot
+module_colors <- df_modules_filt %>% dplyr::select(module, color) %>% distinct %>% deframe

### 2025-01-10 scripts/load_transcripts.R

@@ -1,0 +1,53 @@
+
+### NOTE: This script can only be run after scripts/00.preprocessing/04.transcript_filtering.R is run! ###
+
+################################################################################
+
+# Load transcript-level data; map transcript IDs to genes
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/setup.R"))
+
+### SET DIRECTORY AND LOAD DATA ###
+
+prefix <- "08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1"
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress_filt
+
+
+### MAP TRANSCRIPTS TO GENES ###
+
+### Transcript ensembl ID to gene ID & symbol mappings ###
+ensembl_transcript_ids <- df_vsd_regress_filt %>% dplyr::select(-sample) %>% colnames %>% unique
+
+## Access biomaRt database to map transcript IDs to gene IDs
+mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
+biomart_res <- getBM(attributes = c("ensembl_transcript_id", 
+                                    "external_transcript_name",
+                                    "ensembl_gene_id"),
+                     filters = "ensembl_transcript_id", 
+                     values = ensembl_transcript_ids,
+                     mart = mart
+) %>% 
+    as_tibble() %>% 
+    dplyr::rename("transcript_symbol" = "external_transcript_name")
+rm(list = "mart")
+
+## Identify gene symbol for genes that were not mapped
+df_ensembl_to_symbol <- mapIds(
+    EnsDb.Hsapiens.v79, 
+    keys = biomart_res$ensembl_gene_id,
+    column = "SYMBOL", 
+    keytype = "GENEID", 
+    multiVals = "first"
+) %>% 
+    enframe(name = "ensembl_gene_id", value = "gene_symbol") %>% 
+    mutate(gene_symbol = ifelse(is.na(gene_symbol), ensembl_gene_id, gene_symbol))
+
+## Combine
+df_transcript_to_gene <- biomart_res %>% 
+    left_join(df_ensembl_to_symbol, relationship = "many-to-many", by = join_by(ensembl_gene_id)) %>% 
[…]

### 2025-01-10 scripts/setup.R

@@ -1,0 +1,207 @@
+
+#################################################################################################
+
+# Load libraries, data, and functions that generalize to both gene- and transcript-level analyses
+
+#################################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(tidymodels)
+library(tidytext)
+
+library(janitor)
+
+library(stm)
+
+library(patchwork)
+library(ggrepel)
+library(ggh4x)
+library(ggpubr)
+library(ggwordcloud)
+library(ggvenn)
+library(cowplot)
+library(pheatmap)
+library(RColorBrewer)
+library(pals)
+
+library(fgsea)
+library(biomaRt)
+library(DESeq2)
+library(WGCNA)
+library(clusterProfiler)
+library(msigdbr)
+library(org.Hs.eg.db)
+library(GO.db)
+library(EnsDb.Hsapiens.v79) # hg38
+library(ensembldb)
+
+library(rhdf5)
+
+library(readxl)
+library(writexl)
+
+filter <- dplyr::filter
+select <- dplyr::select
+count <- dplyr::count
+
+
+# Load data ----add_count_()# Load data ---------------------------------------------------------------
+
+
+### CURRENT STUDY DATA ###
+
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/drug_MCA_results.Rdata")) # df_ind_loadings, df_var_loadings
+
+
+### CELL TYPE DATA ###
+
+load(paste0(base_dir, "objects/cell_type_data.RDS")) # df_cell_type
+
+df_lake_cell_type <- df_cell_type %>% 
+    # filter for Lake et al 2018 cell types
+    dplyr::filter(str_detect(type, "lake")) %>% 
+    mutate(type = str_remove(type, "_lake")) %>% 
+    dplyr::filter(!is.na(ensembl_gene_id) & type != "synapses") %>% 
+    # format cell type names for plots
+    mutate(type = str_to_title(type)) %>%
+    mutate(type = ifelse(type == "Opc", toupper(type), type))
+
+cell_types <- df_lake_cell_type %>% 
[…]

### 2025-01-10 scripts/zarchive/TRY-univariate_case_control.R

@@ -1,0 +1,594 @@
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(tidymodels)
+library(vip)
+library(workflowsets)
+library(ggpubr)
+library(WGCNA)
+
+
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 18),
+                  axis.title = element_text(size = 15),
+                  axis.text = element_text(size = 12),
+                  strip.text = element_text(size = 15),
+                  legend.title = element_text(size = 15),
+                  legend.text = element_text(size = 12)))
+
+
+
+
+# data --------------------------------------------------------------------
+
+
+soft_power <- 3
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230228_185samples_19kgenes_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == 40 & cut_height == 0.97) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+load(paste0(base_dir, "data/covariates/185_all_covariates_clean.Rdata"))
+
+# ADD DRUG CLASSES TO COVARIATES
+drugs_of_abuse <- c("smoker", "nicotine", "alcohol", "cannabis", "cocaine", "opioids", "major_stimulants")
+mood_meds <- c("anti_anxiety", "mood_stabilizers", "anti_epileptics")
+other_psych_meds <- c("anticholinergics", "anti_depressant", "antipsychotics" )
+other_drugs <- c("anti_histamines", "non_psychiatric", "other_psychotropic_drug")
+
+df_covariates_all_clean_numeric <- df_covariates_all_clean_numeric %>% 
+  
+  mutate(
+    
[…]

### 2025-01-10 scripts/zarchive/clustering_methods.R

@@ -1,0 +1,269 @@
+
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(WGCNA)
+
+
+# data --------------------------------------------------------------------
+
+
+  prefix <- "20221209_185samples_19kgenes_vsd_qSVA17_resids"
+  
+  load(paste0("objects/", prefix, ".RDS")) # df_vsd_regress
+
+  
+# CONVERT TO MATRIX  
+  m_vsd_regress <- df_vsd_regress %>% 
+    pivot_wider(id_cols = sample,
+                names_from = ensembl_gene_id,
+                values_from = vsd) %>% 
+    as.data.frame %>% 
+    column_to_rownames("sample") 
+
+  
+
+# k-means elbow method ----------------------------------------------------
+
+# Elbow method: percentage of variance explained as a function of the number of clusters
+  
+  set.seed(123)
+  k_seq <- seq(2, 40, 1)
+  
+  df_kmeans <- map(
+    
+    .x = k_seq,
+    .f = ~ tibble(
+      
+      k = .x,
+      tot_within_ss = kmeans(m_vsd_regress, centers = .x)$tot.withinss
+      
+    )
+    
+  ) %>% bind_rows()
+  
+  p_elbow_vst <- df_kmeans %>% 
+    ggplot(aes(x = k, y = tot_within_ss)) +
+    geom_point() +
+    ggtitle("Elbow method VST counts")
+
+  # scaled data
+  scaled_data <- scale(m_vsd_regress)
+
+  df_kmeans_scaled <- map(
+    
+    .x = k_seq,
+    .f = ~ tibble(
+      
+      k = .x,
+      tot_within_ss = kmeans(scaled_data, centers = .x)$tot.withinss
+      
+    )
+    
+  ) %>% bind_rows()
+  
+  p_elbow_scaled <- df_kmeans_scaled %>% 
+    ggplot(aes(x = k, y = tot_within_ss)) +
+    geom_point() +
+    ggtitle("Elbow method scaled data")
+  
+  # TOM
+  adjacency <- adjacency(m_vsd_regress, power = 3)
+  TOM <- TOMsimilarity(adjacency)
+  diss_TOM <- 1 - TOM
+  dist <- as.dist(diss_TOM)
+  
+  library(cluster)
+  pam <- pam(diss_TOM, k = 2, diss = TRUE, variant = "faster")$silinfo$avg.width
+  
+  df_kmediods_tom <- map_dfr(
[…]

### 2025-01-10 scripts/zarchive/combine_akula_result_tables.R

@@ -1,0 +1,136 @@
+
+library(tidyverse)
+library(readxl)
+library(janitor)
+
+df_akula_res <- 
+  
+  # gene level
+  readxl::read_excel("data/Akula_etal_2021_DEGs.xlsx", sheet = 4) %>% 
+  dplyr::select(1:6) %>% 
+  row_to_names(1) %>%
+  mutate(level = "gene",
+         comparison = "schizo_ctrl",
+         .before = 1) %>% 
+  clean_names %>% 
+  dplyr::rename("id" = "ensemble_gene_id",
+                "symbol" = "hgnc_symbol") %>% 
+  
+  bind_rows(
+    
+    readxl::read_excel("data/Akula_etal_2021_DEGs.xlsx", sheet = 5) %>% 
+      dplyr::select(1:6) %>% 
+      row_to_names(1) %>%
+      mutate(level = "gene",
+             comparison = "bipolar_ctrl",
+             .before = 1) %>% 
+      clean_names %>% 
+      dplyr::rename("id" = "ensemble_gene_id",
+                    "symbol" = "hgnc_symbol")
+    
+  ) %>% 
+  
+  bind_rows(
+    
+    readxl::read_excel("data/Akula_etal_2021_DEGs.xlsx", sheet = 6) %>% 
+      dplyr::select(1:6) %>% 
+      row_to_names(1) %>%
+      mutate(level = "gene",
+             comparison = "mdd_ctrl",
+             .before = 1) %>% 
+      clean_names %>% 
+      dplyr::rename("id" = "ensemble_gene_id",
+                    "symbol" = "hgnc_symbol")
+    
+  ) %>% 
+  
+  bind_rows(
+    
+    readxl::read_excel("data/Akula_etal_2021_DEGs.xlsx", sheet = 7) %>% 
+      dplyr::select(1:6) %>% 
+      row_to_names(1) %>%
+      mutate(level = "gene",
+             comparison = "schizo_bipolar",
+             .before = 1) %>% 
+      clean_names %>% 
+      dplyr::rename("id" = "ensemble_gene_id",
+                    "symbol" = "hgnc_symbol")
+    
+  )  %>% 
+  
+  bind_rows(
+    
+    readxl::read_excel("data/Akula_etal_2021_DEGs.xlsx", sheet = 8) %>% 
+      dplyr::select(1:6) %>% 
+      row_to_names(1) %>%
+      mutate(level = "gene",
+             comparison = "schizo_mdd",
+             .before = 1) %>% 
[…]

### 2025-01-10 scripts/zarchive/compare_RCCA_GRCCA.R

@@ -1,0 +1,411 @@
+
+# libraries ---------------------------------------------------------------
+
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(RCCA)
+  library(biomaRt)
+  library(tidymodels)  
+  library(ggrepel)
+  library(raveio)
+
+
+
+
+
+# set theme for plots -----------------------------------------------------
+
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 18),
+                  axis.title = element_text(size = 15),
+                  axis.text = element_text(size = 12),
+                  strip.text = element_text(size = 15),
+                  legend.title = element_text(size = 15),
+                  legend.text = element_text(size = 12)))
+
+
+
+
+# data --------------------------------------------------------------------
+
+
+soft_power <- 3
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230228_185samples_19kgenes_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+  load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+  load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 40 & cut_height == 0.97) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module)
+
+# COVARIATES
+  load("data/covariates/185_all_covariates_clean.Rdata")
+
+# GENE SYMBOL TO ENSEMBL GENE ID MAPPING
+  load("objects/ensembl_id_to_gene_symbol.Rdata") # df_ensembl_to_symbol
+
+# define matrices ---------------------------------------------------------------
+  
+  
+  # Select covariates based on significant covariates from 4.1: 
+  # anti-anxiety, opioids, brain_weight, age_death
+  
+# DEFINE MATRICES  
[…]

### 2025-01-10 scripts/zarchive/depr.benchmarking/differential_expression.R

@@ -1,0 +1,85 @@
+
+########################################################################################
+
+# Benchmark the functional enrichment of Nirmala's DE results (genes & transcripts)
+
+########################################################################################
+
+
+### SETUP ### ------
+
+## Libraries
+library(tidyverse)
+library(cowplot)
+library(clusterProfiler)
+
+## Set project directory
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+
+## Set figures directory
+figures_dir <- paste0(base_dir, "outputs/benchmarking_res/")
+
+## Data
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+
+# genes
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+
+## Plot theme
+theme_set(
+    theme_cowplot() +
+        theme(plot.title = element_text(size = 12),
+              strip.text = element_text(size = 10, color = "black"),
+              axis.title = element_text(size = 10),
+              axis.text = element_text(size = 8),
+              legend.title = element_text(size = 8),
+              legend.text = element_text(size = 8),
+              legend.margin = margin(t = 0, r = 0, b = 0, l = 0),  # Remove legend margin
+              plot.margin = margin(t = 0, r = 0, b = 0, l = 0)
+        )
+)
+
+
+### HIGHLIGHT DEGS ### ------
+
+df_akula_res %>% 
+    filter(comparison == "schizo_ctrl" & level == "gene") %>% 
+    
+    filter(padj < 0.05)
+
+### GO FUNCTIONAL ENRICHMENT OF GENES ### ------
+
+## Set gene universe using our data
+gene_universe <- df_vsd_regress %>% 
+    dplyr::select(-sample) %>% 
+    colnames
+
+## Run functional enrichment
+deg_go <- enrichGO(gene = df_akula_res %>% 
+             filter(comparison == "schizo_ctrl" & level == "gene" & pvalue < 0.05) %>% 
+             pull(id),
+         OrgDb = "org.Hs.eg.db",
+         universe = gene_universe,
+         keyType = "ENSEMBL",
+         ont = "ALL",
+         pAdjustMethod = "BH",
+         pvalueCutoff = 1,
[…]

### 2025-01-10 scripts/zarchive/depr.benchmarking/gene_co_expression.R

@@ -1,0 +1,227 @@
+
+########################################################################################
+
+# Benchmark WGCNA and WTCNA results against the literature
+
+########################################################################################
+
+# setup -------------------------------------------------------------------
+
+## LIBRARIES
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(stm)
+library(ggwordcloud)
+library(ggrepel)
+library(readxl)
+library(writexl)
+library(clusterProfiler)
+
+## SET DIRECTORIES
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+## IDENTIFY MODULE SET OF INTEREST
+soft_power <- 3
+minimum_size <- 40
+tree_cut_height <- 0.98
+
+## LOAD DATA OBJECTS 
+load(paste0(base_dir, "objects/hsapiens_genome_v110.RDS")) # df_hsapiens_genome
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+## FILTER FOR MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+    filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module) %>% 
+    
+    # rename modules to specify gene-level
+    mutate(module = paste0("geneM", module) %>% factor(levels = paste0("geneM", levels(module)))
+    )
+
+## ENSEMBL ID TO GENE SYMBOL MAPPINGS
+df_ensembl_to_symbol <- df_hsapiens_genome %>% 
+    dplyr::select(gene_id, gene_name) %>% 
+    distinct() %>% 
+    dplyr::rename("gene_symbol" = "gene_name", "ensembl_gene_id" = "gene_id")
+
+## PLOT THEME
+theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 12),
+                    axis.title = element_text(size = 12),
+                    axis.text = element_text(size = 10),
[…]

### 2025-01-10 scripts/zarchive/depr.genes/02.DGE/A.DGE_analysis.R

@@ -1,0 +1,254 @@
+
+###################################################################
+
+# Updated DGE analyis including all covariates used in WGCNA/GRCCA
+
+###################################################################
+
+## Set directories & load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/DGE_res/")
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/setup.R"))
+
+
+# Convert count data to matrix for DESeq2 --------------------------------------
+
+m_vsd_regress <- df_vsd_regress %>% 
+    dplyr::filter(str_detect(sample, "schizo|control")) %>% # subset for only schizophrenia-control comparisons
+    column_to_rownames("sample") %>% 
+    t() + 10 # add pseudo-count of 10 so there are no negative values in the matrix
+m_vsd_regress <- round(m_vsd_regress*10^5) # multiply by 10000 and round to convert to integers but maintain precision
+
+
+# Generate covariates matrix ----------------------------------------------
+
+## Load drug MCA results
+load(paste0(base_dir, "objects/drug_MCA_results.Rdata")) # df_ind_loadings, df_var_loadings
+
+## Combine drug MCs with known covariates
+df_covariates_mca <- df_covariates %>% 
+    left_join(
+        df_ind_loadings %>% 
+            dplyr::rename_at(vars(contains("dim")), ~ str_replace(.x, "dim", "MC")),
+        by = join_by(sample)
+    )
+
+## Convert to matrix
+n_mcs <- 8
+paste0("MC", 1:n_mcs)
+m_covariates <- df_covariates_mca %>% 
+    dplyr::filter(dx %in% c("SCZ", "Control")) %>% # SCZ vs controls only
+    mutate(dx = factor(dx, levels = c("Control", "SCZ"))) %>% 
+    dplyr::select(sample, dx, all_of(paste0("MC", 1:n_mcs))) %>% 
+    column_to_rownames("sample")
+
+
+# Run DESeq ---------------------------------------------------------------
+
+## Create DESeq2 object
+dds <- DESeqDataSetFromMatrix(countData = m_vsd_regress, 
+                              colData = m_covariates, 
+                              design = ~ dx + 
[…]

### 2025-01-10 scripts/zarchive/depr.genes/02.DGE/B.benchmark_DEGs.R

@@ -1,0 +1,226 @@
+
+###################################################################
+
+# GENE LEVEL: Describe Akula et al 2021 functional enrichment and #
+# overlap with published SCZ gene lists.                          #
+
+###################################################################
+
+## Set directories & load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/DGE_res/")
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/setup.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+## Current analysis DEG results
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+## GWAS DATA
+df_scz_gwas <- readxl::read_excel(paste0(base_dir, "data/trubetskoy_2023_scz_gwas.xlsx"),
+                                  sheet = 3
+)
+
+# sheet 2 = 95% credible set
+# sheet 3 = 95% credible set k <= 3.5
+# sheet 4 = 95% credible set <= 5 SNPs
+# sheet 5 = 95% credible set <= 5 PIP
+
+## Rare variants
+df_exome_urvs <- readxl::read_excel(paste0(base_dir, "data/singh_2022_exome_URVs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+
+# Identify SCZ-control DEGs -------------------------------
+
+## DEGs
+my_degs <- df_de_res %>% 
+    dplyr::filter(pvalue < 0.05) %>% 
+    pull(ensembl_gene_id)
+
+### DEGs (ranked) ###
+my_degs_ranked <- df_de_res %>% 
+    arrange(-log2fold_change) %>% 
+    dplyr::select(ensembl_gene_id, log2fold_change) %>% 
+    deframe
+
+
+# Consensus DEGs -----------------------------------------------------------
+
+### HYPERGEOMETRIC (overall enrichment) ###
+
+# Pull consensus DEG symbols
+consensus_degs <- df_consensus_degs %>% pull(ensembl_gene_id)
+length(consensus_degs) # n = 161
+consensus_degs %in% ensembl_ids %>% sum # n = 154 in our universe
+
[…]

### 2025-01-10 scripts/zarchive/depr.genes/03.WGCNA/A.case_control_WGCNA.R

@@ -1,0 +1,221 @@
+
+########################################################################################
+
+# Select soft-thresholding power, calculate adjacency & TOM, cluster, and assign modules
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(WGCNA)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(paletteer)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 12),
+                  axis.title = element_text(size = 12),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 10),
+                  legend.text = element_text(size = 10)))
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+# REGRESSED GENE DATA
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+
+# COVARIATES
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+
+
+# select soft thresholding power ------------------------------------------
+
+  
+# CONVERT REGRESSED VSD COUNTS TO TRANSPOSED MATRIX
+  m_vsd_regress <- df_vsd_regress %>% column_to_rownames("sample")
+  
+# CHOOSE A SET OF SOFT-THRESHOLDING POWERS TO TEST
+  powers <- 1:16
+
+# CALL THE NETWORK TOPOLOGY ANALYSIS FUNCTION
+  sft <- pickSoftThreshold(m_vsd_regress, powerVector = powers, verbose = 5)
+  
+  # convert results to tibble
+  df_sft <- sft$fitIndices %>% as_tibble %>% clean_names
+  
+  # save
+  save(df_sft, file = paste0(base_dir, "objects/", prefix, "_SFT.RDS"))
+
+# PLOT RESULTS   
+
+  #load(paste0(base_dir, "objects/", prefix, "_SFT.RDS"))
[…]

### 2025-01-10 scripts/zarchive/depr.genes/03.WGCNA/B.characterize_module_enrichment.R

@@ -1,0 +1,471 @@
+
+
+########################################################################################
+
+# GENE LEVEL: run GO and cell-type enrichment
+
+########################################################################################
+
+
+# setup -------------------------------------------------------------------
+
+
+## LIBRARIES
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(stm)
+library(ggwordcloud)
+library(ggrepel)
+library(readxl)
+library(writexl)
+library(clusterProfiler)
+
+## SET DIRECTORIES
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+## IDENTIFY MODULE SET OF INTEREST
+soft_power <- 3
+minimum_size <- 40
+tree_cut_height <- 0.98
+
+## LOAD DATA OBJECTS 
+load(paste0(base_dir, "objects/hsapiens_genome_v110.RDS")) # df_hsapiens_genome
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+## FILTER FOR MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+      filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+      unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+      arrange(mod_set, module) %>% 
+      
+      # rename modules to specify gene-level
+      mutate(module = paste0("geneM", module) %>% factor(levels = paste0("geneM", levels(module)))
+      )
+  
+ ## ENSEMBL ID TO GENE SYMBOL MAPPINGS
+  df_ensembl_to_symbol <- df_hsapiens_genome %>% 
+      dplyr::select(gene_id, gene_name) %>% 
+      distinct() %>% 
+      dplyr::rename("gene_symbol" = "gene_name", "ensembl_gene_id" = "gene_id")
+  
+  
+## PLOT THEME
+  theme_set(theme_bw() +
+                theme(plot.title = element_text(size = 12),
[…]

### 2025-01-10 scripts/zarchive/depr.genes/03.WGCNA/C2.gene_module_correlations.R

@@ -1,0 +1,73 @@
+
+########################################################################
+
+# Correlate gene expression vectors with module eigengene to calculate
+# gene association with each module
+
+########################################################################
+
+### SETUP ###
+
+## Set directories
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/WGCNA_res/")
+
+## Load data and functions
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/load_WGCNA_res.R"))
+
+# module eigengene analysis results
+load(paste0(base_dir, "objects/kME_analysis_res.RDS"))
+sig_modules <- df_lm_dx %>% 
+    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
+    pull(module)
+
+gene_universe <- df_vsd_regress %>% dplyr::select(-sample) %>% colnames
+modules <- df_kme %>% pull(module) %>% unique
+
+### RUN LOOP ###
+
+df_gene_module_cor <- tibble()
+doParallel::registerDoParallel()
+for (m in 1:length(modules)) {
+    
+    current_module <- modules[m]
+    print(paste0("Calculating gene-module correlations for ", current_module))
+    
+    for (g in 1:length(gene_universe)) {
+        
+        current_gene <- gene_universe[g]
+        
+        if (g %% 100 == 0) {print(paste0("gene ", g, " of ", length(gene_universe)))}
+        
+        df_kme_tmp <- df_kme %>% 
+            ungroup %>% 
+            filter(module == current_module) %>% 
+            dplyr::select(sample, kme)
+        
+        df_gene_tmp <- df_vsd_regress %>% 
+            dplyr::select(sample, all_of(current_gene)) %>% 
+            dplyr::rename_at(2, ~ "gene")
+        
+        df_kme_gene_tmp <- df_kme_tmp %>% 
+            left_join(df_gene_tmp, by = join_by(sample))
+        
+        cor_tmp <- cor.test(df_kme_gene_tmp$kme, df_kme_gene_tmp$gene)$estimate
+        p_tmp <- cor.test(df_kme_gene_tmp$kme, df_kme_gene_tmp$gene)$p.value
+        
+        df_tmp <- tibble(
+            module = current_module,
+            ensembl_gene_id = current_gene,
+            pearsons_r = cor_tmp,
+            p_value = p_tmp
[…]

### 2025-01-10 scripts/zarchive/depr.genes/03.WGCNA/D.benchmark_WGCNA_res.R

@@ -1,0 +1,243 @@
+
+########################################################################
+
+# GENE LEVEL: Describe functional enrichment of modules implicated
+# by univariate analysis and how they overlap with published DEG results
+
+########################################################################
+
+
+# Setup -------------------------------------------------------------------
+
+## Set directories
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/WGCNA_res/")
+
+## Load data and functions
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/load_WGCNA_res.R"))
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/module_hypergeometric_overlap.R")) # f_module_hypergeometric
+
+## Link dx to group number (in the numeric covariates tibble)
+group <- c(0, 1, 2, 3)
+names(group) <- c("BD", "Control", "MDD", "SCZ")
+
+
+# Load data ---------------------------------------------------------------
+
+### Study data ###
+
+## module eigengene analysis results
+load(paste0(base_dir, "objects/kME_analysis_res.RDS")) # df_lm_dx
+sig_modules <- df_lm_dx %>% 
+    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
+    pull(module)
+
+## current analysis DEG results
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
+
+### Benchmarking ###
+
+## Akula DEGs
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+## GWAS DATA
+df_scz_gwas <- readxl::read_excel(paste0(base_dir, "data/trubetskoy_2023_scz_gwas.xlsx"),
+                                  sheet = 3
+)
+
+# sheet 2 = 95% credible set
+# sheet 3 = 95% credible set k <= 3.5
+# sheet 4 = 95% credible set <= 5 SNPs
+# sheet 5 = 95% credible set <= 5 PIP
+
+## Rare variants
+df_exome_urvs <- readxl::read_excel(paste0(base_dir, "data/singh_2022_exome_URVs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+
+
[…]

### 2025-01-10 scripts/zarchive/depr.genes/03.WGCNA/DEPR.module_eigengene_analysis.R

@@ -1,0 +1,547 @@
+
+########################################################################################
+
+# GENE LEVEL: Associate module PCs with covariates 
+
+########################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/03.WGCNA/load_WGCNA_res.R"))
+
+
+# setup ---------------------------------------------------------------
+
+## LIBRARIES
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(tidymodels)
+  library(vip)
+  library(workflowsets)
+  library(ggpubr)
+  library(WGCNA)
+
+
+## PLOT THEME
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 12),
+                  axis.title = element_text(size = 12),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 10),
+                  legend.text = element_text(size = 10)))
+
+# dx colors
+dx_colors <- c("#0072B2", "#E69F00", "#009E73", "#9966FF")
+names(dx_colors) <- c("Control", "BD", "MDD", "SCZ")
+
+## SET DIRECTORIES
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+## IDENTIFY MODULE SET OF INTEREST
+soft_power <- 3
+minimum_size <- 40
+tree_cut_height <- 0.98
+
+# LOAD DATA OBJECTS 
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+    filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module) %>% 
+    
+    # rename modules to specify gene-level
[…]

### 2025-01-10 scripts/zarchive/depr.genes/04.RCCA/A.load_RCCA_results.R

@@ -1,0 +1,184 @@
+
+########################################################################################
+
+# Load gene-level RCCA results to analyze
+
+########################################################################################
+
+
+# Setup -------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/load_WGCNA_res.R"))
+
+
+# Set paths & data split to load results -----------------------------------------------------
+
+
+## Set paths to read in RCCA res (generated in cluster)
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/") # _WITHGRAY, _withGray
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+analysis_dir <- "rcca_permutation_VARx0.1_1_L2xNfeat_withCmat_allEffects"
+model_path <- paste0(cca_dir, "framework/", analysis_dir, "/res/level1/model_1.mat")
+boot_path <- paste0(cca_dir, "framework/", analysis_dir, "/boot/level1/allboot_1.mat")
+
+## Read results table output
+df_results <- read_table(paste0(cca_dir, "framework/", analysis_dir, "/res/level1/results_table.txt"))
+
+## Select the split (defined by variance explained) with the lowest p-value &
+## highest X-Y latent variable correlation
+best_split <- df_results %>% arrange(pval, -correl) %>% pull(set) %>% .[[1]]
+
+
+# Load X.mat & Y.mat ------------------------------------------------------
+
+## Y.mat
+Y.mat <- read.table(paste0(cca_dir, "data/Y.txt"))
+df_y <- Y.mat %>% 
+    rownames_to_column("sample") %>% 
+    as_tibble()
+covariates <- df_y %>% dplyr::select(-sample) %>% colnames
+
+# rename covariates for plotting
+covariates_renamed <- case_when(
+    covariates == "control" ~ "Control",
+    covariates == "bipolar" ~ "BD",
[…]

### 2025-01-10 scripts/zarchive/depr.genes/04.RCCA/C.benchmark_RCCA_genes.R

@@ -1,0 +1,310 @@
+
+################################################################################
+
+# GENE-LEVEL: Validate RCCA results using external data
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/04.RCCA/A.load_RCCA_results.R"))
+figures_dir <- paste0(base_dir, "outputs/RCCA_res/")
+
+## Source plot data
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/plot_functions.R"))
+
+
+# Identify RCCA significant genes -------------------------------
+
+## Significant genes
+sig_rcca_genes <- df_x_res %>% 
+    dplyr::filter(abs(z_score) > 2 & p_adj < 0.05) %>% 
+    pull(ensembl_gene_id)
+
+## Ranked gene vector
+ranked_rcca_genes <- df_x_res %>% 
+    arrange(-pearsons_r) %>% 
+    dplyr::select(ensembl_gene_id, pearsons_r) %>% 
+    deframe
+
+
+# Load benchmarking data ---------------------------------------------------------------
+
+## Current analysis DEGs
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
+
+## Akula DEGs
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+## GWAS DATA
+df_scz_gwas <- readxl::read_excel(paste0(base_dir, "data/trubetskoy_2023_scz_gwas.xlsx"),
+                                  sheet = 3
+)
+
+# sheet 2 = 95% credible set
+# sheet 3 = 95% credible set k <= 3.5
+# sheet 4 = 95% credible set <= 5 SNPs
+# sheet 5 = 95% credible set <= 5 PIP
+
+## Rare variants
+df_exome_urvs <- readxl::read_excel(paste0(base_dir, "data/singh_2022_exome_URVs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+
+# Current analysis DEGs ---------------------------------------------------
+
+### HYPERGEOMETRIC (overall enrichment) ###
+
+# Pull significant DEGs
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/A.define_input_matrices.R

@@ -1,0 +1,104 @@
+
+########################################################################################
+
+# Define X.mat, Y.mat, and C.mat to include in GRCCA analysis
+
+########################################################################################
+
+### X.MAT
+
+# define gene order to align modules & genes
+df_order <- df_modules_filt %>% 
+    dplyr::select(ensembl_gene_id, module) %>% 
+    distinct() %>% 
+    arrange(module, ensembl_gene_id) %>% 
+    
+    # remove gray module!
+    filter(module != 0)
+
+# X grouping vector    
+x_group <- df_order %>% 
+    mutate(module = as.numeric(as.character(module))) %>% 
+    pull(module)
+
+# X 
+X.mat <- df_vsd_regress %>% 
+    dplyr::select(sample, all_of(df_order$ensembl_gene_id)) %>% 
+    column_to_rownames("sample")
+sample_order <- rownames(X.mat)
+
+dim(X.mat) # n = 18677 (_WITHGRAY); n = 12159 (_withGray)
+
+
+### Y.MAT AND C.MAT
+
+# define number of MCA dimensions to include
+n_mca_dim <- 8
+
+# Define variables to include in Y and C matrices
+YC.mat <- df_covariates_numeric %>% 
+    
+    # drugs to keep
+    # dplyr::select(sample, age_death, brain_weight, opioids,
+    #               sedative_hypnotic_anxiolitics, antidepressants, antipsychotics, mood_stabilizers, benzos) %>%
+    # mutate_if(is.numeric, ~ifelse(is.na(.x), 0, .x)) %>%
+    
+    # Add MCA loadings
+    left_join(df_ind_loadings, by = join_by(sample)) %>%
+    dplyr::select(sample, brain_weight, all_of(paste0("dim", seq(1, n_mca_dim)))) %>%
+    
+    # create dx columns
+    mutate(control = ifelse(grepl("control", sample), 1, 0),
+           bipolar = ifelse(grepl("bipolar", sample), 1, 0),
+           mdd = ifelse(grepl("mdd", sample), 1, 0),
+           schizo = ifelse(grepl("schizo", sample), 1, 0)
+    ) %>% 
+    dplyr::select(sample, bipolar, mdd, schizo, everything(), -control) %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/D.characterize_GRCCA_results.R

@@ -1,0 +1,422 @@
+
+################################################################################
+
+# GENE-LEVEL: Characterize RCCA results (gene weights & enrichments)
+
+################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/05.GRCCA/C.load_GRCCA_results.R"))
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/module_hypergeometric_overlap.R")) # f_module_hypergeometric
+figures_dir <- paste0(base_dir, "outputs/GRCCA_res/")
+tables_dir <- paste0(base_dir, "outputs/tables/02Aug_update/")
+
+
+# X-Y LV correlation ------------------------------------------------------
+
+cor.test(df_lvs$lvx, df_lvs$lvy, method = "spearman")
+
+# plot correlation
+df_lvs %>% 
+    ggplot(aes(x = lvx, y = lvy)) +
+    geom_point(aes(color = dx), alpha = 0.7, size = 3) +
+    geom_smooth(method = lm, se = FALSE, color = "black") +
+    stat_cor() +
+    scale_color_manual(values = dx_colors) +
+    guides(color = guide_legend(title = "diagnosis")) +
+    labs(x = "X latent variable (x matrix • x weights)",
+         y = "Y latent variable (y matrix • y weights)",
+         title = "Gene x-y latent variable correlation") +
+    theme(legend.position = c(0.8, 0.15))
+
+# plot X LV
+df_lvs %>% 
+    ggplot(aes(x = lvx)) +
+    geom_density(aes(fill = dx), alpha = 0.7) +
+    scale_fill_manual(values = dx_colors)
+
+
+
+# Plot Y weights ---------------------------------------------------------------
+
+df_y_res %>% 
+    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>% 
+    
+    ggplot(aes(x = reorder(covariate, weight), y = weight)) +
+    geom_col(aes(fill = covariate, alpha = abs(weight)), color = "black") +
+    geom_errorbar(aes(ymin = weight - sd, ymax = weight + sd), width = 0.2) +
+    geom_text(aes(label = significant, y = 0), size = 6, vjust = 1.5) +
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/E.benchmark_GRCCA_genes.R

@@ -1,0 +1,323 @@
+
+################################################################################
+
+# GENE-LEVEL: Validate GRCCA results using external data
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/05.GRCCA/C.load_GRCCA_results.R"))
+figures_dir <- paste0(base_dir, "outputs/GRCCA_res/")
+
+## Source plot data
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/plot_functions.R"))
+
+
+# Identify GRCCA significant genes -------------------------------
+
+## Significant genes
+sig_grcca_genes <- df_x_res %>% 
+    dplyr::filter(abs(z_score) > 2 & p_adj < 0.05) %>% 
+    pull(ensembl_gene_id)
+
+## Ranked gene vector
+ranked_grcca_genes <- df_x_res %>% 
+    arrange(-pearsons_r) %>% 
+    dplyr::select(ensembl_gene_id, pearsons_r) %>% 
+    deframe
+
+
+# Load benchmarking data ---------------------------------------------------------------
+
+## Current analysis DEGs
+load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
+
+## Akula DEGs
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+## GWAS DATA
+df_scz_gwas <- readxl::read_excel(paste0(base_dir, "data/trubetskoy_2023_scz_gwas.xlsx"),
+                                  sheet = 3
+)
+
+# sheet 2 = 95% credible set
+# sheet 3 = 95% credible set k <= 3.5
+# sheet 4 = 95% credible set <= 5 SNPs
+# sheet 5 = 95% credible set <= 5 PIP
+
+## Rare variants
+df_exome_urvs <- readxl::read_excel(paste0(base_dir, "data/singh_2022_exome_URVs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+
+# Current analysis DEGs ---------------------------------------------------
+
+### HYPERGEOMETRIC (overall enrichment) ###
+
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/F.compare_GRCCA_RCCA.R

@@ -1,0 +1,334 @@
+
+################################################################################
+
+# GENE-LEVEL: Compare RCCA vs GRCCA results in the context of the SCZ literature
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/load_WGCNA_res.R"))
+figures_dir <- paste0(base_dir, "outputs/GRCCA_res/")
+tables_dir <- paste0(base_dir, "outputs/tables/02Aug_update/")
+
+
+# Load RCCA & GRCCA data ---------------------------------------------------------
+
+df_rcca_x_res <- read_xlsx(paste0(tables_dir, "gene_RCCA_results.xlsx"), sheet = 3)
+df_grcca_x_res <- read_xlsx(paste0(tables_dir, "gene_GRCCA_results.xlsx"), sheet = 3)
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_ensembl_to_symbol)
+
+
+
+# Combine RCCA and GRCCA z-score and structure correlations ---------------
+
+df_rcca_grcca_res <- df_grcca_x_res %>% 
+    dplyr::select(ensembl_gene_id, gene_symbol, z_score, pearsons_r, p_adj) %>% 
+    dplyr::rename_at(3:5, ~ paste0("grcca_", .)) %>% 
+    left_join(
+        df_rcca_x_res %>% 
+            dplyr::select(ensembl_gene_id, gene_symbol, z_score, pearsons_r, p_adj) %>% 
+            dplyr::rename_at(3:5, ~ paste0("rcca_", .)),
+        by = join_by(ensembl_gene_id, gene_symbol)
+        
+    )
+
+
+# Plot correlation between gene weight vectors ----------------------------
+
+## Correlate z-scores
+df_rcca_grcca_res %>% 
+    ggplot(aes(x = rcca_z_score, y = grcca_z_score)) +
+    geom_point(aes(color = rcca_z_score), size = 0.75) +
+    geom_abline(lty = 2, color = "black") +
+    geom_smooth(method = "lm", color = "black", se = FALSE) +
+    stat_regline_equation(label.y.npc = "top", size = 3) +
+    stat_cor(label.y.npc = "top", size = 3, vjust = 3) +
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/compare_across_mu.R

@@ -1,0 +1,98 @@
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+source(paste0(base_dir, "scripts/full_analysis_scripts/genes/05.GRCCA/f_load_CCA_res.R"))
+
+# Load data ---------------------------------------------------------------
+
+## Identify CCA directory
+n_mca_dim <- 8
+type <- "GENES/"
+project_dir <- paste0(prefix,
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", 
+                      n_mca_dim, "MCA_regressBrainWeight_WITHGRAY/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Identify mu values
+cca_dir_files <- list.files(paste0(cca_dir, "framework"))
+VARx <- "0.1_1"
+analysis_string <- paste0("grcca_permutation_VARx", VARx)
+cca_dir_files_grcca_VARx <- cca_dir_files[str_detect(cca_dir_files, analysis_string)]
+mu_values <- str_remove(cca_dir_files_grcca_VARx, paste0(analysis_string, "_L2xNfeat_groupL2x")) %>%
+    str_remove("_noCmat_allEffects")
+
+## Load across mu values
+l_grcca_res <- map(
+    .x = mu_values,
+    .f = ~ f_load_cca_res(
+        cca_directory = cca_dir,
+        analysis_type = "grcca",
+        mu = .x,
+        VARx = VARx,
+        include_Cmat = FALSE
+    )
+    
+)
+names(l_grcca_res) <- mu_values
+
+
+
+# Compare XY correlations and variance explained --------------------------
+
+df_model_results <- map(l_grcca_res, 
+    ~ .x$model_results %>% 
+        mutate(mu = .x$mu, .before = 1)
+) %>% 
+    bind_rows()
+
+## Plot XY correlation p-values across varX & mu values
+df_model_results %>% 
+    ggplot(aes(x = varx %>% as.factor(), y = -log10(pval), color = mu)) +
+    geom_point() +
+    geom_line(aes(group = mu)) +
+    geom_hline(aes(yintercept = -log10(0.05)), lty = 2)
+
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/explain_GRCCA_hyperparameters.R

@@ -1,0 +1,78 @@
+
+########################################################################################
+
+# Illustrate the influence of mu and lambda on regulation of gene and module weights
+
+########################################################################################
+
+
+df_x_hyperparam %>% 
+    filter(str_detect(framework, "lambda0")) %>% 
+    group_by(ensembl_gene_id) %>% 
+    mutate(mean_weight = mean(weight)) %>% 
+    
+    ggplot(aes(x = mean_weight)) +
+    geom_histogram()
+
+# MU
+tibble(
+    
+    x = 1:1000,
+    group = c(rep("a", 375), 
+              rep("b", 250), 
+              rep("c", 375)),
+    
+    y1 = c(rnorm(375, mean = 5, sd = 2), 
+           rnorm(250, mean = 2, sd = 2), 
+           rnorm(375, mean = -4, sd = 2)),
+    
+    y2 = c(rnorm(375, mean = 3, sd = 1.5), 
+           rnorm(250, mean = 0, sd = 1.5), 
+           rnorm(375, mean = -2, sd = 1.5)),
+    
+) %>% 
+    pivot_longer(starts_with("y"), names_to = "facet", values_to = "y") %>% 
+    
+    ggplot(aes(x = x, y = y)) +
+    geom_col(aes(alpha = y, color = group)) +
+    scale_color_manual(values = c("#ffda80ff", "#80ffd7ff", "#fa86f8f5")) +
+    
+    facet_wrap(vars(facet)) +
+    labs(x = "", y = "weights") +
+    #ggtitle("shrinkage regularization") +
+    theme(strip.text = element_blank(),
+          axis.text = element_blank(),
+          legend.position = "none")
+
+# LAMBDA
+tibble(
+    
+    x = 1:1000,
+    group = c(rep("a", 375), 
+              rep("b", 250), 
+              rep("c", 375)),
+    
+    y1 = c(rnorm(375, mean = 5, sd = 2), 
+           rnorm(250, mean = 2, sd = 2), 
+           rnorm(375, mean = -4, sd = 2)),
+    
+    y2 = c(rnorm(375, mean = 5, sd = 0.5), 
+           rnorm(250, mean = 2, sd = 0.5), 
+           rnorm(375, mean = -4, sd = 0.5)),
+    
+) %>% 
+    pivot_longer(starts_with("y"), names_to = "facet", values_to = "y") %>% 
+    
+    ggplot(aes(x = x, y = y)) +
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/f_load_CCA_res.R

@@ -1,0 +1,194 @@
+
+########################################################################################
+
+## Function that loads results from the RCCA toolkit and formats result into a table ##
+
+## INPUTS
+## - cca_directory = directory that contains CCA data & framework folders
+## - analysis_type = CCA algorithm used, either rcca or grcca
+## - mu = mu (x group regularization parameter) value used in analysis (numeric or character ok)
+## - VARx = was variance explained in the X matrix optimized ("0.1_1") or set at 0.99?
+## - include_Cmat = logical indicating whether Cmat was included in the analysis (were confounds regressed?)
+
+########################################################################################
+
+f_load_cca_res <- function(cca_directory = cca_dir,
+                           analysis_type = c("rcca", "grcca"),
+                           mu,
+                           VARx = c("0.1_1", "0.99"),
+                           include_Cmat = c(TRUE, FALSE)
+                           ) {
+    
+### Set paths & data split to load results ###
+    
+    print("Setting paths")
+    
+    ## Identify correct framework analysis directory
+    if (include_Cmat) {cmat <- "withCmat"} else {cmat <- "noCmat"}
+    analysis_dir <- paste0(analysis_type, "_permutation_VARx", VARx, "_L2xNfeat_groupL2x", as.character(mu), "_", cmat, "_allEffects")
+    
+    ## Combine paths to point to results (and check to make sure directory exists)
+    results_dir <- paste0(cca_dir, "framework/", analysis_dir)
+    if (!dir.exists(results_dir)) {stop("Analysis directory does not exist\n", results_dir)}
+    
+    ## Identify final model path (weights)
+    model_path <- paste0(results_dir, "/res/level1/model_1.mat")
+    
+    # Identify bath to bootstrapped weights (to calculate sd)
+    boot_path <- paste0(results_dir, "/boot/level1/allboot_1.mat")
+    
+    ## Load results table, include X-Y LV correlation across X variance explain values
+    df_results <- read_table(paste0(results_dir, "/res/level1/results_table.txt"))
+    
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/setup.R

@@ -1,0 +1,72 @@
+
+########################################################################################
+
+# Load data and module assignments to run GRCCA & analyze outputs
+
+########################################################################################
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(biomaRt)
+library(tidymodels)  
+library(ggrepel)
+library(ggh4x)
+library(ggpubr)
+library(clusterProfiler)
+library(msigdbr)
+library(org.Hs.eg.db)
+library(readxl)
+library(writexl)
+library(cowplot)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_cowplot() +
+              theme(plot.title = element_text(size = 11),
+                    axis.title = element_text(size = 10),
+                    axis.text = element_text(size = 10),
+                    strip.text = element_text(size = 10),
+                    legend.title = element_text(size = 8),
+                    legend.text = element_text(size = 8)
+              )
+)
+
+# DX COLORS
+dx_colors <- c("#0072B2", "#E69F00", "#009E73", "#9966FF")
+names(dx_colors) <- c("Control", "BD", "MDD", "SCZ")
+
+# ONTOLOGY COLORS
+ontology_colors <- c("BP" = "#F8766D", "CC" = "#00BA38", "MF" = "#619CFF")
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+soft_power <- 3
+minimum_size <- 40
+tree_cut_height <- 0.98
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+    filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    tidyr::unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.genes/05.GRCCA/uhoh.R

@@ -1,0 +1,27 @@
+
+
+df_rcca_x_res <- read_xlsx(paste0(tables_dir, "gene_RCCA_results.xlsx"), sheet = 3)
+df_grcca_x_res <- read_xlsx(paste0(tables_dir, "gene_GRCCA_results.xlsx"), sheet = 3)
+
+df_rcca_grcca_res <- df_grcca_x_res %>% 
+    dplyr::select(ensembl_gene_id, gene_symbol, z_score, pearsons_r, p_adj) %>% 
+    dplyr::rename_at(3:5, ~ paste0("grcca_", .)) %>% 
+    left_join(
+        df_rcca_x_res %>% 
+            dplyr::select(ensembl_gene_id, gene_symbol, z_score, pearsons_r, p_adj) %>% 
+            dplyr::rename_at(3:5, ~ paste0("rcca_", .)),
+        by = join_by(ensembl_gene_id, gene_symbol)
+        
+    )
+df_rcca_grcca_res %>% 
+    ggplot(aes(x = rcca_pearsons_r, y = grcca_pearsons_r)) +
+    geom_point(aes(color = rcca_pearsons_r), size = 0.75) +
+    geom_abline(lty = 2, color = "black") +
+    geom_smooth(method = "lm", color = "black", se = FALSE) +
+    stat_regline_equation(label.y.npc = "top", size = 3) +
+    stat_cor(label.y.npc = "top", size = 3, vjust = 3) +
+    scale_color_gradientn(colors = brewer.rdbu(100) %>% rev, guide = "none") +
+    labs(x = "RCCA structure correlation", y = "GRCCA structure correlation",
+         title = "Relationship between RCCA and GRCCA \ngene weight structure correlations")
+
+

### 2025-01-10 scripts/zarchive/depr.genes/setup.R

@@ -1,0 +1,76 @@
+
+################################################################################
+
+# Load libraries, data, and functions for gene-level analyses
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_GENES_qSVAgeSexRaceGC"
+
+### LIBRARIES ###
+
+library(tidyverse)
+library(tidymodels)
+library(tidytext)
+
+library(janitor)
+
+library(patchwork)
+library(ggrepel)
+library(ggh4x)
+library(ggpubr)
+library(cowplot)
+library(RColorBrewer)
+library(pals)
+
+library(biomaRt)
+library(clusterProfiler)
+library(msigdbr)
+library(org.Hs.eg.db)
+library(EnsDb.Hsapiens.v79) # hg38
+library(ensembldb)
+library(DESeq2)
+
+library(rhdf5)
+
+library(readxl)
+library(writexl)
+
+### DATA ###
+
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+load(paste0(base_dir, "objects/drug_MCA_results.Rdata")) # df_ind_loadings, df_var_loadings
+
+### Ensembl ID to gene symbol mappings ###
+ensembl_ids <- df_vsd_regress %>% dplyr::select(-sample) %>% colnames
+df_ensembl_to_symbol <- mapIds(
+    EnsDb.Hsapiens.v79, 
+    keys = ensembl_ids,
+    column = "SYMBOL", 
+    keytype = "GENEID", 
+    multiVals = "first"
+) %>% 
+    enframe(name = "ensembl_gene_id", value = "gene_symbol") %>% 
+    mutate(gene_symbol = ifelse(is.na(gene_symbol), ensembl_gene_id, gene_symbol))
+
+### SOURCE FUNCTIONS ###
+
+## for plots
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/plot_functions.R"))
+
+## hypergeometric
+source(paste0(base_dir, "scripts/full_analysis_scripts/functions/hypergeometric_function.R"))
+
+### GROUP COVARIATES BASED ON TYPE ###
+
[…]

### 2025-01-10 scripts/zarchive/depr.genes/zarchive/04.2.grcca.R

@@ -1,0 +1,1586 @@
+
+########################################################################################
+
+# Explore GRCCA
+
+########################################################################################
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+# remotes::install_github("ElenaTuzhilina/RCCA")
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(RCCA)
+  library(biomaRt)
+  library(tidymodels)  
+  library(ggrepel)
+  library(raveio)
+  library(ggpubr)
+  library(ggchicklet)
+
+
+
+# set theme for plots -----------------------------------------------------
+
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+soft_power <- 3
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230228_185samples_19kgenes_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+  load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+  load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 40 & cut_height == 0.97) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module)
+
+# COVARIATES
+  load("data/covariates/185_all_covariates_clean.Rdata")
+
+# GENE SYMBOL TO ENSEMBL GENE ID MAPPING
[…]

### 2025-01-10 scripts/zarchive/depr.genes/zarchive/05.dev_expr_trajectories.R

@@ -1,0 +1,152 @@
+
+########################################################################################
+
+# identify windows of development where gene modules are over-/under-expressed
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(janitor)
+  library(tidymodels)
+  library(RColorBrewer)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+# data --------------------------------------------------------------------
+
+prefix <- "20221209_185samples_19kgenes_vsd_qSVA17_resids"
+
+
+# IDENTIFY MODULE SET OF INTEREST
+  load(paste0("objects/", prefix, "_SFT3_MODS.RDS")) # df_modules
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 50 & cut_height == 0.99) %>% 
+    dplyr::select(ensembl_gene_id, module) %>% 
+    mutate(module = factor(module, levels = 0:max(module)))
+
+# PsychENCODE DATA
+  df_psychencode_metadata <- read_table("data/eva/Psychencode_brainIDfiles.txt") %>% 
+    clean_names()
+  
+  df_psychencode_expr <- read_table("data/eva/psychencode_scaledlogtransformed_genexpr.txt")
+  
+  # save as object for future use
+  save(df_psychencode_metadata, df_psychencode_expr,
+       file = "objects/psychencode_development_expr_data.Rdata") #df_psychencode_metadata, df_psychencode_expr
+  
+
+# format PsychENCODE data -------------------------------------------------
+
+
+# IDENTIFY SAMPLES FROM THE REGION OF INTEREST (NEOCORTEX)  
+  neocortex_samples <- df_psychencode_metadata %>% 
+    filter(region_superbroad == "Neocortex") %>% 
+    pull(id)
+  
[…]

### 2025-01-10 scripts/zarchive/depr.genes/zarchive/06.control_WGCNA.R

@@ -1,0 +1,170 @@
+
+########################################################################################
+
+# Define modules based on control data only
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(WGCNA)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+# data --------------------------------------------------------------------
+
+  prefix <- "20221209_185samples_19kgenes_vsd_qSVA17_resids"
+  ctrl_prefix <- "20221209_55samplesCTRL_19kgenes_vsd_qSVA17_resids"
+
+# LOAD REGRESSED DATA
+  load(paste0("objects/", prefix, ".RDS")) # df_vsd_regress
+
+# FILTER FOR CONTROL ONLY
+  df_control <- df_vsd_regress %>% 
+    filter(grepl("control", sample))
+
+# CONVERT REGRESSED VSD COUNTS TO TRANSPOSED MATRIX
+  m_control_t <- df_control %>% 
+    pivot_wider(id_cols = sample, 
+                names_from = ensembl_gene_id, 
+                values_from = resids) %>% 
+    as.data.frame %>% column_to_rownames("sample")
+  
+
+# select soft thresholding power ------------------------------------------
+
+
+# CHOOSE A SET OF SOFT-THRESHOLDING POWERS TO TEST
+  powers <- 1:16
+
+# CALL THE NETWORK TOPOLOGY ANALYSIS FUNCTION
+  sft <- pickSoftThreshold(m_control_t, powerVector = powers, verbose = 5)
+  
+  # convert results to tibble
+  df_sft <- sft$fitIndices %>% as_tibble %>% clean_names
+  
+  # save
+  save(df_sft, file = paste0("objects/", ctrl_prefix, "_SFT.RDS"))
+
+# PLOT RESULTS   
+
+  load(paste0("objects/", ctrl_prefix, "_SFT.RDS"))
+  
[…]

### 2025-01-10 scripts/zarchive/depr.genes/zarchive/08.module_network_analysis.R

@@ -1,0 +1,769 @@
+
+
+########################################################################################
+
+# Characterize network properties of control modules in diagnostic groups
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+
+  library(tidyverse)
+  library(WGCNA)
+  library(igraph)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(corrr)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+  prefix <- "20221209_185samples_19kgenes_vsd_qSVA17_resids"
+  ctrl_prefix <- "20221209_55samplesCTRL_19kgenes_vsd_qSVA17_resids"
+
+# LOAD REGRESSED DATA
+  load(paste0("objects/", prefix, ".RDS")) # df_vsd_regress
+
+# FILTER FOR CONTROL ONLY
+  df_control <- df_vsd_regress %>% 
+    filter(grepl("control", sample))
+
+# LOAD CONTROL MODULE ASSIGNMENTS
+  load(paste0("objects/", ctrl_prefix, "_SFT4_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 50 & cut_height == 0.99) %>% 
+    dplyr::select(ensembl_gene_id, module) %>% 
+    mutate(module = factor(module, levels = 0:max(module)))
+
+# COVARIATE DATA
+  load("data/covariates/185_all_covariates_clean.Rdata")
+  
+
+
+# create graph based on gene correlations (not TOM weights) --------------
+
+  
+  df_mods_vsd <- df_modules_filt %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.genes/zarchive/TRY-covariate_relationships.R

@@ -1,0 +1,300 @@
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(tidymodels)
+  library(vip)
+  library(workflowsets)
+  library(ggpubr)
+  library(corrr)
+  library(pheatmap)
+  library(cluster)
+
+
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 18),
+                  axis.title = element_text(size = 15),
+                  axis.text = element_text(size = 12),
+                  strip.text = element_text(size = 15),
+                  legend.title = element_text(size = 15),
+                  legend.text = element_text(size = 12)))
+
+
+# data --------------------------------------------------------------------
+
+
+  soft_power <- 3
+  base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+  prefix <- "20230228_185samples_19kgenes_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+  load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+  load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 40 & cut_height == 0.97) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module)
+
+# COVARIATES
+  load(paste0(base_dir, "data/covariates/185_all_covariates_clean.Rdata"))
+
+
+
+
+# covariate correlations --------------------------------------------------
+
+
+  technical_covariates <- c("mapped_percent", "five_prime_three_prime_bias", "rin_acsg", 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/common/03.characterize_module_enrichment.R

@@ -1,0 +1,481 @@
+
+## !!!!!!!!!!!!!!!! COMMON !!!!!!!!!!!!
+
+########################################################################################
+
+# run GO and cell-type enrichment
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(stm)
+library(ggwordcloud)
+library(ggrepel)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 18),
+                  axis.title = element_text(size = 15),
+                  axis.text = element_text(size = 12),
+                  strip.text = element_text(size = 15),
+                  legend.title = element_text(size = 15),
+                  legend.text = element_text(size = 12)))
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_COMMON_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+soft_power <- 2
+minimum_size <- 40
+tree_cut_height <- 0.97
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# GENE TO TRANSCRIPT MAPPING
+load(paste0(base_dir, "objects/df_hsapiens_genome.RDS")) # df_hsapiens_genome
+
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/common/04.1.module_pc_analysis.R

@@ -1,0 +1,843 @@
+
+########################################################################################
+
+# Associate module PCs with covariates 
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(tidymodels)
+  library(vip)
+  library(workflowsets)
+  library(ggpubr)
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230314_185samples_26kCOMMONtranscripts_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress_common
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT3_MODS.RDS")) # df_modules
+load(paste0(base_dir, "objects/transcript_gene_go_term.RDS")) # df_transcript_gene_go_term
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(cut_height == 0.99 & min_size == 50) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+  load("data/covariates/185_all_covariates_clean.Rdata")
+  
+
+
+# run PCA on each module --------------------------------------------------
+
+
+# RUN PCA FOR EACH MODULE
+  df_mods_pca <- df_modules_filt %>% 
+    left_join(df_vsd_regress_common) %>% 
+    pivot_wider(id_cols = c(ensembl_transcript_id, module), 
+                names_from = sample, 
+                values_from = resids) %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/common/04.2.grcca.R

@@ -1,0 +1,576 @@
+
+########################################################################################
+
+# Common transcript level GRCCA results from Agoston's toolkit
+
+########################################################################################
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(biomaRt)
+library(tidymodels)  
+library(ggrepel)
+library(raveio)
+library(ggpubr)
+#library(ggchicklet)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+# data --------------------------------------------------------------------
+
+soft_power <- 3
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "20230314_185samples_26kCOMMONtranscripts_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress_common
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == 50 & cut_height == 0.99) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+df_covariates <- df_vsd_regress_common %>% 
+  ungroup %>% 
+  dplyr::select(-ensembl_transcript_id, -vsd, -resids) %>% 
+  distinct()
+
+# GENE SYMBOL TO ENSEMBL GENE ID MAPPING
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/common/archive/04.2.grcca.R

@@ -1,0 +1,1319 @@
+
+########################################################################################
+
+# Explore GRCCA
+
+########################################################################################
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+# remotes::install_github("ElenaTuzhilina/RCCA")
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(RCCA)
+  library(biomaRt)
+  library(tidymodels)  
+  library(ggrepel)
+  library(raveio)
+
+
+
+
+
+# set theme for plots -----------------------------------------------------
+
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230314_185samples_26kCOMMONtranscripts_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+# LOAD OBJECTS  
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress_common
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT3_MODS.RDS")) # df_modules
+load(paste0(base_dir, "objects/transcript_gene_go_term.RDS")) # df_transcript_gene_go_term
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(cut_height == 0.99 & min_size == 50) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+load("data/covariates/185_all_covariates_clean.Rdata")
+
+# GENE SYMBOL TO ENSEMBL GENE ID MAPPING
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall/03.characterize_module_enrichment.R

@@ -1,0 +1,479 @@
+
+
+########################################################################################
+
+# OVERALL TRANSCRIPTS: run GO and cell-type enrichment
+
+########################################################################################
+
+# libraries ---------------------------------------------------------------
+
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(stm)
+library(ggwordcloud)
+library(ggrepel)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 12),
+                  axis.title = element_text(size = 12),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 10),
+                  legend.text = element_text(size = 10)))
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_OVERALL_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+soft_power <- 2
+minimum_size <- 40
+tree_cut_height <- 0.983
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# GENE TO TRANSCRIPT MAPPING
+load(paste0(base_dir, "objects/df_hsapiens_genome.RDS")) # df_hsapiens_genome
+
+df_gene_to_transcript <- df_hsapiens_genome %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall/04.1.module_pc_analysis.R

@@ -1,0 +1,486 @@
+
+########################################################################################
+
+# OVERALL TRANSCRIPTS: Associate module PCs with covariates 
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(R.matlab)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230301_185samples_70ktranscripts_vst_qSVA123567_MP_RNAbatch_Race_resids"
+soft_power <- 3
+
+# LOAD OBJECTS  
+  load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+  load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+  df_modules_filt <- df_modules %>% 
+    filter(min_size == 40 & cut_height == 0.996) %>% 
+    dplyr::select(ensembl_transcript_id, module)
+
+# COVARIATES
+  load("data/covariates/185_all_covariates_clean.Rdata")
+  
+
+
+# run PCA on each module --------------------------------------------------
+
+
+# RUN PCA FOR EACH MODULE
+  df_mods_pca <- df_modules_filt %>% 
+    left_join(df_vsd_regress) %>% 
+    pivot_wider(id_cols = c(ensembl_transcript_id, module), 
+                names_from = sample, 
+                values_from = resids) %>% 
+    group_by(module) %>% 
+    nest() %>% 
+    
+    mutate(pca = map(.x = data,
+                     .f = ~ .x %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall/04.2.grcca.R

@@ -1,0 +1,450 @@
+
+########################################################################################
+
+# OVERALL transcript level GRCCA results from Agoston's toolkit
+
+########################################################################################
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(biomaRt)
+library(tidymodels)  
+library(ggrepel)
+library(raveio)
+library(ggpubr)
+#library(ggchicklet)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 11),
+                  axis.title = element_text(size = 10),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 8),
+                  legend.text = element_text(size = 8)
+            )
+)
+
+# DX COLORS
+dx_colors <- c("#0072B2", "#E69F00", "#009E73", "#9966FF")
+names(dx_colors) <- c("Control", "BD", "MDD", "SCZ")
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_OVERALL_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+soft_power <- 2
+minimum_size <- 40
+tree_cut_height <- 0.984
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/02.case_control_WGCNA.R

@@ -1,0 +1,456 @@
+
+##### !!!!! SUBSET !!!!
+
+##############################################################################################################
+
+# OVERALL TRANSCRIPTS: Select soft-thresholding power, calculate adjacency & TOM, cluster, and assign modules
+
+##############################################################################################################
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(WGCNA)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(ggpubr)
+library(paletteer)
+library(segmented)
+
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 12),
+                  axis.title = element_text(size = 12),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 10),
+                  legend.text = element_text(size = 10)))
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+
+# LOAD GENE NAMES
+load(paste0(base_dir, "objects/08Mar2024_GENES_qSVAgeSexRaceGC.RDS")) # df_vsd_regress (gene_level)
+gene_ids <- df_vsd_regress %>% dplyr::select(-sample) %>% colnames
+length(gene_ids) # n = 18,677
+rm(list = "df_vsd_regress")
+
+# LOAD TRANSCRIPT DATA
+prefix <- "08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC"
+
+# RAW TRANSCRIPT COUNTS
+df_transcript_raw_counts <- read.csv(paste0(base_dir, "data/transcript_count_matrix.csv")) %>% 
+  as_tibble() %>% 
+  dplyr::rename("ensembl_transcript_id" = "X") %>% 
+  rename_all(~str_remove(.x, "_MergedBam_stringtieOutput")) %>% 
+  clean_names() %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/03.characterize_module_enrichment.R

@@ -1,0 +1,569 @@
+
+
+########################################################################################
+
+# OVERALL TRANSCRIPTS: run GO and cell-type enrichment
+
+########################################################################################
+
+
+
+# setup -------------------------------------------------------------------
+
+
+## LIBRARIES
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(tidymodels)
+library(tidytext)
+library(stm)
+library(ggwordcloud)
+library(ggrepel)
+library(readxl)
+library(writexl)
+library(ggalluvial)
+library(clusterProfiler)
+
+## SET DIRECTORIES
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1"
+
+## PARAMETERS FOR MODULE SELECTION
+soft_power <- 2
+minimum_size <- 35
+tree_cut_height <- 0.988
+
+## LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+## IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+    filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+    unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+    arrange(mod_set, module) %>% 
+    
+    # rename modules to specify gene-level
+    mutate(module = paste0("transcriptM", module) %>% factor(levels = paste0("transcriptM", levels(module)))
+    )
+
+## GENE TO TRANSCRIPT MAPPING
+load(paste0(base_dir, "objects/hsapiens_genome_v110.RDS")) # df_hsapiens_genome
+df_gene_to_transcript <- df_hsapiens_genome %>% 
+  dplyr::select(transcript_id, transcript_name, gene_id, gene_name) %>% 
+  distinct() %>% 
+  dplyr::rename("gene_symbol" = "gene_name", "transcript_symbol" = "transcript_name", 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/GRCCA/04.1.define_input_matrices.R

@@ -1,0 +1,109 @@
+
+########################################################################################
+
+# Define X.mat, Y.mat, and C.mat to include in GRCCA analysis
+
+########################################################################################
+
+### X.mat
+
+# define gene order to align modules & genes
+df_order <- df_modules_filt %>% 
+    dplyr::select(ensembl_transcript_id, module) %>% 
+    distinct() %>% 
+    arrange(module, ensembl_transcript_id) #%>% 
+
+## TURN OFF TO KEEP GRAY MODULE GENES
+#filter(module != 0)
+
+# X grouping vector    
+x_group <- df_order %>% 
+    mutate(module = as.numeric(as.character(module))) %>% 
+    pull(module)
+
+# X 
+X.mat <- df_vsd_regress_filt %>% 
+    dplyr::select(sample, all_of(df_order$ensembl_transcript_id)) %>% 
+    column_to_rownames("sample")
+
+sample_order <- rownames(X.mat)
+transcripts <- colnames(X.mat)
+length(transcripts) # n = 54302 (_WITHGRAY CVq1); # n = 24626 (_withGray CVq1)
+
+
+### Y.MAT AND C.MAT
+
+# define number of MCA dimensions to include
+n_mca_dim <- 8
+
+# Define variables to include in Y and C matrices
+YC.mat <- df_covariates_numeric %>% 
+    
+    # Select covariates (MC loadings?)
+    left_join(df_ind_loadings, by = join_by(sample)) %>%
+    dplyr::select(sample, all_of(paste0("dim", seq(1, n_mca_dim)))
+                  
+    ) %>%
+    mutate_if(is.numeric, ~ifelse(is.na(.x), 0, .x)) %>%
+    
+    # create dx columns
+    mutate(control = ifelse(grepl("control", sample), 1, 0),
+           bipolar = ifelse(grepl("bipolar", sample), 1, 0),
+           mdd = ifelse(grepl("mdd", sample), 1, 0),
+           schizo = ifelse(grepl("schizo", sample), 1, 0)
+    ) %>% 
+    dplyr::select(sample, bipolar, mdd, schizo, everything(), -control) %>% 
+    mutate(sample = factor(sample, levels = sample_order)) %>% 
+    arrange(sample) %>% 
+    column_to_rownames("sample")
+
+# Split into Y and C matrices
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/GRCCA/04.4.characterize_results.R

@@ -1,0 +1,676 @@
+
+########################################################################################
+
+# Characterize transcript-level GRCCA results
+
+########################################################################################
+
+
+# X-Y LV correlation ------------------------------------------------------
+
+
+cor.test(df_lvs$lvx, df_lvs$lvy, method = "spearman")
+
+df_lvs %>% 
+  ggplot(aes(x = lvx, y = lvy)) +
+  geom_point(aes(color = dx), alpha = 0.7, size = 3) +
+  geom_smooth(method = lm, se = FALSE, color = "black") +
+  stat_cor() +
+  scale_color_manual(values = dx_colors) +
+  guides(color = guide_legend(title = "diagnosis")) +
+  labs(x = "X latent variable (x matrix • x weights)",
+       y = "Y latent variable (y matrix • y weights)",
+       title = "Transcript x-y latent variable correlation") +
+  theme(legend.position = c(0.8, 0.15))
+
+
+
+
+# Fig 5A | y weights ---------------------------------------------------------------
+
+## WEIGHTS
+df_y_res %>% 
+  mutate(significant = ifelse(p_adj < 0.05 & abs(z_score) >= 2, "*", "")) %>% 
+
+  ggplot(aes(x = reorder(covariate, weight), y = weight)) +
+  geom_col(aes(fill = covariate, alpha = abs(weight)), color = "black") +
+  geom_errorbar(aes(ymin = weight - sd, ymax = weight + sd), width = 0.2) +
+  geom_text(aes(label = significant, y = 0), size = 10, vjust = 1.2) +
+  geom_hline(aes(yintercept = 0), color = "black") +
+  scale_fill_manual(values = c(dx_colors, rep("gray", 6))) +
+  scale_alpha_continuous(range = c(0.1, 1)) +
+  scale_x_discrete(labels = function(x) {str_replace(x, "_", " ")}) +
+  ylim(c(-0.15, 0.15)) +
+  labs(x = "", y = "weight", 
+       title = "Covariate weights",
+       caption = "Error bar shows +/- 1 sd") +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 0.9))
+
+# save to project dir
+map(
+  .x = c(".png", ".pdf"),
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/GRCCA/setup.R

@@ -1,0 +1,106 @@
+
+########################################################################################
+
+# Load data and module assignments to run GRCCA & analyze outputs
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(patchwork)
+library(janitor)
+library(RColorBrewer)
+library(biomaRt)
+library(tidymodels)  
+library(ggrepel)
+library(raveio)
+library(ggpubr)
+library(clusterProfiler)
+library(msigdbr)
+library(org.Hs.eg.db)
+library(ggh4x)
+library(ggVennDiagram)
+library(tidytext)
+library(writexl)
+
+#library(ggchicklet)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 11),
+                    axis.title = element_text(size = 10),
+                    axis.text = element_text(size = 10),
+                    strip.text = element_text(size = 10),
+                    legend.title = element_text(size = 8),
+                    legend.text = element_text(size = 8)
+              )
+)
+
+# DX COLORS
+dx_colors <- c("#0072B2", "#E69F00", "#009E73", "#9966FF")
+names(dx_colors) <- c("Control", "BD", "MDD", "SCZ")
+
+# ONTOLOGY COLORS
+ontology_colors <- c("BP" = "#F8766D", "CC" = "#00BA38", "MF" = "#619CFF")
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1"
+
+soft_power <- 2
+minimum_size <- 35
+tree_cut_height <- 0.988
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/overall_subset/GRCCA/validate_results.R

@@ -1,0 +1,315 @@
+
+################################################################################
+
+# TRANSCRIPT-LEVEL: Validate GRCCA results using external data
+
+################################################################################
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/full_analysis_scripts/transcripts/overall_subset/GRCCA/04.3.load_GRCCA_results.R"))
+figures_dir <- paste0(base_dir, "outputs/benchmarking_res/")
+
+
+# load data ---------------------------------------------------------------
+
+## GWAS DATA
+df_scz_gwas <- readxl::read_excel(paste0(base_dir, "data/trubetskoy_2023_scz_gwas.xlsx"),
+                                  sheet = 2
+) %>% 
+    dplyr::select(gene_symbol, gene_ensembl) %>% 
+    distinct() %>% 
+    left_join(df_gene_to_transcript)
+
+## DE data
+load(paste0(base_dir, "objects/akula_et_al_DE_results.RDS")) # df_akula_res
+
+## Consensus DE data
+df_consensus_degs <- readxl::read_excel(paste0(base_dir, "data/merikangas_2022_consensus_DEGs.xlsx")) %>% 
+    left_join(df_gene_to_transcript)
+
+## Rare variants
+df_exome_urvs <- readxl::read_excel(paste0(base_dir, "data/singh_2022_exome_URVs.xlsx")) %>% 
+    left_join(df_gene_to_transcript)
+
+
+# Akula DET overlap -------------------------------------------------------
+
+## Hypergeometric
+
+# pull genes with transcripts that were in analysis
+transcript_universe <- df_x_res %>% 
+    filter(!is.na(ensembl_transcript_id)) %>% 
+    pull(ensembl_transcript_id) %>% 
+    unique
+length(transcript_universe) # 18031
+
+# pull Akula hits
+akula_transcripts <- df_akula_res %>% 
+    filter(level == "transcript" & comparison == "schizo_ctrl" & padj < 0.05) %>% 
+    pull(id) %>% 
+    unique
+length(akula_transcripts) # 470
+
+# find Akula hits that were part of our analysis
+akula_transcripts_in_universe <- intersect(transcript_universe, akula_transcripts) %>% unique
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/rare/02.case_control_WGCNA.R

@@ -1,0 +1,265 @@
+
+# !!!!!! RARE !!!!!!
+
+########################################################################################
+
+# Select soft-thresholding power, calculate adjacency & TOM, cluster, and assign modules
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(WGCNA)
+  library(sva)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+
+
+# set theme for plots -----------------------------------------------------
+
+theme_set(theme_bw() +
+            theme(plot.title = element_text(size = 12),
+                  axis.title = element_text(size = 12),
+                  axis.text = element_text(size = 10),
+                  strip.text = element_text(size = 10),
+                  legend.title = element_text(size = 10),
+                  legend.text = element_text(size = 10)))
+
+# data --------------------------------------------------------------------
+
+
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_TRANSCRIPTS_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+# RAW TRANSCRIPT COUNTS
+df_transcript_raw_counts <- read.csv(paste0(base_dir, "data/transcript_count_matrix.csv")) %>% 
+  as_tibble() %>% 
+  dplyr::rename("ensembl_transcript_id" = "X") %>% 
+  rename_all(~str_remove(.x, "_MergedBam_stringtieOutput")) %>% 
+  clean_names() %>% 
+  rename_all(~str_remove(.x, "x")) 
+
+# LOAD REGRESSED DATA
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+
+
+
+# filter for rare transcripts ---------------------------------------------
+
+
+# Greater than 10 counts but less than 100 counts in 80% of samples
+
+  rare <- df_transcript_raw_counts %>% 
+    mutate_if(is.numeric, ~ifelse(.x > 10 & .x < 100, 1, 0)) %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/rare/03.characterize_module_enrichment.R

@@ -1,0 +1,481 @@
+
+# !!!!!! RARE !!!!!!
+
+########################################################################################
+
+# run GO and cell-type enrichment
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(stm)
+  library(ggwordcloud)
+  library(ggrepel)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_RARE_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+soft_power <- 2
+minimum_size <- 35
+tree_cut_height <- 0.9936
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# GENE TO TRANSCRIPT MAPPING
+load(paste0(base_dir, "objects/df_hsapiens_genome.RDS")) # df_hsapiens_genome
+
+df_gene_to_transcript <- df_hsapiens_genome %>% 
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/rare/04.1.module_pc_analysis.R

@@ -1,0 +1,932 @@
+
+# !!!!!! RARE !!!!!!
+
+########################################################################################
+
+# Associate module PCs with covariates 
+
+########################################################################################
+
+
+# libraries ---------------------------------------------------------------
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(tidymodels)
+  library(tidytext)
+  library(R.matlab)
+  library(ggpubr)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+prefix <- "15Nov2023_RARE_regress_qSV1.7_ageDeath_RNAbatch_GCperc"
+
+soft_power <- 2
+minimum_size <- 35
+tree_cut_height <- 0.9936
+
+# LOAD OBJECTS 
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_clean
+
+# DRUG MCA RESULTS
+load(paste0(base_dir, "objects/drug_MCA_results.Rdata"))
+
+# TRANSCRIPT SYMBOL TO ENSEMBL TRANSCRIPT ID MAPPING
[…]

### 2025-01-10 scripts/zarchive/depr.transcripts/rare/archive/04.2.grcca.R

@@ -1,0 +1,1924 @@
+
+########################################################################################
+
+# Explore GRCCA
+
+########################################################################################
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+# remotes::install_github("ElenaTuzhilina/RCCA")
+
+  library(tidyverse)
+  library(patchwork)
+  library(janitor)
+  library(RColorBrewer)
+  library(RCCA)
+  library(biomaRt)
+  library(tidymodels)  
+  library(ggrepel)
+  library(raveio)
+  library(viridis)
+  library(ggchicklet)
+  library(ggpubr)
+
+
+
+# set theme for plots -----------------------------------------------------
+
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+  
+
+
+
+# data --------------------------------------------------------------------
+
+
+base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/projects/WGCNA/acsg_wgcna/"
+prefix <- "20230315_185samples_25kRAREtranscripts_vst_qSVA123567_MP_RNAbatch_Race_resids"
+soft_power <- 3
+
+# LOAD OBJECTS  
+load(paste0(base_dir, "objects/", prefix, ".RDS")) # df_vsd_regress_rare
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_MODS.RDS")) # df_modules
+
+# IDENTIFY MODULE SET OF INTEREST  
+df_modules_filt <- df_modules %>% 
+  filter(min_size == 30 & cut_height == 0.9986) %>% 
+  unite("mod_set", c(sft, min_size, cut_height), sep = "_") %>% 
+  arrange(mod_set, module)
+
+# COVARIATES
+load("data/covariates/185_all_covariates_clean.Rdata")
+
+# GENE SYMBOL TO ENSEMBL GENE ID MAPPING
+#load("objects/transcript_gene_go_term.RDS") # df_transcript_gene_go_term
[…]

### 2025-01-10 scripts/zarchive/explore_other_covariates.R

@@ -1,0 +1,514 @@
+
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+    library(tidyverse)
+    library(purrr)
+    library(janitor)
+    library(naniar)
+    library(Hmisc)
+    library(DESeq2)
+
+
+
+# data --------------------------------------------------------------------
+
+
+
+    # COUNTS
+    df_raw_counts <- read.table("data/185samples_allGenes_auto-PARs_noSexChrs_noFilters_KoryAnalysis_21Kgenes.txt") %>% 
+      as_tibble(rownames = "ensembl_gene_id") %>% 
+      rename_at(2:ncol(.), ~substr(.x, start = 2, stop = nchar(.x))) %>% 
+      clean_names() %>% 
+      dplyr::rename_at(2:ncol(.), ~str_remove(.x, "x"))
+
+    # COVARIATES
+    df_covariates_all <- read.table("data/covariates/185samples_multinomialDx_46covariates_10evecs.txt", 
+               header = TRUE,
+               fill = TRUE,
+               sep = "\t") %>% 
+      as_tibble() %>% 
+      clean_names() %>% 
+      rename("sample" = "x") %>% 
+      mutate_if(is.character, ~tolower(.x))
+      
+      # remove outlier!!
+
+
+
+
+# reformat covariates -----------------------------------------------------
+
+
+    # CLEAN UP COVARIATE DATA - COMBINE COVARIATE CATAEGORIES
+    df_covariates_all_clean <- df_covariates_all %>% 
+      
+      dplyr::select(-contains(c("evec", "sq", "_1", "_2", "hallucinogens"))) %>% # no one was doing hallucinogens
+      select(-etoh_results) %>% 
+      
+      dplyr::rename("anti_depressant" = "antidepressants") %>% 
+      select(-anti_depress) %>% 
+      
+      dplyr::rename("anti_anxiety" = "sedative_hypnotic_anxiolitics") %>% 
+      select(-benzos) %>% 
+      
+      dplyr::rename("cannabis" = "cannabinoids") %>% 
+      select(-thc) %>% 
+      
+      mutate(mood_stabilizers = case_when(
+        
+        mood_stabilizers == "negative" & mood_stab == "negative" ~ "negative",
+        mood_stabilizers == "positive" & mood_stab == "positive" ~ "positive",
+        mood_stabilizers == "negative" & mood_stab == "" ~ "negative",
+        mood_stabilizers == "positive" & mood_stab == "" ~ "positive",
[…]

### 2025-01-10 scripts/zarchive/identify_sample_outliers.R

@@ -1,0 +1,733 @@
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+    library(tidyverse)
+    library(janitor)
+    library(plotly)
+    library(RColorBrewer)
+    library(ggdendro)
+    library(microplot)
+    library(tidymodels)            
+    library(purrr)
+    library(grid)
+    library(dendextend)
+    library(themis)
+    library(tidytext)
+
+
+# set theme for plots -----------------------------------------------------
+
+  theme_set(theme_bw() +
+              theme(plot.title = element_text(size = 18),
+                    axis.title = element_text(size = 15),
+                    axis.text = element_text(size = 12),
+                    strip.text = element_text(size = 15),
+                    legend.title = element_text(size = 15),
+                    legend.text = element_text(size = 12)))
+
+
+# data --------------------------------------------------------------------
+
+
+# NORMALIZED AND FILTERED COUNTS
+    # 14k genes, vsd regressed, RIN_GC_race_libraryBatch regressed
+    df_vsd_resids <- read.csv("outputs/20220613_185samples_14kgenes_vsd_RIN_mappedPerc_GC_race_resids.csv") %>% 
+      as_tibble() %>% 
+      dplyr::select(-X) %>% 
+      rename_at(2:ncol(.), ~str_remove(.x, "X"))
+    
+    # transpose
+    df_vsd_resids_t <- df_vsd_resids %>%
+      
+      # transpose & convert to tibble
+      dplyr::select(-1) %>%
+      t %>%
+      as.data.frame %>%
+      as_tibble %>%
+      
+      # rename columns with gene id from original df
+      rename_all(~df_vsd_resids %>% pull(ensembl_gene_id)) %>%
+      
+      # add sample id col using colnames of original df (to maintain order)
+      dplyr::mutate(sample = df_vsd_resids %>% dplyr::select(-1) %>% colnames, .before = 1)
+
+# COVARIATES
+    load("data/covariates/185_all_covariates_clean.Rdata") # df_covariates_all_clean, df_covariates_all_clean_numeric
+
+    
+
+# gene tree ---------------------------------------------------------------
+
+
+    # cluster norm counts
+    sample_tree <- df_vsd_resids_t %>% 
+      as.data.frame %>%
[…]

### 2025-01-10 scripts/zarchive/topic_modeling_demo/topic_modeling.R

@@ -1,0 +1,201 @@
+
+
+
+# libraries ---------------------------------------------------------------
+
+
+library(tidyverse)
+library(tidymodels)
+library(tidylo)
+library(stm)
+library(purrr)
+library(RColorBrewer)
+library(ggwordcloud)
+library(ggrepel)
+library(patchwork)
+
+#devtools::install_github("juliasilge/tidytext", force = TRUE) 
+library(tidytext)
+
+theme_set(theme_bw())
+
+
+# data --------------------------------------------------------------------
+
+soft_power <- 3
+base_dir <- "~/Documents/work/PhD/projects/WGCNA/sgacc_wgcna_grcca/"
+prefix <- "20230228_185samples_19kgenes_vst_qSVA123567_MP_RNAbatch_Race_resids"
+
+load(paste0(base_dir, "objects/", prefix, "_SIGNED_SFT", soft_power, "_GO_RES.RDS")) # df_mods_go
+
+
+# topic modeling ---------------------------------------------------------
+
+
+# Unnest tokens - each word is a separate row in a tibble
+df_go_tokens <- df_mods_go %>% 
+  
+  # select the module set we're using (there's many in this tibble across different parameter combos)
+  filter(mod_set == "3_40_0.97" & p_adj < 0.05) %>% 
+  
+  # remove any rows where the GO term is NA
+  filter(!is.na(term)) %>% 
+  
+  # we are running this for each module, so that should be our grouping variable
+  group_by(module) %>% 
+  
+  # split the term descript up so each word is a row in the tibble
+  unnest_tokens(word, term) %>% 
+  
+  # remove stop words (uninteresting words like of, from, and, the, etc.)
+  anti_join(get_stopwords())
+
+head(df_go_tokens)
+
+# Create a sparse matrix
+go_sparse <- df_go_tokens %>%
+  
+  # count number of times a word appears in each module
+  dplyr::count(module, word) %>%
+  
+  # convert module from factor to numeric
+  mutate(module = as.numeric(as.character(module))) %>% 
+  
+  # filter for words that appear at least 3 times
+  # also, in this case, I've removed words that appear in a lot of path names but are effectively meaningless without context
+  filter(n > 3 & !(word %in% c("positive", "negative", "regulation", "response"))) %>%
+  cast_sparse(module, word, n)
+
+# Fit topic model using Structural Topic Models (http://www.structuraltopicmodel.com/)
[…]

### 2025-01-10 scripts/07.figures_markdown/Fig1_analysisPipeline.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure 1: Study methods overview"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig1_analysisPipeline/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load DE data
load(paste0(base_dir, "outputs/updated_figures/objects/Fig2_DGEres/DE_results.RDS")) # df_de_res
```

## 1. DE box plot

```{r Fig1.DEbox, fig.width = 1, fig.height = 1}
## Select gene for plotting
top_gene <- df_de_res %>% top_n(n = 1, wt = -stat) %>% pull(ensembl_gene_id)

## Plot
df_vsd_regress %>% 
    left_join(df_covariates) %>% 
    dplyr::select(dx, sample, all_of(top_gene)) %>% 
    filter(str_detect(sample, "schizo|control")) %>%

    ggplot(aes(x = dx, y = get(top_gene), color = dx)) +
    geom_point(position = position_jitter(width = 0.2), size = 0.25) +
    geom_boxplot(fill = "transparent", outlier.shape = NA, width = 0.5) +
    scale_color_manual(values = dx_colors, guide = "none") +
    labs(x = NULL, y = NULL, title = NULL) +
    theme(
        axis.text = element_blank(),
        axis.ticks = element_blank()
    )

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "1.DGE_boxplots", .x), width = 1, height = 1)
)

```

## 2. Volcano plot

```{r Fig1.volcano, fig.height = 1, fig.width = 1.25}
## Plot
df_de_res %>% 
    mutate(fill = ifelse(pvalue > 0.05, NA_real_, log2fold_change),
           significant = ifelse(ensembl_gene_id == top_gene, "yes", "no"),
           label = ifelse(ensembl_gene_id == top_gene, gene_symbol, "")
    ) %>%

    ggplot(aes(x = log2fold_change, y = -log10(pvalue))) +
    geom_point(aes(fill = fill, color = significant), shape = 21, size = 0.75) +
    geom_hline(aes(yintercept = -log10(0.05)), linewidth = 0.25, color = "black") +
    geom_text_repel(aes(label = label), min.segment.length = 0, size = 2.25) +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         values = rescale(c(-0.17, 0, 0.17)),
                         limits = c(-0.171, 0.171),
                         na.value = "lightgray", guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), na.value = "transparent") +
    labs(x = NULL, y = NULL, title = NULL) +
    theme(
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none"
    )

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "2.DGE_volcano", .x), width = 1.25, height = 1)
)
```

## 3. WGCNA dendrogram & colors

```{r Fig1.WGCNA, fig.height = 1}
## Convert regressed VSD counts to transposed matrix
m_vsd_regress <- df_vsd_regress %>% column_to_rownames("sample")

## Calculate co-expression similarity and adjacency
set.seed(20240306)
adjacency <- adjacency(m_vsd_regress, type = "signed hybrid", power = soft_power)

## Topological overlap matrix
doParallel::registerDoParallel()
set.seed(20240306)
TOM <- TOMsimilarity(adjacency, TOMType = "signed")
colnames(TOM) <- colnames(adjacency)
rownames(TOM) <- rownames(adjacency)
diag(TOM) <- NA

## Subset for plotting
baby_TOM <- TOM[1:1000, 1:1000]

## Set colors for annotation
annotation <- df_modules_filt %>% 
    dplyr::select(ensembl_gene_id, module) %>% 
    dplyr::filter(ensembl_gene_id %in% colnames(baby_TOM)) %>% 
    mutate(module = as.character(module)) %>% 
    mutate(ensembl_gene_id = factor(ensembl_gene_id, levels = colnames(baby_TOM))) %>% 
    arrange(ensembl_gene_id) %>% 
    column_to_rownames("ensembl_gene_id")
all(rownames(annotation) == rownames(baby_TOM))
module_colors <- df_modules_filt %>% 
    mutate(module = as.character(module)) %>% 
    dplyr::filter(ensembl_gene_id %in% colnames(baby_TOM)) %>% 
    dplyr::select(module, color) %>% 
    distinct %>% 
    deframe
annotation_col = list(module = module_colors)

## Plot
TOM_pheatmap <- pheatmap(
    mat = baby_TOM,
    color = colorRampPalette(c("white", "black"))(100),
    #color = colorRampPalette(brewer.pal(9, "YlOrRd"))(100),
    cluster_rows = FALSE,
    cluster_cols = TRUE,
    #annotation_row = annotation,
    annotation_col = annotation,
    annotation_colors = annotation_col,
    annotation_names_row = FALSE, 
    annotation_names_col = FALSE,
    border_color = NA,
    show_rownames = FALSE,
    show_colnames = FALSE,
    legend = FALSE,
    annotation_legend = FALSE
)

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "3.WGCNA_dendrogram", .x), plot = TOM_pheatmap$gtable, width = fig_width, height = 1)
)
```

## 4. Dummy GSEA distribution

```{r Fig1.GSEA, fig.width = 2.5, fig.height = 2}
## Create dummy data for GSEA distribution
set.seed(124)
df_gsea_dummy <- tibble(
    dist = rnorm(n = 1000, mean = 0, sd = 1),
    points = rnorm(n = 1000, mean = 0.5, sd = .5)
)

## Plot
df_gsea_dummy %>% 
    ggplot() +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density(aes(x = dist)) +
    geom_point(mapping = aes(x = points, y = -0.1, color = points), 
               pos = position_jitter(height = 0.085), size = 0.75
    ) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-2.2, 2.2), guide = "none") +
    labs(x = NULL, y = NULL, title = NULL) +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "none"
    )

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "4.GSEA_dist", .x), width = fig_width, height = fig_height)
)
```

## 5. Dummy risk gene correlation

```{r Fig1.riskGene}

```

### 2025-01-10 scripts/07.figures_markdown/Fig3_characterizeGRCCA.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure 3: GRCCA results benchmarking & function enrichments"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig3_characterizeGRCCA/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig3_characterizeGRCCA/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load df_x_res (in Fig3_GRCCAres)
load(paste0(analysis_objects_dir, "../Fig2_GRCCAres/GRCCA_results.Rdata")) # df_x_res
```

### A | GRCCA results are significantly and specifically enriched for SCZ risk genes

```{r Fig3a.prep}
df_fig3a <- df_grcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))

## Identify genes to label in bottom panel
top_genes <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    top_n(n = 6, wt = abs(pearsons_r)) %>% 
    pull(ensembl_gene_id)

## Subset X res for genes in SCZ common broad list for bottom panel
df_fig3a_bottom <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    mutate(significant = ifelse(p_adj < 0.05, 1, 0) %>% factor) %>% 
    arrange(significant, abs(pearsons_r)) %>% 
    mutate(label = ifelse(ensembl_gene_id %in% top_genes, gene_symbol, ""))
```

```{r Fig3a.plot.top}
fig3a <- df_fig3a %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "A | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
```

```{r Fig3a.plot.bottom}
scale_max <- df_x_res %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)
fig3a_bottom <- df_x_res %>%
    ggplot(aes(x = pearsons_r)) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density() +
    geom_point(data = df_fig3a_bottom, 
               mapping = aes(x = pearsons_r, y = -1, fill = pearsons_r, color = significant), 
               pos = jitter_pos, shape = 21, size = 1
    ) +
    geom_text_repel(data = df_fig3a_bottom,
                    mapping = aes(x = pearsons_r, y = -1, label = label),
                    pos = jitter_pos, 
                    size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
    ) +

    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-scale_max, scale_max), guide = "none") +
    scale_color_manual(values = c("transparent", "black"), guide = "none") +
    labs(x = expression("Structure correlation (" * italic(r)[x] * ")"), y = NULL, 
         title = NULL) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_rect(fill = "transparent", color = "black", size = 1),
          axis.line = element_blank()
    )
```

```{r Fig3a.save, fig.height = 4}
fig3a + plot_spacer() + fig3a_bottom + 
    plot_layout(nrow = 3, heights = c(1, 0.2, 0.65))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.Risk_gene_enrichment", .x), width = fig_width, height = 4)
)
```

### B | GRCCA cell-type enrichments

```{r Fig3b.plot}
fig3b <- df_grcca_fgsea_cell %>% 
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>%

    ggplot(aes(x = NES, y = reorder(cell_type, NES))) +
    geom_col(aes(fill = NES), color = "gray", shape = 21) +
    geom_text(aes(label = sig, x = NES), vjust = 0.75, size = 4, check_overlap = TRUE) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3.39, 3.39)) +
    coord_cartesian(clip = "off") +
    labs(x = "NES", y = NULL,
         title = "B | Cell-type enrichment") +
    theme(legend.position = "none")
```

```{r Fig3b.save, fig.height = 1.5}
fig3b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.cell_type_enrichment", .x), width = fig_width, height = 1.5)
)
```

### C | Developmental expression trajectories of GRCCA deciles

```{r Fig3c.prep}
## Calculate the mean expression for each decile of GRCCA results across development
df_fig3c <- df_psychencode_grcca %>% #dplyr::filter(regioncode == "MFC") %>% 
    group_by(decile, window) %>% 
    summarise(median_expression = median(dev_expression),
              sd_expression = sd(dev_expression)
    )

## Rename windows with time period
psychencode_windows <- c("5-9",
                "12-13",
                "16-18",
                "19-22",
                "PCW35-PY0.3",
                "0.5-2.5",
                "2.8-10.7",
                "13-19",
                "21-64")
names(psychencode_windows) = 1:9

## Define colors for deciles
decile_colors <- colorRampPalette(c(gene_weight_color_scale[5], "gray", gene_weight_color_scale[95]))(10) %>% rev
```

```{r Fig3c.plot}
fig3c <- df_fig3c %>% 
    ggplot(aes(x = window, y = median_expression, color = decile)) +
    geom_vline(xintercept = 5, color = "gray") +
    geom_smooth(method = "loess", se = FALSE) +
    scale_x_continuous(breaks = seq(1, length(psychencode_windows), 1),
                       labels = psychencode_windows) +
    scale_color_manual(values = decile_colors) +
    guides(color = guide_legend()) +
    labs(x = "← PCW | PY →", y = "Median expression",
         title = "C | Developmental \nexpression") +
    theme(
        legend.key.size = unit(0.3, "cm"),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

```{r fig3c.save, fig.width = 2.5, fig.height = 2.5}
fig3c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.PsychENCODE_development", .x), width = fig_width, height = 2.5)
)
```

### D | GRCCA GO enrichments

```{r Fig3d.prep}
## Determine labels
fig3d_labels <- c(
    "cilium movement",
    "motile cilium",
    "axoneme",
    #"axonemal microtubule",
    "axoneme assembly",
    "adaptive immune response",
    "outer dynein arm assembly",
    "MHC class II protein complex",
    "T cell costimulation",
    "MHC class II protein complex binding",
    "minus-end-directed microtubule motor activity",
    "dynein light intermediate chain binding",

    "syntaxin-1 binding",
    "ubiquitin-dependent protein catabolic process",
    "endosome to lysosome transport",
    "cytoplasmic translation",
    "regulation of autophagy",
    "postsynaptic density",
    "presynaptic endocytic zone membrane",
    "ATP binding",
    "ATP hydrolysis activity",
    "protein ubiquitination"
)

## Filter for terms of interest & assign labels
df_fig3d <- df_grcca_fgsea_go %>% 
    dplyr::filter(padj < 0.05) %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(term %in% fig3d_labels, str_wrap(term, 25), ""))
```

```{r Fig3d.plot}
fig3d <- df_fig3d %>%
    ggplot(aes(x = padj, y = NES)) +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 2.75, force = 20,
                    box.padding = 0.25, min.segment.length = 0, max.overlaps = 30) +
    geom_hline(yintercept = 0, color = "gray") +
    facet_wrap(vars(ontology)) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3, 3), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.03, 0.05)) +
    coord_cartesian(clip = "off") +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    labs(x = "Enrichment FDR", y = "NES",
         title = "D | Functional pathway enrichments") +
    theme(legend.position = "bottom",
          legend.justification = "center",
          title.justification = "center",
          legend.key.size = unit(0.3, "cm")
    )
```

```{r Fig3d.save, fig.height = 4, fig.width = 5}
fig3d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "D.GO_enrichment", .x), width = 5, height = 4)
)
```

### 2025-01-10 scripts/07.figures_markdown/Fig4_DGEres.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure 4: Differential gene expression analysis results & comparison with GRCCA"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig4_DGEres/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig4_DGEres/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load GRCCA results (in Fig2_GRCCAres)
load(paste0(analysis_objects_dir, "../Fig2_GRCCAres/GRCCA_results.Rdata")) # df_x_res

## Load GRCCA enrichments
load(paste0(analysis_objects_dir, "../Fig3_characterizeGRCCA/GRCCA_enrichment_benchmark_res.Rdata")) # df_grcca_fgsea...
```

### A | Volcano plot of DEGs

```{r Fig4a.prep}
df_fig4a <- df_de_res %>% 
    mutate(gene_symbol = ifelse(is.na(gene_symbol), ensembl_gene_id, gene_symbol)) %>% 
    mutate(fill = ifelse(pvalue > 0.05, NA_real_, log2fold_change)) %>% 
    mutate(significant = ifelse(padj < 0.05, "yes", "no")) %>% 
    mutate(label = ifelse(-log10(pvalue) > 4.2 | (pvalue < 0.05 & abs(log2fold_change) > 0.11), gene_symbol, NA))
```

```{r Fig4a.plot}
fig4a <- df_fig4a %>% 
    ggplot(aes(x = log2fold_change, y = -log10(pvalue))) +
    geom_point(aes(fill = fill, color = significant), shape = 21, size = 1) +
    geom_hline(aes(yintercept = -log10(0.05)), lty = 2, linewidth = 0.5, color = "black") +
    geom_text_repel(aes(label = label), min.segment.length = 0, size = 2) +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         values = rescale(c(-0.17, 0, 0.17)),
                         limits = c(-0.171, 0.171),
                         na.value = "lightgray", guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), na.value = "transparent") +
    guides(color = guide_legend(title = "FDR < 0.05", title.vjust = -1.5)) +
    labs(x = "log2(fold change)", y = "-log10(p-value)",
         title = "A | DEG volcano plot") +
    theme(#legend.position = c(0.01, 0.91),
          #legend.spacing.x = unit(0.1, "mm"),
          #legend.background = element_blank(),
          #legend.box.background = element_rect(colour = "black")
        legend.position = "none"
    )
```

```{r Fig4a.save, fig.width = 2.17, fig.height = 1.74}
fig4a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.DEG_volcano_plot", .x), width = 2.17, height = 1.74)
)
```

### B | Top n genes by L2FC t-statistic

```{r Fig4b.plot}
fig4b <- df_de_res %>% 
    slice_max(order_by = abs(stat), n = 10) %>%

    ggplot(aes(x = stat, y = reorder(gene_symbol, stat))) +
    geom_point(aes(size = -log10(pvalue), fill = stat), shape = 21) +
    geom_vline(xintercept = 0, color = "gray") +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         limits = c(-5, 5), guide = "none") +
    scale_size_continuous(range = c(1, 4)) +
    guides(size = guide_legend(title = "-log10(p-value)")) +
    xlim(c(-5, 5)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = "t-statistic", 
         title ="B | Top 10 DEGs"
    ) +
    theme(legend.position = c(0.05, 0.7), 
          legend.box = "horizontal",
          legend.key.size = unit(0.35, "cm")
    )
```

```{r Fig4b.save, fig.width = 2.17, fig.height = 1.74}
fig4b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.top_10_DEGs", .x), width = 2.17, height = 1.74)
)
```

### C | GRCCA structure correlations (rx) vs DGE effect size

```{r Fig4c.plot}
fig4c <- df_x_res %>% 
    left_join(df_de_res) %>%

    ggplot(aes(x = stat, y = pearsons_r)) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(color = pearsons_r + stat), size = 0.5) +
    geom_smooth(method = "lm", color = "black", linewidth = 0.5) +
    stat_cor(aes(label = after_stat(r.label)), label.sep = "\n", size = 3,
             label.y.npc = "top", vjust = 0) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-5.4, 5.4)) +
    coord_cartesian(clip = "off") +
    #scale_x_continuous(breaks = c(-0.2, 0, 0.2)) +
    labs(x = "DGE t-statistic", y = expression("Structure correlation (" * italic(r)[x] * ")"),
         title = "C | DGE vs GRCCA results"
    ) +
    theme(legend.position = "none")
```

```{r Fig4c.save, fig.width = 2.17, fig.height = 1.74}
fig4c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.gene_strucCor_vs_DE_stat", .x), width = 2.17, height = 1.74)
)
```

### D | GO enrichments of DEGs

```{r Fig4d.prep}
## Determine labels
fig4d_labels <- c(

    ## BP
    "adaptive immune response",
    "inflammatory response",
    #"positive regulation of T cell activation",
    "positive regulation of tumor necrosis factor production",
    "positive regulation of interleukin-6 production",
    "cellular response to copper ion",
    "nitric oxide transport",
    "carbon dioxide transport",

    ## MF
    "MHC class II protein complex binding",
    "extracellular matrix structural constituent",

    ## CC
    "clathrin-coated endocytic vesicle membrane",
    "TCR signalosome"
)

## Filter for terms of interest & assign labels
df_fig4d <- df_de_fgsea_go %>% 
    dplyr::filter(padj < 0.05) %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(term %in% fig4d_labels, str_wrap(term, 25), ""))
```

```{r Fig4d.plot}
fig4d <- df_fig4d %>%
    ggplot(aes(x = padj, y = NES)) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 2.5, force = 20,
                    box.padding = 0.5, min.segment.length = 0, max.overlaps = 50) +
    facet_wrap(vars(ontology), nrow = 1) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-2.5, 2.5), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.03, 0.05)) +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    coord_cartesian(clip = "off") +
    labs(x = "Enrichment FDR", y = "NES",
         title = "D | Functional pathway enrichments") +
    theme(legend.position = "bottom",
          legend.justification = "center",
          title.justification = "center",
          legend.key.size = unit(0.3, "cm")
    )
```

```{r Fig4d.save, fig.height = 3.48, fig.width = 4.34}
fig4d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "D.DEG_GO_enrichment", .x), width = 4.34, height = 3.48)
)
```

### E | Compare DGE and GRCCA functional enrichments

```{r Fig4e.prep}
df_fig4e <- df_grcca_fgsea_go %>% 
    mutate(analysis = "GRCCA", .before = 1) %>% 
    bind_rows(df_de_fgsea_go %>% 
                  mutate(analysis = "DGE", .before = 1)
    ) %>% 
    mutate(class = ifelse(NES < 0, "NES < 0", "NES > 0")) %>% 
    filter(padj < 0.05) %>% 
    count(analysis, class) %>% 
    mutate(text_color = ifelse(analysis == "DGE", "white", "black"),
           analysis_color = ifelse(analysis == "GRCCA", "white", "black")
    )
```

```{r Fig4e.plot}
fig4e <- df_fig4e %>% 
    ggplot(aes(x = analysis, y = n)) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = class), alpha = 0.35) +
    geom_col(aes(fill = I(analysis_color)), color = "black") +
    geom_text(aes(label = n, color = I(text_color)), size = 3, vjust = 1.2) +
    facet_wrap(vars(class)) +
    scale_fill_manual(values = c("NES < 0" = gene_weight_color_scale[15], "NES > 0" = gene_weight_color_scale[85])) +
    labs(x = NULL, y = "n pathways (FDR < 0.05)",
         title = "E | DGE vs GRCCA GO results"
    ) +
    theme(
        #panel.background = element_rect(fill = "transparent"),
        legend.position = "none",
        axis.text = element_text(size = 8)
    )
```

```{r Fig4e.save, fig.width = 2.17, fig.height = 1.74}
fig4e
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "E.GRCCA_vs_DGE_GO", .x), width = 2.17, height = 1.74)
)
```

### F | DEG enrichment for GWAS results

```{r Fig4f.prep}
df_fig4f <- df_de_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
```

```{r Fig4f.plot}
fig4f <- df_fig4f %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 2.75) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "F | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          strip.text = element_text(size = 9),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.60, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
```

```{r Fig4f.save, fig.width = 2.17, fig.height = 1.74}
fig4f
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "F.Risk_gene_enrichment", .x), width = 2.17, height = 1.74)
)
```

### 2025-01-10 scripts/07.figures_markdown/Fig5_GRCCAtranscriptsRes.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure 5: GRCCA transcript-level results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig5_GRCCAtranscriptsRes/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load GRCCA results (in Fig2_GRCCAres)
load(paste0(analysis_objects_dir, "../Fig2_GRCCAres/GRCCA_results.Rdata")) # df_x_res
df_x_res_genes <- df_x_res
rm(list = "df_x_res")
```

### A | Covariate structure correlations (ry)

```{r Fig5a.plot}
fig5a <- df_y_res_transcripts %>% 
    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(covariate, pearsons_r))) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    #geom_text(aes(label = significant), size = 6, vjust = 0.75, hjust = -1.5, color = "black") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    xlim(c(-1, 1)) +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[y] * ")"), 
         title = "A | Covariate structure correlations") +
    theme(legend.position = c(0.04, 0.55), 
          legend.box = "horizontal",
          #legend.justification = "center",
          #legend.key.size = unit(0.2, 'cm')
          legend.key.size = unit(0.1, "cm")
    )
```

```{r Fig5a.save, fig.width = 2.5}
fig5a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.covariate_structCor", .x), width = fig_width, height = fig_height)
)
```

### B | Correlation with gene-level GRCCA results

```{r Fig5b.prep}
## Identify all SCZ risk genes across lists
scz_risk_genes <- benchmarking_lists[names(benchmarking_lists)[str_detect(names(benchmarking_lists), "SCZ")]] %>% 
    unlist %>% 
    unique

## Write function to calculate distance from point to line
f_distance_to_line <- function(x, y, m, b) {
    abs(m * x - y + b) / sqrt(m^2 + 1)
}

## Combine gene & transcript-level gene weights, highlight SCZ risk genes for plot
df_fig5b <- df_x_res_transcripts %>% 
    #filter(!is.na(ensembl_gene_id)) %>% 
    group_by(ensembl_gene_id) %>% 
    summarise(pearsons_r = max(pearsons_r)) %>% 
    #summarise(pearsons_r = ifelse(pearsons_r < 0, min(pearsons_r), max(pearsons_r))) %>% 
    dplyr::rename("transcript_res" = "pearsons_r") %>%

    left_join(df_x_res_genes %>% 
                  dplyr::select(ensembl_gene_id, gene_symbol, pearsons_r) %>% 
                  dplyr::rename("gene_res" = "pearsons_r")
    ) %>%

    # calculate distance to y = x
    mutate(distance = f_distance_to_line(gene_res, transcript_res, 1, 0)) %>%

    # highlight SCZ risk genes
    mutate(
        risk_gene = ifelse(ensembl_gene_id %in% scz_risk_genes, "yes", "no"),
        label = ifelse(risk_gene == "yes" & distance < 0.1 & abs(gene_res) > 0.1, gene_symbol, "")
    ) %>% 
    arrange(risk_gene)
```

```{r Fig5b.plot}
fig5b <- df_fig5b %>% 
    #filter(risk_gene == "yes") %>%

    ggplot(aes(x = gene_res, y = transcript_res)) +
    geom_abline(lty = 2) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_hline(yintercept = 0, color = "gray") +
    geom_point(aes(fill = gene_res + transcript_res, color = risk_gene, size = risk_gene), shape = 21) +
    geom_text_repel(aes(label = label), size = 2.5, max.overlaps = 100, min.segment.length = 0) +
    stat_cor(size = 2.5) +
    geom_smooth(method = "lm", color = "black", linewidth = 0.5) +
    guides(color = guide_legend(title = "SCZ risk gene?")) +
    xlim(c(-0.6, 0.6)) +
    ylim(c(-0.6, 0.6)) +
    scale_fill_gradientn(colors = gene_weight_color_scale, guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_size_manual(values = c("yes" = 0.75, "no" = 0.25), guide = "none") +
    labs(x = expression("Gene " * italic(r)[x]), y = expression("Transcript " * italic(r)[x]),
         title = "B | Gene vs transcript GRCCA results") +
    theme(legend.position = c(0.55, 0.15),
          legend.key.size = unit(0.1, "cm"),
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 6)
    )
```

```{r Fig5b.save, fig.width = 2.5}
fig5b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.gene_vs_transcript_GRCCA_res", .x), width = fig_width, height = fig_height)
)
```

### C | Directionality of transcripts associated with significant and non-significant risk genes

```{r Fig5c.prep}
## Select illustrative genes to plot
fig5c_sig_genes <- c(        
    "FMNL1", # sig, one transcript
    "C3", # sig, subset of transcripts,
    "EZH1" # sig, many transcripts
)
fig5c_non_sig_genes <- c(
    "PLCL2", # non sig, 0 transcripts
    "TMEM127", # non-sig, opposite directions
    "CDK14" # non-sig, opposite directions
)

## Set order
fig5c_gene_order <- df_x_res_genes %>% 
    filter(gene_symbol %in% c(fig5c_sig_genes, fig5c_non_sig_genes)) %>% 
    arrange(-abs(pearsons_r)) %>% 
    pull(gene_symbol)

## Prep df for plotting
df_fig5c <- df_x_res_transcripts %>% 
    filter(gene_symbol %in% fig5c_gene_order) %>% 
    mutate(
        gene_symbol = factor(gene_symbol, levels = fig5c_gene_order),
        significant = ifelse(p_adj < 0.05, "yes", "no"),
        gene_sig = ifelse(gene_symbol %in% fig5c_sig_genes, "significant", "non-significant")
    )
```

```{r Fig5c.function}
f_plot_fig5c <- function(df, gene_level_sig, max_color_scale = 1) {
    df %>% 
        filter(gene_sig == gene_level_sig) %>%

        ggplot(aes(x = pearsons_r, y = reorder(transcript_symbol, pearsons_r))) +
        geom_vline(xintercept = 0, color = "gray") +
        geom_point(aes(fill = pearsons_r, size = -log10(p_adj), color = significant), shape = 21) +
        facet_wrap(vars(gene_symbol), scales = "free_y") +
        scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-max_color_scale, max_color_scale), guide = "none") +
        scale_color_manual(values = c("yes" = "black", "no" = "gray"), guide = "none") +
        scale_size_continuous(range = c(1, 4)) +
        guides(size = guide_legend(title = "-log10(FDR)")) +
        coord_cartesian(clip = "off") +
        labs(x = expression("Transcript " * italic(r)[x]), y = NULL)
}
```

```{r Fig5c.plot}
fig5c_sig <- f_plot_fig5c(df_fig5c, gene_level_sig = "significant", max_color_scale = 0.457) +
    labs(title = "Significant")
fig5c_non_sig <- f_plot_fig5c(df_fig5c, gene_level_sig = "non-significant", max_color_scale = 0.146) +
    labs(title = "Non-significant")
```

```{r Fig5b.save, fig.width = 6, fig.height = 4.25}
fig5c_sig / fig5c_non_sig
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.directionality_of_transcripts", .x), width = 6, height = 4.25)
)
```

############ OTHER GENES TO PLOT? ##############

```{r}
genes_to_plot <- c("AKAP9", "NRXN1", "DCTN2", "EHBP1", "KIF1B", "SMARCE1", "TCF4", "AKT3", "GRIN1", "FMNL1", "RBFOX1")
genes_to_plot <- c("NRXN1", "GRIN1", "FMNL1", "RBFOX1", "C3", "PKN2-AS1")
genes_to_plot <- c("NSF", "KANSL1", "CRHR1", "ARHGAP27", "LRRC37A", "CCDC43", "FMNL1", 
                   "SPPL2C", "MAPT", "PLEKHM1", "STH", "LINC002210-CRHR1") # Eva's paper
genes_to_plot <- c("ZNF804A") # Owen et al 2023
#genes_to_plot <- benchmarking_lists[["SCZ (rare)"]]
p <- df_x_res_transcripts %>% 
    #filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]]) %>% 
    filter(gene_symbol %in% genes_to_plot) %>%
    #filter(ensembl_gene_id %in% intersect(sig_risk_genes, more_than_one_transcript)) %>%
    mutate(
        #gene_symbol = factor(gene_symbol, levels = sig_risk_gene_order),
        significant = ifelse(p_adj < 0.05, "yes", "no")
    ) %>%

    ggplot(aes(x = pearsons_r, y = reorder(ensembl_transcript_id, pearsons_r))) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(fill = pearsons_r, size = -log10(p_adj), color = significant), shape = 21) +
    facet_wrap(vars(gene_symbol), scales = "free_y") +
    scale_fill_gradientn(colors = gene_weight_color_scale) +
    scale_color_manual(values = c("yes" = "black", "no" = "gray")) +
    coord_cartesian(clip = "off")
df_x_res_genes %>% filter(gene_symbol %in% genes_to_plot)
df_de_res %>% filter(gene_symbol %in% genes_to_plot)
```

### 2025-01-10 scripts/07.figures_markdown/archive/FigS1_preproc.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure S1: Preprocessing steps & justification"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/preproc/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/preproc/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### A | Variance explained by qSVs

```{r FigS1a.plot}
figS1a <- df_qsv_var_expl %>% #dplyr::count(feature)
    dplyr::filter(feature == "Proportion of Variance" & qSV %in% 1:16) %>%

    ggplot(aes(x = qSV, y = value*100)) +
    geom_col(aes(fill = value), color = "black") +
    geom_hline(yintercept = 0.02*100, color = "maroon") +
    scale_fill_gradient(low = "white", high = "black", 
                        limits = c(0, 0.06), guide = "none") +
    #scale_y_continuous(breaks = seq(0, 0.06, 0.01)) +
    coord_cartesian(clip = "off") +
    labs(y = "% variance explained",
         title = "A | qSV variance explained") +
    theme(axis.text.x = element_text(angle = 45))
```

```{r FigS1a.save}
figS1a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.qSV_variance_explained", .x), width = fig_width, height = fig_height)
)
```

### B | Correlate qSVs with known covariates

```{r FigS1b.prep}
m_covariates <- df_covariates_numeric %>% column_to_rownames("sample") %>% t()
covariate_order <- m_covariates[hclust(dist(m_covariates))$order,] %>% rownames
```

```{r FigS1b.plot}
figS1b <- df_covariate_qsv_cor %>% 
    mutate(significant = ifelse(p_adj < 0.05, 1, 0) %>% factor,
           covariate = factor(covariate, levels = covariate_order)) %>% 
    ggplot(aes(x = qsv, y = covariate)) +
    geom_tile(aes(fill = pearsons_r, color = significant), width = 0.98, height = 0.98, linewidth = 0.5) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100))) +
    scale_color_manual(values = c("1" = "black", "0" = "transparent"), guide = "none") +
    coord_cartesian(clip = "off") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.hjust = 0.5, title.position = "left")) +
    labs(x = "qSV", y = NULL, 
         title = "B | Correlate qSVs and known covariates"
    ) +
    theme(legend.key.height = unit(1.5, "cm"),
          legend.key.width = unit(0.1, "cm"),
          legend.title = element_text(angle = 90),
          legend.margin = margin(l = -15),
          axis.text.x = element_text(angle = 45)
    )
```

```{r FigS1b.save, fig.height = 5, fig.width = 5.1}
figS1b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.qSV_covariate_cor", .x), width = 5.1, height = 5)
)
```

### C | PC-covariate correlations before and after covariate regressions

```{r FigS1c.prep}
df_figS1c <- df_pca1_covar_cor %>% 
    mutate(regression = "Before", .before = 1) %>% 
    bind_rows(
        df_pca3_covar_cor %>% 
            mutate(regression = "After", .before = 1) 
    ) %>% 
    mutate(regression = factor(regression, levels = c("Before", "After")))
```

```{r FigS1c.plot}
figS1c <- df_figS1c %>% 
    ggplot(aes(x = pc, y = covariate)) +
    geom_tile(aes(fill = pearsons_r, color = significant), width = 0.98, height = 0.98, linewidth = 0.5) +
    facet_wrap(vars(regression), nrow = 1, scales = "free_x") +
    force_panelsizes(cols = c(1, 4)) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-0.43, 0.43)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    coord_cartesian(clip = "off") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.hjust = 0.5, title.position = "left")) +
    labs(x = "PC (> 2% explained var)", y = NULL,
         title = "D | PC-covariate correlations \nbefore vs after regression") +
    theme(legend.key.height = unit(1.5, "cm"),
          legend.key.width = unit(0.1, "cm"),
          legend.title = element_text(angle = 90),
          legend.margin = margin(l = -15)
    )
```

```{r FigS1c.save, fig.height = 6, fig.width = 5}
figS1c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.PCA_covariate_cor", .x), width = 5, height = 6)
)
```

### 2025-01-10 scripts/07.figures_markdown/archive/FigS3_4_WGCNAenrichment.Rmd
added grcca_analysis.txt from cluster

---
title: "Figure S3 & S4: WGCNA module characterization"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/FigS3_4_WGCNAenrichment/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/FigS3_4_WGCNAenrichment/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### A | Module sizes

```{r FigS3a.plot}
figS3a <- df_modules_filt %>% 
    dplyr::count(module, color) %>% 
    ggplot(aes(x = module, y = n)) +
    geom_col(aes(fill = module), color = "black", linewidth = 0.25) +
    geom_text(aes(label = n), size = 3, angle = 45, hjust = 0, vjust = 0) +
    scale_fill_manual(values = module_colors) +
    coord_cartesian(clip = "off") +
    labs(title = "A | Module sizes",
         x = NULL, y = "Number of genes") +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 0.9, size = 8)
    )
```

```{r FigS3a.save, fig.width = 3.5, fig.height = 2.5}
figS3a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.Module_sizes", .x), width = 3.5, height = 2.5)
)
```

### B | Semantic themes across module GO results

```{r FigS3b.prep}
## Extract topic descriptions (frequent words)
# FREX weights words by frequency and exclusivity to the topic
df_word_beta <- tidy(topic_model, matrix = "beta") %>%
    group_by(topic) %>%
    arrange(topic, -beta) %>% 
    top_n(n = 10, wt = beta) %>%
    mutate(topic = paste0("topic ", topic)) %>% 
    dplyr::filter(!str_detect(term, "cell"))

## Extract topic probabilities for each module (document)
df_group_gamma <- tidy(
    topic_model, 
    matrix = "gamma",
    document_names = rownames(go_sparse)
) %>% 
    mutate(module = factor(document)) %>% 
    dplyr::select(module, topic, gamma) %>% 
    mutate(topic = factor(topic, levels = seq(n_topics, 1, by = -1)),
           module = paste0("geneM", module) %>% factor(levels = paste0("geneM", levels(module)))
    ) %>% 
    dplyr::filter(gamma > 0.01)
```

```{r FigS3b.plot}
## Plot word strength of association with each topic
p_figS3b_beta <- df_word_beta %>% 
    ggplot(aes(x = reorder_within(term, beta, topic), y = beta, fill = beta)) +
    geom_col(color = "#222222") +
    geom_text(aes(label = term, y = 0.001, size = beta), color = "black", angle = 90, hjust = 0) +
    facet_wrap(vars(topic), ncol = 1, scales = "free_x") +
    scale_x_reordered() +
    scale_y_continuous(limits = c(0, 0.06), breaks = c(0, 0.05)) +
    scale_size_continuous(range = c(2, 3), guide = "none") +
    scale_fill_gradient(low = "white", high = "blue", guide = "none", limits = c(0, 0.058)) +
    labs(x = NULL, y = "Word strength of association with each topic") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()
    )

## Plot module strength of association with each topic
p_figS3b_gamma <- df_group_gamma %>%
    ggplot(aes(x = gamma, y = topic, color = module)) +
    geom_jitter(aes(alpha = gamma), position = position_jitter(seed = 2, width = 0.15)) +
    geom_text_repel(aes(alpha = gamma, label = module), size = 3, box.padding = 0.2, min.segment.length = 0,
                    max.overlaps = 15, position = position_jitter(seed = 2, width = 0.15)) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), color = "black", linewidth = 0.25) +
    scale_color_manual(values = module_colors) +
    labs(x = "Module strength of association \nwith each topic", y = NULL) +
    theme(legend.position = "none",
          axis.text.y = element_blank()
    )

```

```{r FigS3b.save, fig.width = 4, fig.height = 6.5}
p_figS3b_beta + p_figS3b_gamma + plot_layout(widths = c(1, 2)) + plot_annotation(title = "B | Semantic themes across module GO results")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.topic_modeling", .x), width = 4, height = 6.5)
)
```

### C | Cell-type enrichment of GO modules

```{r FigS3c.prep}
df_figS3c <- df_mods_cell_type_hypergeometric %>%
    mutate(
        p_adj = ifelse(p_adj == 0, 10^-15, p_adj),
        color = ifelse(p_value < 0.05, paste0(module), NA) %>% factor,
        label = case_when(
            p_adj < 0.05 & p_adj > 0.01 ~ "*",
            p_adj < 0.01 & p_adj > 0.001 ~ "**",
            p_adj < 0.001 ~ "***",
            TRUE ~ ""
        )
    )
```

```{r FigS3c.plot}
figS3c <- df_figS3c %>% 
    ggplot(aes(x = cell_type, y = module)) +
    geom_tile(aes(fill = color, alpha = -log10(p_adj), color = color), 
              width = 0.95, height = 0.95, linewidth = 0.5) +
    geom_text(aes(label = label), vjust = 0.75, size = 3.5) +
    #annotate(geom = "text", label = "* = FDR < 0.05", x = 6.5, y = 1, size = 3) +
    scale_alpha_continuous(range = c(0.1, 1), na.value = 1) +
    scale_fill_manual(values = module_colors, na.value = "transparent", guide = "none") +
    scale_color_manual(values = module_colors, na.value = "transparent", guide = "none") +
    guides(alpha = guide_legend(title = "-log10(FDR)", title.position = "top", title.hjust = 0.5)) +
    labs(title =  "C | Cell-type enrichment", 
         x = NULL, y = NULL
    ) +
    theme(legend.position = "bottom",
          legend.margin = margin(t = -12),
          legend.key.size = unit(0.3, "cm"),
          legend.justification = "center",
          axis.text.x = element_text(angle = 45, hjust = 0.9)
    )
```

```{r FigS3c.save, fig.height = 4, fig.width = 3.5}
figS3c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.cell_type_enrichment", .x), width = 3.5, height = 4)
)
```

### D | Developmental trajectory of each module per PsychENCODE data

```{r FigS3d.plot}
figS3d <- df_expr_mod_median %>% 
    ggplot(aes(x = window, y = median_expr, color = module, fill = module)) +
    geom_smooth(method = "loess") +
    geom_vline(xintercept = 5, linewidth = 0.2) +
    facet_wrap( ~ module, scales = "free_y", nrow = 3) +
    scale_fill_manual(values = module_colors) +
    scale_color_manual(values = module_colors) +
    labs(x = "Developmental window", y = "Median expression (PsychENCODE)", 
         title = "D | Developmental expression trajectory of each module") +
    theme(legend.position = "none",
          strip.text.x = element_text(size = 9),
          strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank()
    )
```

```{r FigS3d.save, fig.width = 7.5, fig.height = 3}
figS3d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "D.developmental_trajectories", .x), width = 7.5, height = 3)
)
```

### Fig S4: full module functional enrichments

```{r FigS4.prep}
df_figS4 <- df_mods_go %>% 
    clean_names %>% 
    dplyr::filter(!is.na(description) & module != 0) %>%

    # take top 5 paths by p-value per module to plot
    dplyr::group_by(module) %>% 
    arrange(module, -p_adjust) %>% 
    mutate(row = row_number(),
           gene_ratio = parse(text = gene_ratio) %>% eval
    ) %>% 
    top_n(n = 4, wt = row) %>% 
    #slice_max(order_by = -weight_fisher, n = 5) %>% 
    mutate(`survives FDR` = ifelse(p_adjust < 0.05, "yes", "no")) %>% 
    mutate(description = str_wrap(description, width = 18))
```

```{r FigS4.plot}
figS4 <- df_figS4 %>% 
    ggplot(aes(x = p_adjust, y = reorder_within(description, within = module, by = -p_adjust))) +
    #y = reorder_within(str_wrap(description, width = 35), -pvalue, module))) +
    geom_point(aes(size = gene_ratio, fill = ontology, color = `survives FDR`),
               shape = 21, stroke = 1) +
    geom_vline(aes(xintercept = 0.05), linewidth = 1, color = "gray") +
    facet_wrap(vars(module), scales = "free", ncol = 5) +
    scale_size_continuous(range = c(1, 4)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_y_reordered() +
    scale_x_reverse(breaks = c(0.25, 0.15, 0.05, 0)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = "Enrichment FDR",
         title = "S4 | WGCNA module GO enrichments"
    ) +
    guides(size = guide_legend(title = "Gene ratio")) +
    theme(legend.position = c(0.85, 0.03),
          legend.direction = "vertical",
          legend.title.position = "top",
          #legend.justification = "center",
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(size = 8),
          axis.text = element_text(size = 8)
    )
```

```{r FigS4.save, fig.height = 10, fig.width = 8}
figS4
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "S4.full_module_enrichments", .x), width = 8, height = 10)
)
```

### 2025-01-10 scripts/07.figures_markdown/poster.Rmd
added grcca_analysis.txt from cluster

---
title: "GRCCA paper figures for poster"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
poster_figures_dir <- paste0(base_dir, "outputs/figures/poster/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load data objects for plots
directories <- list.files(analysis_objects_dir)
for (dir in directories) {
    print(dir)
    objects <- list.files(paste0(analysis_objects_dir, dir))
    #objects <- objects[!str_detect(objects, "null|archive")]
    for (obj in objects){
        print(paste0("loading ", obj, "...."))
        load(paste0(analysis_objects_dir, dir, "/", obj))
    }
}
```

# Fig 1

## A | Covariate (Y) structure correlations

```{r Fig1a.plot}
fig1a <- df_y_res %>% 
    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(covariate, pearsons_r))) +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_text(aes(label = significant), size = 6, vjust = 0.75, hjust = -1.5, color = "black") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    xlim(c(-1, 1)) +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[y] * ")"), 
         title = "A | Covariate structure correlations") +
    theme(legend.position = c(0.075, 0.6), 
          legend.box = "horizontal",
          #legend.justification = "center",
          legend.key.size = unit(0.1, 'cm')
          #legend.key.height = unit(0.29, "cm"),
          #legend.key.width = unit(0.30, "cm")
    )
```

```{r Fig1a.save, fig.width = 2.5, fig.height = 2}
fig1a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "1A.covariate_structCor", .x), width = fig_width, height = fig_height)
)
```

## B | Top 10 genes (X) by structure correlation

```{r Fig1b.prep}
df_fig1b <- df_x_res %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) >= 2) %>%
    top_n(n = 10, wt = abs(pearsons_r)) %>%
    mutate(direction = ifelse(pearsons_r < 0, "2neg", "1pos"))
```

```{r Fig1b.plot}
fig1b <- df_fig1b %>% 
    ggplot(aes(x = pearsons_r, y = reorder(gene_symbol, pearsons_r))) +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    facet_wrap(vars(direction), ncol = 1, scales = "free_y") +
    force_panelsizes(rows = c(16, 4)) +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         limits = c(-0.57, 0.57), guide = "none") +
    scale_size_continuous(range = c(1, 4)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    xlim(c(-1, 1)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[x] * ")"), 
         title = "B | Top 10 genes by \nstructure correlation"
    ) +
    theme(strip.text = element_blank(), 
          panel.spacing = unit(0.5, "lines"),
          legend.position = c(0.05, 0.7), 
          legend.box = "horizontal",
          legend.key.size = unit(0.35, "cm")
    )
```

```{r Fig1b.save, fig.height = 2, fig.width = 2.5}
fig1b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "1B.gene_structCor", .x), width = fig_width, height = fig_height)
)
```

## C | Expression directions of top positively- and negatively-weighted genes

```{r Fig1c.prep}
top_gene <- df_x_res %>% arrange(-abs(pearsons_r)) %>% head(1) %>% pull(ensembl_gene_id) # BCL7A
df_fig1c <- df_vsd_regress %>% 
    dplyr::select(sample, all_of(c(top_gene))) %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expr") %>% 
    left_join(df_ensembl_to_symbol) %>% 
    mutate(dx = case_when(
        str_detect(sample, "control") ~ "Control",
        str_detect(sample, "bipolar") ~ "BD",
        str_detect(sample, "mdd") ~ "MDD",
        str_detect(sample, "schizo") ~ "SCZ"
    ) %>% 
        factor(levels = names(dx_colors)))
```

```{r Fig1c.plot}
fig1c <- df_fig1c %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.5) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "darkgrey") +
    facet_wrap(vars(gene_symbol), scales = "free_y", nrow = 1) +
    #force_panelsizes(cols = c(1, 1)) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    labs(x = NULL, y = "Corrected expression value",
         title = "C | Gene expression values") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

```{r Fig1c.save, fig.width = 2, fig.height = 2}
fig1c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "1C.top_gene_expression_values", .x), width = 2, height = fig_height)
)
```

## D | GRCCA structure correlations (r) as they relate to expression values

```{r Fig1d.prep}
df_fig1d <- df_vsd_regress %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expression_value") %>% 
    mutate(dx = case_when(
        str_detect(sample, "control") ~ "Control",
        str_detect(sample, "bipolar") ~ "BD",
        str_detect(sample, "mdd") ~ "MDD",
        str_detect(sample, "schizo") ~ "SCZ"
    ) %>% factor(levels = names(dx_colors))) %>% 
    group_by(ensembl_gene_id, dx) %>% 
    summarise(mean_expr = mean(expression_value)) %>% 
    left_join(df_x_res)
```

```{r Fig1d.plot}
fig1d <- df_fig1d %>% 
    ggplot(aes(x = mean_expr, y = pearsons_r)) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(color = "black", size = 0.25) +
    geom_smooth(aes(color = dx), method = "lm", linewidth = 0.75) +
    stat_cor(aes(label = after_stat(r.label), color = dx), label.sep = "\n", size = 3,
             label.y.npc = "top", hjust = -0.1, vjust = -1.75, face = "bold") +
    facet_wrap(vars(dx), nrow = 1) +
    scale_color_manual(values = dx_colors) +
    scale_x_continuous(breaks = c(-0.4, 0, 0.4)) +
    coord_cartesian(clip = "off") +
    labs(x = "Mean expression (corrected)", y = expression("Structure correlation (" * italic(r)[x] * ")"),
         title = "D | Relationship between gene structure correlation & expression"
    ) +
    theme(legend.position = "none")
```

```{r Fig1d.save, fig.width = 5, fig.height = 2}
fig1d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "1D.gene_strucCor_vs_Expr", .x), width = 5, height = fig_height)
)
```

# Fig 2

## A | GRCCA results are significantly and specifically enriched for SCZ risk genes

```{r Fig2a.prep}
df_fig2a <- df_grcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))

## Identify genes to label in bottom panel
top_genes <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    top_n(n = 6, wt = abs(pearsons_r)) %>% 
    pull(ensembl_gene_id)

## Subset X res for genes in SCZ common broad list for bottom panel
df_fig2a_bottom <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    mutate(significant = ifelse(p_adj < 0.05, 1, 0) %>% factor) %>% 
    arrange(significant, abs(pearsons_r)) %>% 
    mutate(label = ifelse(ensembl_gene_id %in% top_genes, gene_symbol, ""))
```

```{r Fig2a.plot.top}
fig2a <- df_fig2a %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color), alpha = abs(NES)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.58, 1.58)) +
    scale_alpha_continuous(guide = "none", limits = c(0.83, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "top")) +
    labs(x = NULL, y = NULL,
         title = "A | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          legend.margin = margin(r = -15),
          legend.position = "bottom",
          legend.justification = "center",
          #legend.title = element_text(angle = 90),
          legend.key.width = unit(0.60, "cm"),
          legend.key.height = unit(0.15, "cm")
    )
```

```{r Fig2a.top.save, fig.height = 3, fig.width = 2.25}
fig2a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "2A.Risk_gene_enrichment_grid", .x), width = 2.25, height = 3)
)
```

```{r Fig2a.plot.right}
scale_max <- df_x_res %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)
fig2a_right <- df_x_res %>%
    ggplot(aes(x = pearsons_r)) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density() +
    geom_point(data = df_fig2a_bottom, 
               mapping = aes(x = pearsons_r, y = -1, fill = pearsons_r, color = significant), 
               pos = jitter_pos, shape = 21, size = 0.75
    ) +
    geom_text_repel(data = df_fig2a_bottom,
                    mapping = aes(x = pearsons_r, y = -1, label = label),
                    pos = jitter_pos, 
                    size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
    ) +

    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-scale_max, scale_max), guide = "none") +
    scale_color_manual(values = c("transparent", "black"), guide = "none") +
    scale_x_continuous(limits = c(-0.6, 0.6), breaks = c(-0.5, 0, 0.5)) +
    labs(x = expression(italic(r)[x]), y = NULL, 
         title = NULL) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.border = element_rect(fill = "transparent", color = "black", size = 1),
          axis.line = element_blank()
    )
```

```{r Fig2a.right.save, fig.height = 1.5, fig.width = 2}
fig2a_right
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "2A.Risk_gene_enrichment_distribution", .x), width = 2, height = 1.5)
)
```

## B | GRCCA cell-type enrichments

```{r Fig2b.plot}
fig2b <- df_grcca_fgsea_cell %>% 
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>%

    ggplot(aes(x = NES, y = reorder(cell_type, NES))) +
    geom_col(aes(fill = NES), color = "gray", shape = 21) +
    geom_text(aes(label = sig, x = NES), vjust = 0.75, size = 4, check_overlap = TRUE) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3.39, 3.39)) +
    coord_cartesian(clip = "off") +
    labs(x = "NES", y = NULL,
         title = "B | Cell-type enrichment") +
    theme(legend.position = "none")
```

```{r Fig2b.save, fig.height = 1.5, fig.width = 2}
fig2b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "2B.cell_type_enrichment", .x), width = 2, height = 1.5)
)
```

## C | Developmental expression trajectories of GRCCA deciles

```{r Fig2c.prep}
## Calculate the mean expression for each decile of GRCCA results across development
df_fig2c <- df_psychencode_grcca %>% #dplyr::filter(regioncode == "MFC") %>% 
    group_by(decile, window) %>% 
    summarise(median_expression = median(dev_expression),
              sd_expression = sd(dev_expression)
    )

## Rename windows with time period
psychencode_windows <- c("5-9",
                "12-13",
                "16-18",
                "19-22",
                "PCW35-PY0.3",
                "0.5-2.5",
                "2.8-10.7",
                "13-19",
                "21-64")
names(psychencode_windows) = 1:9

## Define colors for deciles
decile_colors <- colorRampPalette(c(gene_weight_color_scale[5], "gray", gene_weight_color_scale[95]))(10) %>% rev
```

```{r Fig2c.plot}
fig2c <- df_fig2c %>% 
    ggplot(aes(x = window, y = median_expression, color = decile)) +
    geom_vline(xintercept = 5, color = "gray") +
    geom_smooth(method = "loess", se = FALSE) +
    scale_x_continuous(breaks = seq(1, length(psychencode_windows), 1),
                       labels = psychencode_windows) +
    scale_color_manual(values = decile_colors) +
    guides(color = guide_legend()) +
    labs(x = "← PCW | PY →", y = "Median expression",
         title = "C | Developmental expression") +
    theme(
        legend.key.size = unit(0.3, "cm"),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

```{r fig2c.save, fig.width = 2.5, fig.height = 3}
fig2c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "2C.PsychENCODE_development", .x), width = fig_width, height = 3)
)
```

## D | GRCCA GO enrichments

```{r Fig2d.prep}
## Determine labels
fig2d_labels <- c(
    "cilium movement",
    "motile cilium",
    "axoneme",
    #"axonemal microtubule",
    "axoneme assembly",
    "adaptive immune response",
    "outer dynein arm assembly",
    "MHC class II protein complex",
    "T cell costimulation",
    "MHC class II protein complex binding",
    "minus-end-directed microtubule motor activity",
    "dynein light intermediate chain binding",

    "syntaxin-1 binding",
    "ubiquitin-dependent protein catabolic process",
    "endosome to lysosome transport",
    "cytoplasmic translation",
    "regulation of autophagy",
    "postsynaptic density",
    "presynaptic endocytic zone membrane",
    "ATP binding",
    "ATP hydrolysis activity",
    "protein ubiquitination"
)

## Filter for terms of interest & assign labels
df_fig2d <- df_grcca_fgsea_go %>% 
    dplyr::filter(padj < 0.05) %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(term %in% fig2d_labels, str_wrap(term, 25), ""))
```

```{r Fig2d.plot}
fig2d <- df_fig2d %>%
    ggplot(aes(x = padj, y = NES)) +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 2.75, force = 20,
                    box.padding = 0.25, min.segment.length = 0, max.overlaps = 30) +
    geom_hline(yintercept = 0, color = "gray") +
    facet_wrap(vars(ontology)) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3, 3), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.03, 0.05)) +
    coord_cartesian(clip = "off") +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    labs(x = "Enrichment FDR", y = "NES",
         title = "D | Functional pathway enrichments") +
    theme(#legend.position = "bottom",
          #legend.justification = "center",
          title.justification = "center",
          legend.key.size = unit(0.3, "cm")
    )
```

```{r Fig2d.save, fig.height = 3, fig.width = 6}
fig2d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "2D.GO_enrichment", .x), width = 6, height = 3.0)
)
```

# Fig 3

## A | Top n genes by L2FC t-statistic

```{r Fig3a.plot}
fig3a <- df_de_res %>% 
    slice_max(order_by = abs(stat), n = 10) %>%

    ggplot(aes(x = stat, y = reorder(gene_symbol, stat))) +
    geom_point(aes(size = -log10(pvalue), fill = stat), shape = 21) +
    geom_vline(xintercept = 0, color = "gray") +
    scale_fill_gradientn(colors = gene_weight_color_scale, 
                         limits = c(-5, 5), guide = "none") +
    scale_size_continuous(range = c(1, 4)) +
    guides(size = guide_legend(title = "-log10(p-value)")) +
    xlim(c(-5, 5)) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = "t-statistic", 
         title ="A | Top 10 DEGs"
    ) +
    theme(legend.position = c(0.05, 0.7), 
          legend.box = "horizontal",
          legend.key.size = unit(0.35, "cm")
    )
```

```{r Fig3a.save, fig.width = 2.5, fig.height = 2}
fig3a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "3A.top_10_DEGs", .x), width = fig_width, height = fig_height)
)
```

## B | GRCCA structure correlations (rx) vs DGE effect size

```{r Fig3b.plot}
fig3b <- df_x_res %>% 
    left_join(df_de_res) %>%

    ggplot(aes(x = stat, y = pearsons_r)) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(color = pearsons_r + stat), size = 0.5) +
    geom_smooth(method = "lm", color = "black", linewidth = 0.5) +
    stat_cor(aes(label = after_stat(r.label)), label.sep = "\n", size = 3,
             label.y.npc = "top", vjust = 0) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-5.4, 5.4)) +
    coord_cartesian(clip = "off") +
    #scale_x_continuous(breaks = c(-0.2, 0, 0.2)) +
    labs(x = "DGE t-statistic", y = expression("Structure correlation (" * italic(r)[x] * ")"),
         title = "B | DGE vs GRCCA results"
    ) +
    theme(legend.position = "none")
```

```{r Fig3b.save, fig.width = 2.5, fig.height = 2}
fig3b
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "3B.gene_strucCor_vs_DE_stat", .x), width = fig_width, height = fig_height)
)
```

## C | Compare DGE and GRCCA functional enrichments

```{r Fig3c.prep}
df_fig3c <- df_grcca_fgsea_go %>% 
    mutate(analysis = "GRCCA", .before = 1) %>% 
    bind_rows(df_de_fgsea_go %>% 
                  mutate(analysis = "DGE", .before = 1)
    ) %>% 
    mutate(class = ifelse(NES < 0, "NES < 0", "NES > 0")) %>% 
    filter(padj < 0.05) %>% 
    count(analysis, class) %>% 
    mutate(text_color = ifelse(analysis == "DGE", "white", "black"),
           analysis_color = ifelse(analysis == "GRCCA", "white", "black")
    )
```

```{r Fig3c.plot}
fig3c <- df_fig3c %>% 
    ggplot(aes(x = analysis, y = n)) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = class), alpha = 0.35) +
    geom_col(aes(fill = I(analysis_color)), color = "black") +
    geom_text(aes(label = n, color = I(text_color)), size = 3, vjust = 1.2) +
    facet_wrap(vars(class)) +
    scale_fill_manual(values = c("NES < 0" = gene_weight_color_scale[15], "NES > 0" = gene_weight_color_scale[85])) +
    labs(x = NULL, y = "n pathways (FDR < 0.05)",
         title = "C | DGE vs GRCCA GO results"
    ) +
    theme(
        #panel.background = element_rect(fill = "transparent"),
        legend.position = "none",
        axis.text = element_text(size = 8)
    )
```

```{r Fig3c.save, fig.width = 2.5, fig.height = 2}
fig3c
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "3C.GRCCA_vs_DGE_GO", .x), width = 2.5, height = 2)
)
```

## D | DEG enrichment for GWAS results

```{r Fig3d.prep}
df_fig3d <- df_de_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
```

```{r Fig3d.plot}
fig3d <- df_fig3d %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color), alpha = abs(NES)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 2.75) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.58, 1.58)) +
    scale_alpha_continuous(guide = "none", limits = c(0.83, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "D | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          strip.text = element_text(size = 9),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.60, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
```

```{r Fig3d.save, fig.width = 2.5, fig.height = 2.0}
fig3d
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(poster_figures_dir, "3D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
)
```

### 2025-07-02 scripts/07.figures_markdown/Fig1_analysisPipeline.Rmd
Update scripts with response to reviewer comments

@@
 base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
-figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/Fig1_analysisPipeline/")
+figures_dir <- paste0(base_dir, "outputs/figures/Fig1_analysisPipeline/")
 
 source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
-
-## Load DE data
-load(paste0(base_dir, "outputs/updated_figures/objects/Fig2_DGEres/DE_results.RDS")) # df_de_res
 ```

### 2025-07-02 scripts/07.figures_markdown/Fig2_GRCCAres.Rmd
Update scripts with response to reviewer comments

--- 2025-01-10 scripts/08.figures_markdown/Fig2_GRCCAres.Rmd
@@
 tables_dir <- paste0(base_dir, "outputs/tables/")
-analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig2_GRCCAres/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
 
-source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
 
@@
 }
-
-## Load DE res (to compare to GRCCA)
-load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res
 ```

### 2025-07-02 scripts/07.figures_markdown/Fig3_characterizeGRCCA_v2.Rmd
Update scripts with response to reviewer comments

---
title: "Figure 3: GRCCA results benchmarking & function enrichments"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig3_characterizeGRCCA/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

source(paste0(base_dir, "/scripts/load_WGCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### A | GRCCA results are significantly and specifically enriched for SCZ risk genes

```{r Fig3a.plot.top, fig.height = 2, fig.width = 2.5}
## Prep
df_fig3a_top <- df_grcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3),
           label = ifelse(padj < 0.05, paste0(label, "*"), label)
    )
#mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))

## Plot
df_fig3a_top %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)[1:50]), limits = c(1.0, 1.58)) +
    #scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-1.58, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "A | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "Atop.Risk_gene_enrichment", .x), width = 2.5, height = 2)
)
```

```{r Fig3a.plot.bottom, fig.width = 1.3, fig.height = 1.3}
## Prep

# Identify genes to label in bottom panel
top_genes <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    top_n(n = 6, wt = abs(pearsons_r)) %>% 
    pull(ensembl_gene_id)

# Subset X res for genes in SCZ common broad list for bottom panel
df_fig3a_bottom <- df_x_res %>% 
    dplyr::filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]]) %>% 
    mutate(significant = ifelse(p_adj < 0.05, 1, 0) %>% factor) %>% 
    arrange(significant, abs(pearsons_r)) %>% 
    mutate(label = ifelse(ensembl_gene_id %in% top_genes, gene_symbol, ""))

## Plot
scale_max <- df_x_res %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)
df_x_res %>%
    ggplot(aes(x = pearsons_r)) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_density() +
    geom_point(data = df_fig3a_bottom, 
               mapping = aes(x = pearsons_r, y = -0.5, fill = pearsons_r, color = significant), 
               pos = jitter_pos, shape = 21, size = 1
    ) +
    geom_text_repel(data = df_fig3a_bottom,
                    mapping = aes(x = pearsons_r, y = -0.5, label = label),
                    pos = jitter_pos, 
                    size = 2.5, box.padding = 0.75, max.overlaps = 100, force = 10, min.segment.length = 0
    ) +

    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-scale_max, scale_max), guide = "none") +
    scale_color_manual(values = c("transparent", "black"), guide = "none") +
    labs(x = expression("Structure correlation (" * italic(r)[x] * ")"), y = NULL, 
         title = NULL) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(size = 7),
          axis.title.x = element_text(size = 8),
          panel.border = element_rect(fill = "transparent", color = "black", size = 1),
          axis.line = element_blank()
    )

## Save
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "Abottom.Risk_gene_enrichment", .x), width = 1.3, height = 1.3)
)
```

### B | GRCCA cell-type enrichments

```{r Fig3b.plot, fig.height = 2.25, fig.width = 2.5}
df_grcca_fgsea_cell %>% 
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>%

    ggplot(aes(x = NES, y = reorder(cell_type, NES))) +
    geom_col(aes(fill = NES), color = "gray", shape = 21) +
    geom_text(aes(label = sig, x = NES), vjust = 0.75, size = 4, check_overlap = TRUE) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3.39, 3.39)) +
    coord_cartesian(clip = "off") +
    labs(x = "NES", y = NULL,
         title = "B | Cell-type enrichment") +
    theme(legend.position = "none")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.cell_type_enrichment", .x), width = 2.5, height = 2.25)
)
```

### C | GRCCA GO enrichments

```{r Fig3c.prep}
## Determine labels
fig3c_labels <- c(
    "cilium movement",
    #"motile cilium",
    "axoneme",
    #"axonemal microtubule",
    #"axoneme assembly",
    "adaptive immune response",
    #"outer dynein arm assembly",
    "MHC class II protein complex",
    #"T cell costimulation",
    #"MHC class II protein complex binding",
    "minus-end-directed microtubule motor activity",
    "toll-like receptor signaling pathway",
    #"dynein light intermediate chain binding",

    "syntaxin-1 binding",
    #"ubiquitin-dependent protein catabolic process",
    #"endosome to lysosome transport",
    #"cytoplasmic translation",
    "regulation of autophagy",
    "postsynaptic density",
    "presynaptic endocytic zone membrane",
    "ATP binding",
    #"ATP hydrolysis activity",
    "protein ubiquitination"
)

## Filter for terms of interest & assign labels
df_fig3c <- df_grcca_fgsea_go %>% 
    dplyr::filter(padj < 0.05) %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(term %in% fig3c_labels, str_wrap(term, 25), ""))
```

```{r Fig3c.plot, fig.width = 6.5, fig.height = 4}
df_fig3c %>%
    ggplot(aes(x = padj, y = NES)) +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 2.75, force = 20,
                    box.padding = 0.25, min.segment.length = 0, max.overlaps = 50) +
    geom_hline(yintercept = 0, color = "gray") +
    facet_wrap(vars(ontology)) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3, 3), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.03, 0.05)) +
    coord_cartesian(clip = "off") +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    labs(x = "Enrichment FDR", y = "NES",
         title = "C | Functional pathway enrichments") +
    theme(legend.position = "bottom",
          legend.justification = "center",
          title.justification = "center",
          legend.key.size = unit(0.3, "cm")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.GO_enrichment", .x), width = 6.5, height = 4)
)
```

### 2025-07-02 scripts/07.figures_markdown/Fig5_GRCCAtranscriptsRes_v2.Rmd
Update scripts with response to reviewer comments

---
title: "Figure 5: GRCCA transcript-level results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

source(paste0(base_dir, "scripts/setup.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load GRCCA results (in Fig2_GRCCAres)
#load(paste0(analysis_objects_dir, "../Fig2_GRCCAres/GRCCA_results.Rdata")) # df_x_res
df_x_res_genes <- df_x_res
rm(list = "df_x_res")
```

### A | Covariate structure correlations (ry)

```{r Fig5a.plot}
df_y_res_transcripts %>% 
    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(covariate, pearsons_r))) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
    #geom_text(aes(label = significant), size = 6, vjust = 0.75, hjust = -1.5, color = "black") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    guides(size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    xlim(c(-1, 1)) +
    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[y] * ")"), 
         title = "A | Covariate structure correlations") +
    theme(legend.position = c(0.04, 0.55), 
          legend.box = "horizontal",
          #legend.justification = "center",
          #legend.key.size = unit(0.2, 'cm')
          legend.key.size = unit(0.025, "cm")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.covariate_structCor", .x), width = fig_width, height = fig_height)
)
```

### B | Correlation with gene-level GRCCA results

```{r Fig5.genes_to_label}
## Identify genes to label in panel B, and plot in panel C
fig5_sig_genes <- c("CSMD1", "PTK2B") # "SLC4A10"
fig5_non_sig_genes <- c("ARHGAP44", "CSDE1") # "SOX2-OT" "HMOX2"
```

```{r Fig5b.prep}
## Identify all SCZ risk genes across lists
scz_risk_genes <- benchmarking_lists[names(benchmarking_lists)[str_detect(names(benchmarking_lists), "SCZ")]] %>%
    unlist %>%
    unique

# ## Write function to calculate distance from point to line
# f_distance_to_line <- function(x, y, m, b) {
#     abs(m * x - y + b) / sqrt(m^2 + 1)
# }

## Combine gene & transcript-level gene weights, highlight SCZ risk genes for plot
df_fig5b <- df_x_res_transcripts %>% 
    #filter(!is.na(ensembl_gene_id)) %>% 
    group_by(ensembl_gene_id) %>% 
    summarise(pearsons_r = pearsons_r[which.max(abs(pearsons_r))], .groups = "drop") %>% 
    dplyr::rename("transcript_res" = "pearsons_r") %>%

    left_join(df_x_res_genes %>% 
                  dplyr::select(ensembl_gene_id, gene_symbol, pearsons_r) %>% 
                  dplyr::rename("gene_res" = "pearsons_r")
    ) %>%

    # calculate distance to y = x
    #mutate(distance = f_distance_to_line(gene_res, transcript_res, 1, 0)) %>%

    # highlight SCZ risk genes
    mutate(
        risk_gene = ifelse(ensembl_gene_id %in% scz_risk_genes, "yes", "no"),
        #label = ifelse(risk_gene == "yes" & distance < 0.1 & abs(gene_res) > 0.1, gene_symbol, "")
        label = ifelse(gene_symbol %in% c(fig5_sig_genes, fig5_non_sig_genes), gene_symbol, "")
    ) %>% 
    arrange(risk_gene)
```

```{r Fig5b.plot}
df_fig5b %>% 
    ggplot(aes(x = gene_res, y = transcript_res)) +
    geom_abline(lty = 2) +
    geom_hline(yintercept = 0, color = "gray") +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(fill = gene_res + transcript_res, color = risk_gene, size = risk_gene), 
               shape = 21, stroke = 0.25) +
    geom_text_repel(aes(label = label), max.overlaps = 1000, min.segment.length = 0, size = 2.5) +
    geom_smooth(method = "lm", color = "black", linewidth = 0.25) +
    guides(color = guide_legend(title = "SCZ risk gene?")) +
    # xlim(c(-0.6, 0.6)) +
    # ylim(c(-0.6, 0.6)) +
    stat_cor(aes(label = after_stat(r.label)), label.sep = "\n", size = 3,
             label.y.npc = "top") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.95, 0.95), guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_size_manual(values = c("yes" = 1, "no" = 0.5), guide = "none") +
    coord_cartesian(clip = "off") +
    #scale_x_continuous(breaks = c(-0.2, 0, 0.2)) +
    labs(x = expression("Gene " * italic(r)[x]), y = expression("Transcript " * italic(r)[x]),
         title = "B | Gene vs transcript GRCCA results"
    ) +
    theme(legend.position = c(0.60, 0.11),
          legend.key.size = unit(0.1, "cm"),
          legend.title = element_text(size = 8, margin = margin(b = 0)),
          legend.text = element_text(size = 6)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.gene_vs_transcript_GRCCA_res", .x), width = fig_width, height = fig_height)
)
```

### C | Directionality of transcripts associated with significant and non-significant risk genes

```{r Fig5c.prep}
## Set order
fig5c_gene_order <- df_x_res_genes %>% 
    filter(gene_symbol %in% c(fig5_sig_genes, fig5_non_sig_genes)) %>% 
    arrange(-abs(pearsons_r)) %>% 
    pull(gene_symbol)

## Prep df for plotting
df_fig5c <- df_x_res_transcripts %>% 
    filter(gene_symbol %in% fig5c_gene_order) %>% 
    mutate(
        gene_symbol = factor(gene_symbol, levels = fig5c_gene_order),
        significant = ifelse(p_adj < 0.05, "yes", "no"),
        gene_sig = ifelse(gene_symbol %in% fig5_sig_genes, "significant", "non-significant")
    )
```

```{r Fig5c.plot, fig.width = 3.75, fig.height = 4.15}
df_fig5c %>%
    #mutate(transcript_symbol = str_remove(transcript_symbol, "^.+-")) %>%

    ggplot(aes(x = pearsons_r, y = reorder(transcript_symbol, pearsons_r))) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(fill = pearsons_r, size = -log10(p_adj), color = significant), shape = 21) +
    facet_wrap(vars(gene_symbol), scales = "free_y") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.278, 0.278), guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "gray"), guide = "none") +
    scale_size_continuous(range = c(1, 4)) +
    scale_y_discrete(labels = function(x) str_remove(x, "^.+-")) +
    guides(size = guide_legend(title = "-log10(FDR)", title.position = "top")) +
    coord_cartesian(clip = "off") +
    labs(x = expression("Transcript " * italic(r)[x]), y = NULL,
         title = expression("C | Transcript " * italic(r)[x] * " of SCZ risk genes")) +
    theme(#legend.position = "bottom",
        legend.position = c(0.95, 0.12),
        legend.title = element_text(hjust = 0.5, margin = margin(b = 0)),
        legend.justification = "center",
        legend.margin = margin(l = -15)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "C.directionality_of_transcripts", .x), width = 3.75, height = 4.15)
)
```

### D | Gene isoform annotation plots

```{r Fig5d.plot, fig.width = 6.5, fig.height = 4.5}
## See 05.GRCCA_transcripts/03c.transcript_annotation.R for figure generation
fig5c
map(
    .x = c(".png", ".pdf"),
    .f = ~ ggsave(paste0(figures_dir, "D.CSMD1_example", .x), width = 6.5, height = 4.5)
)
```

############ OTHER GENES TO PLOT? ##############

```{r}
genes_to_plot <- c("AKAP9", "NRXN1", "DCTN2", "EHBP1", "KIF1B", "SMARCE1", "TCF4", "AKT3", "GRIN1", "FMNL1", "RBFOX1")
genes_to_plot <- c("NRXN1", "GRIN1", "FMNL1", "RBFOX1", "C3", "PKN2-AS1")
genes_to_plot <- c("NSF", "KANSL1", "CRHR1", "ARHGAP27", "LRRC37A", "CCDC43", "FMNL1", 
                   "SPPL2C", "MAPT", "PLEKHM1", "STH", "LINC002210-CRHR1") # Eva's paper
genes_to_plot <- c("ZNF804A") # Owen et al 2023
#genes_to_plot <- benchmarking_lists[["SCZ (rare)"]]
p <- df_x_res_transcripts %>% 
    #filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]]) %>% 
    filter(gene_symbol %in% genes_to_plot) %>%
    #filter(ensembl_gene_id %in% intersect(sig_risk_genes, more_than_one_transcript)) %>%
    mutate(
        #gene_symbol = factor(gene_symbol, levels = sig_risk_gene_order),
        significant = ifelse(p_adj < 0.05, "yes", "no")
    ) %>%

    ggplot(aes(x = pearsons_r, y = reorder(ensembl_transcript_id, pearsons_r))) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_point(aes(fill = pearsons_r, size = -log10(p_adj), color = significant), shape = 21) +
    facet_wrap(vars(gene_symbol), scales = "free_y") +
    scale_fill_gradientn(colors = gene_weight_color_scale) +
    scale_color_manual(values = c("yes" = "black", "no" = "gray")) +
    coord_cartesian(clip = "off")
df_x_res_genes %>% filter(gene_symbol %in% genes_to_plot)
df_de_res %>% filter(gene_symbol %in% genes_to_plot)
```

### 2025-07-02 scripts/07.figures_markdown/supplement_v2.Rmd
Update scripts with response to reviewer comments

--- 2025-01-10 scripts/08.figures_markdown/supplement.Rmd
@@
 ---
-title: "Supplemental figures for RLSmith Transl Psych 2024"
+title: "Supplemental figures for RLSmith Transl Psych 2025"
 author: "Rachel L Smith"
@@
 base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
-supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
+supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/panels/")
 tables_dir <- paste0(base_dir, "outputs/tables/")
@@
 }
-
-## Load Nirmala's DEG results
-df_akula_res <- read_xlsx(path = paste0(base_dir, "data/Akula2021_geneLevel_21kgenes_baseMeanGe5_cqn-Rin-Race-GC_DESeq2-pvalsCorrected_betaPriorT-lfcShrink_092518.xlsx"),
-          sheet = "SczVsCtrls") %>% 
-    clean_names %>% 
-    dplyr::select(ensembl_gene_id, hgnc_symbol, log2fold_change, fold_change, lfc_se, stat, pvalue, padj)
 ```
@@
 ```{r TableS1}
-write_xlsx(df_covariates, path = paste0(tables_dir, "TableS1.covariate_information.xlsx"))
+write_xlsx(df_covariates %>% 
+               mutate(dx = factor(dx, levels = names(dx_colors))) %>% 
+               arrange(dx, sex_at_birth, race, age_death) %>% 
+               mutate_at(vars(c(bmi, height, weight)), ~ as.numeric(.x)) %>% 
+               dplyr::select(-suicide, -sample), 
+           path = paste0(tables_dir, "TableS1.covariate_information.xlsx"))
 ```
@@
 ```{r TableS2}
-write_xlsx(list("A.GRCCA results" = df_results,
-                "B.Covariate (Y) results" = df_y_res,
-                "C.Gene (X) results" = df_x_res), 
+write_xlsx(list("A.GRCCA results" = df_results %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
+                "B.Covariate (Y) results" = df_y_res %>% dplyr::rename("structure_correlation" = "pearsons_r"),
+                "C.Gene (X) results" = df_x_res %>% dplyr::rename("structure_correlation" = "pearsons_r")
+                ), 
            path = paste0(tables_dir, "TableS2.GRCCA_results.xlsx")
@@
 ```{r TableS4}
-write_xlsx(df_de_res, path = paste0(tables_dir, "TableS4.DEGs.xlsx"))
+write_xlsx(
+    list("TableS4" = df_limma_res %>% arrange(-abs(t))), 
+    path = paste0(tables_dir, "TableS4.DEGs.xlsx")
+)
 ```
@@
 ```{r TableS6}
-write_xlsx(list("A.GRCCA results" = df_results_transcripts,
-                "B.Covariate (Y) results" = df_y_res_transcripts,
-                "C.Gene (X) results" = df_x_res_transcripts), 
+write_xlsx(list("A.GRCCA results" = df_results_transcripts %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
+                "B.Covariate (Y) results" = df_y_res_transcripts %>% dplyr::rename("structure_correlation" = "pearsons_r"),
+                "C. Transcript (X) results" = df_x_res_transcripts %>% dplyr::rename("structure_correlation" = "pearsons_r")), 
            path = paste0(tables_dir, "TableS6.GRCCA_transcript_results.xlsx")
@@
     "A. Module gene assignments" = df_modules_filt %>% left_join(df_ensembl_to_symbol) %>% dplyr::select(module, color, ensembl_gene_id, gene_symbol) %>% as.data.frame,
-    "B. Module GO results" = df_mods_go %>% dplyr::filter(pvalue < 0.05 & module != 0) %>% as.data.frame
+    "B. Module GO results" = df_mods_go %>% dplyr::filter(pvalue < 0.05 & module != 0) %>% dplyr::select(-qvalue) %>% as.data.frame
 ),
@@
 ```{r TableS8}
-write_xlsx(df_lm_dx %>% dplyr::select(-data) %>% dplyr::filter(p_value < 0.05),
+write_xlsx(df_kme %>% 
+               dplyr::filter(p_value < 0.05 & pc == "PC1") %>% 
+               dplyr::select(-c(data, pc, proportion_of_variance, estimate, std_error)),
            path = paste0(tables_dir, "TableS8.kME_analysis_significant_results.xlsx")
@@
 ```{r TableS9}
-write_xlsx(list("A.RCCA results" = df_results_rcca,
-                "B.Covariate (Y) results" = df_y_res_rcca,
-                "C.Gene (X) results" = df_x_res_rcca), 
+write_xlsx(list("A.RCCA results" = df_results_rcca %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
+                "B.Covariate (Y) results" = df_y_res_rcca %>% dplyr::rename("structure_correlation" = "pearsons_r"),
+                "C.Gene (X) results" = df_x_res_rcca %>% dplyr::rename("structure_correlation" = "pearsons_r")), 
            path = paste0(tables_dir, "TableS9.RCCA_results.xlsx")
@@
 ### B | Correlate qSVs with known covariates
-
 ```{r FigS2b.prep}
@@
     } else (stop("Covariate type must be technical, biological, or toxicological"))
-    matrix_tmp <- covariate_matrix[subset,]
-   order <- matrix_tmp[hclust(dist(matrix_tmp))$order,] %>% rownames
+    matrix_tmp <- covariate_matrix[,subset]
+   order <- matrix_tmp[hclust(dist(matrix_tmp))$order,] %>% colnames
    return(order)
 }
+m_covariates <- df_covariates_numeric %>% dplyr::select(technical_covariates, biological_covariates, drug_covariates) %>%  as.matrix()
 covariate_order <- c(
@@
 
-# FIG S3: Drug Multiple Correspondence Analysis results #
+# FIG S3: Toxicology Multiple Correspondence Analysis results #
 
@@
 ```{r FigS3d.plot, fig.width = 3.75, fig.height = 5}
-df_figS3d %>% 
+df_figS3d %>% filter(type != "Diagnostic group") %>%
+
     ggplot(aes(x = dimension, y = covariate)) +
@@
 ```{r FigS5b.prep}
+n_topics <- topic_model$settings$dim$K
 ## Extract topic descriptions (frequent words)
@@
 
-# FIG S8: Vector of significant GRCCA genes (FDR_r < 0.05 & z-score > 2) #
+# FIG S8: Vector of significant GRCCA genes (FDR_r < 0.05 & Z-score > 2) #
 
@@
 
-# FIG S9: Compare current DEGs to Akula et al 2021 DEGs #
+# FIG S9: Compare across WGCNA module solutions
+
+### A | Model significance
+```{r FigS9a, fig.width = 6.5, fig.height = 2}
+df_model_res_wgcna_groups %>% 
+    group_by(mod_set) %>% 
+    mutate(best = ifelse(pval == min(pval), 1, 0)) %>% 
+    mutate(best = ifelse(set == 5, 0, best)) %>% 
+    mutate(best = factor(best)) %>%
+    mutate(mod_set = str_replace(mod_set, "_", ", ")) %>%
+
+    ggplot(aes(x = factor(varx), y = -log10(pval))) +
+    geom_hline(yintercept = -log10(0.005), color = "grey", linewidth = 0.5) +
+    geom_segment(aes(y = 0, yend = -log10(pval), color = mod_set, linetype = mod_set),
+                 position = position_dodge(width = 0.6)
+    ) +
+    geom_point(aes(fill = mod_set, size = correl, shape = best), 
+               position = position_dodge(width = 0.6)
+    ) +
+    scale_size_continuous(range = c(1, 4)) +
+    scale_shape_manual(values = c("0" = 21, "1" = 23), guide = "none") +
+    scale_color_manual(values = c("actual (minSize40, cutHeight0.98)" = "#F8766D", rep("gray", 3))) +
+    scale_fill_manual(values = c("actual (minSize40, cutHeight0.98)" = "#F8766D", rep("gray", 3))) +
+    scale_linetype_manual(values = c("actual (minSize40, cutHeight0.98)" = "solid", 
+                                     "minSize30, cutHeight0.95" = "dashed", 
+                                     "minSize30, cutHeight0.98" = "dotted", 
+                                     "minSize40, cutHeight0.95" = "twodash")
+    ) +
+    guides(fill = guide_legend(title = NULL), color = guide_legend(title = NULL),
+           size = guide_legend(title = "X-Y cor"), linetype = guide_legend(title = NULL)) +
+    coord_cartesian(clip = "off") +
+    labs(x = "X variance explained", y = "-log10(model P value)",
+         title = "A | Model significance across WGCNA solutions") +
+    theme(
+        legend.key.height = unit(0.2, "cm"),
+        legend.title = element_text(size = 8),
+        legend.position = c(0.025, 0.68),
+        legend.spacing = unit(0.2, "lines")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S9A.wgcna_groups_model_significance", .x), width = 6.5, height = 2)
+)
+```
+
+### B | X and Y results correlations
+```{r FigS9b, fig.width = 4.0, fig.height = 3.75}
+# Plot
+plot_y_func <- function(data, mapping, ...) {
+    ggplot(data = data, mapping = mapping) +
+        geom_point(alpha = 0.5) +
+        #geom_smooth(method = "lm", color = "maroon", se = FALSE, linewdith = 0.25) +
+        stat_cor(aes(label = ..r.label..), size = 3, color = "maroon")
+}
+plot_x_func <- function(data, mapping, second_df, ...) {
+    ggplot(data = second_df, mapping = mapping) +  
+        geom_point(alpha = 0.5, size = 0.5) +
+        stat_cor(aes(label = ..r.label..), size = 3, label.x.npc = 0.4, label.y.npc = 0.1, color = "maroon")
+}
+
+ggpairs(df_y_res_wgcna_groups %>% 
+            dplyr::rename_all(~str_replace(.x, "_", ",\n")) %>% 
+            dplyr::rename_at(4, ~"actual"),
+        lower = list(continuous = plot_y_func),  # Scatter + best fit line
+        upper = list(continuous = wrap(plot_x_func, 
+                                       second_df = df_x_res_wgcna_groups %>% 
+                                           dplyr::rename_all(~str_replace(.x, "_", ",\n")) %>% 
+                                           dplyr::rename_at(4, ~"actual")
+                                       )
+        ),  # Correlation values in upper triangle
+        diag = list(continuous = wrap("blank"))  # Remove density plot from diagonal
+) + 
+    theme_bw() +
+    labs(title = "B | Model feature correlations") +
+    theme(plot.title = element_text(face = "bold", size = 11),
+          strip.text = element_text(size = 8.25),
+          strip.clip = "off",
+          strip.background = element_blank(),
+          panel.spacing = unit(0.2, "lines")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S9B.wgcna_groups_x_y_cor", .x), width = 4.0, height = 3.75)
+)
+```
+
+### C | Enrichment for risk genes
+```{r FigS9c, fig.height = 3.75, fig.width = 2.4}
+df_benchmarking_res_wgcna_groups %>%
+    mutate(significant = ifelse(padj < 0.05, "yes", "no")) %>%
+    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), 
+                          "SCZ",
+                          benchmark_list),
+           benchmark_list = ifelse(str_detect(benchmark_list, "SCZ"),
+                                   str_remove(benchmark_list, "SCZ \\(") %>% str_remove("\\)") %>% str_replace(", ", ",\n"), 
+                                   ""
+           ),
+           mod_set = str_replace(mod_set, "_", ",\n")
+    ) %>%
+
+    ggplot(aes(x = benchmark_list, y = mod_set)) +
+    geom_point(aes(fill = NES, size = -log10(padj), color = significant), 
+               stroke = 1, shape = 21) +
+
+    facet_wrap(vars(class), ncol = 1, scales = "free_y") +
+    force_panelsizes(rows = c(1,1,1,3)) +
+    scale_size_continuous(range = c(0.5, 7)) +
+    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.61, 1.61)) +
+    guides(fill = guide_colorbar(title = "NES", title.position = "left", title.hjust = 0.5), 
+           size = guide_legend(title = "-log10(FDR)", title.position = "left", title.hjust = 0.5)) +
+    coord_flip(clip = "off") +
+    labs(x = NULL, y = NULL,
+         title = "C | Risk gene enrichments \nacross WGCNA solutions") +
+    theme(
+        legend.key.height = unit(0.5, "cm"),
+        legend.key.width = unit(0.25, "cm"),
+        legend.title = element_text(size = 8, angle = 90, margin = margin(r = 0)),
+        legend.text = element_text(margin = margin(r = -1)),
+        legend.spacing = unit(0.2, "lines"),
+        legend.position = c(-0.32, 0.75),
+        panel.spacing = unit(0.05, "lines"),
+        strip.text = element_text(size = 9),
+        axis.text.x = element_text(size = 9, angle = 45, hjust = 0.9),
+        plot.margin = margin(b = -10)
+        #legend.margin = margin(l = -10)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S9C.wgcna_groups_benchmarking", .x), width = 2.4, height = 3.75)
+)
+```
+
+# FIG S10: Compare across random module solutions
+
+### A | Model significance
+```{r FigS10a, fig.width = 6.5, fig.height = 2}
+seeds <- df_model_res_random_groups %>% filter(seed != "actual") %>% pull(seed) %>% unique
+df_model_res_random_groups %>% 
+    group_by(seed) %>% 
+    mutate(best = ifelse(pval == min(pval), 1, 0)) %>%
+    mutate(best = ifelse(seed %in% c("seed1856", "seed3144", "seed3578", "seed9883") & set == 4, 0, best)) %>%
+    mutate(best = factor(best)) %>%
+    arrange(desc(seed)) %>%
+
+    ggplot(aes(x = factor(varx), y = -log10(pval))) +
+    geom_hline(yintercept = -log10(0.005), color = "grey", linewidth = 0.5) +
+    geom_segment(aes(y = 0, yend = -log10(pval), color = seed),
+                 position = position_dodge(width = 0.80)
+    ) +
+    geom_point(aes(fill = seed, size = correl, shape = best), 
+               position = position_dodge(width = 0.80)
+    ) +
+    scale_color_manual(values = c("actual" = "#F8766D", rep("gray", 10))) +
+    scale_fill_manual(values = c("actual" = "#F8766D", 
+                                 setNames(rep(adjustcolor("gray", alpha.f = 0.5), 10), seeds)),
+                      guide = "none"
+    ) +
+    scale_size_continuous(range = c(1, 4)) +
+    scale_shape_manual(values = c("0" = 21, "1" = 23), guide = "none") +
+    guides(color = guide_legend(title = NULL), size = guide_legend(title = "X-Y cor")) +
+    coord_cartesian(clip = "off") +
+    labs(x = "X variance explained", y = "-log10(model P value)",
+         title = "A | Model significance across random groupings") +
+    theme(
+        legend.key.height = unit(0.2, "cm"),
+        legend.title = element_text(size = 8),
+        legend.position = c(0.025, 0.60),
+        legend.spacing = unit(0.5, "lines")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S10A.random_groups_model_significance", .x), width = 6.5, height = 2)
+)
+```
+
+### B | X and Y results correlations
+```{r FigS10b, fig.width = 6.5, fig.height = 2}
+df_y_res_random_groups %>% 
+    pivot_longer(contains("seed"), names_to = "seed", values_to = "random") %>%
+    mutate(type = "Covariates \n(ry)") %>%
+
+    bind_rows(
+        df_x_res_random_groups %>% 
+            pivot_longer(contains("seed"), names_to = "seed", values_to = "random") %>% 
+            mutate(type = "Genes \n(rx)")
+    ) %>%
+
+    ggplot(aes(x = random, y = actual)) +
+    geom_point(alpha = 0.5, size = 0.5) +
+    stat_cor(aes(label = ..r.label..), label.y = 0.8, color = "maroon", size = 2.75) +
+    facet_grid(type ~ seed) +
+    scale_x_continuous(breaks = c(-0.5, 0.5)) +
+    labs(x = "Random grouping structure correlations", y = "Actual structure correlations",
+         title = "B | Model feature correlations") +
+    coord_cartesian(clip = "off") +
+    theme_bw() +
+    theme(
+        strip.text.x = element_blank(),
+        strip.text.y = element_text(size = 8.5),
+        strip.background = element_blank(),
+        strip.clip = "off",
+        plot.title = element_text(face = "bold", size = 11),
+        axis.title = element_text(size = 10),
+        axis.text = element_text(size = 8),
+        panel.border = element_rect(linewidth = 0.30),
+        plot.margin = margin(r = -5)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S10B.random_groups_x_y_cor", .x), width = 6.5, height = 2)
+)
+```
+
+### C | Enrichment for risk genes
+```{r FigS10c, fig.height = 2, fig.width = 6.5}
+df_benchmarking_res_random_groups %>%
+    mutate(significant = ifelse(padj < 0.05, "yes", "no")) %>%
+    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), 
+                          "SCZ",
+                          benchmark_list),
+           benchmark_list = ifelse(str_detect(benchmark_list, "SCZ"),
+                                   str_remove(benchmark_list, "SCZ \\(") %>% str_remove("\\)") %>% str_replace(", ", ",\n"), 
+                                   ""
+           )
+    ) %>%
+
+    ggplot(aes(x = benchmark_list, y = seed)) +
+    geom_point(aes(fill = NES, size = -log10(padj), color = significant), 
+               stroke = 1, shape = 21) +
+
+    facet_wrap(vars(class), nrow = 1, scales = "free_x") +
+    force_panelsizes(cols = c(1,1,1,3)) +
+    scale_size_continuous(range = c(0.1, 5)) +
+    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.61, 1.61)) +
+    guides(fill = guide_colorbar(title = "NES", title.position = "left", title.hjust = 0.5), 
+           size = guide_legend(title = "-log10(FDR)", title.position = "left", title.hjust = 0.5)) +
+    coord_cartesian(clip = "off") +
+    labs(x = NULL, y = NULL,
+         title = "C | Risk gene enrichments across random groupings") +
+    theme(
+        legend.key.height = unit(0.5, "cm"),
+        legend.key.width = unit(0.25, "cm"),
+        legend.title = element_text(size = 8, angle = 90),
+        legend.spacing = unit(0.2, "lines"),
+        legend.margin = margin(l = -10)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S10C.random_groups_benchmarking", .x), width = 6.5, height = 2)
+)
+```
+
+# FIG S11: Compare GRCCA and RCCA directly #
+
+### A | Correlate GRCCA and RCCA covariate (y) structure correlations
+```{r FigS11a.plot}
+df_y_res_all %>% 
+    pivot_wider(id_cols = covariate, names_from = analysis, values_from = pearsons_r) %>% 
+    ggplot(aes(x = rcca, y = grcca)) +
+    geom_point(aes(fill = rcca + grcca), shape = 21, size = 2.5) +
+    geom_vline(aes(xintercept = 0), color = "gray") +
+    geom_hline(aes(yintercept = 0), color = "gray") +
+    geom_abline(lty = 2) +
+    geom_text_repel(aes(label = covariate), size = 2.5, max.overlaps = 20, box.padding = 1.25,
+                     nudge_y = 0.3, force = 150) +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.84, 1.84), guide = "none") +
+    labs(title = "A | Covariate structure correlations",
+         x = "RCCA", y = "GRCCA")
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.covariate_structCor_correlation", .x), width = fig_width, height = fig_height)
+)
+```
+
+### B | Correlate GRCCA and RCCA gene (x) structure correlations
+```{r FigS11b.plot}
+df_x_res_all %>% 
+    pivot_wider(id_cols = ensembl_gene_id, names_from = analysis, values_from = pearsons_r) %>% 
+    ggplot(aes(x = rcca, y = grcca)) +
+    geom_point(aes(color = rcca + grcca), size = 1) +
+    geom_vline(aes(xintercept = 0), color = "gray") +
+    geom_hline(aes(yintercept = 0), color = "gray") +
+    geom_abline(lty = 2) +
+    stat_cor(label.sep = "\n") +
+    #geom_label_repel(aes(label = gene_symbol), size = 3, max.overlaps = 50, box.padding = 1.5) +
+    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-1.14, 1.14), guide = "none") +
+    labs(title = "B | Gene structure correlations",
+         x = "RCCA", y = "GRCCA")
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S11B.gene_structCor_correlation", .x), width = fig_width, height = fig_height)
+)
+```
+
+### C | Venn diagram of significant gene overlap (made in PowerPoint/Inkscape)
+```{r FigS11c.prep}
+## Venn diagram (overlap with Nirmala's results)
+sig_grcca_genes <- df_x_res %>%
+    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>%
+    pull(ensembl_gene_id)
+sig_rcca_genes <- df_x_res_rcca %>% 
+    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
+    pull(ensembl_gene_id)
+length(sig_grcca_genes) # N = 1211
+length(sig_rcca_genes) # N = 2026
+hypergeometric_overlap <- f_hypergeometric(sig_grcca_genes, sig_rcca_genes, gene_universe = ensembl_gene_ids)
+length(hypergeometric_overlap$overlap_genes) # N = 855; P < 0.0001
+```
+
+### D | RCCA results risk gene enrichment
+```{r FigS11d.prep}
+df_fig13d <- df_rcca_fgsea_benchmark %>% 
+    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
+    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
+               str_trim %>% 
+               str_replace(", ", ",\n")
+    ) %>% 
+    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
+    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
+    mutate(label = round(NES, 2) %>% format(digits = 3))
+    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
+   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))
+```
+
+```{r FigS11d.plot}
+df_fig13d %>% 
+    ggplot(aes(x = benchmark_list, y = 1)) +
+    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
+    geom_text(aes(label = label, color = I(label_color)), size = 3) +
+    facet_wrap(vars(class), scales = "free", nrow = 2) +
+    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
+    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
+    labs(x = NULL, y = NULL,
+         title = "D | Risk gene enrichment") +
+    theme(axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line = element_blank(),
+          axis.text.x = element_text(margin = margin(t = -3)),
+          strip.background = element_rect(fill = "white", color = "black"),
+          legend.margin = margin(r = -15),
+          legend.position = "left",
+          legend.justification = "top",
+          legend.title = element_text(angle = 90),
+          legend.key.height = unit(0.75, "cm"),
+          legend.key.width = unit(0.15, "cm")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S11D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
+)
+```
+
+# FIG S12: Compare limme DEG results with Akula et al 2021 #
+
+### A | Venn diagram of significant DEG overlap
+```{r FigS12a.prep}
+## Venn diagram (overlap with Nirmala's results)
+my_degs <- df_limma_res %>% dplyr::filter(p_value < 0.05) %>% pull(ensembl_gene_id)
+length(my_degs) # N = 1389
+
+akula_degs <- df_akula_res %>% 
+    dplyr::filter(pvalue < 0.05) %>% 
+    pull(ensembl_gene_id)
+length(akula_degs) # N = 1373
+
+sig_grcca_genes <- df_x_res %>% 
+    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
+    pull(ensembl_gene_id)
+length(sig_grcca_genes) # N = 1211
+
+# Akula DEGs vs my DEGs
+hypergeometric_overlap <- f_hypergeometric(my_degs, akula_degs, gene_universe = ensembl_gene_ids)
+length(hypergeometric_overlap$overlap_genes) # N = 385; P = 1.9*10^-145; OR = 7.17
+
+# GRCCA vs my DEGs
+hypergeometric_overlap <- f_hypergeometric(my_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
+length(hypergeometric_overlap$overlap_genes) # N = 203; P = 1.5*10^-29; OR = 2.77
+
+# GRCCA vs Akula DEGs
+hypergeometric_overlap <- f_hypergeometric(akula_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
+length(hypergeometric_overlap$overlap_genes) # N = 285; P = 2.3*10^-85; OR = 5.19
+
+# Combine the data into a named list for ggvenn
+FigS12a_data <- list(
+  `Current study DEGs` = my_degs,
+  `Akula et al 2021 DEGs` = akula_degs,
+  `Significant GRCCA genes` = sig_grcca_genes
+)
+```
+
+```{r FigS12a.plot}
+ggvenn(
+    FigS12a_data, 
+    fill_color = c("#E0E0E0", "#E0E0E0", "#E0E0E0"),
+    stroke_size = 0.75,
+    text_size = 3,
+    set_name_size = 4,
+    show_percentage = FALSE
+) +
+    coord_fixed(clip = "off") +
+    theme_void() +
+    theme(
+        text = element_text(size = 8)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S12A.DEG_GRCCA_venn_diagram", .x), width = fig_width, height = fig_height)
+)
+```
+
+### B | Correlation plot of L2FC
+```{r FigS12b.stats}
+
+### PERMUTE L2FC CORRELATION ###
+n_perm <- 10000
+df_limma_akula_cor_perm <- tibble()
+set.seed(0406)
+for (i in 1:n_perm) {
+
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+
+    # Permute LFC within module
+    df_tmp <- df_de_res_all %>% 
+        left_join(df_modules_filt) %>%
+        filter(!is.na(Akula) & metric == "l2fc") %>% 
+        group_by(module) %>%
+        mutate(permuted_l2fc = sample(Limma, replace = FALSE))
+
+    # Calculate correlation between permuted data and Nirmala's LFC
+    test <- cor.test(df_tmp$permuted_l2fc, df_tmp$Akula)
+    correlation <- test$estimate
+    p_value <- test$p.value
+
+    # Append to tibble
+    df_limma_akula_cor_perm <- df_limma_akula_cor_perm %>% 
+        bind_rows(
+            tibble(
+                perm = i,
+                correlation = correlation,
+                p_value = p_value
+            )
+        )
+
+}
+
+## Calculate permutation p-values
+de_limma_akula_perm_p <- df_limma_akula_cor_perm %>% 
+    mutate(obs_corr = cor.test(
+        df_de_res_all %>% filter(!is.na(Akula) & metric == "l2fc") %>% pull(Limma), 
+        df_de_res_all %>% filter(!is.na(Akula) & metric == "l2fc") %>% pull(Akula), 
+        )$estimate
+    ) %>%
+    summarise(
+        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
+    )
+#Pperm=1*10^-4
+
+### PERMUTE T-STAT CORRELATION ###
+n_perm <- 10000
+df_limma_akula_cor_perm <- tibble()
+set.seed(0406)
+for (i in 1:n_perm) {
+
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+
+    # Permute LFC within module
+    df_tmp <- df_de_res_all %>% 
+        left_join(df_modules_filt) %>%
+        filter(!is.na(Akula) & metric == "t_stat") %>% 
+        group_by(module) %>%
+        mutate(permuted_t = sample(Limma, replace = FALSE))
+
+    # Calculate correlation between permuted data and Nirmala's LFC
+    test <- cor.test(df_tmp$permuted_t, df_tmp$Akula)
+    correlation <- test$estimate
+    p_value <- test$p.value
+
+    # Append to tibble
+    df_limma_akula_cor_perm <- df_limma_akula_cor_perm %>% 
+        bind_rows(
+            tibble(
+                perm = i,
+                correlation = correlation,
+                p_value = p_value
+            )
+        )
+
+}
+
+## Calculate permutation p-values
+de_limma_akula_perm_p <- df_limma_akula_cor_perm %>% 
+    mutate(obs_corr = cor.test(
+        df_de_res_all %>% filter(!is.na(Akula) & metric == "t_stat") %>% pull(Limma), 
+        df_de_res_all %>% filter(!is.na(Akula) & metric == "t_stat") %>% pull(Akula), 
+        )$estimate
+    ) %>%
+    summarise(
+        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
+    )
+#Pperm=1*10^-4
+
+```
+
+```{r FigS12b.prep}
+## Create df for plot
+df_figS12b <- df_de_res_all %>%
+
+    # remove genes that were not included in Nirmala's analysis
+    dplyr::filter(!is.na(Akula)) %>%
+
+    # normalize values to compare across analyses
+    #mutate(akula = scale(akula), current = scale(current)) %>%
+
+    # add gene labels
+    mutate(label = ifelse(abs(Limma) > 1.5 & abs(Akula) > 1.5, gene_symbol, "")) %>%
+
+    # arrange to plot
+    arrange(-abs(Limma))
+
+```
+
+```{r FigS12b.plot, fig.width = 4, fig.height = 2.25}
+df_de_res_all %>% 
+    #filter(metric == "l2fc") %>%
+    #mutate(color = scale(Akula)[,1] + scale(Limma)[,1]) %>% 
+    mutate(metric = ifelse(metric == "l2fc", "L2FC", "t-statistic")) %>%
+
+    ggplot(aes(x = Akula, y = Limma)) +
+    geom_hline(yintercept = 0, color = "gray") +
+    geom_vline(xintercept = 0, color = "gray") +
+    geom_abline(lty = 2) +
+    geom_point(color = "darkgray", size = 0.25) +
+    # geom_text_repel(aes(label = label), min.segment.length = 0, max.overlaps = 15, size = 2.5) +
+    geom_smooth(method = "lm", linewidth = 0.5, color = "maroon") +
+    stat_cor(label.sep = "\n", size = 3, vjust = 1) +
+    facet_wrap(vars(metric), scales = "free") +
+    #scale_color_gradientn(colors = rev(brewer.rdbu(100)), guide = "none") +
+    labs(
+        x = "Akula DGE (DESeq2)", y = "Current study DGE (Limma-voom)",
+        title = "B | Correlation between DGE analysis results"
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S12B.DEG_correlation", .x), width = 4, height = 2.25)
+)
+```
+
+# FIG S13: Transcript model results (subset and full) #
+
+### A | Optimal model for all transcripts
+```{r FigS13a.plot}
+df_results_ALLtranscripts %>% 
+    mutate(significant = ifelse(pval == 0.001, "yes", "no")) %>%
+
+    ggplot(aes(x = factor(varx*100), y = -log10(pval))) +
+    geom_segment(aes(y = 0, yend = -log10(pval))) +
+    geom_point(aes(size = correl, shape = significant, fill = -log10(pval))) +
+    geom_hline(yintercept = 0, color = "darkgrey") +
+    scale_shape_manual(values = c("yes" = 23, "no" = 21), guide = "none") +
+    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0, 3), guide = "none") +
+    guides(size = guide_legend(title = "X-Y LV correlation", override.aes = list(shape = 21))) +
+    coord_cartesian(clip = "off") +
+    labs(x = "X variance explained (%)", y = "-log10(model p-value)",
+         title = "A | Optimal variance explained \nby model (all transcripts)"
+    ) +
+    theme(
+        legend.position = c(0.55, 0.75),
+        legend.key.size = unit(0.1, "cm")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S13A.optimal_model_all_transcripts", .x), width = fig_width, height = fig_height)
+)
+```
+
+### B | Covariate correlations with the latent variable (all transcripts)
+```{r FigS13b.plot}
+df_y_res_ALLtranscripts %>% 
+    mutate(significant = ifelse(abs(z_score) >= 2 & p_adj < 0.05, "*", "")) %>%
+
+    ggplot(aes(x = pearsons_r, y = reorder(covariate, pearsons_r))) +
+    geom_point(aes(size = -log10(p_adj), fill = pearsons_r), shape = 21) +
+    geom_vline(xintercept = 0, color = "darkgrey") +
+    #geom_text(aes(label = significant), size = 6, vjust = 0.75, hjust = -1.5, color = "black") +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1, 1), guide = "none") +
+    scale_size_continuous(range = c(2, 6)) +
+    guides(size = guide_legend(title = "-log10(FDR)")) +
+    coord_cartesian(clip = "off") +
+    xlim(c(-1, 1)) +
+    labs(y = NULL, x = expression("Structure correlation (" * italic(r)[y] * ")"), 
+         title = "B | Covariate structure \ncorrelations (all transcripts)") +
+    theme(legend.position = c(0.05, 0.55), 
+          legend.box = "horizontal",
+          #legend.justification = "center",
+          legend.key.size = unit(0.1, 'cm')
+          #legend.key.height = unit(0.29, "cm"),
+          #legend.key.width = unit(0.30, "cm")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S13B.covariate_correlations_all_transcripts", .x), width = fig_width, height = fig_height)
+)
+```
+
+### C | Optimal model for subset transcripts
+```{r FigS13c.plot}
+df_results_transcripts %>% 
+    mutate(significant = ifelse(pval == 0.001, "yes", "no")) %>% 
+    ggplot(aes(x = factor(varx*100), y = -log10(pval))) +
+    geom_segment(aes(y = 0, yend = -log10(pval))) +
+    geom_point(aes(size = correl, shape = significant, fill = -log10(pval))) +
+    geom_hline(yintercept = 0, color = "darkgrey") +
+    scale_shape_manual(values = c("yes" = 23, "no" = 21), guide = "none") +
+    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0, 3), guide = "none") +
+    guides(size = guide_legend(title = "X-Y LV correlation", override.aes = list(shape = 21))) +
+    coord_cartesian(clip = "off") +
+    labs(x = "X variance explained (%)", y = "-log10(model p-value)",
+         title = "C | Optimal variance explained \nby model (subset transcripts)"
+    ) +
+    theme(
+        legend.position = c(0.55, 0.75),
+        legend.key.size = unit(0.1, "cm")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S13C.optimal_model_subset_transcripts", .x), width = fig_width, height = fig_height)
+)
+```
+
+# FIG S14: Isoform structures of PTK2B and ARHGAP44 (schizophrenia risk genes) #
+
+### A | PTK2B
+```{r FigS14a.plot}
+## See 05.GRCCA_transcripts/03c.transcript_annotation.R for figure generation
+figS14b
+map(
+    .x = c(".png", ".pdf"),
+    .f = ~ ggsave(paste0(supplement_figures_dir, "S14A.PTK2B", .x), width = 6.5, height = 4.5)
+)
+```
+
+### B | ARHGAP44
+```{r FigS14a.plot}
+## See 05.GRCCA_transcripts/03c.transcript_annotation.R for figure generation
+figS14a
+map(
+    .x = c(".png", ".pdf"),
+    .f = ~ ggsave(paste0(supplement_figures_dir, "S14B.ARHGAP44", .x), width = 6.5, height = 4.5)
+)
+```
+
+# FIG S15: GRCCA results are related to single-cell expression gradients implicated in SCZ #
+
+```{r FigS15.prep}
+df_figS15 <- df_lf4_grcca %>% 
+    left_join(df_lake_cell_type) %>% 
+    dplyr::filter((type == "Astro" & cell_type == "astrocyte") |
+                      (type == "Neuro-Ex" & cell_type == "glutamatergic") |
+                      (type == "Neuro-In" & cell_type == "gabaergic")
+    )
+```
+
+```{r FigS15.plot, fig.width = 6}
+df_figS15 %>% 
+    ggplot(aes(x = PEER_4, y = pearsons_r)) +
+    geom_vline(aes(xintercept = 0), color = "gray") +
+    geom_hline(aes(yintercept = 0), color = "gray") +
+    geom_point(aes(color = pearsons_r), size = 0.5) +
+    geom_smooth(color = "black", method = "lm", linewidth = 0.75) +
+    stat_cor(label.sep = "\n", size = 3, 
+             label.y.npc = "top", label.x.npc = "right", hjust = 1, vjust = 1) +
+    facet_wrap(vars(cell_type), scales = "free") +
+    scale_color_gradientn(colors = gene_weight_color_scale, guide = "none") +
+    scale_x_continuous(breaks = c(-2000, 0, 2000)) +
+    labs(x = "Ling et al LF4", y = expression("Structure correlation (" * italic(r)[x] * ")"),
+         title = "GRCCA correlation with SNAP program by cell type")
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S15.correlation_with_SNAP", .x), width = 6, height = fig_height)
+)
+```
+
+# FIG S16: WGCNA module - phenotype associations determined through traditional module eigengene analysis #
+
+### A | Module eigengene - covariate significant associations
+```{r FigS16a.prep}
+df_figS16a <- df_lm_dx_allPCs %>% 
+    filter(module != "geneM0" & pc == "PC1") %>% 
+    mutate(term = str_replace(term, "MC", "Dim")) %>% 
+    mutate(term = factor(term, levels = c(paste0("Dim", 8:1), names(dx_colors)))) %>%
+    mutate(significant = ifelse(p_value < 0.05, "yes", "no")) %>% 
+    arrange(abs(statistic))
+```
+
+```{r FigS16a.plot, fig.height = 3, fig.width = 6.5}
+df_figS16a %>% 
+    ggplot(aes(x = module, y = term)) +
+    geom_hline(yintercept = 8.5, color = "gray") +
+    geom_point(aes(fill = statistic, size = abs(statistic), color = significant), shape = 21) +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-3.3, 3.3)) +
+    scale_color_manual(values = c("yes" = "black", "no" = "grey")) +
+    scale_size_continuous(limits = c(0, 3.3), range = c(0.5, 5), guide = "none") +
+    guides(fill = guide_colorbar(title = "t-statistic", title.position = "top", title.hjust = 0.5),
+           color = guide_legend(title = "P < 0.05", title.position = "top", title.hjust = 0.5)
+    ) +
+    coord_cartesian(clip = "off") +
+    labs(x = NULL, y = NULL,
+         title = "A | Module eigengene - covariate associations"
+    ) +
+    theme(
+        legend.position = "bottom",
+        legend.justification = "center",
+        legend.margin = margin(t = -10),
+        legend.key.height = unit(0.15, "cm"),
+        legend.key.width = unit(0.75, "cm"),
+        axis.text.x = element_text(size = 8.5, angle = 45, hjust = 0.9)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S16a.module_kme_covar_matrix", .x), width = 6.5, height = 3)
+)
+```
+
+### B | Modules significantly associated with SCZ per module eigengene analysis
+```{r FigS16b.prep}
+df_figS16b <- df_lm_dx_allPCs %>% 
+    dplyr::filter(p_value < 0.05 & term == "SCZ" & pc == "PC1") %>% 
+    unnest(cols = c(data)) %>% 
+    mutate(label = case_when(
+        dx == "SCZ" & p_value < 0.05 & p_adj > 0.05 ~ "*",
+        dx == "SCZ" & p_value < 0.05 & p_adj < 0.05 ~ "**",
+        TRUE ~ ""
+    )
+    )
+```
+
+```{r FigS16b.plot, fig.height = 3, fig.width = 6.5}
+df_figS16b %>% 
+    ggplot(aes(x = pc_resids, y = dx, color = dx, fill = dx)) +
+    geom_point(position = position_jitter(width = 0.1), size = 0.5, alpha = 0.75) +
+    geom_boxplot(alpha = 0.1, outlier.shape = NA) +
+    # geom_text(aes(label = label, x = -Inf), color = "black", size = 6,
+    #           vjust = 0.75, hjust = -0.1, check_overlap = TRUE) +
+    facet_wrap(vars(module), ncol = 3, scales = "free_x") +
+    scale_fill_manual(values = dx_colors) +
+    scale_color_manual(values = dx_colors) +
+    labs(x = "kME (corrected for covariates)", y = NULL,
+         #caption = "* = p < 0.05; ** = FDR < 0.05",
+         title = "B | Module - SCZ associations"
+    ) +
+    theme(legend.position = "none",
+          axis.text.y = element_text(size = 10)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S16B.sig_SCZ_kME_modules", .x), width = 6.5, height = 3)
+)
+```
+
+# FIG S17: WGCNA module enrichment for risk gene lists #
+
+```{r FigS17.prep}
+df_figS17 <- df_mods_risk_gene_hypergeometric %>% 
+    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", benchmark_list) %>% factor(levels = c("SCZ", "ASD", "BD", "MDD"))) %>% 
+    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
+               str_trim %>% 
+               str_replace(", ", ",\n")
+    ) %>% 
+    filter(module != "geneM0") %>% 
+    mutate(significant = ifelse(p_adj < 0.05, "yes", "no"),
+           very_significant = ifelse(p_adj < 0.001, "***", NA_character_),
+           benchmark_list = ifelse(class != "SCZ", "", benchmark_list),
+           module = factor(module, levels = names(rev(module_colors)))
+    )
+```
+
+```{r FigS17.plot, fig.width = 5, fig.height = 5}
+df_figS17 %>% 
+    ggplot(aes(x = benchmark_list, y = module)) +
+    # Layer 1: fill with variable alpha
+    geom_point(aes(fill = module, size = -log10(p_adj), alpha = -log10(p_adj)), color = "transparent", shape = 21, stroke = 0.75) +
+    # Layer 2: Border only, no alpha applied
+    geom_point(aes(size = -log10(p_adj), color = significant), fill = "transparent", shape = 21, stroke = 0.75) +
+
+    geom_text(aes(label = very_significant), size = 4, color = "black", vjust = 0.75) +
+    facet_wrap(vars(class), scales = "free_x", nrow = 1) +
+    force_panelsizes(cols = c(3,1,1,1)) +
+    scale_alpha_continuous(guide = "none") +
+    scale_fill_manual(values = module_colors, guide = "none") +
+    scale_color_manual(values = c("yes" = "black", "no" = "grey"), guide = "none") +
+    guides(color = guide_legend(title = "FDR < 0.05"),
+           size = guide_legend(title = "-log10(FDR)")
+    ) +
+    coord_cartesian(clip = "off") +
+    labs(x = NULL, y = NULL, 
+         title = "Risk gene list enrichment across WGCNA modules") +
+    theme(legend.margin = margin(l = -15))
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S17.risk_gene_module_hypergeometric", .x), width = 5, height = 5)
+)
+```
+
+# FIG S18: Functional enrichment of GRCCA overrepresented WGCNA modules #
+
+### A | Hypergeometric overlap of WGCNA modules and significant GRCCA genes
+```{r FigS18a.prep}
+## Identify significant GRCCA genes (FDR < 0.05 & |z| > 2)
+sig_grcca_genes <- df_x_res %>% 
+    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
+    pull(ensembl_gene_id)
+# n = 1211
+
+## Hypergeometric test for significant overlap with modules
+df_grcca_module_hypergeometric <- map_dfr(
+    .x = names(module_colors),
+    .f = ~ f_module_hypergeometric(
+        gene_list = sig_grcca_genes, 
+        gene_list_name = "", 
+        wgcna_module = .x, 
+        gene_universe = ensembl_gene_ids
+    ) %>% 
+        dplyr::select(-gene_list)
+) %>% 
+    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>% 
+    arrange(p_adj) %>% 
+    mutate(module = factor(module, levels = names(module_colors))) %>%
+
+    # specify modules to label
+    mutate(module_label = ifelse(p_adj < 0.05, paste0("N = ", overlap_n), ""))
+```
+
+```{r FigS18a.plot, fig.height = 5, fig.width = 2.15}
+f_plot_module_hypergeometric(df_grcca_module_hypergeometric %>% mutate(p_adj = ifelse(p_adj == 0, 10^-5, p_adj)) %>% dplyr::filter(module != "geneM0"),
+                             x_axis = "p-adj",
+                             plot_title = "A | GRCCA significant modules",
+                             #subset_x_labs = TRUE,
+                             include_module_labels = TRUE,
+                             module_text_size = 3,
+                             legend_position = "none",
+                             p_value_threshold = 0.05,
+                             include_threshold_label = TRUE
+)
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S18A.GRCCA_module_hypergeometric", .x), width = 2.15, height = 5)
+)
+```
+
+### B | Direction of significant genes in top modules
+```{r FigS18b.prep}
+## Identify significant modules
+sig_modules <- df_grcca_module_hypergeometric %>% dplyr::filter(module != "geneM0" & p_adj < 0.05) %>% pull(module) %>% sort()
+
+## Identify panel sizes (number of significant genes in each module)
+panel_widths <- df_x_res %>% 
+    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
+    dplyr::count(module) %>% 
+    pull(n)
+```
+
+```{r FigS18b.plot, fig.height = 5, fig.width = 2.15}
+df_x_res %>% 
+    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
+    mutate(weight = ifelse(significant == 1, weight, 0)) %>%
+
+    ggplot(aes(x = reorder_within(ensembl_gene_id, within = module, by = pearsons_r), y = pearsons_r)) +
+    geom_col(aes(fill = module, alpha = abs(pearsons_r))) +
+    facet_wrap(vars(module), ncol = 1, scales = "free_y") +
+    force_panelsizes(rows = panel_widths) +
+    scale_fill_manual(values = module_colors) +
+    scale_alpha_continuous(range = c(0.3, 1)) +
+    labs(title = "B | Significant module genes",
+         x = "Gene", y = expression("Structure correlation (" * italic(r)[x] * ")")) +
+    guides(fill = guide_legend(nrow = 2)) +
+    coord_flip() +
+    theme(axis.text.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          strip.text = element_blank(),
+          legend.position = "none",
+          panel.spacing = unit(0.3, "lines")
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S18B.sig_module_direction", .x), width = 2.15, height = 5)
+)
+
+```
+
+### C | GO BP enrichments of significant modules
+```{r FigS18c.prep}
+# Filter for GRCCA modules of interest & take top 5 pathways for each
+df_figS18c <- df_mods_go %>% 
+    dplyr::filter(module %in% sig_modules) %>% 
+    group_by(module) %>% 
+    top_n(5, wt = -log10(pvalue)) %>% 
+    mutate(Description = str_wrap(Description, width = 23.5))
+```
+
+```{r FigS18c.bubbleplot, fig.height = 5, fig.width = 2.15}
+df_figS18c %>% 
+    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
+    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
+    facet_wrap(vars(module), scales = "free_y", ncol = 1) +
+    scale_fill_manual(values = module_colors, guide = "none") +
+    scale_y_reordered() +
+    scale_x_reverse(breaks = c(0.00005, 0.00)) +
+    scale_size_continuous(range = c(1, 4)) +
+    coord_cartesian(clip = "off") +
+    labs(x = "Enrichment FDR", y = NULL,
+         title = "C | Module enrichments") +
+    theme(
+        legend.position = c(-1.6, 1.05),
+        legend.key.size = unit(0.2, "cm"),
+        axis.text = element_text(size = 8.5)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S18C.sig_GRCCA_module_enrichments", .x), width = 2.15, height = 5)
+)
+```
+
+# FIG S19: Developmental expression trajectories of GRCCA genes
+
+```{r FigS19.prep}
+## Calculate the mean expression for each decile of GRCCA results across development
+df_figS19 <- df_psychencode_grcca %>% #dplyr::filter(regioncode == "MFC") %>% 
+    group_by(decile, window) %>% 
+    summarise(median_expression = median(dev_expression),
+              sd_expression = sd(dev_expression)
+    )
+
+## Rename windows with time period
+psychencode_windows <- c("5-9",
+                "12-13",
+                "16-18",
+                "19-22",
+                "PCW35-PY0.3",
+                "0.5-2.5",
+                "2.8-10.7",
+                "13-19",
+                "21-64")
+names(psychencode_windows) = 1:9
+
+## Define colors for deciles
+decile_colors <- colorRampPalette(c(gene_weight_color_scale[5], "gray", gene_weight_color_scale[95]))(10) %>% rev
+```
+
+```{r FigS19, fig.height = 3, fig.width = 4}
+df_figS19 %>% 
+    ggplot(aes(x = window, y = median_expression, color = decile)) +
+    geom_vline(xintercept = 5, color = "gray") +
+    geom_smooth(method = "loess", se = FALSE) +
+    scale_x_continuous(breaks = seq(1, length(psychencode_windows), 1),
+                       labels = psychencode_windows) +
+    scale_color_manual(values = decile_colors) +
+    guides(color = guide_legend()) +
+    labs(x = "← PCW | PY →", y = "Median expression",
+         title = "Developmental expression trends") +
+    theme(
+        legend.key.size = unit(0.3, "cm"),
+        axis.text.x = element_text(angle = 45, hjust = 1)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S19.developmental_trajectory", .x), width = 4, height = 3)
+)
+```
+
+# GRAVEYARD #
+
+## Fig S4B alt: Module stability (alluvial)
+
+```{r FigS4b.alluvial, fig.width = 6.5, fig.height = 4}
+df_summary <- df_modules %>%
+  group_by(min_size, cut_height, module, color) %>%
+  summarise(n_genes = n(), .groups = "drop")
+
+# Create the alluvial plot
+library(ggalluvial)
+df_summary %>% 
+    mutate(
+        cut_height = factor(cut_height),
+        min_size = factor(min_size)
+    ) %>%
+
+    ggplot(aes(x = min_size, y = n_genes, stratum = module, alluvium = module, fill = color)) +
+    geom_flow(stat = "alluvium", lode.guidance = "frontback", aes(fill = color)) +
+    geom_stratum(width = 0.2, color = "black") +
+    scale_fill_identity() +  # Use the hex codes from the 'color' column
+    facet_grid(cols = vars(cut_height)) +  # Facet into rows by min_size
+    theme_minimal() +
+    labs(x = "Cut Height", 
+         y = "Number of Genes",
+         title = "Module stability across parameters") +
+    theme(
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        strip.text = element_text(size = 10)
+    )
+```
+
+## Fig SXX: Full WGCNA module functional enrichments
+
+```{r FigS5.prep}
+df_figS5 <- df_mods_go %>% 
+    clean_names %>% 
+    dplyr::filter(!is.na(description) & module != 0) %>%
+
+    # take top 5 paths by p-value per module to plot
+    dplyr::group_by(module) %>% 
+    arrange(module, -p_adjust) %>% 
+    mutate(row = row_number(),
+           gene_ratio = parse(text = gene_ratio) %>% eval
+    ) %>% 
+    top_n(n = 4, wt = row) %>% 
+    #slice_max(order_by = -weight_fisher, n = 5) %>% 
+    mutate(`survives FDR` = ifelse(p_adjust < 0.05, "yes", "no")) %>% 
+    mutate(description = str_wrap(description, width = 20))
+```
+
+```{r FigS5.plot, fig.height = 9, fig.width = 6.5}
+df_figS5 %>% 
+    ggplot(aes(x = p_adjust, y = reorder_within(description, within = module, by = -p_adjust))) +
+    #y = reorder_within(str_wrap(description, width = 35), -pvalue, module))) +
+    geom_point(aes(size = gene_ratio, fill = ontology, color = `survives FDR`),
+               shape = 21, stroke = 0.5) +
+    geom_vline(aes(xintercept = 0.05), linewidth = 0.5, color = "gray") +
+    facet_wrap(vars(module), scales = "free", ncol = 5) +
+    scale_size_continuous(range = c(1, 3.5)) +
+    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
+    scale_y_reordered() +
+    scale_x_reverse(breaks = c(0.25, 0.15, 0.05, 0)) +
+    coord_cartesian(clip = "off") +
+    labs(y = NULL, x = "Enrichment FDR",
+         title = "WGCNA module GO enrichments"
+    ) +
+    guides(size = guide_legend(title = "Gene ratio")) +
+    theme(
+        legend.position = c(0.85, 0.075),
+        #legend.direction = "vertical",
+        legend.title.position = "top",
+        legend.title = element_text(size = 8),
+        legend.key.height = unit(0.05, "cm"),
+        legend.spacing.y = unit(0.1, "cm"),
+        #legend.justification = "center",
+        strip.background = element_rect(fill = "white"),
+        strip.clip = "off",
+        strip.text = element_text(size = 8),
+        panel.spacing = unit(0.005, "lines"),
+        axis.text.y = element_text(size = 7.5),
+        axis.text.x = element_text(size = 7.0)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S5.full_GO_enrichments", .x), width = 6.5, height = 9)
+)
+```
+
+Figure SXX.
+Full GO enrichments of gene-level modules.
+Top 4 Gene Ontology pathways associated with each module, determined by p-value, across all three ontologies (BP = biological process, CC = cellular component, MF = molecular function).
+The x-axis shows the enrichment FDR, while the y-axis shows pathway names.
+Each point is colored by its ontology and sized by the gene ratio (number of genes in the module in the pathway divided by the total number of genes in the pathway); points that survived correction for multiple comparisons (FDR < 0.05; vertical gray line) are outlined in black.
+
+## Fig SXX: Module eigengene explanation
+
+### A | What is an eigengene?
+```{r FigSXXa.prep}
+df_figSXXa <- df_mods_pca_covariates %>% 
+    filter(module == "geneM1") %>% 
+    filter(PC %in% c("PC1", "PC2")) %>% 
+    mutate(PC = paste0(PC, " (", round(proportion_of_variance, 3)*100, "%)")) %>% 
+    dplyr::select(sample, PC, score) %>% distinct %>% 
+    pivot_wider(id_cols = sample, names_from = PC, values_from = score) %>%
+
+    mutate(dx = case_when(
+        str_detect(sample, "control") ~ "Control",
+        str_detect(sample, "bipolar") ~ "BD",
+        str_detect(sample, "mdd") ~ "MDD",
+        str_detect(sample, "schizo") ~ "SCZ"
+    ) %>% 
+        factor(levels = names(dx_colors))
+    )
+```
+
+```{r FigSXXa.plot}
+df_figSXXa %>% 
+    ggplot(aes(x = `PC1 (14.9%)`,
+               y = `PC2 (4.1%)`,
+               fill = dx)) +
+    geom_vline(aes(xintercept = 0), color = "darkgrey") +
+    geom_hline(aes(yintercept = 0), color = "darkgrey") +
+    geom_point(shape = 21, size = 2.0, color = "black") +
+    guides(fill = guide_legend(title = NULL)) +
+    scale_fill_manual(values = dx_colors) +
+    labs(title = "A | Module 1 PCs") +
+    theme(legend.position = "bottom",
+          legend.key.width = unit(0.2, "cm"),
+          legend.margin = margin(t = -15)
+    )
+
+```
+
+### B | Covariate-module PC correlations
+
+```{r FigSXXb.prep}
+df_figSXXb <- df_lm_dx_allPCs %>% 
+    filter(module != "geneM0" & proportion_of_variance >= 0.02) %>% 
+    mutate(term = str_replace(term, "MC", "Dim")) %>% 
+    mutate(term = factor(term, levels = c(paste0("Dim", 8:1), names(dx_colors)))) %>%
+    mutate(significant = ifelse(p_adj < 0.05, "yes", "no")) %>%
+    arrange(abs(statistic)) 
+```
+
+```{r FigSXXb.plot, fig.width = 6.5, fig.height = 8.5}
+df_figSXXb %>% 
+    ggplot(aes(x = pc, y = term)) +
+    geom_hline(yintercept = 8.5, color = "darkgrey", linewidth = 0.25) +
+    geom_point(aes(fill = statistic, size = abs(statistic), color = significant), shape = 21) +
+    facet_wrap(vars(module), scales = "free_x") +
+    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-4.59, 4.59)) +
+    scale_color_manual(values = c("yes" = "black", "no" = "grey")) +
+    scale_size_continuous(limits = c(0, 4.59), range = c(0.25, 4), guide = "none") +
+        guides(fill = guide_colorbar(title = "t-statistic", title.position = "top", title.hjust = 0.5),
+           color = guide_legend(title = "FDR < 0.05", title.position = "top", title.hjust = 0.5)
+    ) +
+    coord_cartesian(clip = "off") +
+    labs(x = NULL, y = NULL,
+         title = "Module PC-covariate associations"
+    ) +
+    theme(
+        legend.direction = "horizontal",
+        legend.box = "vertical",
+        legend.position = c(0.82, 0.08),
+        legend.justification = "center",
+        legend.key.height = unit(0.15, "cm"),
+        legend.key.width = unit(0.75, "cm"),
+        axis.text.x = element_text(size = 8.0, angle = 45, hjust = 0.9),
+        axis.text.y = element_text(size = 8.0),
+        strip.text = element_text(size = 9)
+    )
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "SXX.all_module_PC_covariate_associations", .x), width = 6.5, height = 8.5)
+)
+```
+
+### C | Significant module eigengene SCZ module enrichments
+
+```{r FigS12c.prep}
+## Identify significant modules
+sig_modules <- df_lm_dx_allPCs %>% 
+    dplyr::filter(p_value < 0.05 & term == "SCZ" & pc == "PC1") %>% 
+    pull(module)
+
+# Filter for modules of interest & take top 5 pathways for each
+df_figS12c <- df_mods_go %>% 
+    dplyr::filter(module %in% sig_modules) %>% 
+    group_by(module) %>% 
+    top_n(5, wt = -log10(pvalue)) %>% 
+    mutate(Description = str_wrap(Description, width = 30))
+```
+
+```{r FigS12c.bubbleplot, fig.height = 5, fig.width = 4.5}
+df_figS12c %>% 
+    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
+    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
+    facet_wrap(vars(module), scales = "free_y", ncol = 2) +
+    scale_fill_manual(values = module_colors, guide = "none") +
+    scale_y_reordered() +
+    scale_x_reverse(breaks = c(0.05, 0.00)) +
+    scale_size_continuous(range = c(1, 4)) +
+    coord_cartesian(clip = "off") +
+    labs(x = "Enrichment FDR", y = NULL,
+         title = "C | SCZ module enrichments") +
+    theme(legend.position = c(-1.9, 0.85))
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S12B.sig_SCZ_kME_modules", .x), width = 3.5, height = 5)
+)
+```
+
+## Fig S10B alternate (word clouds)
+
+```{r FigS10b.wordcloud, fig.height = 5}
+## Identify uninteresting words
+data(stop_words)
+vague_terms <- c("positive", "negative", "regulation", "response", "protein", "activity",
+                 "pathway", "process", "involved", "signaling")
+
+## Identify top words in each significant (FDR < 0.05) module pathway
+module_tokens_freq <- df_sig_go_res %>% 
+    dplyr::filter(p.adjust < 0.05) %>% 
+    group_by(module) %>% 
+    unnest_tokens(word, Description) %>% 
+    anti_join(stop_words) %>% 
+    dplyr::filter(!(word %in% vague_terms)) %>% 
+    dplyr::count(word, sort = TRUE)
+
+## Create wordcloud plot for each module
+l_figS10b <- map(
+    .x = as.character(sig_modules),
+    .f = ~ module_tokens_freq %>% 
+        dplyr::filter(module == .x) %>% 
+        top_n(10, wt = n) %>% 
+        ggplot(aes(label = word, size = n, color = module)) +
+        geom_text_wordcloud() +
+        facet_wrap(vars(module)) +
+        scale_color_manual(values = module_colors) +
+        scale_size_continuous(range = c(2.5, 5)) +
+        theme_void() +
+        theme(strip.text = element_blank(),
+              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
+        )
+)
+
+## Combine and save
+wrap_plots(l_figS10b, ncol = 1) & plot_annotation(title = "B | SCZ module enrichments")
+map(
+  .x = c(".png", ".pdf"),
+  .f = ~ ggsave(paste0(supplement_figures_dir, "S10B.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
+)
+```
+
+GO biological process pathway enrichments for each module.
+The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module.
+Words are colored by respective module.
+
+## Fig S12C alternate (word clouds)
+```{r FigS12c.prep}
+## Load module GO enrichment data
+#load(paste0(base_dir, "objects/", prefix, sep = "_", "SIGNED_SFT", soft_power, "_GO_RES.RDS")) # df_mods_go
+
+## Identify words to not include
+data(stop_words)
+vague_terms <- c("positive", "negative", "regulation", "response", "activity",
+                 "pathway", "process", "involved", "signaling")
+
+# Filter for modules of interest
+df_go_res <- df_mods_go %>% 
+    dplyr::filter(module %in% sig_modules)
+
+## Identify top words in each significant (FDR < 0.05) module pathway
+module_tokens_freq <- df_go_res %>% 
+    dplyr::filter(p.adjust < 0.05) %>% 
+    group_by(module) %>% 
+    unnest_tokens(word, Description) %>% 
+    anti_join(stop_words) %>% 
+    dplyr::filter(!(word %in% vague_terms)) %>% 
+    dplyr::count(word, sort = TRUE)
+```
+
+```{r FigS12c.wordcloud}
+l_module_wordclouds <- map(
+    .x = as.character(sig_modules),
+    .f = ~ module_tokens_freq %>% 
+        dplyr::filter(module == .x) %>% 
+        top_n(10, wt = n) %>% 
+        ggplot(aes(label = word, size = n, color = module)) +
+        geom_text_wordcloud() +
+        facet_wrap(vars(module)) +
+        scale_color_manual(values = module_colors) +
+        scale_size_continuous(range = c(2.5, 5)) +
+        theme_void() +
+        theme(strip.text = element_blank(),
+              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
+        )
+)
+```
+
+GO biological process pathway enrichments for each module.
+The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module.
+Words are colored by respective module.
+
+## FIG S9: Compare current DEGs to Akula et al 2021 DEGs #
 
@@
 hypergeometric_overlap <- f_hypergeometric(my_degs, akula_degs, gene_universe = ensembl_gene_ids)
-length(hypergeometric_overlap$overlap_genes) # N = 433; p < 0.0001
+length(hypergeometric_overlap$overlap_genes) # N = 433; P = 5.2*10^-152; OR = 6.7
 
@@
 hypergeometric_overlap <- f_hypergeometric(my_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
-length(hypergeometric_overlap$overlap_genes) # N = 211; p < 0.0001
+length(hypergeometric_overlap$overlap_genes) # N = 211; p < 0.0001; OR = 2.30
 
@@
 hypergeometric_overlap <- f_hypergeometric(akula_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
-length(hypergeometric_overlap$overlap_genes) # N = 285; p < 0.0001
+length(hypergeometric_overlap$overlap_genes) # N = 285; p < 0.0001; OR = 5.21
 
@@
     arrange(-abs(current))
-```
-
-```{r FigS3b.plot}
+
+```
+
+```{r FigS9b.stats}
+n_perm <- 10000
+df_de_akula_cor_perm <- tibble()
+set.seed(0206)
+for (i in 1:n_perm) {
+
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+
+    # Permute LFC within module
+    df_tmp <- df_figS9b %>% 
+        left_join(df_modules_filt) %>% 
+        group_by(module) %>% 
+        mutate(permuted_lfc = sample(current, replace = FALSE))
+
+    # Calculate correlation between permuted data and Nirmala's LFC
+    test <- cor.test(df_tmp$permuted_lfc, df_tmp$akula)
+    correlation <- test$estimate
+    p_value <- test$p.value
+
+    # Append to tibble
+    df_de_akula_cor_perm <- df_de_akula_cor_perm %>% 
+        bind_rows(
+            tibble(
+                perm = i,
+                correlation = correlation,
+                p_value = p_value
+            )
+        )
+
+}
+
+## Calculate permutation p-values
+de_akula_perm_p <- df_de_akula_cor_perm %>% 
+    mutate(obs_corr = cor.test(df_figS9b$current, df_figS9b$akula)$estimate) %>%
+    summarise(
+        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
+    )
+```
+
+```{r FigS9b.plot}
 df_figS9b %>% 
@@
 
-# FIG S10: GRCCA results are related to single-cell expression gradients implicated in SCZ #
-
-```{r FigS10.prep}
-df_figS10 <- df_lf4_grcca %>% 
-    left_join(df_lake_cell_type) %>% 
-    dplyr::filter((type == "Astro" & cell_type == "astrocyte") |
-                      (type == "Neuro-Ex" & cell_type == "glutamatergic") |
-                      (type == "Neuro-In" & cell_type == "gabaergic")
-    )
-```
-
-```{r FigS10.plot, fig.width = 6}
-df_figS10 %>% 
-    ggplot(aes(x = PEER_4, y = pearsons_r)) +
-    geom_vline(aes(xintercept = 0), color = "gray") +
-    geom_hline(aes(yintercept = 0), color = "gray") +
-    geom_point(aes(color = pearsons_r), size = 0.5) +
-    geom_smooth(color = "black", method = "lm", linewidth = 0.75) +
-    stat_cor(label.sep = "\n", size = 3, 
-             label.y.npc = "top", label.x.npc = "right", hjust = 1, vjust = 1) +
-    facet_wrap(vars(cell_type), scales = "free") +
-    scale_color_gradientn(colors = gene_weight_color_scale, guide = "none") +
-    scale_x_continuous(breaks = c(-2000, 0, 2000)) +
-    labs(x = "Ling et al LF4", y = expression("Structure correlation (" * italic(r)[x] * ")"),
-         title = "GRCCA correlation with SNAP program by cell type")
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S10.correlation_with_SNAP", .x), width = 6, height = fig_height)
-)
-```
-
-# FIG S11: SCZ-WGCNA module associations determined through traditional module eigengene analysis #
-
-### A | Modules significantly associated with SCZ per module eigengene analysis
-
-```{r FigS11a.prep}
-df_figS11a <- df_lm_dx %>% 
-    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
-    unnest(cols = c(data)) %>% 
-    mutate(label = case_when(
-        dx == "SCZ" & p_value < 0.05 & p_adj > 0.05 ~ "*",
-        dx == "SCZ" & p_value < 0.05 & p_adj < 0.05 ~ "**",
-        TRUE ~ ""
-    )
-    )
-```
-
-```{r FigS11a.plot, fig.height = 5}
-df_figS11a %>% 
-    ggplot(aes(x = kme_resids, y = dx, color = dx, fill = dx)) +
-    geom_point(position = position_jitter(width = 0.1), size = 0.5) +
-    geom_boxplot(alpha = 0.1, outlier.shape = NA) +
-    # geom_text(aes(label = label, x = -Inf), color = "black", size = 6,
-    #           vjust = 0.75, hjust = -0.1, check_overlap = TRUE) +
-    facet_wrap(vars(module), ncol = 1, scales = "free_x") +
-    scale_fill_manual(values = dx_colors) +
-    scale_color_manual(values = dx_colors) +
-    labs(x = "kME (corrected for covariates)", y = NULL,
-         #caption = "* = p < 0.05; ** = FDR < 0.05",
-         title = "A | SCZ kME-diagnosis \nassociations"
-    ) +
-    theme(legend.position = "none",
-          axis.text.y = element_text(size = 10)
-    )
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
-)
-```
-
-### B | Significant SCZ module enrichments
-
-```{r FigS11b.prep}
-## Identify significant modules
-sig_modules <- df_lm_dx %>% 
-    dplyr::filter(p_value < 0.05 & term == "SCZ") %>% 
-    pull(module)
-
-# Filter for modules of interest & take top 5 pathways for each
-df_figS11b <- df_mods_go %>% 
-    dplyr::filter(module %in% sig_modules) %>% 
-    group_by(module) %>% 
-    top_n(5, wt = -log10(pvalue)) %>% 
-    mutate(Description = str_wrap(Description, width = 40))
-```
-
-```{r FigS11b.bubbleplot, fig.height = 5, fig.width = 3.5}
-df_figS11b %>% 
-    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
-    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
-    facet_wrap(vars(module), scales = "free_y", ncol = 1) +
-    scale_fill_manual(values = module_colors, guide = "none") +
-    scale_y_reordered() +
-    scale_x_reverse(breaks = c(0.015, 0.00)) +
-    scale_size_continuous(range = c(1, 4)) +
-    coord_cartesian(clip = "off") +
-    labs(x = "Enrichment FDR", y = NULL,
-         title = "B | SCZ module enrichments") +
-    theme(legend.position = c(-1.9, 0.85))
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S11B.sig_SCZ_kME_modules", .x), width = 3.5, height = 5)
-)
-```
-
-# FIG S12: WGCNA module enrichment for risk gene lists #
-
-```{r FigS12.prep}
-df_figS12 <- df_mods_risk_gene_hypergeometric %>% 
-    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
-    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
-               str_trim %>% 
-               str_replace(", ", ",\n")
-    ) %>% 
-    mutate(sig = case_when(p_adj < 0.001 ~ "***", p_adj < 0.01 ~ "**", p_adj < 0.05 ~ "*", TRUE ~ ""))
-```
-
-```{r FigS12.plot, fig.width = 5, fig.height = 4}
-df_figS12 %>% 
-    ggplot(aes(x = benchmark_list, y = module)) +
-    geom_tile(aes(fill = module, alpha = -log10(p_adj))) +
-    geom_text(aes(label = sig), size = 6, color = "black", vjust = 0.75) +
-    facet_wrap(vars(class), scales = "free_x", nrow = 1) +
-    scale_fill_manual(values = module_colors, guide = "none") +
-    guides(alpha = guide_legend(title = "-log10(FDR)")) +
-    labs(x = NULL, y = NULL, 
-         title = "Risk gene list enrichment in WGCNA modules")
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S12.risk_gene_module_hypergeometric", .x), width = 5, height = 4)
-)
-```
-
-# FIG S13: Functional enrichment of GRCCA overrepresented WGCNA modules #
-
-### A | Hypergeometric overlap of WGCNA modules and significant GRCCA genes
-
-```{r FigS13a.prep}
-## Identify significant GRCCA genes (FDR < 0.05 & |z| > 2)
-sig_grcca_genes <- df_x_res %>% 
-    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
-    pull(ensembl_gene_id)
-# n = 1211
-
-## Hypergeometric test for significant overlap with modules
-df_grcca_module_hypergeometric <- map_dfr(
-    .x = names(module_colors),
-    .f = ~ f_module_hypergeometric(
-        gene_list = sig_grcca_genes, 
-        gene_list_name = "", 
-        wgcna_module = .x, 
-        gene_universe = ensembl_gene_ids
-    ) %>% 
-        dplyr::select(-gene_list)
-) %>% 
-    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>% 
-    arrange(p_adj) %>% 
-    mutate(module = factor(module, levels = names(module_colors))) %>%
-
-    # specify modules to label
-    mutate(module_label = ifelse(p_adj < 0.05, paste0("N = ", overlap_n), ""))
-```
-
-```{r FigS10a.plot, fig.height = 5, fig.width = 2.15}
-f_plot_module_hypergeometric(df_grcca_module_hypergeometric %>% mutate(p_adj = ifelse(p_adj == 0, 10^-5, p_adj)) %>% dplyr::filter(module != "geneM0"),
-                             x_axis = "p-adj",
-                             plot_title = "A | GRCCA significant modules",
-                             #subset_x_labs = TRUE,
-                             include_module_labels = TRUE,
-                             module_text_size = 3,
-                             legend_position = "none",
-                             p_value_threshold = 0.05,
-                             include_threshold_label = TRUE
-)
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S13A.GRCCA_module_hypergeometric", .x), width = 2.15, height = 5)
-)
-```
-
-### B | Direction of significant genes in top modules
-
-```{r FigS13b.prep}
-## Identify significant modules
-sig_modules <- df_grcca_module_hypergeometric %>% dplyr::filter(module != "geneM0" & p_adj < 0.05) %>% pull(module) %>% sort()
-
-## Identify panel sizes (number of significant genes in each module)
-panel_widths <- df_x_res %>% 
-    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
-    dplyr::count(module) %>% 
-    pull(n)
-```
-
-```{r FigS13b.plot, fig.height = 5, fig.width = 2.15}
-df_x_res %>% 
-    dplyr::filter(module %in% sig_modules & ensembl_gene_id %in% sig_grcca_genes) %>% 
-    mutate(weight = ifelse(significant == 1, weight, 0)) %>%
-
-    ggplot(aes(x = reorder_within(ensembl_gene_id, within = module, by = pearsons_r), y = pearsons_r)) +
-    geom_col(aes(fill = module, alpha = abs(pearsons_r))) +
-    facet_wrap(vars(module), ncol = 1, scales = "free_y") +
-    force_panelsizes(rows = panel_widths) +
-    scale_fill_manual(values = module_colors) +
-    scale_alpha_continuous(range = c(0.3, 1)) +
-    labs(title = "B | Significant module genes",
-         x = "Gene", y = expression("Structure correlation (" * italic(r)[x] * ")")) +
-    guides(fill = guide_legend(nrow = 2)) +
-    coord_flip() +
-    theme(axis.text.y = element_blank(),
-          axis.ticks.y = element_blank(),
-          strip.text = element_blank(),
-          legend.position = "none",
-          panel.spacing = unit(0.3, "lines")
-    )
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S13B.sig_module_direction", .x), width = 2.15, height = 5)
-)
-
-```
-
-### C | GO BP enrichments of significant modules
-
-```{r FigS13c.prep}
-# Filter for GRCCA modules of interest & take top 5 pathways for each
-df_figS13c <- df_mods_go %>% 
-    dplyr::filter(module %in% sig_modules) %>% 
-    group_by(module) %>% 
-    top_n(5, wt = -log10(pvalue)) %>% 
-    mutate(Description = str_wrap(Description, width = 23.5))
-```
-
-```{r FigS13c.bubbleplot, fig.height = 5, fig.width = 2.15}
-df_figS13c %>% 
-    ggplot(aes(x = p.adjust, y = reorder_within(Description, -p.adjust, module))) +
-    geom_point(aes(size = RichFactor, fill = module), shape = 21) +
-    facet_wrap(vars(module), scales = "free_y", ncol = 1) +
-    scale_fill_manual(values = module_colors, guide = "none") +
-    scale_y_reordered() +
-    scale_x_reverse(breaks = c(0.00005, 0.00)) +
-    scale_size_continuous(range = c(1, 4)) +
-    coord_cartesian(clip = "off") +
-    labs(x = "Enrichment FDR", y = NULL,
-         title = "C | Module enrichments") +
-    theme(
-        legend.position = c(-1.6, 1.05),
-        legend.key.size = unit(0.2, "cm"),
-        axis.text = element_text(size = 8.5)
-    )
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S13C.sig_GRCCA_module_enrichments", .x), width = 2.15, height = 5)
-)
-```
-
-# FIG S14: Compare GRCCA and RCCA directly #
-
-### A | Correlate GRCCA and RCCA covariate (y) structure correlations
-
-```{r FigS14a.plot}
-df_y_res_all %>% 
-    pivot_wider(id_cols = covariate, names_from = analysis, values_from = pearsons_r) %>% 
-    ggplot(aes(x = rcca, y = grcca)) +
-    geom_point(aes(fill = rcca + grcca), shape = 21, size = 2.5) +
-    geom_vline(aes(xintercept = 0), color = "gray") +
-    geom_hline(aes(yintercept = 0), color = "gray") +
-    geom_abline(lty = 2) +
-    geom_text_repel(aes(label = covariate), size = 2.5, max.overlaps = 20, box.padding = 1.25,
-                     nudge_y = 0.3, force = 150) +
-    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.84, 1.84), guide = "none") +
-    labs(title = "A | Covariate structure correlations",
-         x = "RCCA", y = "GRCCA")
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S14A.covariate_structCor_correlation", .x), width = fig_width, height = fig_height)
-)
-```
-
-### B | Correlate GRCCA and RCCA gene (x) structure correlations
-
-```{r FigS14b.plot}
-df_x_res_all %>% 
-    pivot_wider(id_cols = ensembl_gene_id, names_from = analysis, values_from = pearsons_r) %>% 
-    ggplot(aes(x = rcca, y = grcca)) +
-    geom_point(aes(color = rcca + grcca), size = 1) +
-    geom_vline(aes(xintercept = 0), color = "gray") +
-    geom_hline(aes(yintercept = 0), color = "gray") +
-    geom_abline(lty = 2) +
-    stat_cor(label.sep = "\n") +
-    #geom_label_repel(aes(label = gene_symbol), size = 3, max.overlaps = 50, box.padding = 1.5) +
-    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-1.14, 1.14), guide = "none") +
-    labs(title = "B | Gene structure correlations",
-         x = "RCCA", y = "GRCCA")
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S14B.gene_structCor_correlation", .x), width = fig_width, height = fig_height)
-)
-```
-
-### C | Venn diagram of significant gene overlap (made in PowerPoint/Inkscape)
-
-```{r FigS14c.prep}
-## Venn diagram (overlap with Nirmala's results)
-sig_rcca_genes <- df_x_res_rcca %>% 
-    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
-    pull(ensembl_gene_id)
-length(sig_grcca_genes) # N = 1211
-length(sig_rcca_genes) # N = 2026
-hypergeometric_overlap <- f_hypergeometric(sig_grcca_genes, sig_rcca_genes, gene_universe = ensembl_gene_ids)
-length(hypergeometric_overlap$overlap_genes) # N = 855; P < 0.0001
-```
-
-### D | RCCA results risk gene enrichment
-
-```{r FigS14d.prep}
-df_fig13d <- df_rcca_fgsea_benchmark %>% 
-    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
-    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
-               str_trim %>% 
-               str_replace(", ", ",\n")
-    ) %>% 
-    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
-    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
-    mutate(label = round(NES, 2) %>% format(digits = 3))
-    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
-   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))
-```
-
-```{r FigS14d.plot}
-df_fig13d %>% 
-    ggplot(aes(x = benchmark_list, y = 1)) +
-    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
-    geom_text(aes(label = label, color = I(label_color)), size = 3) +
-    facet_wrap(vars(class), scales = "free", nrow = 2) +
-    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
-    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
-    labs(x = NULL, y = NULL,
-         title = "D | Risk gene enrichment") +
-    theme(axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          axis.line = element_blank(),
-          axis.text.x = element_text(margin = margin(t = -3)),
-          strip.background = element_rect(fill = "white", color = "black"),
-          legend.margin = margin(r = -15),
-          legend.position = "left",
-          legend.justification = "top",
-          legend.title = element_text(angle = 90),
-          legend.key.height = unit(0.75, "cm"),
-          legend.key.width = unit(0.15, "cm")
-    )
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S14D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
-)
-```
-
-# GRAVEYARD #
-
-## Fig S4B alt: Module stability (alluvial)
-
-```{r FigS4b.alluvial, fig.width = 6.5, fig.height = 4}
-df_summary <- df_modules %>%
-  group_by(min_size, cut_height, module, color) %>%
-  summarise(n_genes = n(), .groups = "drop")
-
-# Create the alluvial plot
-library(ggalluvial)
-df_summary %>% 
-    mutate(
-        cut_height = factor(cut_height),
-        min_size = factor(min_size)
-    ) %>%
-
-    ggplot(aes(x = min_size, y = n_genes, stratum = module, alluvium = module, fill = color)) +
-    geom_flow(stat = "alluvium", lode.guidance = "frontback", aes(fill = color)) +
-    geom_stratum(width = 0.2, color = "black") +
-    scale_fill_identity() +  # Use the hex codes from the 'color' column
-    facet_grid(cols = vars(cut_height)) +  # Facet into rows by min_size
-    theme_minimal() +
-    labs(x = "Cut Height", 
-         y = "Number of Genes",
-         title = "Module stability across parameters") +
-    theme(
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        strip.text = element_text(size = 10)
-    )
-```
-
-## Fig SXX: Full WGCNA module functional enrichments
-
-```{r FigS5.prep}
-df_figS5 <- df_mods_go %>% 
-    clean_names %>% 
-    dplyr::filter(!is.na(description) & module != 0) %>%
-
-    # take top 5 paths by p-value per module to plot
-    dplyr::group_by(module) %>% 
-    arrange(module, -p_adjust) %>% 
-    mutate(row = row_number(),
-           gene_ratio = parse(text = gene_ratio) %>% eval
-    ) %>% 
-    top_n(n = 4, wt = row) %>% 
-    #slice_max(order_by = -weight_fisher, n = 5) %>% 
-    mutate(`survives FDR` = ifelse(p_adjust < 0.05, "yes", "no")) %>% 
-    mutate(description = str_wrap(description, width = 10))
-```
-
-```{r FigS5.plot, fig.height = 8.75, fig.width = 6.5}
-df_figS5 %>% 
-    ggplot(aes(x = p_adjust, y = reorder_within(description, within = module, by = -p_adjust))) +
-    #y = reorder_within(str_wrap(description, width = 35), -pvalue, module))) +
-    geom_point(aes(size = gene_ratio, fill = ontology, color = `survives FDR`),
-               shape = 21, stroke = 1) +
-    geom_vline(aes(xintercept = 0.05), linewidth = 0.75, color = "gray") +
-    facet_wrap(vars(module), scales = "free", ncol = 5) +
-    scale_size_continuous(range = c(0.5, 3)) +
-    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
-    scale_y_reordered() +
-    scale_x_reverse(breaks = c(0.25, 0.15, 0.05, 0)) +
-    coord_cartesian(clip = "off") +
-    labs(y = NULL, x = "Enrichment FDR",
-         title = "WGCNA module GO enrichments"
-    ) +
-    guides(size = guide_legend(title = "Gene ratio")) +
-    theme(legend.position = c(0.85, 0.0),
-          legend.direction = "vertical",
-          legend.title.position = "top",
-          #legend.justification = "center",
-          strip.background = element_rect(fill = "white"),
-          strip.clip = "off",
-          axis.text = element_text(size = 7.0)
-    )
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S5.full_GO_enrichments", .x), width = 6.5, height = 8.75)
-)
-```
-
-Figure SXX.
-Full GO enrichments of gene-level modules.
-Top 4 Gene Ontology pathways associated with each module, determined by p-value, across all three ontologies (BP = biological process, CC = cellular component, MF = molecular function).
-The x-axis shows the enrichment FDR, while the y-axis shows pathway names.
-Each point is colored by its ontology and sized by the gene ratio (number of genes in the module in the pathway divided by the total number of genes in the pathway); points that survived correction for multiple comparisons (FDR < 0.05; vertical gray line) are outlined in black.
-
-## Fig S10B alternate (word clouds)
-
-```{r FigS10b.wordcloud, fig.height = 5}
-## Identify uninteresting words
-data(stop_words)
-vague_terms <- c("positive", "negative", "regulation", "response", "protein", "activity",
-                 "pathway", "process", "involved", "signaling")
-
-## Identify top words in each significant (FDR < 0.05) module pathway
-module_tokens_freq <- df_sig_go_res %>% 
-    dplyr::filter(p.adjust < 0.05) %>% 
-    group_by(module) %>% 
-    unnest_tokens(word, Description) %>% 
-    anti_join(stop_words) %>% 
-    dplyr::filter(!(word %in% vague_terms)) %>% 
-    dplyr::count(word, sort = TRUE)
-
-## Create wordcloud plot for each module
-l_figS10b <- map(
-    .x = as.character(sig_modules),
-    .f = ~ module_tokens_freq %>% 
-        dplyr::filter(module == .x) %>% 
-        top_n(10, wt = n) %>% 
-        ggplot(aes(label = word, size = n, color = module)) +
-        geom_text_wordcloud() +
-        facet_wrap(vars(module)) +
-        scale_color_manual(values = module_colors) +
-        scale_size_continuous(range = c(2.5, 5)) +
-        theme_void() +
-        theme(strip.text = element_blank(),
-              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
-        )
-)
-
-## Combine and save
-wrap_plots(l_figS10b, ncol = 1) & plot_annotation(title = "B | SCZ module enrichments")
-map(
-  .x = c(".png", ".pdf"),
-  .f = ~ ggsave(paste0(supplement_figures_dir, "S10B.sig_SCZ_kME_modules", .x), width = fig_width, height = 5)
-)
-```
-
-GO biological process pathway enrichments for each module.
-The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module.
-Words are colored by respective module.
-
-## Fig S12C alternate (word clouds)
-```{r FigS12c.prep}
-## Load module GO enrichment data
-#load(paste0(base_dir, "objects/", prefix, sep = "_", "SIGNED_SFT", soft_power, "_GO_RES.RDS")) # df_mods_go
-
-## Identify words to not include
-data(stop_words)
-vague_terms <- c("positive", "negative", "regulation", "response", "activity",
-                 "pathway", "process", "involved", "signaling")
-
-# Filter for modules of interest
-df_go_res <- df_mods_go %>% 
-    dplyr::filter(module %in% sig_modules)
-
-## Identify top words in each significant (FDR < 0.05) module pathway
-module_tokens_freq <- df_go_res %>% 
-    dplyr::filter(p.adjust < 0.05) %>% 
-    group_by(module) %>% 
-    unnest_tokens(word, Description) %>% 
-    anti_join(stop_words) %>% 
-    dplyr::filter(!(word %in% vague_terms)) %>% 
-    dplyr::count(word, sort = TRUE)
-```
-
-```{r FigS12c.wordcloud}
-l_module_wordclouds <- map(
-    .x = as.character(sig_modules),
-    .f = ~ module_tokens_freq %>% 
-        dplyr::filter(module == .x) %>% 
-        top_n(10, wt = n) %>% 
-        ggplot(aes(label = word, size = n, color = module)) +
-        geom_text_wordcloud() +
-        facet_wrap(vars(module)) +
-        scale_color_manual(values = module_colors) +
-        scale_size_continuous(range = c(2.5, 5)) +
-        theme_void() +
-        theme(strip.text = element_blank(),
-              plot.margin = margin(t = -75, r = -10, b = -75, l = -10)
-        )
-)
-```
-
-GO biological process pathway enrichments for each module.
-The top 10 words per module are shown, and the size indicates the number of times the word appears in a GO term significantly associated with that module.
-Words are colored by respective module.
-
 ## Fig SXX: Cell-type enrichment of DEGs

### 2025-07-02 functions/f_load_CCA_res.R

@@ -20,7 +20,8 @@ f_load_cca_res <- function(cca_directory = cca_dir,
                            lambda = "Nfeat",
                            VARx = c("0.1_1", "0.99"),
                            include_Cmat = c(TRUE, FALSE),
-                           rename_covariates = c(TRUE, FALSE)
+                           rename_covariates = c(TRUE, FALSE),
+                           all_effects = c(TRUE, FALSE)
                            ) {
     
 ### Set paths & data split to load results ###
@@ -30,7 +31,8 @@ f_load_cca_res <- function(cca_directory = cca_dir,
     ## Identify correct framework analysis directory
     if (include_Cmat) {cmat <- "withCmat"} else {cmat <- "noCmat"}
     if (analysis_type == "grcca") {group <- paste0("_groupL2x", as.character(mu))} else {group <- ""}
-    analysis_dir <- paste0(analysis_type, "_permutation_VARx", VARx, "_L2x", lambda, group, "_", cmat, "_allEffects")
+    if (all_effects == TRUE) {analysis_dir <- paste0(analysis_type, "_permutation_VARx", VARx, "_L2x", lambda, group, "_", cmat, "_allEffects")
+    } else {analysis_dir <- paste0(analysis_type, "_permutation_VARx", VARx, "_L2x", lambda, group, "_", cmat)}
     
     ## Combine paths to point to results (and check to make sure directory exists)
     results_dir <- paste0(cca_dir, "framework/", analysis_dir)

### 2025-07-02 functions/hypergeometric_functions.R

@@ -1,4 +1,58 @@
 
+
+# Calculate hypergeometric odds ratio -------------------------------------
+
+#=====================================================================================================================#
+
+### GOAL: Calculate the hypergeometric overlap odds ratio (for use in hypergeometric function, below)
+
+### INPUTS ###
+
+## q = Overlap size minus 1
+## m = Size of Set 1
+## n = Complement of Set 1 size (N - m)
+## k = Size of Set 2
+
+### OUTPUTS ###
+
+## odds_ratio = likelihood of overlap, given the respective population sizes
+
+#=====================================================================================================================#
+
+
+calculate_odds_ratio <- function(q, m, n, k) {
+    
+    # Total population size
+    N <- m + n  # Since n = N - m by definition
+    
+    # Check for invalid inputs
+    if (q + 1 > m || q + 1 > k || q + 1 > N) {
+        stop("Invalid values: Overlap q + 1 cannot exceed set or population sizes.")
+    }
+    if (m > N || k > N) {
+        stop("Invalid values: Set sizes cannot exceed population size.")
+    }
+    
+    # 2x2 table components
+    a <- q
+    b <- k - q
+    c <- m - q
+    d <- N - (a + b + c)
+    
+    # Add pseudocount to avoid zero division
+    a <- a + 0.5
+    b <- b + 0.5
+    c <- c + 0.5
+    d <- d + 0.5
+    
+    # Odds ratio formula
+    odds_ratio <- (a * d) / (b * c)
+    return(odds_ratio)
+    
+}
+
+
+
 # General hypergeometric function -----------------------------------------
 
 #=====================================================================================================================#
@@ -15,6 +69,7 @@
 
 ## overlap_genes = list of genes that were in both gene_list1 and gene_list2
 ## p_value = hypergeometric p-value (significant of overlap given total pool of genes)
+## odds_ratio = likelihood of overlap, given the respective population sizes
 
 #=====================================================================================================================#
 
@@ -27,15 +82,31 @@ f_hypergeometric <- function(gene_list1, gene_list2, gene_universe) {
     # Identify the overlap between the published gene list (including only those in our universe) and the gene list of interest
     overlap_genes <- intersect(gene_list1_in_universe, gene_list2_in_universe)
     
+    # Determine variables for hypergeometric test
+    q <- length(overlap_genes)
[…]

### 2025-07-02 scripts/00.preprocessing/00.clean_covariates.R

@@ -10,6 +10,7 @@
 
 library(tidyverse)
 library(readxl)
+library(janitor)
 
 
 # data --------------------------------------------------------------------
@@ -19,7 +20,7 @@ base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
 # complete covariates from Ajeet
 df_complete_covariates <- read_xlsx(paste0(base_dir, "data/14Nov2023_all_covariates.xlsx"))
 
-# other covariates from Nirmala (includes additional technical covariates)
+# other covariates from Nirmala (includes additional technical covariates, like sgACC RIN)
 df_nirmala_covariates <- read.table(paste0(base_dir, "data/185samples_multinomialDx_46covariates_10evecs.txt"), 
                                     sep = "\t", header = TRUE) %>% 
   dplyr::select(-contains("evec")) %>% 
@@ -33,79 +34,95 @@ df_nirmala_covariates <- read.table(paste0(base_dir, "data/185samples_multinomia
   mutate(sample = tolower(sample))
 
 # define covariate types
-technical_covariates <- c("source", "pmi", "ph", "pmi_confidence", "max_rin", "max_rine", "mapped_percent", "five_prime_three_prime_bias",
-                          "rna_extraction_batch", "library_batch", "gc_percent")
-biological_covariates <- c("age_death", "sex_at_birth", "race", "marital_status", "education", "manner_death", "suicide", "brain_weight",
-                           "height", "weight", "bmi")
-drug_covariates <- c("smoker", "nicotine_cotinine", "alcohol", "sedative_hypnotic_anxiolitics", "opioids", "cannabinoids",
+technical_covariates <- c("mapped_percent", "gc_percent", "five_prime_three_prime_bias", "rin_acsg", "rna_extraction_batch",
+                          "library_batch", "pmi", "pmi_confidence", "source", "ph" # "max_rin", "max_rine"
+                           )
+biological_covariates <- c("age_death", "sex_at_birth", "race", "bmi", "height", "weight",
+                           "marital_status", "manner_death", "education", "suicide", "brain_weight")
+drug_covariates <- c("smoker", "nicotine_cotinine", "alcohol", "opioids", "cannabinoids", "beta_blockers",
                      "major_stimulants_cocaine_included", "minor_stimulants", "anticholinergics", "antidepressants",
-                     "anti_epileptics", "anti_histamines", "antipsychotics", "mood_stabilizers", "non_psychiatric",
[…]

### 2025-07-02 scripts/01.DGE/01.characterize_DEGs.R

@@ -13,13 +13,15 @@ figures_dir <- paste0(base_dir, "outputs/figures/Fig2_DGEres/")
 analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
 
 source(paste0(base_dir, "scripts/load_genes.R"))
-load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res (generated in A.DGE_analysis.R)
+load(paste0(base_dir, "objects/DE_results.Rdata")) # df_limma_res (generated in 00.DGE_analysis_Limma_DESeq2.R)
 
 
-### Get ranked gene list ###
-my_degs_ranked <- df_de_res %>% 
-    arrange(-stat) %>% 
-    dplyr::select(ensembl_gene_id, log2fold_change) %>% 
+### Get ranked gene list ### (switching to limma!)
+my_degs_ranked <- df_limma_res %>% 
+    arrange(-log_fc) %>%
+    dplyr::select(ensembl_gene_id, log_fc) %>%
+    # arrange(-t) %>%
+    # dplyr::select(ensembl_gene_id, t) %>% 
     deframe
 
 

### 2025-07-02 scripts/01.DGE/DEPR.DGE_analysis.R

@@ -35,7 +35,6 @@ df_covariates_mca <- df_covariates %>%
 
 ## Convert to matrix
 n_mcs <- 8
-paste0("MC", 1:n_mcs)
 m_covariates <- df_covariates_mca %>% 
     dplyr::filter(dx %in% c("SCZ", "Control")) %>% # SCZ vs controls only
     mutate(dx = factor(dx, levels = c("Control", "SCZ"))) %>% 

### 2025-07-02 scripts/02.WGCNA/02.module_eigengene_analysis.R

@@ -34,12 +34,18 @@ df_covariates_mca <- df_covariates_numeric %>%
     ) %>% 
     
     # remove drug covariates (since we have the MCs)
-    dplyr::select(-any_of(drug_covariates))
+    dplyr::select(-any_of(drug_covariates)) %>% 
+    
+    # bring sample to column 1
+    dplyr::select(sample, everything())
 
 
 # Run PCA on each module -------------------------------------------------
 
+
+## Run PCA on samples x gene expression matrix for each module
 df_mods_pca <- df_modules_filt %>% 
+    dplyr::select(module, ensembl_gene_id) %>% 
     
     # combine with sample expression information
     left_join(df_vsd_regress %>% 
@@ -47,56 +53,58 @@ df_mods_pca <- df_modules_filt %>%
               by = join_by(ensembl_gene_id)
     ) %>% 
     
-    # create gene x sample expression matrix for each module
-    pivot_wider(id_cols = c(ensembl_gene_id, module), 
-                names_from = sample, 
-                values_from = resids) %>% 
+    # create sample x gene expression matrix for each module
     group_by(module) %>% 
     nest() %>% 
+    mutate(
+        data = map(
+            .x = data,
+            .f = ~ .x %>% 
+                pivot_wider(id_cols = sample, 
+                            names_from = ensembl_gene_id, 
+                            values_from = resids)
+        )
+    ) %>% 
     
     # run PCA on gene x sample expression matrix
-    mutate(pca = map(.x = data,
-                     .f = ~ .x %>% 
-                         as.data.frame %>%
-                         column_to_rownames("ensembl_gene_id") %>% 
-                         prcomp()
-    )) %>% 
+    mutate(pca = map(
+        .x = data,
+        .f = ~ .x %>% 
+            as.data.frame %>%
+            column_to_rownames("sample") %>% 
+            prcomp(center = TRUE, scale = TRUE)
+    )
+    )
+
+
+## Extract sample scores in PC space
+df_mods_pca_covariates <- df_mods_pca %>% 
     
-    # combine PCA results
-    mutate(sample_pcs = map(.x = pca,
-                            .f = ~.x$rotation %>% 
-                                as.data.frame %>% 
-                                rownames_to_column("sample") %>% 
-                                as_tibble() %>% 
-                                clean_names() %>% 
[…]

### 2025-07-02 scripts/03.CCA_toolkit_setup/00b.define_input_matrices_SENS.R

@@ -1,0 +1,178 @@
+
+######################################################################################
+
+# Define X.mat, Y.mat, and C.mat to include in (G)RCCA analysis
+
+######################################################################################
+
+
+### Setup ###
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+### SELECT samples to include ###
+dx_groups_to_analyze <- c("Control", "BD", "SCZ", "MDD") # select diagnostic groups of interest
+samples_to_include <- df_covariates %>% 
+    filter(dx %in% dx_groups_to_analyze) %>% 
+    pull(sample) %>% 
+    unique
+
+
+### SENSITIVITY ANALYSES: define grouping structure (module set) ###
+
+# In main text, we chose: sft = 3, min_size = 40, cut_height = 0.98
+df_modules %>% count(sft, min_size, cut_height) %>% print(n=nrow(.)) # 33 potential module sets
+# Try:
+    # min_size = 30, cut_height = 0.98
+    # min_size = 30, cut_height = 0.95
+    # min_size = 40, cut_height = 0.95
+    # random: Scramble group label across genes (genes not related to each other, but same group number and size distribution)
+
+# Plot
+df_modules %>% 
+    group_by(sft, min_size, cut_height, module) %>% 
+    count() %>% 
+    filter( min_size %in% c(30, 40) & cut_height %in% c(0.98, 0.95) ) %>% 
+    
+    ggplot(aes(x = module, y = n)) +
+    geom_point() +
+    facet_grid(cut_height ~ min_size)
+
+
+
+######### SELECT NEW MODULE PARAMETERS HERE #############
+
+# Try:
+# min_size = 30, cut_height = 0.98 (withCmat JOBID 49259979; noCmat JOBID 49293558)
+# min_size = 30, cut_height = 0.95 (withCmat JOBID 49260784; noCmat JOBID 49296943)
+# min_size = 40, cut_height = 0.95 (withCmat JOBID 49261067; noCmat JOBID 49298399)
+# random: Scramble group label across genes (genes not related to each other, but same group number and size distribution) (withCmat JOBID 49262317; noCmat JOBID 49289235)
+
+minimum_size <- 40
+tree_cut_height <- 0.98
+
[…]

### 2025-07-02 scripts/04.GRCCA/01.GRCCA_results.R

@@ -50,17 +50,98 @@ df_x_res <- l_grcca_res$x_res %>%
     mutate(gene_symbol = ifelse(is.na(gene_symbol), ensembl_gene_id, gene_symbol))
 
 
+
+# Test the significance of the structure correlation alignment with expression in each diagnostic group ------------------
+
+## Combine structure correlation and expression data
+df_rx_expr <- df_vsd_regress %>% 
+    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expression_value") %>% 
+    mutate(dx = case_when(
+        str_detect(sample, "control") ~ "Control",
+        str_detect(sample, "bipolar") ~ "BD",
+        str_detect(sample, "mdd") ~ "MDD",
+        str_detect(sample, "schizo") ~ "SCZ"
+    ) %>% factor(levels = names(dx_colors))) %>% 
+    group_by(ensembl_gene_id, dx) %>% 
+    summarise(mean_expr = mean(expression_value)) %>% 
+    left_join(df_x_res)
+
+
+## Calculate the actual (observed) correlation between rx and mean expression in each diagnostic group
+df_rx_expr_cor <- df_rx_expr %>% 
+    group_by(dx) %>% 
+    group_modify( ~ {
+        
+        test <- cor.test(.x$mean_expr, .x$pearsons_r)
+        
+        tibble(
+            correlation = test$estimate,
+            p_value = test$p.value
+        )
+        
+    }) %>% 
+    ungroup() %>% 
+    mutate(p_adj = p.adjust(p_value, method = "BH"))
+
+## Permute the rx vector and re-calculate the correlation
+n_perm <- 10000
+df_rx_expr_cor_perm <- tibble()
+set.seed(0206)
+for (i in 1:n_perm) {
+    
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+    
+    # 1: Shuffle structure correlations within module
+    df_perm <- df_rx_expr %>%
+        group_by(module) %>%
+        mutate(permuted_r = sample(pearsons_r, replace = FALSE)) %>%
+        ungroup()
+    
+    # For each diagnostic group, correlate permuted_r with mean_expr
+    df_tmp <- df_perm %>% 
+        group_by(dx) %>% 
+        group_modify( ~ {
+            
+            test <- cor.test(.x$mean_expr, .x$permuted_r)
+            
[…]

### 2025-07-02 scripts/04.GRCCA/03b.compare_to_DEGs.R

@@ -1,0 +1,65 @@
+
+#########################################################################################
+
+# Determine the relationship between DEG test statistics and GRCCA structure correlations
+
+#########################################################################################
+
+
+## Set paths and load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig4_DGEres/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "/scripts/setup.R"))
+
+
+## Combine GRCCA and DEG results
+df_de_grcca <- df_x_res %>% 
+    left_join(df_limma_res, by = join_by(ensembl_gene_id, gene_symbol))
+
+## Calculate the observed correlation between rx and DEG t stat
+cor.test(df_de_grcca$pearsons_r, df_de_grcca$log_fc)
+
+
+## Run permutation testing (resampling within module) to determine the significance of the correlation
+n_perm <- 10000
+df_limma_grcca_cor_perm <- tibble()
+set.seed(0406)
+for (i in 1:n_perm) {
+    
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+    
+    # Permute LFC within module
+    df_tmp <- df_de_grcca %>% 
+        group_by(module) %>%
+        mutate(permuted_r = sample(pearsons_r, replace = FALSE))
+    
+    # Calculate correlation between permuted data and Nirmala's LFC
+    test <- cor.test(df_tmp$permuted_r, df_tmp$log_fc)
+    correlation <- test$estimate
+    p_value <- test$p.value
+    
+    # Append to tibble
+    df_limma_grcca_cor_perm <- df_limma_grcca_cor_perm %>% 
+        bind_rows(
+            tibble(
+                perm = i,
+                correlation = correlation,
+                p_value = p_value
+            )
+        )
+    
+}
+
+
+## Calculate permutation p-values
+de_limma_grcca_perm_p <- df_limma_grcca_cor_perm %>% 
+    mutate(obs_corr = cor.test(df_de_grcca$pearsons_r, df_de_grcca$log_fc)$estimate) %>%
+    summarise(
+        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
+    )
+# Pperm = 1*10^-4
+
+

### 2025-07-02 scripts/04.GRCCA/05.group_structure_sensitivity.R

@@ -1,0 +1,298 @@
+
+#######################################################################################
+
+# GENE-LEVEL: Sensitivity analyses of robustness of GRCCA results to grouping structure
+
+#######################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+
+# Load sensitivity GRCCA data ---------------------------------------------
+
+sensitivity_dir <- paste0(base_dir, "objects/GRCCA_results_objects/")
+sensitivity_objs <- list.files(paste0(base_dir, "objects/GRCCA_results_objects/"))
+
+
+## Real WGCNA groupings
+wgcna_groups <- sensitivity_objs[!str_detect(sensitivity_objs, "seed")]
+
+l_wgcna_groups_res <- list()
+for (obj in wgcna_groups){
+    
+    ## Identify WGCNA parameters used
+    parameters <- str_extract(obj, "sft[^_]*_minSize[^_]*_cutHeight[^_]*")
+    minimum_size <- str_extract(parameters, "(?<=minSize)\\d+") %>% as.numeric
+    tree_cut_height <- str_extract(parameters, "(?<=cutHeight)\\d+\\.\\d+") %>% as.numeric
+    
+    ## Identify respective module assignments
+    df_modules_filt_sens <- df_modules %>% 
+        filter(min_size == minimum_size & cut_height == tree_cut_height) %>% 
+        unite(mod_set, c(sft, min_size, cut_height), sep = "_") %>% 
+        mutate(module = paste0("geneM", module) %>% factor(levels = paste0("geneM", module) %>% unique))
+    
+    ## Load GRCCA results
+    load(paste0(sensitivity_dir, obj))
+    
+    ## Negate weights/structure correlations if SCZ is negative (want everything to be the same sign)
+    SCZ_r <- df_y_res %>% filter(covariate == "SCZ") %>% pull(pearsons_r)
+    
+    if (SCZ_r < 0) {
+        df_lvs <- df_lvs %>% mutate_at(vars(c(lvx, lvy)), ~ -.x)
[…]

### 2025-07-02 scripts/05.GRCCA_transcripts/02a.GRCCA_transcript_results_subset.R

@@ -1,0 +1,79 @@
+
+########################################################################################
+
+# Load transcript-level GRCCA results for subset results
+# Include: transcript derivatives of 1) common variant-associated genes of any psychiatric
+# disorder, or 2) genes significantly associated with the latent variable in the gene-
+# level GRCCA (at FDR < 0.05)
+
+########################################################################################
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WTCNA_res.R"))
+
+
+# Load data ---------------------------------------------------------------
+
+# 08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1_sft2_minSize35_cutHeight0.988_ControlBDMDDSCZ_8MCA_ControlBDMDDSCZ_regressBrainWeight_WITHGRAY
+# 08Mar2024_TRANSCRIPTS_qSVAgeSexRaceGC_CVq1_sft2_minSize35_cutHeight0.988_ControlBDMDDSCZ_8MCA_regressDim2_WITHGRAY
+
+
+## Identify CCA directory
+dx_groups_to_analyze <- c("Control", "BD", "MDD", "SCZ") # select diagnostic groups of interest
+n_mca_dim <- 8
+type <- "OVERALLsubs/"
+project_dir <- paste0(prefix, # date & preprocessing info
+                      "_sft", soft_power, "_minSize", minimum_size, "_cutHeight", tree_cut_height, "_", # WTCNA parameters
+                      paste0(dx_groups_to_analyze, collapse = ""), "_", # diagnostic groups included
+                      n_mca_dim, "MCA_regressDim2_", # covariates included
+                      "broadRareSCZBDMDDASD_sigGRCCAfdr0_05/" # define subset of transcripts included (based on gene-level results)
+                      #"WITHGRAY" # includes all transcripts
+)
+cca_dir <- paste0(base_dir, "RCCA_toolkit/", type, project_dir)
+
+## Read in GRCCA analysis results
+analysis_type <- "grcca"
[…]

### 2025-07-02 scripts/05.GRCCA_transcripts/03b.GRCCA_transcript_gene_cor.R

@@ -1,0 +1,75 @@
+
+############################################################################################
+
+# Determine the relationship between GRCCA transcript- and gene-level structure correlations
+
+############################################################################################
+
+
+## Set paths and load data
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptRes/")
+tables_dir <- paste0(base_dir, "outputs/tables/")
+analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "/scripts/setup.R"))
+
+
+## Combine GRCCA gene and transcript results
+df_grcca_tx_gene <- df_x_res_transcripts %>% 
+    #filter(!is.na(ensembl_gene_id)) %>% 
+    group_by(ensembl_gene_id) %>% 
+    summarise(pearsons_r = max(pearsons_r)) %>% 
+    #summarise(pearsons_r = ifelse(pearsons_r < 0, min(pearsons_r), max(pearsons_r))) %>% 
+    dplyr::rename("transcript_res" = "pearsons_r") %>% 
+    
+    # add gene level results
+    left_join(df_x_res_genes %>% 
+                  dplyr::select(module, ensembl_gene_id, gene_symbol, pearsons_r) %>% 
+                  dplyr::rename("gene_res" = "pearsons_r")
+    )
+
+## Calculate the observed correlation between rx and DEG t stat
+cor.test(df_grcca_tx_gene$transcript_res, df_grcca_tx_gene$gene_res)
+
+
+## Run permutation testing (resampling within module) to determine the significance of the correlation
+n_perm <- 10000
+df_grcca_tx_gene_cor_perm <- tibble()
+set.seed(0206)
+for (i in 1:n_perm) {
+    
+    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}
+    
+    # Permute transcript rx within module
+    df_tmp <- df_grcca_tx_gene %>% 
+        group_by(module) %>% 
+        mutate(permuted_r = sample(transcript_res, replace = FALSE))
+    
+    # Calculate correlation between permuted data and Nirmala's LFC
+    test <- cor.test(df_tmp$permuted_r, df_tmp$gene_res)
+    correlation <- test$estimate
+    p_value <- test$p.value
+    
+    # Append to tibble
[…]

### 2025-07-02 scripts/setup.R

@@ -24,6 +24,8 @@ library(ggh4x)
 library(ggpubr)
 library(ggwordcloud)
 library(ggvenn)
+library(ggarchery)
+library(GGally)
 library(cowplot)
 library(pheatmap)
 library(RColorBrewer)
@@ -39,6 +41,7 @@ library(org.Hs.eg.db)
 library(GO.db)
 library(EnsDb.Hsapiens.v79) # hg38
 library(ensembldb)
+library(gtexr)
 
 library(rhdf5)
 
@@ -50,12 +53,12 @@ select <- dplyr::select
 count <- dplyr::count
 
 
-# Load data ----add_count_()# Load data ---------------------------------------------------------------
+# Load data ---------------------------------------------------------------
 
 
 ### CURRENT STUDY DATA ###
 
-load(paste0(base_dir, "objects/14Nov2023_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
+load(paste0(base_dir, "objects/22Jan2025_covariates.Rdata")) # df_covariates, df_covariates_numeric (generated in clean_covariates.R)
 load(paste0(base_dir, "objects/drug_MCA_results.Rdata")) # df_ind_loadings, df_var_loadings
 
 
@@ -148,14 +151,21 @@ benchmarking_lists <- list(
 
 # Group covariates by type ------------------------------------------------
 
-technical_covariates <- c("source", "pmi", "ph", "pmi_confidence", "max_rin", "max_rine", "mapped_percent", "five_prime_three_prime_bias",
-                          "rna_extraction_batch", "library_batch", "gc_percent")
-biological_covariates <- c("age_death", "sex_at_birth", "race", "marital_status", "education", "manner_death", "suicide", "brain_weight",
-                           "height", "weight", "bmi")
-drug_covariates <- c("smoker", "nicotine_cotinine", "alcohol", "sedative_hypnotic_anxiolitics", "opioids", "cannabinoids",
-                     "major_stimulants_cocaine_included", "minor_stimulants", "anticholinergics", "antidepressants",
-                     "anti_epileptics", "anti_histamines", "antipsychotics", "mood_stabilizers", "non_psychiatric",
-                     "other_psychotropic_drug", "benzos", "beta_blockers")
+
+technical_covariates <- c(
+    "mapped_percent", "gc_percent", "five_prime_three_prime_bias", "rin_acsg", "rna_extraction_batch",
[…]

### 2025-09-29 scripts/00.preprocessing/05.R2-MCAv2.R
Round 2 reviews: R3C3 cross-disorder convergence

@@ -1,0 +1,6 @@
+
+###############################################################################
+
+# Rerun MCA treating missing variables
+
+###############################################################################

### 2025-09-29 scripts/03.CCA_toolkit_setup/00c.define_input_matrices_COMP.R

@@ -1,0 +1,132 @@
+
+######################################################################################
+
+# Define X.mat, Y.mat, and C.mat to include in (G)RCCA analysis -- COMPOSITE CASE
+
+######################################################################################
+
+
+### Setup ###
+
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+### SELECT samples to include ###
+dx_groups_to_analyze <- c("Control", "BD", "SCZ", "MDD") # select diagnostic groups of interest
+samples_to_include <- df_covariates %>% 
+    filter(dx %in% dx_groups_to_analyze) %>% 
+    pull(sample) %>% 
+    unique
+
+
+
+### X.MAT ###
+
+# define gene order to align modules & genes
+df_order <- df_modules_filt %>% 
+    dplyr::select(ensembl_gene_id, module) %>% 
+    distinct() %>% 
+    arrange(module, ensembl_gene_id) #%>% 
+    
+    # remove gray module!
+    #filter(module != 0)
+
+# X grouping vector    
+x_group <- df_order %>% 
+    mutate(module = as.numeric(as.character(module %>% str_remove("geneM")))) %>% 
+    pull(module)
+
+# X 
+X.mat <- df_vsd_regress %>% 
+    filter(sample %in% samples_to_include) %>% 
+    dplyr::select(sample, all_of(df_order$ensembl_gene_id)) %>% 
+    column_to_rownames("sample")
+sample_order <- rownames(X.mat)
+
+dim(X.mat) # n = 18677
+
+
+
+##### COMPOSITE CASE: Combine SCZ, BD, and MDD into 1 diagnosis column (1 for psych diagnosis, 0 for no known diagnosis) #####
+
+### Y.MAT AND C.MAT
+
+# define number of MCA dimensions to include
+n_mca_dim <- 8
+
+# Define variables to include in Y and C matrices
+YC.mat <- df_covariates_numeric %>% 
+    
+    # include only samples selected
+    filter(sample %in% samples_to_include) %>% 
+    
+    # drugs to keep
+    # dplyr::select(sample, age_death, brain_weight, opioids,
[…]

### 2025-09-29 scripts/04.GRCCA/06.R2-cross_disorder_composite_GRCCA.R

@@ -1,0 +1,436 @@
+
+#######################################################################################
+
+# GENE-LEVEL: Run GRCCA with 1 diagnosis column indicating if the individual was
+# diagnosed with -any- of SCZ, BD, or MDD (1) or no psychiatric diagnosis (0)
+
+#######################################################################################
+
+
+## Load data & functions
+base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
+figures_dir <- paste0(base_dir, "outputs/figures/response_to_reviewers/round2/")
+#tables_dir <- paste0(base_dir, "outputs/tables/")
+#analysis_objects_dir <- paste0(base_dir, "outputs/objects/")
+
+source(paste0(base_dir, "scripts/load_WGCNA_res.R"))
+
+
+
+# Load single disorder and composite case GRCCA results -----------------------------------------------
+
+
+## Identify analysis directories for separate diagnosis runs
+diagnosis_runs <- c("BD", "MDD") #"BDMDD")
+project_dirs <- paste0("08Mar2024_GENES_qSVAgeSexRaceGC_sft3_minSize40_cutHeight0.98_Control", 
+                       diagnosis_runs, 
+                       "_8MCA_regressBrainWeight_withGray/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/GENES/", project_dirs)
+
+
+## Load GRCCA results for each
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res <- 
+    map(
+        .x = cca_dir,
+        .f = ~ f_load_cca_res(
+            cca_directory = .x,
+            level = "gene",
+            analysis_type = analysis_type,
+            mu = 0.1,
+            VARx = VARx,
+            include_Cmat = FALSE,
+            rename_covariates = FALSE,
+            all_effects = TRUE
+        )
+    )
+names(l_grcca_res) <- diagnosis_runs
+
+
+## Load composite analysis results
+project_dirs <- paste0("26Sept2025_COMPOSITE_sft3_minSize40_cutHeight0.98/")
+cca_dir <- paste0(base_dir, "RCCA_toolkit/GENES/", project_dirs)
+analysis_type <- "grcca"
+VARx <- "0.1_1"
+l_grcca_res_comp <- f_load_cca_res(
+    cca_directory = cca_dir,
+    level = "gene",
[…]
